### 2022-01-07 paper/issta22/abstract.tex
abstract

\begin{abstract}
  Evaluating natural language model (\Nlp) on testset does with
  held-out accuracy is limited to show its quality assurance because
  the held-out datasets are often not comprehensive. While the
  behavioral testing over multiple general linguistic capabilities are
  employed, it relies on manually created test cases, and is still
  limited to measure its comprehensive performance for each linguistic
  capability. In this work, we introduce \OurModelName, an \Nlp model
  testing methodology. Given a linguistic capability, The
  \OurModelName finds relevant testcases to test the linguistic
  capability from extisting datasets as seed inputs, generates
  sufficient number of new test cases by fuzzing the seed inputs based
  on their context-free grammar (\Cfg). We illustrate the usefulness
  of the \OurModelName by showing input diversity and identifying
  critical failures in \sota models for \Nlp task. In our experiment,
  we show that the \OurModelName generates {} more test cases with {}
  higher diversity, and finds {} more bugs.

\end{abstract}