### 2022-01-07 paper/issta22/intro.tex

\section{Introduction}
\label{sec:intro}

\Nlp testing is crucial part for developing a reliable \Nlp model. In
\Nlp, testing is mainly used to check the ML model's performance on
\ho set with respect to the accuracy of the model. The \ho data refers
to a portion of dataset that is held out of the datasets used for
training \ml models. Generally, the \ho set is extracted via
Train-Validation-Test split.

%% One of the primary goals of training NLP models is
%% generalization. Since testing “in the wild” is expensive and does not
%% allow for fast iterations, the standard paradigm for evaluation is
%% using trainvalidation-test splits to estimate the accuracy of the
%% model, including the use of leader boards to track progress on a task
%% (Rajpurkar et al., 2016).

%% While performance on held-out data is a
%% useful indicator, held-out datasets are often not comprehensive, and
%% contain the same biases as the training data (Rajpurkar et al., 2018),
%% such that real-world performance may be overestimated (Patel et al.,
%% 2008; Recht et al., 2019). Further, by summarizing the performance as
%% a single aggregate statistic, it becomes difficult to figure out where
%% the model is failing, and how to fix it (Wu et al., 2019).