### 2022-03-03 paper/fse22/conclusion.tex
change conf name

\section{Conclusion}

We present methodology of code structure understanding to learn code
contextual representation using control flow graph. To the best of our
knowledge, this is the first pre-trained model that leverages the
control flow graph. We show that our schemes for feeding graph edge
distance and graph node type into \Trsfm self-attention operation
outperformed the state-of-the-art pretraing model on code-related \ds
tasks. Furthermore, our qualitive analysis shows the performance
improvements results from understanding \Cfg code structure. Our study
stands to motivate other researchers and practitioners to develop code
related software engineering techniques based on deep learning for \pl
to provide powerful tools to software engineers.