### 2022-03-15 paper/ase22/background.tex

\section{Background}
\label{sec:background}

\sw{We should motivate the testing of linguistic capabilities. Give
  background of the work that has been done in CheckList: what
  capabilities they support and their limitations. Maybe find a
  motivating example?}