### 2022-03-15 paper/ase22/related.tex

\section{Related Work}
\label{sec:related}

\textbf{NLP Testing.} With the increasing use of NLP models,
evaluation of NLP models is becoming more important. Wang \textit{et
  al.}~\cite{wang2018glue} propose multiple diagnostic datasets to
evaluate NLP models. Few recent works have also considered model
robustness as an aspect for model evaluation. Different methods like
adversarial set generation~\cite{ribeiro2018sear,belinkov2018breaknmt,
  rychalska2019wildnlp,iyyer2018adversarial}, fairness
evaluation~\cite{prabhakaran2019fairness,rottger2020hatecheck},
logical consistency evaluation~\cite{ribeiro2019consistencyeval},
prediction interpretations~\cite{ribeiroSG16lime} and interactive
error analysis~\cite{wu2019errudite} have been proposed to evaluate
model robustness. More recently, \Chlst introduces input-output
behaviors of linguistic capabilities and generates behavior-guided
inputs for validating the behaviors.~\cite{marcoACL2020checklist}
However, the approach only relies on manually generated input
templates, thus the template generation becomes expensive and time
consuming. Also, it does not guarantee the comprehensive evaluation.

\begin{comment}
Not only that, model is evaluated on different aspects such as
robustness of the model on adversarial
sets~\cite{ribeiro2018sear,belinkov2018breaknmt,
  rychalska2019wildnlp,iyyer2018adversarial},
fairness~\cite{prabhakaran2019fairness,rottger2020hatecheck}, logical
consistancy~\cite{ribeiro2019consistencyeval}, prediction
interpretations~\cite{ribeiroSG16lime} and interactive error
analysis~\cite{wu2019errudite}. Especially, \Chlst implements
behavioral testing methodolgy for evaluating multiple linguistic
capabilities of NLP model~\cite{marcoACL2020checklist}. \Chlst
introduces input-output behaviors of linguistic capabilities and
generates behavior-guided inputs for validating the behaviors. It
provides comprehensive behavioral testing of NLP models through a
number of generated inputs. However, the approach only relies on
manually generated input templates, thus the template generation
becomes expensive and time consuming. In addition, the generated
templates are selective and often too simple, and it is limited to
provide restricted evaluation of linguistic capabilities. Thus, it
does not garauntee the comprehensive evaluation.
\end{comment}