### 2022-03-15 paper/ase22/tables/lc-requirement-numbers.tex
debugging

%% Automatically generated by pyutil.latex

\DefMacro{lc_1_desc}{Short sentences with neutral adjectives and nouns}
\DefMacro{lc_1_search}{search: \{length: <10; include: neutral adjs \& neutral nouns;}
\DefMacro{lc_1_exclude}{exclude: pos adjs \& neg adjs \& pos nouns \& neg nouns; label: neutral\}}
\DefMacro{lc_1_transform}{transform: -}
\DefMacro{lc_2_desc}{Short sentences with sentiment-laden adjectives}
\DefMacro{lc_2_search}{search: \{length: <10; include: pos adjs;}
\DefMacro{lc_2_exclude}{exclude: neg adjs \& neg verbs \& neg nouns; label: pos\} | \{length: <10; include: neg adjs;exclude: pos adjs \& pos verbs \& pos nouns \& neg verbs \& neg nouns; label: neg\}}
\DefMacro{lc_2_transform}{transform: -}
\DefMacro{lc_3_desc}{Sentiment change over time, present should prevail}
\DefMacro{lc_3_search}{search: \{label: pos\} | \{label: neg\}}
\DefMacro{lc_3_transform}{transform:\{['Previously, I used to like it saying that','Last time, I agreed with saying that','I liked it much as to say that']+[neg sent]+['but', 'although', 'on the other hand']+['now I don't like it.', 'now I hate it.']\} | \{['I used to disagree with saying that','Last time, I didn't like it saying that','I hated it much as to say that']+[neg sent]+['but', 'although', 'on the other hand']+['now I like it.']\}}
\DefMacro{lc_4_desc}{Negated negative should be positive or neutral}
\DefMacro{lc_4_search}{search: \{start: [This, That, These, Those] X [is, are]; label: neg\}}
\DefMacro{lc_4_transform}{transform: negation of demonstratives}
\DefMacro{lc_5_desc}{Negated neutral should still be neutral}
\DefMacro{lc_5_search}{search: \{start: [This, That, These, Those] X [is, are]; label: neutral\}}
\DefMacro{lc_5_transform}{transform: negation of demonstratives}
\DefMacro{lc_6_desc}{Negation of negative at the end, should be positive or neutral}
\DefMacro{lc_6_search}{search: \{label: neg\}}
\DefMacro{lc_6_transform}{transform:\{['I agreed that', 'I thought that']+[neg sent]+['but it wasn't', 'but I didn't']\}}
\DefMacro{lc_7_desc}{Negated positive with neutral content in the middle}
\DefMacro{lc_7_search}{search: \{length: <20; label: pos\} | \{length: <20; label: neutral\}}
\DefMacro{lc_7_transform}{transform:\{['I wouldn't say,', 'I do not think,', 'I don't agree with,']+[neutral sent]+[pos sent]\}}
\DefMacro{lc_8_desc}{Author sentiment is more important than others}
\DefMacro{lc_8_search}{search: \{label:pos\} | \{label: neg\}}
\DefMacro{lc_8_transform}{transform:\{['Some people think that', 'Many people agree with that', 'They think that', 'You agree with that']+[pos sent | neg sent]+['but I think that']+[neg sent | pos sent]\}}
\DefMacro{lc_9_desc}{Parsing sentiment in (question, yes) form}
\DefMacro{lc_9_search}{search: \{label: pos\} | \{label: neg\}}
\DefMacro{lc_9_transform}{transform:\{['Do I think that', 'Do I agree that']+[pos sent | neg sent]+['? yes']\}}
\DefMacro{lc_10_desc}{Parsing positive sentiment in (question, no) form}
\DefMacro{lc_10_search}{search: \{label: pos\}}
\DefMacro{lc_10_transform}{transform:\{['Do I think that', 'Do I agree that']+[pos sent]+['? no']\}}
\DefMacro{lc_11_desc}{Parsing negative sentiment in (question, no) form}
\DefMacro{lc_11_search}{search: \{label: neg\}}
\DefMacro{lc_11_transform}{transform:\{['Do I think that', 'Do I agree that']+[neg sent]+['? no']\}}