### 2022-03-16 paper/ase22/tables/lc-requirement-numbers.tex
update table

%% Automatically generated by pyutil.latex

\DefMacro{lc_1_desc}{Short sentences with neutral adjectives and nouns}
\DefMacro{lc_1_search}{\textbf{Search}~seed=\{length: <10; include: neutral adjs \& neutral nouns;~Exlcude: pos adjs \& neg adjs \& pos nouns \& neg nouns; label: neutral\}}
\DefMacro{lc_1_transform}{\textbf{Transform}~N/A}
\DefMacro{lc_2_desc}{Short sentences with sentiment-laden adjectives}
\DefMacro{lc_2_search}{\textbf{Search}~seed=\{length: <10; include: pos adjs;~Exlcude: neg adjs \& neg verbs \& neg nouns; label: pos\} | \{length: <10; include: neg adjs;exclude: pos adjs \& pos verbs \& pos nouns \& neg verbs \& neg nouns; label: neg\}}
\DefMacro{lc_2_transform}{\textbf{Transform}~N/A}
\DefMacro{lc_3_desc}{Sentiment change over time, present should prevail}
\DefMacro{lc_3_search}{\textbf{Search}~pos\_sent=\{label: pos\}, neg\_sent=\{label: neg\}}
\DefMacro{lc_3_transform}{\textbf{Transform}~seed=\{['Previously, I used to like it saying that','Last time, I agreed with saying that','I liked it much as to say that']+[neg sent]+['but', 'although', 'on the other hand']+['now I don't like it.', 'now I hate it.']\} | \{['I used to disagree with saying that','Last time, I didn't like it saying that','I hated it much as to say that']+[neg sent]+['but', 'although', 'on the other hand']+['now I like it.']\}}
\DefMacro{lc_4_desc}{Negated negative should be positive or neutral}
\DefMacro{lc_4_search}{\textbf{Search}~demonstrative\_sent=\{start: [This, That, These, Those] X [is, are]; label: neg\}}
\DefMacro{lc_4_transform}{\textbf{Transform}~seed=negation of demonstrative\_sent (['is'] -> ['is not', 'isn't'], ['are'] -> ['are not', 'aren't'])}
\DefMacro{lc_5_desc}{Negated neutral should still be neutral}
\DefMacro{lc_5_search}{\textbf{Search}~demonstrative\_sent=\{start: [This, That, These, Those] X [is, are]; label: neutral\}}
\DefMacro{lc_5_transform}{\textbf{Transform}~negation of demonstrative\_sent}
\DefMacro{lc_6_desc}{Negation of negative at the end, should be positive or neutral}
\DefMacro{lc_6_search}{\textbf{Search}~neg\_sent=\{label: neg\}}
\DefMacro{lc_6_transform}{\textbf{Transform}~seed=\{['I agreed that', 'I thought that']+[neg sent]+['but it wasn't', 'but I didn't']\}}
\DefMacro{lc_7_desc}{Negated positive with neutral content in the middle}
\DefMacro{lc_7_search}{\textbf{Search}~pos\_sent=\{length: <20; label: pos\}, neutral\_sent=\{length: <20; label: neutral\}}
\DefMacro{lc_7_transform}{\textbf{Transform}~seed=\{['I wouldn't say,', 'I do not think,', 'I don't agree with,']+[neutral sent]+[pos sent]\}}
\DefMacro{lc_8_desc}{Author sentiment is more important than others}
\DefMacro{lc_8_search}{\textbf{Search}~pos\_sent=\{label:pos\}, neg\_sent=\{label: neg\}}
\DefMacro{lc_8_transform}{\textbf{Transform}~seed=\{[temp1]+[pos\_sent]+[temp2]+[neg\_sent]\} | \{[temp1]+[neg\_sent]+[temp2]+[pos\_sent]\} where temp1=\{['Some people think that', 'Many people agree with that', 'They think that', 'You agree with that'], temp2=['but I think that']\}}
\DefMacro{lc_9_desc}{Parsing sentiment in (question, yes) form}
\DefMacro{lc_9_search}{\textbf{Search}~pos\_sent=\{label: pos\}, neg\_sent=\{label: neg\}}
\DefMacro{lc_9_transform}{\textbf{Transform}~seed=\{['Do I think that', 'Do I agree that']+[pos sent | neg sent]+['? yes']\}}
\DefMacro{lc_10_desc}{Parsing positive sentiment in (question, no) form}
\DefMacro{lc_10_search}{\textbf{Search}~pos\_sent=\{label: pos\}}
\DefMacro{lc_10_transform}{\textbf{Transform}~seed=\{['Do I think that', 'Do I agree that']+[pos sent]+['? no']\}}
\DefMacro{lc_11_desc}{Parsing negative sentiment in (question, no) form}
\DefMacro{lc_11_search}{\textbf{Search}~neg\_sent=\{label: neg\}}
\DefMacro{lc_11_transform}{\textbf{Transform}~seed=\{['Do I think that', 'Do I agree that']+[neg sent]+['? no']\}}