### 2022-04-04 paper/ase22/tables/retrain-debug-numbers.tex
paper

%% Automatically generated by pyutil.latex

\DefMacro{retrain_debug_approach_0}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_0}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_0}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_0}{0}
\DefMacro{retrain_debug_num_fail_orig_0}{860}
\DefMacro{retrain_debug_num_pass_orig_0}{140}
\DefMacro{retrain_debug_num_fail_retrained_0}{1000}
\DefMacro{retrain_debug_num_pass_retrained_0}{0}
\DefMacro{retrain_debug_approach_1}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_1}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_1}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_1}{0}
\DefMacro{retrain_debug_num_fail_orig_1}{26}
\DefMacro{retrain_debug_num_pass_orig_1}{8632}
\DefMacro{retrain_debug_num_fail_retrained_1}{8658}
\DefMacro{retrain_debug_num_pass_retrained_1}{0}
\DefMacro{retrain_debug_approach_2}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_2}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_2}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_2}{1082}
\DefMacro{retrain_debug_num_fail_orig_2}{1793}
\DefMacro{retrain_debug_num_pass_orig_2}{7411}
\DefMacro{retrain_debug_num_fail_retrained_2}{1203}
\DefMacro{retrain_debug_num_pass_retrained_2}{8001}
\DefMacro{retrain_debug_approach_3}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_3}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_3}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_3}{0}
\DefMacro{retrain_debug_num_fail_orig_3}{3741}
\DefMacro{retrain_debug_num_pass_orig_3}{4787}
\DefMacro{retrain_debug_num_fail_retrained_3}{8528}
\DefMacro{retrain_debug_num_pass_retrained_3}{0}
\DefMacro{retrain_debug_approach_4}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_4}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_4}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_4}{1330}
\DefMacro{retrain_debug_num_fail_orig_4}{1330}
\DefMacro{retrain_debug_num_pass_orig_4}{386}
\DefMacro{retrain_debug_num_fail_retrained_4}{0}
\DefMacro{retrain_debug_num_pass_retrained_4}{1716}
\DefMacro{retrain_debug_approach_5}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_5}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_5}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_5}{2427}
\DefMacro{retrain_debug_num_fail_orig_5}{2427}
\DefMacro{retrain_debug_num_pass_orig_5}{69}
\DefMacro{retrain_debug_num_fail_retrained_5}{0}
\DefMacro{retrain_debug_num_pass_retrained_5}{2496}
\DefMacro{retrain_debug_approach_6}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_6}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_eval_lc_6}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_6}{0}
\DefMacro{retrain_debug_num_fail_orig_6}{1680}
\DefMacro{retrain_debug_num_pass_orig_6}{6320}
\DefMacro{retrain_debug_num_fail_retrained_6}{8000}
\DefMacro{retrain_debug_num_pass_retrained_6}{0}
\DefMacro{retrain_debug_approach_7}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_7}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_7}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_7}{0}
\DefMacro{retrain_debug_num_fail_orig_7}{860}
\DefMacro{retrain_debug_num_pass_orig_7}{140}
\DefMacro{retrain_debug_num_fail_retrained_7}{1000}
\DefMacro{retrain_debug_num_pass_retrained_7}{0}
\DefMacro{retrain_debug_approach_8}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_8}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_8}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_8}{0}
\DefMacro{retrain_debug_num_fail_orig_8}{26}
\DefMacro{retrain_debug_num_pass_orig_8}{8632}
\DefMacro{retrain_debug_num_fail_retrained_8}{8658}
\DefMacro{retrain_debug_num_pass_retrained_8}{0}
\DefMacro{retrain_debug_approach_9}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_9}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_9}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_9}{1540}
\DefMacro{retrain_debug_num_fail_orig_9}{1793}
\DefMacro{retrain_debug_num_pass_orig_9}{7411}
\DefMacro{retrain_debug_num_fail_retrained_9}{7644}
\DefMacro{retrain_debug_num_pass_retrained_9}{1560}
\DefMacro{retrain_debug_approach_10}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_10}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_10}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_10}{0}
\DefMacro{retrain_debug_num_fail_orig_10}{3741}
\DefMacro{retrain_debug_num_pass_orig_10}{4787}
\DefMacro{retrain_debug_num_fail_retrained_10}{8528}
\DefMacro{retrain_debug_num_pass_retrained_10}{0}
\DefMacro{retrain_debug_approach_11}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_11}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_11}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_11}{1330}
\DefMacro{retrain_debug_num_fail_orig_11}{1330}
\DefMacro{retrain_debug_num_pass_orig_11}{386}
\DefMacro{retrain_debug_num_fail_retrained_11}{0}
\DefMacro{retrain_debug_num_pass_retrained_11}{1716}
\DefMacro{retrain_debug_approach_12}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_12}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_12}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_12}{2427}
\DefMacro{retrain_debug_num_fail_orig_12}{2427}
\DefMacro{retrain_debug_num_pass_orig_12}{69}
\DefMacro{retrain_debug_num_fail_retrained_12}{0}
\DefMacro{retrain_debug_num_pass_retrained_12}{2496}
\DefMacro{retrain_debug_approach_13}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_13}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_eval_lc_13}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_13}{0}
\DefMacro{retrain_debug_num_fail_orig_13}{1680}
\DefMacro{retrain_debug_num_pass_orig_13}{6320}
\DefMacro{retrain_debug_num_fail_retrained_13}{8000}
\DefMacro{retrain_debug_num_pass_retrained_13}{0}
\DefMacro{retrain_debug_approach_14}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_14}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_14}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_14}{0}
\DefMacro{retrain_debug_num_fail_orig_14}{860}
\DefMacro{retrain_debug_num_pass_orig_14}{140}
\DefMacro{retrain_debug_num_fail_retrained_14}{1000}
\DefMacro{retrain_debug_num_pass_retrained_14}{0}
\DefMacro{retrain_debug_approach_15}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_15}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_15}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_15}{0}
\DefMacro{retrain_debug_num_fail_orig_15}{26}
\DefMacro{retrain_debug_num_pass_orig_15}{8632}
\DefMacro{retrain_debug_num_fail_retrained_15}{8658}
\DefMacro{retrain_debug_num_pass_retrained_15}{0}
\DefMacro{retrain_debug_approach_16}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_16}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_16}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_16}{1540}
\DefMacro{retrain_debug_num_fail_orig_16}{1793}
\DefMacro{retrain_debug_num_pass_orig_16}{7411}
\DefMacro{retrain_debug_num_fail_retrained_16}{7644}
\DefMacro{retrain_debug_num_pass_retrained_16}{1560}
\DefMacro{retrain_debug_approach_17}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_17}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_17}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_17}{0}
\DefMacro{retrain_debug_num_fail_orig_17}{3741}
\DefMacro{retrain_debug_num_pass_orig_17}{4787}
\DefMacro{retrain_debug_num_fail_retrained_17}{8528}
\DefMacro{retrain_debug_num_pass_retrained_17}{0}
\DefMacro{retrain_debug_approach_18}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_18}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_18}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_18}{1330}
\DefMacro{retrain_debug_num_fail_orig_18}{1330}
\DefMacro{retrain_debug_num_pass_orig_18}{386}
\DefMacro{retrain_debug_num_fail_retrained_18}{0}
\DefMacro{retrain_debug_num_pass_retrained_18}{1716}
\DefMacro{retrain_debug_approach_19}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_19}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_19}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_19}{2427}
\DefMacro{retrain_debug_num_fail_orig_19}{2427}
\DefMacro{retrain_debug_num_pass_orig_19}{69}
\DefMacro{retrain_debug_num_fail_retrained_19}{0}
\DefMacro{retrain_debug_num_pass_retrained_19}{2496}
\DefMacro{retrain_debug_approach_20}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_20}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_eval_lc_20}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_20}{0}
\DefMacro{retrain_debug_num_fail_orig_20}{1680}
\DefMacro{retrain_debug_num_pass_orig_20}{6320}
\DefMacro{retrain_debug_num_fail_retrained_20}{8000}
\DefMacro{retrain_debug_num_pass_retrained_20}{0}
\DefMacro{retrain_debug_approach_21}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_21}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_21}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_21}{0}
\DefMacro{retrain_debug_num_fail_orig_21}{860}
\DefMacro{retrain_debug_num_pass_orig_21}{140}
\DefMacro{retrain_debug_num_fail_retrained_21}{1000}
\DefMacro{retrain_debug_num_pass_retrained_21}{0}
\DefMacro{retrain_debug_approach_22}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_22}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_22}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_22}{0}
\DefMacro{retrain_debug_num_fail_orig_22}{26}
\DefMacro{retrain_debug_num_pass_orig_22}{8632}
\DefMacro{retrain_debug_num_fail_retrained_22}{8658}
\DefMacro{retrain_debug_num_pass_retrained_22}{0}
\DefMacro{retrain_debug_approach_23}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_23}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_23}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_23}{1540}
\DefMacro{retrain_debug_num_fail_orig_23}{1793}
\DefMacro{retrain_debug_num_pass_orig_23}{7411}
\DefMacro{retrain_debug_num_fail_retrained_23}{7644}
\DefMacro{retrain_debug_num_pass_retrained_23}{1560}
\DefMacro{retrain_debug_approach_24}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_24}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_24}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_24}{0}
\DefMacro{retrain_debug_num_fail_orig_24}{3741}
\DefMacro{retrain_debug_num_pass_orig_24}{4787}
\DefMacro{retrain_debug_num_fail_retrained_24}{8528}
\DefMacro{retrain_debug_num_pass_retrained_24}{0}
\DefMacro{retrain_debug_approach_25}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_25}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_25}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_25}{1330}
\DefMacro{retrain_debug_num_fail_orig_25}{1330}
\DefMacro{retrain_debug_num_pass_orig_25}{386}
\DefMacro{retrain_debug_num_fail_retrained_25}{0}
\DefMacro{retrain_debug_num_pass_retrained_25}{1716}
\DefMacro{retrain_debug_approach_26}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_26}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_26}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_26}{2427}
\DefMacro{retrain_debug_num_fail_orig_26}{2427}
\DefMacro{retrain_debug_num_pass_orig_26}{69}
\DefMacro{retrain_debug_num_fail_retrained_26}{0}
\DefMacro{retrain_debug_num_pass_retrained_26}{2496}
\DefMacro{retrain_debug_approach_27}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_27}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_eval_lc_27}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_27}{0}
\DefMacro{retrain_debug_num_fail_orig_27}{1680}
\DefMacro{retrain_debug_num_pass_orig_27}{6320}
\DefMacro{retrain_debug_num_fail_retrained_27}{7888}
\DefMacro{retrain_debug_num_pass_retrained_27}{112}
\DefMacro{retrain_debug_approach_28}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_28}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_28}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_28}{840}
\DefMacro{retrain_debug_num_fail_orig_28}{860}
\DefMacro{retrain_debug_num_pass_orig_28}{140}
\DefMacro{retrain_debug_num_fail_retrained_28}{20}
\DefMacro{retrain_debug_num_pass_retrained_28}{980}
\DefMacro{retrain_debug_approach_29}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_29}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_29}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_29}{0}
\DefMacro{retrain_debug_num_fail_orig_29}{26}
\DefMacro{retrain_debug_num_pass_orig_29}{8632}
\DefMacro{retrain_debug_num_fail_retrained_29}{8658}
\DefMacro{retrain_debug_num_pass_retrained_29}{0}
\DefMacro{retrain_debug_approach_30}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_30}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_30}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_30}{1540}
\DefMacro{retrain_debug_num_fail_orig_30}{1793}
\DefMacro{retrain_debug_num_pass_orig_30}{7411}
\DefMacro{retrain_debug_num_fail_retrained_30}{7644}
\DefMacro{retrain_debug_num_pass_retrained_30}{1560}
\DefMacro{retrain_debug_approach_31}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_31}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_31}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_31}{0}
\DefMacro{retrain_debug_num_fail_orig_31}{3741}
\DefMacro{retrain_debug_num_pass_orig_31}{4787}
\DefMacro{retrain_debug_num_fail_retrained_31}{8528}
\DefMacro{retrain_debug_num_pass_retrained_31}{0}
\DefMacro{retrain_debug_approach_32}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_32}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_32}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_32}{1330}
\DefMacro{retrain_debug_num_fail_orig_32}{1330}
\DefMacro{retrain_debug_num_pass_orig_32}{386}
\DefMacro{retrain_debug_num_fail_retrained_32}{0}
\DefMacro{retrain_debug_num_pass_retrained_32}{1716}
\DefMacro{retrain_debug_approach_33}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_33}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_33}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_33}{2427}
\DefMacro{retrain_debug_num_fail_orig_33}{2427}
\DefMacro{retrain_debug_num_pass_orig_33}{69}
\DefMacro{retrain_debug_num_fail_retrained_33}{0}
\DefMacro{retrain_debug_num_pass_retrained_33}{2496}
\DefMacro{retrain_debug_approach_34}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_34}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_eval_lc_34}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_34}{0}
\DefMacro{retrain_debug_num_fail_orig_34}{1680}
\DefMacro{retrain_debug_num_pass_orig_34}{6320}
\DefMacro{retrain_debug_num_fail_retrained_34}{8000}
\DefMacro{retrain_debug_num_pass_retrained_34}{0}
\DefMacro{retrain_debug_approach_35}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_35}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_35}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_35}{0}
\DefMacro{retrain_debug_num_fail_orig_35}{860}
\DefMacro{retrain_debug_num_pass_orig_35}{140}
\DefMacro{retrain_debug_num_fail_retrained_35}{1000}
\DefMacro{retrain_debug_num_pass_retrained_35}{0}
\DefMacro{retrain_debug_approach_36}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_36}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_36}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_36}{0}
\DefMacro{retrain_debug_num_fail_orig_36}{26}
\DefMacro{retrain_debug_num_pass_orig_36}{8632}
\DefMacro{retrain_debug_num_fail_retrained_36}{8658}
\DefMacro{retrain_debug_num_pass_retrained_36}{0}
\DefMacro{retrain_debug_approach_37}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_37}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_37}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_37}{1540}
\DefMacro{retrain_debug_num_fail_orig_37}{1793}
\DefMacro{retrain_debug_num_pass_orig_37}{7411}
\DefMacro{retrain_debug_num_fail_retrained_37}{7644}
\DefMacro{retrain_debug_num_pass_retrained_37}{1560}
\DefMacro{retrain_debug_approach_38}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_38}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_38}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_38}{0}
\DefMacro{retrain_debug_num_fail_orig_38}{3741}
\DefMacro{retrain_debug_num_pass_orig_38}{4787}
\DefMacro{retrain_debug_num_fail_retrained_38}{8528}
\DefMacro{retrain_debug_num_pass_retrained_38}{0}
\DefMacro{retrain_debug_approach_39}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_39}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_39}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_39}{1330}
\DefMacro{retrain_debug_num_fail_orig_39}{1330}
\DefMacro{retrain_debug_num_pass_orig_39}{386}
\DefMacro{retrain_debug_num_fail_retrained_39}{0}
\DefMacro{retrain_debug_num_pass_retrained_39}{1716}
\DefMacro{retrain_debug_approach_40}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_40}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_40}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_40}{2427}
\DefMacro{retrain_debug_num_fail_orig_40}{2427}
\DefMacro{retrain_debug_num_pass_orig_40}{69}
\DefMacro{retrain_debug_num_fail_retrained_40}{0}
\DefMacro{retrain_debug_num_pass_retrained_40}{2496}
\DefMacro{retrain_debug_approach_41}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_41}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_eval_lc_41}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_41}{0}
\DefMacro{retrain_debug_num_fail_orig_41}{1680}
\DefMacro{retrain_debug_num_pass_orig_41}{6320}
\DefMacro{retrain_debug_num_fail_retrained_41}{8000}
\DefMacro{retrain_debug_num_pass_retrained_41}{0}
\DefMacro{retrain_debug_approach_42}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_42}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_42}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_42}{0}
\DefMacro{retrain_debug_num_fail_orig_42}{860}
\DefMacro{retrain_debug_num_pass_orig_42}{140}
\DefMacro{retrain_debug_num_fail_retrained_42}{1000}
\DefMacro{retrain_debug_num_pass_retrained_42}{0}
\DefMacro{retrain_debug_approach_43}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_43}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_43}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_43}{0}
\DefMacro{retrain_debug_num_fail_orig_43}{26}
\DefMacro{retrain_debug_num_pass_orig_43}{8632}
\DefMacro{retrain_debug_num_fail_retrained_43}{8658}
\DefMacro{retrain_debug_num_pass_retrained_43}{0}
\DefMacro{retrain_debug_approach_44}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_44}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_44}{parsing sentiment in (question, yes) form}
\DefMacro{retrain_debug_num_fail2pass_44}{1540}
\DefMacro{retrain_debug_num_fail_orig_44}{1793}
\DefMacro{retrain_debug_num_pass_orig_44}{7411}
\DefMacro{retrain_debug_num_fail_retrained_44}{7644}
\DefMacro{retrain_debug_num_pass_retrained_44}{1560}
\DefMacro{retrain_debug_approach_45}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_45}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_45}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_45}{0}
\DefMacro{retrain_debug_num_fail_orig_45}{3741}
\DefMacro{retrain_debug_num_pass_orig_45}{4787}
\DefMacro{retrain_debug_num_fail_retrained_45}{8528}
\DefMacro{retrain_debug_num_pass_retrained_45}{0}
\DefMacro{retrain_debug_approach_46}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_46}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_46}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_46}{1330}
\DefMacro{retrain_debug_num_fail_orig_46}{1330}
\DefMacro{retrain_debug_num_pass_orig_46}{386}
\DefMacro{retrain_debug_num_fail_retrained_46}{0}
\DefMacro{retrain_debug_num_pass_retrained_46}{1716}
\DefMacro{retrain_debug_approach_47}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_47}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_47}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_47}{2427}
\DefMacro{retrain_debug_num_fail_orig_47}{2427}
\DefMacro{retrain_debug_num_pass_orig_47}{69}
\DefMacro{retrain_debug_num_fail_retrained_47}{0}
\DefMacro{retrain_debug_num_pass_retrained_47}{2496}
\DefMacro{retrain_debug_approach_48}{Retrain:Ours::Test:Checklist}
\DefMacro{retrain_debug_retrained_lc_48}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_eval_lc_48}{Sentiment change over time, present should prevail}
\DefMacro{retrain_debug_num_fail2pass_48}{0}
\DefMacro{retrain_debug_num_fail_orig_48}{1680}
\DefMacro{retrain_debug_num_pass_orig_48}{6320}
\DefMacro{retrain_debug_num_fail_retrained_48}{8000}
\DefMacro{retrain_debug_num_pass_retrained_48}{0}
\DefMacro{retrain_debug_approach_50}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_50}{used to  but now}
\DefMacro{retrain_debug_eval_lc_50}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_50}{0}
\DefMacro{retrain_debug_num_fail_orig_50}{1185}
\DefMacro{retrain_debug_num_pass_orig_50}{1188}
\DefMacro{retrain_debug_num_fail_retrained_50}{2373}
\DefMacro{retrain_debug_num_pass_retrained_50}{0}
\DefMacro{retrain_debug_approach_51}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_51}{used to  but now}
\DefMacro{retrain_debug_eval_lc_51}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_51}{956}
\DefMacro{retrain_debug_num_fail_orig_51}{956}
\DefMacro{retrain_debug_num_pass_orig_51}{79}
\DefMacro{retrain_debug_num_fail_retrained_51}{0}
\DefMacro{retrain_debug_num_pass_retrained_51}{1035}
\DefMacro{retrain_debug_approach_52}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_52}{used to  but now}
\DefMacro{retrain_debug_eval_lc_52}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_52}{0}
\DefMacro{retrain_debug_num_fail_orig_52}{1480}
\DefMacro{retrain_debug_num_pass_orig_52}{204}
\DefMacro{retrain_debug_num_fail_retrained_52}{1684}
\DefMacro{retrain_debug_num_pass_retrained_52}{0}
\DefMacro{retrain_debug_approach_53}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_53}{used to  but now}
\DefMacro{retrain_debug_eval_lc_53}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_53}{4}
\DefMacro{retrain_debug_num_fail_orig_53}{12}
\DefMacro{retrain_debug_num_pass_orig_53}{38}
\DefMacro{retrain_debug_num_fail_retrained_53}{34}
\DefMacro{retrain_debug_num_pass_retrained_53}{16}
\DefMacro{retrain_debug_approach_54}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_54}{used to  but now}
\DefMacro{retrain_debug_eval_lc_54}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_54}{197}
\DefMacro{retrain_debug_num_fail_orig_54}{197}
\DefMacro{retrain_debug_num_pass_orig_54}{32}
\DefMacro{retrain_debug_num_fail_retrained_54}{0}
\DefMacro{retrain_debug_num_pass_retrained_54}{229}
\DefMacro{retrain_debug_approach_55}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_55}{used to  but now}
\DefMacro{retrain_debug_eval_lc_55}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_55}{0}
\DefMacro{retrain_debug_num_fail_orig_55}{19}
\DefMacro{retrain_debug_num_pass_orig_55}{425}
\DefMacro{retrain_debug_num_fail_retrained_55}{444}
\DefMacro{retrain_debug_num_pass_retrained_55}{0}
\DefMacro{retrain_debug_approach_56}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_56}{used to  but now}
\DefMacro{retrain_debug_eval_lc_56}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_56}{0}
\DefMacro{retrain_debug_num_fail_orig_56}{39}
\DefMacro{retrain_debug_num_pass_orig_56}{1384}
\DefMacro{retrain_debug_num_fail_retrained_56}{1423}
\DefMacro{retrain_debug_num_pass_retrained_56}{0}
\DefMacro{retrain_debug_approach_57}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_57}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_57}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_57}{0}
\DefMacro{retrain_debug_num_fail_orig_57}{1185}
\DefMacro{retrain_debug_num_pass_orig_57}{1188}
\DefMacro{retrain_debug_num_fail_retrained_57}{2373}
\DefMacro{retrain_debug_num_pass_retrained_57}{0}
\DefMacro{retrain_debug_approach_58}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_58}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_58}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_58}{956}
\DefMacro{retrain_debug_num_fail_orig_58}{956}
\DefMacro{retrain_debug_num_pass_orig_58}{79}
\DefMacro{retrain_debug_num_fail_retrained_58}{0}
\DefMacro{retrain_debug_num_pass_retrained_58}{1035}
\DefMacro{retrain_debug_approach_59}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_59}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_59}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_59}{0}
\DefMacro{retrain_debug_num_fail_orig_59}{1480}
\DefMacro{retrain_debug_num_pass_orig_59}{204}
\DefMacro{retrain_debug_num_fail_retrained_59}{1684}
\DefMacro{retrain_debug_num_pass_retrained_59}{0}
\DefMacro{retrain_debug_approach_60}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_60}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_60}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_60}{0}
\DefMacro{retrain_debug_num_fail_orig_60}{12}
\DefMacro{retrain_debug_num_pass_orig_60}{38}
\DefMacro{retrain_debug_num_fail_retrained_60}{50}
\DefMacro{retrain_debug_num_pass_retrained_60}{0}
\DefMacro{retrain_debug_approach_61}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_61}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_61}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_61}{197}
\DefMacro{retrain_debug_num_fail_orig_61}{197}
\DefMacro{retrain_debug_num_pass_orig_61}{32}
\DefMacro{retrain_debug_num_fail_retrained_61}{0}
\DefMacro{retrain_debug_num_pass_retrained_61}{229}
\DefMacro{retrain_debug_approach_62}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_62}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_62}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_62}{0}
\DefMacro{retrain_debug_num_fail_orig_62}{19}
\DefMacro{retrain_debug_num_pass_orig_62}{425}
\DefMacro{retrain_debug_num_fail_retrained_62}{444}
\DefMacro{retrain_debug_num_pass_retrained_62}{0}
\DefMacro{retrain_debug_approach_63}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_63}{simple negations: not neutral is still neutral}
\DefMacro{retrain_debug_eval_lc_63}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_63}{0}
\DefMacro{retrain_debug_num_fail_orig_63}{39}
\DefMacro{retrain_debug_num_pass_orig_63}{1384}
\DefMacro{retrain_debug_num_fail_retrained_63}{1423}
\DefMacro{retrain_debug_num_pass_retrained_63}{0}
\DefMacro{retrain_debug_approach_64}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_64}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_64}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_64}{0}
\DefMacro{retrain_debug_num_fail_orig_64}{1185}
\DefMacro{retrain_debug_num_pass_orig_64}{1188}
\DefMacro{retrain_debug_num_fail_retrained_64}{2373}
\DefMacro{retrain_debug_num_pass_retrained_64}{0}
\DefMacro{retrain_debug_approach_65}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_65}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_65}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_65}{956}
\DefMacro{retrain_debug_num_fail_orig_65}{956}
\DefMacro{retrain_debug_num_pass_orig_65}{79}
\DefMacro{retrain_debug_num_fail_retrained_65}{0}
\DefMacro{retrain_debug_num_pass_retrained_65}{1035}
\DefMacro{retrain_debug_approach_66}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_66}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_66}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_66}{0}
\DefMacro{retrain_debug_num_fail_orig_66}{1480}
\DefMacro{retrain_debug_num_pass_orig_66}{204}
\DefMacro{retrain_debug_num_fail_retrained_66}{1684}
\DefMacro{retrain_debug_num_pass_retrained_66}{0}
\DefMacro{retrain_debug_approach_67}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_67}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_67}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_67}{0}
\DefMacro{retrain_debug_num_fail_orig_67}{12}
\DefMacro{retrain_debug_num_pass_orig_67}{38}
\DefMacro{retrain_debug_num_fail_retrained_67}{50}
\DefMacro{retrain_debug_num_pass_retrained_67}{0}
\DefMacro{retrain_debug_approach_68}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_68}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_68}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_68}{197}
\DefMacro{retrain_debug_num_fail_orig_68}{197}
\DefMacro{retrain_debug_num_pass_orig_68}{32}
\DefMacro{retrain_debug_num_fail_retrained_68}{0}
\DefMacro{retrain_debug_num_pass_retrained_68}{229}
\DefMacro{retrain_debug_approach_69}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_69}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_69}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_69}{0}
\DefMacro{retrain_debug_num_fail_orig_69}{19}
\DefMacro{retrain_debug_num_pass_orig_69}{425}
\DefMacro{retrain_debug_num_fail_retrained_69}{444}
\DefMacro{retrain_debug_num_pass_retrained_69}{0}
\DefMacro{retrain_debug_approach_70}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_70}{neutral words in context}
\DefMacro{retrain_debug_eval_lc_70}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_70}{0}
\DefMacro{retrain_debug_num_fail_orig_70}{39}
\DefMacro{retrain_debug_num_pass_orig_70}{1384}
\DefMacro{retrain_debug_num_fail_retrained_70}{1423}
\DefMacro{retrain_debug_num_pass_retrained_70}{0}
\DefMacro{retrain_debug_approach_71}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_71}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_71}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_71}{0}
\DefMacro{retrain_debug_num_fail_orig_71}{1185}
\DefMacro{retrain_debug_num_pass_orig_71}{1188}
\DefMacro{retrain_debug_num_fail_retrained_71}{2373}
\DefMacro{retrain_debug_num_pass_retrained_71}{0}
\DefMacro{retrain_debug_approach_72}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_72}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_72}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_72}{956}
\DefMacro{retrain_debug_num_fail_orig_72}{956}
\DefMacro{retrain_debug_num_pass_orig_72}{79}
\DefMacro{retrain_debug_num_fail_retrained_72}{0}
\DefMacro{retrain_debug_num_pass_retrained_72}{1035}
\DefMacro{retrain_debug_approach_73}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_73}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_73}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_73}{0}
\DefMacro{retrain_debug_num_fail_orig_73}{1480}
\DefMacro{retrain_debug_num_pass_orig_73}{204}
\DefMacro{retrain_debug_num_fail_retrained_73}{1684}
\DefMacro{retrain_debug_num_pass_retrained_73}{0}
\DefMacro{retrain_debug_approach_74}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_74}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_74}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_74}{0}
\DefMacro{retrain_debug_num_fail_orig_74}{12}
\DefMacro{retrain_debug_num_pass_orig_74}{38}
\DefMacro{retrain_debug_num_fail_retrained_74}{50}
\DefMacro{retrain_debug_num_pass_retrained_74}{0}
\DefMacro{retrain_debug_approach_75}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_75}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_75}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_75}{197}
\DefMacro{retrain_debug_num_fail_orig_75}{197}
\DefMacro{retrain_debug_num_pass_orig_75}{32}
\DefMacro{retrain_debug_num_fail_retrained_75}{0}
\DefMacro{retrain_debug_num_pass_retrained_75}{229}
\DefMacro{retrain_debug_approach_76}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_76}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_76}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_76}{0}
\DefMacro{retrain_debug_num_fail_orig_76}{19}
\DefMacro{retrain_debug_num_pass_orig_76}{425}
\DefMacro{retrain_debug_num_fail_retrained_76}{444}
\DefMacro{retrain_debug_num_pass_retrained_76}{0}
\DefMacro{retrain_debug_approach_77}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_77}{my opinion is what matters}
\DefMacro{retrain_debug_eval_lc_77}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_77}{0}
\DefMacro{retrain_debug_num_fail_orig_77}{39}
\DefMacro{retrain_debug_num_pass_orig_77}{1384}
\DefMacro{retrain_debug_num_fail_retrained_77}{1423}
\DefMacro{retrain_debug_num_pass_retrained_77}{0}
\DefMacro{retrain_debug_approach_78}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_78}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_78}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_78}{0}
\DefMacro{retrain_debug_num_fail_orig_78}{1185}
\DefMacro{retrain_debug_num_pass_orig_78}{1188}
\DefMacro{retrain_debug_num_fail_retrained_78}{2373}
\DefMacro{retrain_debug_num_pass_retrained_78}{0}
\DefMacro{retrain_debug_approach_79}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_79}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_79}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_79}{956}
\DefMacro{retrain_debug_num_fail_orig_79}{956}
\DefMacro{retrain_debug_num_pass_orig_79}{79}
\DefMacro{retrain_debug_num_fail_retrained_79}{0}
\DefMacro{retrain_debug_num_pass_retrained_79}{1035}
\DefMacro{retrain_debug_approach_80}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_80}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_80}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_80}{0}
\DefMacro{retrain_debug_num_fail_orig_80}{1480}
\DefMacro{retrain_debug_num_pass_orig_80}{204}
\DefMacro{retrain_debug_num_fail_retrained_80}{1684}
\DefMacro{retrain_debug_num_pass_retrained_80}{0}
\DefMacro{retrain_debug_approach_81}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_81}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_81}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_81}{0}
\DefMacro{retrain_debug_num_fail_orig_81}{12}
\DefMacro{retrain_debug_num_pass_orig_81}{38}
\DefMacro{retrain_debug_num_fail_retrained_81}{50}
\DefMacro{retrain_debug_num_pass_retrained_81}{0}
\DefMacro{retrain_debug_approach_82}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_82}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_82}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_82}{197}
\DefMacro{retrain_debug_num_fail_orig_82}{197}
\DefMacro{retrain_debug_num_pass_orig_82}{32}
\DefMacro{retrain_debug_num_fail_retrained_82}{0}
\DefMacro{retrain_debug_num_pass_retrained_82}{229}
\DefMacro{retrain_debug_approach_83}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_83}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_83}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_83}{0}
\DefMacro{retrain_debug_num_fail_orig_83}{19}
\DefMacro{retrain_debug_num_pass_orig_83}{425}
\DefMacro{retrain_debug_num_fail_retrained_83}{444}
\DefMacro{retrain_debug_num_pass_retrained_83}{0}
\DefMacro{retrain_debug_approach_84}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_84}{Sentiment-laden words in context}
\DefMacro{retrain_debug_eval_lc_84}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_84}{0}
\DefMacro{retrain_debug_num_fail_orig_84}{39}
\DefMacro{retrain_debug_num_pass_orig_84}{1384}
\DefMacro{retrain_debug_num_fail_retrained_84}{1423}
\DefMacro{retrain_debug_num_pass_retrained_84}{0}
\DefMacro{retrain_debug_approach_85}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_85}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_85}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_85}{0}
\DefMacro{retrain_debug_num_fail_orig_85}{1185}
\DefMacro{retrain_debug_num_pass_orig_85}{1188}
\DefMacro{retrain_debug_num_fail_retrained_85}{2373}
\DefMacro{retrain_debug_num_pass_retrained_85}{0}
\DefMacro{retrain_debug_approach_86}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_86}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_86}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_86}{956}
\DefMacro{retrain_debug_num_fail_orig_86}{956}
\DefMacro{retrain_debug_num_pass_orig_86}{79}
\DefMacro{retrain_debug_num_fail_retrained_86}{0}
\DefMacro{retrain_debug_num_pass_retrained_86}{1035}
\DefMacro{retrain_debug_approach_87}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_87}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_87}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_87}{3}
\DefMacro{retrain_debug_num_fail_orig_87}{1480}
\DefMacro{retrain_debug_num_pass_orig_87}{204}
\DefMacro{retrain_debug_num_fail_retrained_87}{1673}
\DefMacro{retrain_debug_num_pass_retrained_87}{11}
\DefMacro{retrain_debug_approach_88}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_88}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_88}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_88}{0}
\DefMacro{retrain_debug_num_fail_orig_88}{12}
\DefMacro{retrain_debug_num_pass_orig_88}{38}
\DefMacro{retrain_debug_num_fail_retrained_88}{50}
\DefMacro{retrain_debug_num_pass_retrained_88}{0}
\DefMacro{retrain_debug_approach_89}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_89}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_89}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_89}{197}
\DefMacro{retrain_debug_num_fail_orig_89}{197}
\DefMacro{retrain_debug_num_pass_orig_89}{32}
\DefMacro{retrain_debug_num_fail_retrained_89}{0}
\DefMacro{retrain_debug_num_pass_retrained_89}{229}
\DefMacro{retrain_debug_approach_90}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_90}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_90}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_90}{0}
\DefMacro{retrain_debug_num_fail_orig_90}{19}
\DefMacro{retrain_debug_num_pass_orig_90}{425}
\DefMacro{retrain_debug_num_fail_retrained_90}{444}
\DefMacro{retrain_debug_num_pass_retrained_90}{0}
\DefMacro{retrain_debug_approach_91}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_91}{Hard: Negation of positive with neutral stuff in the middle (should be negative)}
\DefMacro{retrain_debug_eval_lc_91}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_91}{0}
\DefMacro{retrain_debug_num_fail_orig_91}{39}
\DefMacro{retrain_debug_num_pass_orig_91}{1384}
\DefMacro{retrain_debug_num_fail_retrained_91}{1423}
\DefMacro{retrain_debug_num_pass_retrained_91}{0}
\DefMacro{retrain_debug_approach_92}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_92}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_92}{Author sentiment is more important than of others}
\DefMacro{retrain_debug_num_fail2pass_92}{0}
\DefMacro{retrain_debug_num_fail_orig_92}{1185}
\DefMacro{retrain_debug_num_pass_orig_92}{1188}
\DefMacro{retrain_debug_num_fail_retrained_92}{2373}
\DefMacro{retrain_debug_num_pass_retrained_92}{0}
\DefMacro{retrain_debug_approach_93}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_93}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_93}{Negated neutral should still be neutral}
\DefMacro{retrain_debug_num_fail2pass_93}{956}
\DefMacro{retrain_debug_num_fail_orig_93}{956}
\DefMacro{retrain_debug_num_pass_orig_93}{79}
\DefMacro{retrain_debug_num_fail_retrained_93}{0}
\DefMacro{retrain_debug_num_pass_retrained_93}{1035}
\DefMacro{retrain_debug_approach_94}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_94}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_94}{Negated positive with neutral content in the middle}
\DefMacro{retrain_debug_num_fail2pass_94}{0}
\DefMacro{retrain_debug_num_fail_orig_94}{1480}
\DefMacro{retrain_debug_num_pass_orig_94}{204}
\DefMacro{retrain_debug_num_fail_retrained_94}{1684}
\DefMacro{retrain_debug_num_pass_retrained_94}{0}
\DefMacro{retrain_debug_approach_95}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_95}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_95}{Sentiment change over time  present should prevail}
\DefMacro{retrain_debug_num_fail2pass_95}{0}
\DefMacro{retrain_debug_num_fail_orig_95}{12}
\DefMacro{retrain_debug_num_pass_orig_95}{38}
\DefMacro{retrain_debug_num_fail_retrained_95}{50}
\DefMacro{retrain_debug_num_pass_retrained_95}{0}
\DefMacro{retrain_debug_approach_96}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_96}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_96}{Short sentences with neutral adjectives and nouns}
\DefMacro{retrain_debug_num_fail2pass_96}{197}
\DefMacro{retrain_debug_num_fail_orig_96}{197}
\DefMacro{retrain_debug_num_pass_orig_96}{32}
\DefMacro{retrain_debug_num_fail_retrained_96}{0}
\DefMacro{retrain_debug_num_pass_retrained_96}{229}
\DefMacro{retrain_debug_approach_97}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_97}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_97}{Short sentences with sentiment-laden adjectives}
\DefMacro{retrain_debug_num_fail2pass_97}{0}
\DefMacro{retrain_debug_num_fail_orig_97}{19}
\DefMacro{retrain_debug_num_pass_orig_97}{425}
\DefMacro{retrain_debug_num_fail_retrained_97}{444}
\DefMacro{retrain_debug_num_pass_retrained_97}{0}
\DefMacro{retrain_debug_approach_98}{Retrain:Checklist::Test:Ours}
\DefMacro{retrain_debug_retrained_lc_98}{['Q & A: yes'  'Q & A: yes (neutral)']}
\DefMacro{retrain_debug_eval_lc_98}{parsing sentiment in (question  yes) form}
\DefMacro{retrain_debug_num_fail2pass_98}{2}
\DefMacro{retrain_debug_num_fail_orig_98}{39}
\DefMacro{retrain_debug_num_pass_orig_98}{1384}
\DefMacro{retrain_debug_num_fail_retrained_98}{1240}
\DefMacro{retrain_debug_num_pass_retrained_98}{183}