### 2022-04-28 paper/ase22/abstract.tex
modify abstract

\begin{abstract}
  \Nlp (NLP) technique becomes one of the core techniques for
  developing text analytics applications.  For developing the NLP
  applications, the applications are required to achieve high
  reliability before it goes to market.  The trustworthiness of the
  prevalent NLP applications is obtained by measuring the accuracy of
  the applications on held-out dataset.  However, evaluating NLP on
  testset with the held-out accuracy is limited to validate its
  quality because the held-out datasets are often not
  comprehensive. Along with this, Evaluating a NLP model on
  task-specific behaviours defined on empirical linguistic
  capabilities has been introduced. However, it relies on manually
  created test cases, and is still limited to measure the model
  performance on biased testcases. In this work, we introduce \tool,
  an NLP model testing methodology.  Given a linguistic capability,
  The \tool finds suitable testcases to test the linguistic capability
  from existing datasets as seed inputs, generates sufficient number
  of new test cases by fuzzing the seed inputs based on their \cfg
  (CFG).  We illustrate the usefulness of the \tool by showing its
  reliability on generated testcases and its generalization ability.
  In our experiment, we also show that \tool facilitates
  identification of critical failures and its origins in NLP models
  for \sa task.
\end{abstract}