### 2022-04-30 paper/ase22/abstract.tex
abstract

\begin{abstract}
  \Nlp (NLP) technique has become one of the core techniques for
  developing text analytic applications.  These applications are
  required to achieve high reliability to be useful in practice.  The
  trustworthiness of the prevalent NLP applications is obtained by
  measuring the accuracy of the applications on held-out dataset.
  However, evaluating NLP on testset with the held-out accuracy is
  limited in validating its overall quality because the held-out
  datasets are often not comprehensive. Along with this, evaluating an
  NLP model on task-specific behaviors defined on empirical linguistic
  capabilities has been introduced. However, such evaluation relies on
  manually created test cases, and is still limited to measure the
  model performance on biased dataset. In this work, we introduce
  \tool, an NLP model testing infrastructure.  Given a linguistic
  capability that users want to evaluate for a NLP model, \tool finds
  suitable seed inputs from existing datasets, generates sufficient
  number of new test inputs by fuzzing the seed inputs based on their
  \cfg (CFG).  We evaluate \tool by showing its reliability on
  generated inputs and its generalization ability.  In our
  experiments, we also show that \tool facilitates identification of
  critical failures and its origins in the NLP models for \sa task.
\end{abstract}