### 2022-05-01 paper/ase22/intro.tex
intro

\section{Introduction}
\label{sec:intro}

%% [Thursday 3:01 PM] Yang, Wei

%% 1. existing NLP testing practice.

%% 2. Limitation in three perspectives:
%% (i) automated input generation,
%% (ii) test oracle (of generated input),
%% (iii) metrics/capabilities to debug the issues

%% 3. Issues with existing techniques: cannot satisfy all three requirements. Give some example, checklist (ii) and (iii), adversarial
%% testing (i) (ii) ....

%% 4. To address the limitation of existing work, we propose....

%% 5. insight that enable the three requirements

In the early stage of the software development process, automated
software testing and debugging can identify and fix
defects. Therefore, effective software testing assures software
quality, and it meets needs of the end user. Nowdays, software at work
has elements of \ml (ML). Especially, \Nlp (NLP) applications in
software are growing exponentially. Therefore, as quality assurance of
software is an essential process in the software development
processes, trustworthiness in quality of NLP application has become an
important component for its practical use in real world.
%% \sw{Should we lead with this paragraph? We are talking about
%%   functional testing here but next we say NLP uses
%%   train-validation-test.}
Traditionally, the prevalent models of NLP are evaluated via
train-validation-test splits. Training and validation set are used to
train the NLP models, and the test set, also named as \ho set, is used
for testing the trained model. Meanwhile, the model performance is
estimated into numbers used as performance metric. Especially,
accuracy, the fraction of model output that the model correctly
predicts, is the most widely the performance metric for classification
models. Accordingly, a model is considered better if it has higher
accuracy than another model.

Despite its simplicity and usefulness, there are several limitations
for the prevalent testing paradigm: First, the testing paradigm mostly
requires manual work for collecting and generating the data. The
manual work is costly with respect to time consumption and its impact
on market price. Therefore, it necessitates automated data generation
for improving model testing approach. Second, this testing paradigm
often overestimates the model performances from the \ho
set~\cite{patel2018mlevalforsoftware, recht2019imagenetbias,
  marcoACL2020checklist}.
%% \sw{Not sure what overestimate means here}
The overestimation comes from the discrepancy between distribution of
data collected and actual distribution in real world.  Oftentimes, the
\ho dataset is not representative and it is likely to introduce
specific biases. In addition, the inconsistency between test input and
its oracle causes biases. Such biases increase the discrepancy of
distribution between the dataset and real-world data. Consequently,
representiveness of the test data is required for testing approach.
Not only that, forced aggregation static into a single number such as
average makes user to difficult to localize and fix the bug found in
\ho set. Therefore, making this testing method fails to validate model
behaviors resulting in localization of the causes of the inaccuracy at
a high. In cost~\cite{wu2019errudite}. Therefore behavioral testing
approach provides the better ability to examine the inaccuracy and
find the bugs in the model.

To address these challenges, several approaches have been proposed to
evaluate different aspects of the NLP models, such as robustness on
adversarial examples \cite{ribeiro2018sear,belinkov2018breaknmt,
  rychalska2019wildnlp,iyyer2018adversarial}, model coverage
\cite{rottger2020hatecheck}, and fairness
\cite{prabhakaran2019fairness,rottger2020hatecheck}. In addition,
Ribeiro~\etal introduced \Cklst, a behavioral testing framework for
evaluating NLP model on multiple linguistic
capabilities~\cite{marcoACL2020checklist}. \Cklst defines
task-relevant linguistic capabilities and generates testcases for each
\lc. In spite of their effectiveness, they have limitations with
respective to model evaluation. For example, Testing model coverage
only shows model behaviors, but not the model performance on existing
testing set. The fairness and robustness testing only focus on only
one model capability resulting in its limited comprehensiveness. In
addition, \Cklst relies on manually generated input templates; thus,
the template generation requires manual work, and it needs to be
preset before test data generation. Consequently, the templates
becomes distributed in limited range of their structure. It restricts
its ability to comprehensively test the linguistic capabilities.

Despite the limitations, \Cklst approach of testing the linguistic
capabilities is a promising direction. Each \lc explains the
functionality of the input and output behavior for the NLP model under
test. Typically, it describes certain type of input and outputs
observed in real world for the target NLP task ranging from simple to
complicated behaviors so that they are able to evaluate the NLP model
comprehensively. For exmaple, a linguistic capability of ``Negated
neutral should still be neutral.'' measures how accurately the \sa
model understands the negative neutral input as an neutral
sentiment. Therefore, it requires the \sa model to output neural
sentiment on the negated neutral input.  Such methodology of
evaluation on the specified functionalities avoids the overestimation
of the model performance as it equivalently measures the model
performance on each functionality, and the separate model performance
explain distribution of the performance over the linguistic
capabilities. In the end, it provides not only the overall model
performance, but also the malfunction facets of the model.

To address the limitations of existing works, we present \tool, an
automated NLP model evaluation method for comprehensive behavioral
testing of NLP models on \sa task. We first address increasing input
representiveness. Compared with templates used in \Cklst, \tool
instead establishes input requirement for evaluating a linguistic
capability and finds suitable inputs that meet the requirement from
existing public dataset. In this process, \tool applies the fuzzing
testing principle to generate inputs by mutating the selected inputs
as seed inputs. Fuzzer in \tool first expands seed input grammar
structures and determines its available \pos to maintain structural
naturalness. After that, to hold contextual naturalness of the mutated
inputs, the fuzzer completes the expanded new structures via
data-driven context-aware word suggestion. Additionally,
sentiment-independent words in the inputs are replaced with rule-based
word suggestion. Further, we address the manual input generation
process by automating the process mentioned above. Lastly, we adopts
behavioral model testing method by introducing multiple behavior of
linguistic capabilities on \sa task and generating inputs relevant to
each linguistic capability.

We demonstrate its generality and utility as a NLP model evaluation
tool by evaluating well-known \sa models: \Bert~\cite{devlin2019bert},
\Roberta~\cite{liu2019roberta} and \Dbert~\cite{sanh2019distilbert}.
We show that ...