### 2022-05-02 paper/ase22/experiment.tex
Testcases.py

\section{Experimental Setup}
\label{sec:experiment}
%

% \InputWithSpace{tables/retrain-debug-table}

In this section, we present the setup of the experiments to evaluate the effectiveness
of \tool{}. We address the
following research questions (RQs):

\sw{We may miss a RQ for the test results using \tool. In the setup,
  we have not said which sentiment analysis models we tested, and how
  we measure the results (e.g., number of misclassified test cases).}

\begin{enumerate}[label=\textbf{RQ\arabic*}]
\item \label{rq:one}: Can \tool generate test cases with the correct \lc categorization and
  sentiment label?
\item \label{rq:two}: Can \tool generate more diverse test cases than \Cklst?
\item \label{rq:three}: Can \tool be useful to find root causes of bugs in the \sa models?
%% \item \label{rq:three}: How effective is our new test case generation
%%   using \cfg expansion? % ablation study
\end{enumerate}

%To answer the RQs, we need to show: (\romnum{1}) correctness of testcase to
%evaluate its target \lc.  (\romnum{2}) effect of testcase distribution on
%finding bugs; (\romnum{3}) degree of execution of a \sa model on testcases;
%and (\romnum{4}) ability to guide to find cause of bug in a \sa
%model.

\sw{Make clear (earlier in the paper): a test case is a sentence in a linguistic capability with a sentiment label.}

\paragraph{RQ1.} As described in Section \ref{sec:approach}, \tool generates test cases in
two steps: specification-based seed generation and syntax-based sentence expansion. These
automated steps may generate seed/expanded sentences marked with incorrect sentiment labels or categorized into
wrong linguistic capabilities. For example, the search rule and template defined in a linguistic capability
may not always generate seed sentences in that capability or with the correct label.
To answer {\bf RQ1}, we perform a manual study to measure the correctness of the sentiment labels and linguistic capabilities
associated with the seed/expanded sentences, produced by \tool

\sw{@Jaeseong: rewrite the next two paragraphs to reflect the updated manual study. Also, be explicit about the setup: sample size, population size, number of participants, protocol (each given how many sentences, and how disagreement is resolved), etc.}
In the manual study, we sample a set of test cases and corresponding \lc randomly sampled
from seed sentences and expanded sentences from \tool. 
also provided with the sampled test cases. Each subject is asked for
scoring relevancy between the linguistic capability of the
sentences. The score is discrete and it ranges from 1 to 5
representing ``strongly not relevant'' to ``strongly relevant''
respectively. Additionally, the subject scores the relevancy between
sentences and their sentiment labels. The score is also discreate and
ranges from 1 to 5. It represents ``strongly negative'' to ``strongly
positive'' respectively. In this work, we collect manual study scores
from 8 subjects attended. From the collected scores, we measure the
following metrics:

\begin{eqnarray}
  &Reported\:bugs &=\sum_{i} \delta(label_{S^2LCT}!=pred_{model}) \label{metric:rb}\\
  &Incorrect\:inputs &= \sum_{i} \delta(pred_{model}!=label_{human}) \label{metric:ii}
  %% &Label\:inconsistency &= \sum_{i} \delta(label_{S^2LCT}!=label_{human}) \label{metric:li}
\end{eqnarray}

The metrics represents the number of data instance that its label
assigned from are different between \tool and its model prediction
(eq~\ref{metric:rb}) and between the model prediction and human
(eq~\ref{metric:ii}). We show the correspondence between \tool and
human by computing the precision and recall from the reported bugs and
incorrect inputs.
%% and between \tool and human (eq~\ref{metric:li}).

\paragraph{RQ2.}
Recall that a key limitation of \Cklst is that its template-based
approach that relies on significant manual efforts may not generate test cases that comprehensively cover the sentences
in a linguistic capability. \tool, instead, automatically generates test cases based on a search
dataset and the syntax in a large reference corpus. We expect \tool can generate a more diverse test suite
than \Cklst. To measure diversity, we follow the approach presented by Ma et al. \cite{ma2018deepgauge},
where the authors measure the coverage of NLP model intermediate states as corner-case neurons.
\sw{Why is this a good metric for diversity?}
Specifically, we use the  \sw{are these metrics we define or are they from ma2018deepgauge?} two coverage metrics in existing work \cite{ma2018deepgauge}, \textit{boundary coverage} (BoundCov) and \textit{strong activation coverage} (SActCov), as our metrics to evaluate the test suite diversity.

\begin{equation}
\begin{split}
    \text{UpperCornerNeuron}(\mathcal{X}) = \{n \in N | \exists x \in \mathcal{X}: f_n(x) \in (high_n, +\infty)\}; \\
    \text{LowerCornerNeuron}(\mathcal{X}) = \{n \in N | \exists x \in \mathcal{X}: f_n(x) \in (-\infty, low_n)\}; \\
\end{split}
    \label{eq:corner}
\end{equation}

\noindent Eq. \ref{eq:corner} shows the formal definition of the corner-case neuron of the NLP model $f(\cdot)$, where $\mathcal{X}$ is the given test suite, $N$ is the number of neurons in model $f(\cdot)$, $f_n(\cdot)$ is the $n^{th}$ neuron's output, and $high_n, low_n$ are the $n^{th}$ neurons' output bounds on the model training dataset.
Eq. \ref{eq:corner} can be interpreted as the collection of neurons that emit outputs beyond the model's numerical boundary.

\begin{equation}
\begin{split}
     & BoundCov(\mathcal{X}) = \frac{|UpperCornerNeuron(\mathcal{X})| + |LowerCornerNeuron| }{2 \times |N|} \\ 
     & SActCov(\mathcal{X}) = \frac{|UpperCornerNeuron(\mathcal{X})|} {|N|} \\ 
\end{split}
    \label{eq:coverage}
\end{equation}

\noindent The formal definition of our coverage metrics are shown in Eq.\ref{eq:coverage}, where BoundCov measures the coverage of neurons that produce outputs that exceed the upper or lower bounds, and SActCov measures the coverage of neurons that create outputs that exceed the lower bound.
Higher coverage indicates the test suite is better for triggering the corner-case neurons, thus better test suite diversity.

% ($Cov(\mathcal{X})$ in \equref{eq:coverage}),
% where $N$ is the total number blocks, $\mathbb I(\cdot)$ is the indicator function, and
% $(B_i(x) > \tau_i))$ represents whether $i^{th}$ block is activated by input $x$~(the definition of $B_i$ and $\tau_i$ are the same with \equref{eq:new2} and \equref{eq:new3}).
% Because AdNNs activate different blocks for decision making, then a higher block coverage indicates the test samples cover more decision behaviors.

To answer {\bf RQ2}, for each NLP model under test, we first feed its training dataset to compute each neuron's lower and upper bounds. After that, we randomly select 100 \sw{Why 100, not all?} test cases from \tool and \Cklst as the test suite and compute the corresponding coverage metrics. 
\cm{we repeat this process and record both the average and variance value of each coverage}
% For each subject, we randomly select 100 seed samples from the test dataset as seed inputs. We then feed the same seed inputs into \tool and \texttt{ILFO} to generate test samples. 
% Finally, we feed the generated test samples to AdNNs and measure block coverage.
% We repeat this process 10 times and record the average coverage and the variance.
% The results are shown in \tabref{tab:coverage} last two columns.

%
%The amount of \sa model
%components \sw{?} executed during testing is a critical measurement for
%assessing quality of software testing. A high software coverage
%results in higher chances of unidentified bugs in the \sa model. On
%the other hand, limited distribution only represents narrow portion of
%real world covering limited execution behaviors in a \sa model. It leads
%to detect bugs within the restricted execution behaviors.  Therefore,
%test cases more representative of real-world data result in more generalized
%distribution and higher coverage of the \sa model.
%Therefore, we answer the {\bf RQ2} by
%measuring the neural coverage of the \sa model.
%
%Specifically, we implemented DeepXplore to measure the \sa model
%coverage~\cite{pei2017deepxplore}. \Dxp is the first efficient
%white-box testing framework for large-scale \dl systems. It introduces
%neuron coverage of a set of test inputs as the ratio of the number of
%unique activated neurons and the total number of neurons in input \dl
%system. In this experiment, we compute the neuron coverage of a test
%cases from \tool and \Cklst on the fine-tuned \sw{first time mention fine-tuning: reader does not know how we fine-tune} \sa model of
%\bertsamodel and compare the coverage between \tool and \Cklst.

\paragraph{RQ3.} To demonstrate that \tool can help developers to understand the bugs in the sentiment analysis modes, we conduct experiments to visualize

\sw{@Simin: add the setup of the bug explanation case study here.}
%In addition to the detection of bugs \sw{have we defined ``bug" in this context?} in the model,
%explanation of the bugs is also important for debugging and
%repairing the model. Therefore, we answer {\bf RQ3} by analyzing the \sa model based on \tool test cases to find the root causes of the bugs.
%
%Specifically, we adapt \Denas{} for this experiment. \sw{Say more what DENAS does.} Rules generated from
%\Denas{} are interpretable functions mapping certain features of the
%input to the expected output of a deep \nn system, and the generated
%rules are considered as the behaviors of the deep \nn system. To
%analyze the bug of a \sa model, we generates rules over the test
%inputs using \Denas{} and identifies faulty rules for the failed test
%inputs.
%\sw{Previous sounds vague. Any specific adaptation of DENAS we did?}
%We manually identify the root of the faulty rules and find the
%root causes of bug in the end.

\noindent\textbf{Implementation Details.}

\sw{Missing: environment running these experiments.}

%% \MyPara{Seed Input Selection}
%% %
%% For each linguistic capability, we first search all sentences that
%% meet its requirement. Among found sentences, we randomly select 10
%% sentences due to memory constraint.

%% \MyPara{Word Sentiment}
%% %
%% we extract sentiments of words using the
%% \Swn~\cite{baccianella2010sentiwordnet}. The \Swn is a publicly
%% available lexical resource of words on Wordnet with three numerical
%% scores of objectivity, positivity and negativity. Sentiment word
%% labels from the scores are classified from the algorithm from Mihaela
%% \etal~\cite{mihaela2017sentiwordnetlabel}.

%% \MyPara{\Cfg Expansion}
%% %
%% We build a reference \Cfg of natural language from the English Penn
%% \Trb corpora~\cite{mitchell1993treebank,nltkTreebankCorporaWebPage}.
%% The corpus is sampled from 2,499 stories from a tree year \Wsj
%% collection The \Trb provides a parsed text corpus with annotation of
%% syntactic and semantic structure. In this experiment We implement the
%% \trb corpora available through \Nltk, which is a suite of libraries
%% and programs for \Nlp for English. In addition, we parse the seed
%% input using into its CFG using the Berkeley Neural
%% Parser~\cite{kitaev2018constituency, kitaev2019multilingual}, a
%% high-accuracy parser with models for 11 languages. The input is a raw
%% text in natural language and the output is the string representation of
%% parse tree. Next after comparing CFGs between reference and seed input,
%% we randomly select 10 expansions for generating templates due to
%% memory constraint.

%% \MyPara{Synonyms}
%% %
%% \Model searches synonyms of each token from synonym sets extracted
%% from \Wrdnt using \Spacy open-source library for NLP.

%% \MyPara{Models}
%% %
%% We evaluate the following \sa models via \Model:
%% \Bert~\cite{devlin2019bert}, \Roberta~\cite{liu2019roberta} and
%% \Dbert~\cite{sanh2019distilbert}. These models are fine-tuned on \Sstt
%% and their accuracies are \BertAcc, \RobertaAcc and \DbertAcc.

%% \MyPara{Retraining}
%% %
%% We retrain \sa models. we split \Model generated test cases into
%% train/validation/test sets with the ratio of 8:1:1. The number of
%% epochs and batch size for retraining are 1 and 16 respectively.

%\input{CM/coverage}

%\input{CM/explain}