### 2022-05-04 paper/ase22/result.tex
add table in result tex

\section{Result}
\label{sec:result}

\subsection{RQ1: Can \tool generate consistent test sentence and
  its oracle?}

\InputWithSpace{tables/manual-study-table}