### 2022-05-04 paper/ase22/result.tex
table add

\section{Result}
\label{sec:result}

\section{Experimental Results}
\label{sec:result}

\subsection{RQ1: \tool Sentiment Label Consistency}
\InputWithSpace{tables/manual-study-table}

\subsection{RQ2: Correctness of Linguistic Capability Categorization}

\subsection{RQ3: Test Suite Diversity}

Table \ref{tab:coverage} shows that ....

\subsection{RQ4: Use \tool for Debugging}