### 2022-05-05 paper/ase22/tables/test-results-numbers.tex
table

%% Automatically generated by pyutil.latex

\DefMacro{test-results-lc0}{Author sentiment is more important than of others}
\DefMacro{test-results-bl-lc0-num-sents}{8528}
\DefMacro{test-results-bl-model0-lc0-num-fail}{3741}
\DefMacro{test-results-lc1}{Negated negative should be positive or neutral}
\DefMacro{test-results-bl-lc1-num-sents}{6786}
\DefMacro{test-results-bl-model0-lc1-num-fail}{799}
\DefMacro{test-results-lc2}{Negated neutral should still be neutral}
\DefMacro{test-results-bl-lc2-num-sents}{2496}
\DefMacro{test-results-bl-model0-lc2-num-fail}{2427}
\DefMacro{test-results-lc3}{Negated positive with neutral content in the middle}
\DefMacro{test-results-bl-lc3-num-sents}{1000}
\DefMacro{test-results-bl-model0-lc3-num-fail}{860}
\DefMacro{test-results-lc4}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-bl-lc4-num-sents}{2124}
\DefMacro{test-results-bl-model0-lc4-num-fail}{1871}
\DefMacro{test-results-lc5}{Sentiment change over time, present should prevail}
\DefMacro{test-results-bl-lc5-num-sents}{8000}
\DefMacro{test-results-bl-model0-lc5-num-fail}{1680}
\DefMacro{test-results-lc6}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-bl-lc6-num-sents}{1716}
\DefMacro{test-results-bl-model0-lc6-num-fail}{1330}
\DefMacro{test-results-lc7}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-bl-lc7-num-sents}{8658}
\DefMacro{test-results-bl-model0-lc7-num-fail}{26}
\DefMacro{test-results-lc8}{Parsing positive and negative sentiment in (question, no) form}
\DefMacro{test-results-bl-lc8-num-sents}{7644}
\DefMacro{test-results-bl-model0-lc8-num-fail}{4056}
\DefMacro{test-results-lc9}{parsing sentiment in (question, yes) form}
\DefMacro{test-results-bl-lc9-num-sents}{9204}
\DefMacro{test-results-bl-model0-lc9-num-fail}{1793}
\DefMacro{test-results-bl-model1-lc0-num-fail}{2692}
\DefMacro{test-results-bl-model1-lc1-num-fail}{218}
\DefMacro{test-results-bl-model1-lc2-num-fail}{2304}
\DefMacro{test-results-bl-model1-lc3-num-fail}{416}
\DefMacro{test-results-bl-model1-lc4-num-fail}{445}
\DefMacro{test-results-bl-model1-lc5-num-fail}{829}
\DefMacro{test-results-bl-model1-lc6-num-fail}{1391}
\DefMacro{test-results-bl-model1-lc7-num-fail}{139}
\DefMacro{test-results-bl-model1-lc8-num-fail}{4576}
\DefMacro{test-results-bl-model1-lc9-num-fail}{1262}
\DefMacro{test-results-bl-model2-lc0-num-fail}{3535}
\DefMacro{test-results-bl-model2-lc1-num-fail}{734}
\DefMacro{test-results-bl-model2-lc2-num-fail}{2450}
\DefMacro{test-results-bl-model2-lc3-num-fail}{865}
\DefMacro{test-results-bl-model2-lc4-num-fail}{2124}
\DefMacro{test-results-bl-model2-lc5-num-fail}{2532}
\DefMacro{test-results-bl-model2-lc6-num-fail}{1661}
\DefMacro{test-results-bl-model2-lc7-num-fail}{125}
\DefMacro{test-results-bl-model2-lc8-num-fail}{6440}
\DefMacro{test-results-bl-model2-lc9-num-fail}{1557}
\DefMacro{test-results-lc0-num-seeds}{50}
\DefMacro{test-results-lc0-num-exps}{2323}
\DefMacro{test-results-model0-lc0-num-seed-fail}{22}
\DefMacro{test-results-model0-lc0-num-exp-fail}{1163}
\DefMacro{test-results-model0-lc0-num-pass-to-fail}{134}
\DefMacro{test-results-lc1-num-seeds}{50}
\DefMacro{test-results-lc1-num-exps}{1784}
\DefMacro{test-results-model0-lc1-num-seed-fail}{46}
\DefMacro{test-results-model0-lc1-num-exp-fail}{1680}
\DefMacro{test-results-model0-lc1-num-pass-to-fail}{49}
\DefMacro{test-results-lc2-num-seeds}{26}
\DefMacro{test-results-lc2-num-exps}{1009}
\DefMacro{test-results-model0-lc2-num-seed-fail}{24}
\DefMacro{test-results-model0-lc2-num-exp-fail}{932}
\DefMacro{test-results-model0-lc2-num-pass-to-fail}{62}
\DefMacro{test-results-lc3-num-seeds}{50}
\DefMacro{test-results-lc3-num-exps}{1634}
\DefMacro{test-results-model0-lc3-num-seed-fail}{42}
\DefMacro{test-results-model0-lc3-num-exp-fail}{1438}
\DefMacro{test-results-model0-lc3-num-pass-to-fail}{40}
\DefMacro{test-results-lc4-num-seeds}{50}
\DefMacro{test-results-lc4-num-exps}{1486}
\DefMacro{test-results-model0-lc4-num-seed-fail}{50}
\DefMacro{test-results-model0-lc4-num-exp-fail}{1486}
\DefMacro{test-results-model0-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-lc8-num-seeds}{100}
\DefMacro{test-results-lc8-num-exps}{2379}
\DefMacro{test-results-model0-lc8-num-seed-fail}{71}
\DefMacro{test-results-model0-lc8-num-exp-fail}{1723}
\DefMacro{test-results-model0-lc8-num-pass-to-fail}{40}
\DefMacro{test-results-lc5-num-seeds}{50}
\DefMacro{test-results-lc5-num-exps}{-}
\DefMacro{test-results-model0-lc5-num-seed-fail}{12}
\DefMacro{test-results-model0-lc5-num-exp-fail}{-}
\DefMacro{test-results-model0-lc5-num-pass-to-fail}{-}
\DefMacro{test-results-lc6-num-seeds}{19}
\DefMacro{test-results-lc6-num-exps}{210}
\DefMacro{test-results-model0-lc6-num-seed-fail}{15}
\DefMacro{test-results-model0-lc6-num-exp-fail}{182}
\DefMacro{test-results-model0-lc6-num-pass-to-fail}{19}
\DefMacro{test-results-lc7-num-seeds}{50}
\DefMacro{test-results-lc7-num-exps}{394}
\DefMacro{test-results-model0-lc7-num-seed-fail}{2}
\DefMacro{test-results-model0-lc7-num-exp-fail}{17}
\DefMacro{test-results-model0-lc7-num-pass-to-fail}{4}
\DefMacro{test-results-lc9-num-seeds}{50}
\DefMacro{test-results-lc9-num-exps}{1373}
\DefMacro{test-results-model0-lc9-num-seed-fail}{2}
\DefMacro{test-results-model0-lc9-num-exp-fail}{37}
\DefMacro{test-results-model0-lc9-num-pass-to-fail}{43}
\DefMacro{test-results-model1-lc0-num-seed-fail}{13}
\DefMacro{test-results-model1-lc0-num-exp-fail}{725}
\DefMacro{test-results-model1-lc0-num-pass-to-fail}{78}
\DefMacro{test-results-model1-lc1-num-seed-fail}{41}
\DefMacro{test-results-model1-lc1-num-exp-fail}{1537}
\DefMacro{test-results-model1-lc1-num-pass-to-fail}{64}
\DefMacro{test-results-model1-lc2-num-seed-fail}{24}
\DefMacro{test-results-model1-lc2-num-exp-fail}{939}
\DefMacro{test-results-model1-lc2-num-pass-to-fail}{38}
\DefMacro{test-results-model1-lc3-num-seed-fail}{21}
\DefMacro{test-results-model1-lc3-num-exp-fail}{664}
\DefMacro{test-results-model1-lc3-num-pass-to-fail}{54}
\DefMacro{test-results-model1-lc4-num-seed-fail}{50}
\DefMacro{test-results-model1-lc4-num-exp-fail}{1486}
\DefMacro{test-results-model1-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-model1-lc8-num-seed-fail}{79}
\DefMacro{test-results-model1-lc8-num-exp-fail}{1947}
\DefMacro{test-results-model1-lc8-num-pass-to-fail}{18}
\DefMacro{test-results-model1-lc5-num-seed-fail}{28}
\DefMacro{test-results-model1-lc5-num-exp-fail}{-}
\DefMacro{test-results-model1-lc5-num-pass-to-fail}{-}
\DefMacro{test-results-model1-lc6-num-seed-fail}{17}
\DefMacro{test-results-model1-lc6-num-exp-fail}{160}
\DefMacro{test-results-model1-lc6-num-pass-to-fail}{9}
\DefMacro{test-results-model1-lc7-num-seed-fail}{2}
\DefMacro{test-results-model1-lc7-num-exp-fail}{19}
\DefMacro{test-results-model1-lc7-num-pass-to-fail}{5}
\DefMacro{test-results-model1-lc9-num-seed-fail}{1}
\DefMacro{test-results-model1-lc9-num-exp-fail}{57}
\DefMacro{test-results-model1-lc9-num-pass-to-fail}{46}
\DefMacro{test-results-model2-lc0-num-seed-fail}{20}
\DefMacro{test-results-model2-lc0-num-exp-fail}{1040}
\DefMacro{test-results-model2-lc0-num-pass-to-fail}{31}
\DefMacro{test-results-model2-lc1-num-seed-fail}{44}
\DefMacro{test-results-model2-lc1-num-exp-fail}{1579}
\DefMacro{test-results-model2-lc1-num-pass-to-fail}{26}
\DefMacro{test-results-model2-lc2-num-seed-fail}{24}
\DefMacro{test-results-model2-lc2-num-exp-fail}{961}
\DefMacro{test-results-model2-lc2-num-pass-to-fail}{74}
\DefMacro{test-results-model2-lc3-num-seed-fail}{38}
\DefMacro{test-results-model2-lc3-num-exp-fail}{1280}
\DefMacro{test-results-model2-lc3-num-pass-to-fail}{9}
\DefMacro{test-results-model2-lc4-num-seed-fail}{50}
\DefMacro{test-results-model2-lc4-num-exp-fail}{1486}
\DefMacro{test-results-model2-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-model2-lc8-num-seed-fail}{93}
\DefMacro{test-results-model2-lc8-num-exp-fail}{2216}
\DefMacro{test-results-model2-lc8-num-pass-to-fail}{12}
\DefMacro{test-results-model2-lc5-num-seed-fail}{39}
\DefMacro{test-results-model2-lc5-num-exp-fail}{-}
\DefMacro{test-results-model2-lc5-num-pass-to-fail}{-}
\DefMacro{test-results-model2-lc6-num-seed-fail}{19}
\DefMacro{test-results-model2-lc6-num-exp-fail}{198}
\DefMacro{test-results-model2-lc6-num-pass-to-fail}{0}
\DefMacro{test-results-model2-lc7-num-seed-fail}{1}
\DefMacro{test-results-model2-lc7-num-exp-fail}{11}
\DefMacro{test-results-model2-lc7-num-pass-to-fail}{8}
\DefMacro{test-results-model2-lc9-num-seed-fail}{3}
\DefMacro{test-results-model2-lc9-num-exp-fail}{76}
\DefMacro{test-results-model2-lc9-num-pass-to-fail}{14}