### 2022-05-05 paper/ase22/tables/test-results-table.tex

%% Automatically generated by pyutil.latex

\begin{table*}[t]
\begin{small}
\begin{center}
\caption{\TestResultsTableCaption}
\begin{adjustbox}{angle=90}
\begin{tabular}{p{5cm}||ccccccc}
\toprule
\tLc & \tNumBlSents & \tNumBlFail & \tNumSeeds & \tNumSeedFail & \tNumExps & \tNumExpFail & \tNumPasstoFail\\
\midrule
\multirow{3}{*}{\parbox{5cm}{LC1: \UseMacro{test-results-lc0}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc0-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc0-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc0-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc0-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc0-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc0-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc0-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc0-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc0-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc0-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc0-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc0-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc0-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc0-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc0-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC2: \UseMacro{test-results-lc1}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc1-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc1-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc1-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc1-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc1-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc1-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc1-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc1-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc1-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc1-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc1-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc1-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc1-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc1-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc1-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC3: \UseMacro{test-results-lc2}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc2-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc2-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc2-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc2-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc2-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc2-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc2-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc2-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc2-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc2-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc2-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc2-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc2-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc2-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc2-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC4: \UseMacro{test-results-lc3}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc3-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc3-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc3-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc3-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc3-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc3-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc3-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc3-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc3-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc3-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc3-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc3-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc3-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc3-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc3-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC5: \UseMacro{test-results-lc4}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc4-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc4-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc4-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc4-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc4-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc4-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc4-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc4-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc4-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc4-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc4-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc4-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc4-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc4-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc4-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC6: \UseMacro{test-results-lc5}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc5-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc5-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc5-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc5-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc5-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc5-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc5-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc5-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc5-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc5-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc5-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc5-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc5-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc5-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc5-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC7: \UseMacro{test-results-lc6}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc6-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc6-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc6-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc6-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc6-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc6-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc6-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc6-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc6-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc6-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc6-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc6-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc6-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc6-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc6-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC8: \UseMacro{test-results-lc7}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc7-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc7-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc7-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc7-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc7-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc7-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc7-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc7-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc7-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc7-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc7-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc7-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc7-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc7-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc7-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC9: \UseMacro{test-results-lc8}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc8-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc8-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc8-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc8-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc8-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc8-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc8-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc8-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc8-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc8-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc8-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc8-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc8-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc8-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc8-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{5cm}{LC10: \UseMacro{test-results-lc9}}}
 & \multirow{3}{*}{\UseMacro{test-results-bl-lc9-num-sents}}
 & Model1$\colon$\UseMacro{test-results-bl-model0-lc9-num-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc9-num-seeds}}
 & Model1$\colon$\UseMacro{test-results-model0-lc9-num-seed-fail}
 & \multirow{3}{*}{\UseMacro{test-results-lc9-num-exps}}
 & Model1$\colon$\UseMacro{test-results-model0-lc9-num-exp-fail}
 & Model1$\colon$\UseMacro{test-results-model0-lc9-num-pass-to-fail}\\
 & & Model2$\colon$\UseMacro{test-results-bl-model1-lc9-num-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc9-num-seed-fail}
 & & Model2$\colon$\UseMacro{test-results-model1-lc9-num-exp-fail}
 & Model2$\colon$\UseMacro{test-results-model1-lc9-num-pass-to-fail}\\
 & & Model3$\colon$\UseMacro{test-results-bl-model2-lc9-num-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc9-num-seed-fail}
 & & Model3$\colon$\UseMacro{test-results-model2-lc9-num-exp-fail}
 & Model3$\colon$\UseMacro{test-results-model2-lc9-num-pass-to-fail}\\
\hline
\bottomrule
\end{tabular}
\end{adjustbox}
\end{center}
\end{small}
\vspace{\TestResultsTableVSpace}
\end{table*}