### 2022-05-06 paper/ase22/result.tex

\section{Experimental Results}
\label{sec:result}

This section presents experiment results and answer the RQs by
studying the results quantatively and qualitatively.

\InputWithSpace{tables/manual-study-table}
\InputWithSpace{tables/test-results-table}

\subsection{RQ1: \tool Sentiment Label Consistency}
Table~\ref{table:ManualStudy} shows the statistics of scores collected
from the manual study of \tool. First column represents type of test
case. The number of test cases used for the study is represented in
second column. Label consistency score and \lc relevancy score defined
at equation~\ref{metric:srel} and~\ref{metric:lcrel} are shown at the
remaining columns. First, it is shown that high label consistency
score over 0.8. It means that \tool generates test oracles consistent
to human understanding. It is also observed that there is no
difference of the scores between seed and expanded sentences. The
observation implies that the expansion method in \tool preserves
sentiment of expanded sentences as its seed, and provides its
reliability. These observations answer RQ1 and conclude that \tool
generates test case with high consistency between test sentence and
its oracle.

\subsection{RQ2: Correctness of Linguistic Capability Categorization}
The result in table~\ref{table:ManualStudy} also shows that \tool
achieves high order of agreement with human assessment on \lc
relevancy. In parallel, it is observed that the expanded sentences
generated from \tool also have same level of \lc
relevancy. Accordingly, we answer the RQ2 by concluding that
\tool generates \lc relevant test cases with the agreement with human.

\subsection{RQ3: Test Suite Diversity}
\input{CM/coverage}

\subsection{RQ4: Use \tool for Debugging}