### 2022-05-06 paper/ase22/tables/test-results-numbers.tex
revise table3

%% Automatically generated by pyutil.latex

\DefMacro{test-results-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-bl-lc7-num-sents}{8528}
\DefMacro{test-results-bl-model0-lc7-num-fail}{43.87}
\DefMacro{test-results-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-bl-lc3-num-sents}{6786}
\DefMacro{test-results-bl-model0-lc3-num-fail}{11.77}
\DefMacro{test-results-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-bl-lc4-num-sents}{2496}
\DefMacro{test-results-bl-model0-lc4-num-fail}{97.24}
\DefMacro{test-results-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-bl-lc6-num-sents}{1000}
\DefMacro{test-results-bl-model0-lc6-num-fail}{86.00}
\DefMacro{test-results-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-bl-lc5-num-sents}{2124}
\DefMacro{test-results-bl-model0-lc5-num-fail}{88.09}
\DefMacro{test-results-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-bl-lc2-num-sents}{8000}
\DefMacro{test-results-bl-model0-lc2-num-fail}{21.00}
\DefMacro{test-results-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-bl-lc0-num-sents}{1716}
\DefMacro{test-results-bl-model0-lc0-num-fail}{77.51}
\DefMacro{test-results-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-bl-lc1-num-sents}{8658}
\DefMacro{test-results-bl-model0-lc1-num-fail}{0.30}
\DefMacro{test-results-lc9}{Parsing positive and negative sentiment in (question, no) form}
\DefMacro{test-results-bl-lc9-num-sents}{7644}
\DefMacro{test-results-bl-model0-lc9-num-fail}{53.06}
\DefMacro{test-results-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-bl-lc8-num-sents}{9204}
\DefMacro{test-results-bl-model0-lc8-num-fail}{19.48}
\DefMacro{test-results-bl-model1-lc7-num-fail}{31.57}
\DefMacro{test-results-bl-model1-lc3-num-fail}{3.21}
\DefMacro{test-results-bl-model1-lc4-num-fail}{92.31}
\DefMacro{test-results-bl-model1-lc6-num-fail}{41.60}
\DefMacro{test-results-bl-model1-lc5-num-fail}{20.95}
\DefMacro{test-results-bl-model1-lc2-num-fail}{10.36}
\DefMacro{test-results-bl-model1-lc0-num-fail}{81.06}
\DefMacro{test-results-bl-model1-lc1-num-fail}{1.61}
\DefMacro{test-results-bl-model1-lc9-num-fail}{59.86}
\DefMacro{test-results-bl-model1-lc8-num-fail}{13.71}
\DefMacro{test-results-bl-model2-lc7-num-fail}{41.45}
\DefMacro{test-results-bl-model2-lc3-num-fail}{10.82}
\DefMacro{test-results-bl-model2-lc4-num-fail}{98.16}
\DefMacro{test-results-bl-model2-lc6-num-fail}{86.50}
\DefMacro{test-results-bl-model2-lc5-num-fail}{100.00}
\DefMacro{test-results-bl-model2-lc2-num-fail}{31.65}
\DefMacro{test-results-bl-model2-lc0-num-fail}{96.79}
\DefMacro{test-results-bl-model2-lc1-num-fail}{1.44}
\DefMacro{test-results-bl-model2-lc9-num-fail}{84.25}
\DefMacro{test-results-bl-model2-lc8-num-fail}{16.92}
\DefMacro{test-results-lc7-num-seeds}{50}
\DefMacro{test-results-lc7-num-exps}{2323}
\DefMacro{test-results-model0-lc7-num-seed-fail}{44.00}
\DefMacro{test-results-model0-lc7-num-exp-fail}{50.06}
\DefMacro{test-results-model0-lc7-num-pass-to-fail}{134}
\DefMacro{test-results-lc3-num-seeds}{50}
\DefMacro{test-results-lc3-num-exps}{1784}
\DefMacro{test-results-model0-lc3-num-seed-fail}{92.00}
\DefMacro{test-results-model0-lc3-num-exp-fail}{94.17}
\DefMacro{test-results-model0-lc3-num-pass-to-fail}{49}
\DefMacro{test-results-lc4-num-seeds}{26}
\DefMacro{test-results-lc4-num-exps}{1009}
\DefMacro{test-results-model0-lc4-num-seed-fail}{92.31}
\DefMacro{test-results-model0-lc4-num-exp-fail}{92.37}
\DefMacro{test-results-model0-lc4-num-pass-to-fail}{62}
\DefMacro{test-results-lc6-num-seeds}{50}
\DefMacro{test-results-lc6-num-exps}{1634}
\DefMacro{test-results-model0-lc6-num-seed-fail}{84.00}
\DefMacro{test-results-model0-lc6-num-exp-fail}{88.00}
\DefMacro{test-results-model0-lc6-num-pass-to-fail}{40}
\DefMacro{test-results-lc5-num-seeds}{50}
\DefMacro{test-results-lc5-num-exps}{1486}
\DefMacro{test-results-model0-lc5-num-seed-fail}{100.00}
\DefMacro{test-results-model0-lc5-num-exp-fail}{100.00}
\DefMacro{test-results-model0-lc5-num-pass-to-fail}{0}
\DefMacro{test-results-lc9-num-seeds}{100}
\DefMacro{test-results-lc9-num-exps}{2379}
\DefMacro{test-results-model0-lc9-num-seed-fail}{71.00}
\DefMacro{test-results-model0-lc9-num-exp-fail}{72.43}
\DefMacro{test-results-model0-lc9-num-pass-to-fail}{40}
\DefMacro{test-results-lc2-num-seeds}{50}
\DefMacro{test-results-lc2-num-exps}{-}
\DefMacro{test-results-model0-lc2-num-seed-fail}{24.00}
\DefMacro{test-results-model0-lc2-num-exp-fail}{-}
\DefMacro{test-results-model0-lc2-num-pass-to-fail}{-}
\DefMacro{test-results-lc0-num-seeds}{19}
\DefMacro{test-results-lc0-num-exps}{210}
\DefMacro{test-results-model0-lc0-num-seed-fail}{78.95}
\DefMacro{test-results-model0-lc0-num-exp-fail}{86.67}
\DefMacro{test-results-model0-lc0-num-pass-to-fail}{19}
\DefMacro{test-results-lc1-num-seeds}{50}
\DefMacro{test-results-lc1-num-exps}{394}
\DefMacro{test-results-model0-lc1-num-seed-fail}{4.00}
\DefMacro{test-results-model0-lc1-num-exp-fail}{4.31}
\DefMacro{test-results-model0-lc1-num-pass-to-fail}{4}
\DefMacro{test-results-lc8-num-seeds}{50}
\DefMacro{test-results-lc8-num-exps}{1373}
\DefMacro{test-results-model0-lc8-num-seed-fail}{4.00}
\DefMacro{test-results-model0-lc8-num-exp-fail}{2.69}
\DefMacro{test-results-model0-lc8-num-pass-to-fail}{43}
\DefMacro{test-results-model1-lc7-num-seed-fail}{26.00}
\DefMacro{test-results-model1-lc7-num-exp-fail}{31.21}
\DefMacro{test-results-model1-lc7-num-pass-to-fail}{78}
\DefMacro{test-results-model1-lc3-num-seed-fail}{82.00}
\DefMacro{test-results-model1-lc3-num-exp-fail}{86.15}
\DefMacro{test-results-model1-lc3-num-pass-to-fail}{64}
\DefMacro{test-results-model1-lc4-num-seed-fail}{92.31}
\DefMacro{test-results-model1-lc4-num-exp-fail}{93.06}
\DefMacro{test-results-model1-lc4-num-pass-to-fail}{38}
\DefMacro{test-results-model1-lc6-num-seed-fail}{42.00}
\DefMacro{test-results-model1-lc6-num-exp-fail}{40.64}
\DefMacro{test-results-model1-lc6-num-pass-to-fail}{54}
\DefMacro{test-results-model1-lc5-num-seed-fail}{100.00}
\DefMacro{test-results-model1-lc5-num-exp-fail}{100.00}
\DefMacro{test-results-model1-lc5-num-pass-to-fail}{0}
\DefMacro{test-results-model1-lc9-num-seed-fail}{79.00}
\DefMacro{test-results-model1-lc9-num-exp-fail}{81.84}
\DefMacro{test-results-model1-lc9-num-pass-to-fail}{18}
\DefMacro{test-results-model1-lc2-num-seed-fail}{56.00}
\DefMacro{test-results-model1-lc2-num-exp-fail}{-}
\DefMacro{test-results-model1-lc2-num-pass-to-fail}{-}
\DefMacro{test-results-model1-lc0-num-seed-fail}{89.47}
\DefMacro{test-results-model1-lc0-num-exp-fail}{76.19}
\DefMacro{test-results-model1-lc0-num-pass-to-fail}{9}
\DefMacro{test-results-model1-lc1-num-seed-fail}{4.00}
\DefMacro{test-results-model1-lc1-num-exp-fail}{4.82}
\DefMacro{test-results-model1-lc1-num-pass-to-fail}{5}
\DefMacro{test-results-model1-lc8-num-seed-fail}{2.00}
\DefMacro{test-results-model1-lc8-num-exp-fail}{4.15}
\DefMacro{test-results-model1-lc8-num-pass-to-fail}{46}
\DefMacro{test-results-model2-lc7-num-seed-fail}{40.00}
\DefMacro{test-results-model2-lc7-num-exp-fail}{44.77}
\DefMacro{test-results-model2-lc7-num-pass-to-fail}{31}
\DefMacro{test-results-model2-lc3-num-seed-fail}{88.00}
\DefMacro{test-results-model2-lc3-num-exp-fail}{88.51}
\DefMacro{test-results-model2-lc3-num-pass-to-fail}{26}
\DefMacro{test-results-model2-lc4-num-seed-fail}{92.31}
\DefMacro{test-results-model2-lc4-num-exp-fail}{95.24}
\DefMacro{test-results-model2-lc4-num-pass-to-fail}{74}
\DefMacro{test-results-model2-lc6-num-seed-fail}{76.00}
\DefMacro{test-results-model2-lc6-num-exp-fail}{78.34}
\DefMacro{test-results-model2-lc6-num-pass-to-fail}{9}
\DefMacro{test-results-model2-lc5-num-seed-fail}{100.00}
\DefMacro{test-results-model2-lc5-num-exp-fail}{100.00}
\DefMacro{test-results-model2-lc5-num-pass-to-fail}{0}
\DefMacro{test-results-model2-lc9-num-seed-fail}{93.00}
\DefMacro{test-results-model2-lc9-num-exp-fail}{93.15}
\DefMacro{test-results-model2-lc9-num-pass-to-fail}{12}
\DefMacro{test-results-model2-lc2-num-seed-fail}{78.00}
\DefMacro{test-results-model2-lc2-num-exp-fail}{-}
\DefMacro{test-results-model2-lc2-num-pass-to-fail}{-}
\DefMacro{test-results-model2-lc0-num-seed-fail}{100.00}
\DefMacro{test-results-model2-lc0-num-exp-fail}{94.29}
\DefMacro{test-results-model2-lc0-num-pass-to-fail}{0}
\DefMacro{test-results-model2-lc1-num-seed-fail}{2.00}
\DefMacro{test-results-model2-lc1-num-exp-fail}{2.79}
\DefMacro{test-results-model2-lc1-num-pass-to-fail}{8}
\DefMacro{test-results-model2-lc8-num-seed-fail}{6.00}
\DefMacro{test-results-model2-lc8-num-exp-fail}{5.54}
\DefMacro{test-results-model2-lc8-num-pass-to-fail}{14}