### 2022-05-07 paper/ase22/CM/coverage.tex
writing

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/covergae.pdf}
    \caption{The coverage results of the generated test
      samples. \sw{Make the figures larger.}}
    \label{fig:coverage}
\end{figure}

Next, Fig. \ref{fig:coverage} shows the coverage results of the
generated test samples and reliability of the former evaluation
results. In the figure, the red line represents \tool and the black
line represents \Cklst. Each column in Fig. \ref{fig:coverage}
represent the results for one NLP model, the first row is the
\textit{BoundCov} results and the second row is the \textit{SActCov}
results.  From the results, we make three observations
observations. \TODO{First, as the number of test cases increases, the
  test suite can achieve better coverage. Such observation is
  intuitive. However, generating a more extensive test suite is not
  easy, particularly for \Cklst, which is a manually template-based
  approach.}  Second, for \emph{all} experimental settings (\eg NLP
model, coverage metric), \tool achieves high coverage than
\Cklst. Recall that a higher coverage implies the test case in the
test suite is more diverse and rarely to have a statistical
distribution similar to the model training data. As a result, a test
suite with greater coverage complements the model training data
distribution (\ie hold-out testing data) better.  \TODO{For example,
  for the first NLP model under test, \tool can achieve a higher
  coverage than \Cklst with only half test cases.}  Such experimental
results confirm that \tool can generate more diverse test cases to
complement the hold-out testing data for testing NLP models.  \sw{What
  does the growth trend in each figure indicate?} \sw{Does the
  absolute numbers or relative difference of the two lines on y axis
  mean anything concrete? E.g., how significant is the improvement in
  diversity?}

Another interesting finding is that for each NLP model, there is no
fixed relationship between \textit{BoundCov} and \textit{SActCov}. In
other words, while a test suite may produce higher \textit{BoundCov}
for some models, the same test suite may get higher \textit{SActCov}
for other NLP models.  Recall that \textit{BoundCov} measures both the
upper and lower corner neurons and \textit{SActCov} measures only the
upper corner neurons.  Such observation implies that the upper and
lower corner neurons are distributed unevenly, and measuring only one
of them is not enough.