### 2022-05-07 paper/ase22/concl.tex

\section{Conclusions}
\label{sec:concl}

In this paper, we present \tool automatically that finds and generates
suitable seed test cases for each \lc. \tool fuzz the seeds into a
sufficient number of test cases by expanding \cfg of the seed inputs.
We evaluated \sa models on the generated test cases and anlyzed its
impact on testing the models. Further, we examined failure-inducing
cases for seeking the root of the bug.  We also studied the quality of
\tool test cases by measuring their consistencies with their test
oracles and \lc. These led to show generalization ability of \tool for
model evaluation and its reliability in the end. Our work stands to
provide automated and comprehensive NLP model testing tool.