### 2022-05-07 paper/ase22/threats.tex

\section{Threats to Validity}
\label{sec:threats}

Use of publicly available datasets poses an external validity threat
with respect to the generalizability of our results. We implemented
domain-spefic knowledge on publically available word sentiments
dataset~\cite{baccianella2010sentiwordnet}. We also constructed
reference CFG from publically available corpus dataset
\cite{mitchell1993treebank, nltkTreebankCorporaWebPage}. The dataset
used in our study might not be completely representative of all
English grammatical structures and English word sentiments. We
mitigate this threat by observing that they are widely used in NLP
domain~\cite{husnain2021swnvalidity} and reviewed each script and the
output logs.