### 2022-05-07 paper/ase22/related.tex
related

\section{Related Work}
\label{sec:related}

\MyPara{NLP Testing.}
%
With the increasing use of NLP models, evaluation of NLP models is
becoming more important. Apart from the task accuracy based testing
scheme, recent works have also considered model robustness as an
aspect for model evaluation. Belinkov
\textit{\etal}~\cite{belinkov2018breaknmt} aims to fail neural machine
translation model by intentionally introducing noise in the input
text.Pinjia
\textit{\etal}~\cite{pinjia2020structinvtestingnmt,pinjia2020testnmtrt}
measures the robustness by assuming syntatic and semantic relation
between input and output of neural machine translation model.  Ribeiro
\textit{\etal}~\cite{ribeiro2018sear} proposed an approach to
generalize semantically equivalent adversarial rules. In addition,
Rychalska \textit{\etal}~\cite{rychalska2019wildnlp} measures drops in
BLEU scores by corruption operation, and compare model robustness
based on the amount of the drops. In addition, Iyyer
\textit{\etal}~\cite{iyyer2018adversarial} introduced learning based
model for adversarial data augmentation. Not only evaluating the
robustness on adversarial set, but various aspets on the NLP model are
considered for the robustness evaluation. Prabhakaran
\textit{\etal}~\cite{prabhakaran2019fairness} developed an evaluation
framework to detect unintended societal bias in NLP model. Rottger
\textit{\etal}~\cite{rottger2020hatecheck} introduced a functional
test suite for hate speech detection in the NLP model.  Ribeiro
\textit{\etal}~\cite{ribeiro2019consistencyeval} measures logical
consistency of NLP model. These techniques evaluate the robustness of
the NLP model. However, we focused on evaluation of model capability
over multiple perspectives and give debugging information by comparing
seed and expanded test cases.

\MyPara{Linguistic Capability Evaluation}
%
Wang \textit{\etal}~\cite{wang2018glue, wang2019superglue} propose
multiple diagnostic datasets to evaluate NLP models. This dataset
evaluates NLP model's ability to understand input sentence via natural
language inference problems. More recently, \Cklst proposes evaluation
method of input-output behavior defined as \lcs. \Cklst generates
behavior-guided inputs for validating the
behaviors.~\cite{marcoACL2020checklist}. Unlike prior work on manual
data generation method, we used structural information in text to
generate data which enables automated data generation.

\MyPara{NLP Model Debugging}
%
Futher, researchers have been explaining NLP model prediction and
analyze it for debugging the model. Ribeiro
\textit{\etal}~\cite{ribeiroSG16lime} computes the word relevance for
a model prediction are evaluated and implemented crowdcourcing for
model inspection with the word relevance explanation.  Interactive
error analysis~\cite{wu2019errudite} have been proposed to evaluate
model robustness. Zylberajch~\cite{zylberajch2021hildif} used
influence functions for model explanation generation so as to enable
interactive debugging using humans effort as feedback with the
explanation. Lertvittaya~\cite{lertvittayakumjorn2020find} proposed
approach to understand behavior of text classifier model and improve
model by disabling irrelevant hidden features. In this work, \tool
provides source of model failure by comparing performance on seed and
expanded test cases.

%% On the subject of the limitation of traditional testing paradigm, a
%% number of methods have been proposed. First, multiple diagnostic
%% datasets for evaluating NLP model were introduced for obtaining
%% generalized evaluation of the NLP model~\cite{wang2018glue}. Not only
%% that, model is evaluated on different aspects such as robustness of
%% the model on adversarial
%% sets~\cite{ribeiro2018sear,belinkov2018breaknmt,
%%   rychalska2019wildnlp,iyyer2018adversarial},
%% fairness~\cite{prabhakaran2019fairness,rottger2020hatecheck}, logical
%% consistancy~\cite{ribeiro2019consistencyeval}, prediction
%% interpretations~\cite{ribeiroSG16lime} and interactive error
%% analysis~\cite{wu2019errudite}. Especially, \Cklst implements
%% behavioral testing methodolgy for evaluating multiple linguistic
%% capabilities of NLP model~\cite{marcoACL2020checklist}. \Cklst
%% introduces input-output behaviors of linguistic capabilities and
%% generates behavior-guided inputs for validating the behaviors. It
%% provides comprehensive behavioral testing of NLP models through a
%% number of generated inputs. However, the approach only relies on
%% manually generated input templates, thus the template generation
%% becomes expensive and time consuming. In addition, the generated
%% templates are selective and often too simple, and it is limited to
%% provide restricted evaluation of linguistic capabilities. Thus, it
  %% does not garauntee the comprehensive evaluation.