### 2022-06-07 paper/ase22/concl.tex

\vspace{-6pt}
\section{Conclusions}
\label{sec:concl}

In this paper, we present \tool that automatically searches and generates seed test cases for each \lc. \tool expands the seeds into a
sufficient number of test cases based on \cfg, covering diverse syntactic structures.
We evaluated \sa models on the generated test cases. It is shown that \tool ensures quality of test cases by showing consistency of test oracle with its input sentence and its \lc. In addition, we showed that diversity in \tool test cases leads to cover more region in \sa models, resulting in higher reliability on the testing. We also examined failure-inducing
cases to understand the root causes of the bugs. These results demonstrate the correctness and usefulness \tool for
model evaluation.