### 2022-06-07 paper/ase22/experiment.tex

\section{Experimental Setup}
\label{sec:experiment}
%
In this section, we present the setup of the experiments to evaluate
the effectiveness of \tool{}. We address the following research
questions (RQs):

\begin{enumerate}[label=\textbf{RQ\arabic*}]
\item \label{rq:three}: Can \tool generate more diverse test cases
  than \Cklst?
\item \label{rq:one}: Can \tool generate test sentences with correct sentiment labels?
\item \label{rq:two}: Are \tool generated test sentences correctly categorized into a \lc?
\item \label{rq:four}: Can \tool be useful to find root causes of bugs
  in the \sa models?
%% \item \label{rq:three}: How effective is our new test case generation
%%   using \cfg expansion? % ablation study
\end{enumerate}

%To answer the RQs, we need to show: (\romnum{1}) correctness of testcase to
%evaluate its target \lc.  (\romnum{2}) effect of testcase distribution on
%finding bugs; (\romnum{3}) degree of execution of a \sa model on testcases;
%and (\romnum{4}) ability to guide to find cause of bug in a \sa
%model.

\subsection{Experimental Subjects}

\paragraph{\textbf{NLP Models \& Dataset.}}
We evaluate our approach on three learning-based \sa models
implemented in Transformer library from the Hugging Face centralized
Model Hub:\footnote{https://huggingface.co/models} \texttt{\Bert}
(\bertsamodel), \texttt{\Roberta} (\robertasamodel), and \texttt{\Dbert} (\disbertsamodel). These models were pre-trained on English
language using a masked language modeling (MLM) objective, and
fine-tuned on the \sa task. In this experiment, we use \Sstt \cite{socher2013sst} dataset for
searching the seeds in \tool. \Sstt is a
corpus of movie review, and consists of 11,855 sentences
with sentiment scores. We split the
scores into ranges [0, 0.4], (0.4, 0.6] and (0.6, 1.0] to assign 
negative, neutral and positive labels, respectively. In addition, 
we use \Swn as the word sentiment dataset~\cite{baccianella2010sentiwordnet}. 
\Swn is a publicly available English sentiment lexicons.  
It provides lexical sentiment scores and the sentiment word labels 
are categorized by implementing the rules in~\cite{mihaela2017sentiwordnetlabel}.
%% \TODO{introduce details about each model}

\paragraph{\textbf{Comparison Baseline.}}
We compare \tool with
\Cklst,\footnote{https://github.com/marcotcr/checklist} a
manual template-based approach to generate test cases. In this
experiment, we used the \Cklst released \sa test cases which are
generated from its publicly available jupyter notebook implementation.
%% \TODO{introduce details about CHECKLIST}

\subsection{Experimental Process}

\paragraph{\textbf{RQ1.}}
Recall that a key limitation of \Cklst is that its template-based
approach relies on significant manual efforts and may not generate test cases that comprehensively cover the sentences
in a linguistic capability. \tool, instead, automatically generates test cases based on a search
dataset and the syntax in a large reference corpus. 
We expect \tool can generate a more diverse test suite than \Cklst, and achieve
more probable evaluation results.

We first evaluate the three \sa
models by testing them on the \tool generated test cases, reporting number of test cases and each model's failure rate in each linguistic capability.

Next,
to measure diversity, we follow the approach presented by Ma et al. \cite{ma2018deepgauge},
where the authors measure the coverage of NLP model intermediate states as corner-case neurons.
Because the matrix computation of intermediate states impacts NLP model decision-making, a test suite that covers a greater number of intermediate states can represent more NLP model decision-making, making it more diverse.
Specifically, we used two coverage metrics by Ma et al. \cite{ma2018deepgauge}, \textit{boundary coverage} (BoundCov) and \textit{strong activation coverage} (SActCov), to evaluate the test suite diversity.
It is worth noting that a test sample with a statistical distribution similar to the training data would rarely be found in the corner case region. As a result, covering a larger corner case region means that the test suite is more likely to be buggy.

\begin{equation}
\begin{split}
    \text{UpperCorner}(\mathcal{X}) = \{n \in N | \exists x \in \mathcal{X}: f_n(x) \in (high_n, +\infty)\}; \\
    \text{LowerCorner}(\mathcal{X}) = \{n \in N | \exists x \in \mathcal{X}: f_n(x) \in (-\infty, low_n)\}; \\
\end{split}
    \label{eq:corner}
\end{equation}

\noindent Equation \ref{eq:corner} shows the formal definition of the corner-case neuron of the NLP model $f(\cdot)$, where $\mathcal{X}$ is the given test suite, $N$ is the number of neurons in model $f(\cdot)$, $f_n(\cdot)$ is the $n^{th}$ neuron's output, and $high_n$ and $low_n$ are the $n^{th}$ neurons' output bounds on the model training dataset.
Eq. \ref{eq:corner} can be interpreted as the collection of neurons that emit outputs beyond the model's numerical boundary.

\begin{equation}
\begin{split}
     & BoundCov(\mathcal{X}) = \frac{|UpperCorner(\mathcal{X})| + |LowerCorner(\mathcal{X})| }{2 \times |N|} \\ 
     &\quad  \qquad \qquad  SActCov(\mathcal{X}) = \frac{|UpperCorner(\mathcal{X})|} {|N|} \\ 
\end{split}
    \label{eq:coverage}
\end{equation}

\noindent The formal definition of our coverage metrics are shown in Equation \ref{eq:coverage}, where BoundCov measures the coverage of neurons that produce outputs that exceed the upper or lower bounds, and SActCov measures the coverage of neurons that create outputs that exceed the lower bound.
Higher coverage indicates the test suite is better for triggering the corner-case neurons, thus better test suite diversity.

% ($Cov(\mathcal{X})$ in \equref{eq:coverage}),
% where $N$ is the total number blocks, $\mathbb I(\cdot)$ is the indicator function, and
% $(B_i(x) > \tau_i))$ represents whether $i^{th}$ block is activated by input $x$~(the definition of $B_i$ and $\tau_i$ are the same with \equref{eq:new2} and \equref{eq:new3}).
% Because AdNNs activate different blocks for decision making, then a higher block coverage indicates the test samples cover more decision behaviors.

To answer RQ1, for each NLP model under test, we first feed its training dataset to compute each neuron's lower and upper bounds. After that, we select the same number of test cases from \tool and \Cklst as the test suite and compute the corresponding coverage metrics.

% For each subject, we randomly select 100 seed samples from the test dataset as seed inputs. We then feed the same seed inputs into \tool and \texttt{ILFO} to generate test samples. 
% Finally, we feed the generated test samples to AdNNs and measure block coverage.
% We repeat this process 10 times and record the average coverage and the variance.
% The results are shown in \tabref{tab:coverage} last two columns.

%
%The amount of \sa model
%components \sw{?} executed during testing is a critical measurement for
%assessing quality of software testing. A high software coverage
%results in higher chances of unidentified bugs in the \sa model. On
%the other hand, limited distribution only represents narrow portion of
%real world covering limited execution behaviors in a \sa model. It leads
%to detect bugs within the restricted execution behaviors.  Therefore,
%test cases more representative of real-world data result in more generalized
%distribution and higher coverage of the \sa model.
%Therefore, we answer the {\bf RQ2} by
%measuring the neural coverage of the \sa model.
%
%Specifically, we implemented DeepXplore to measure the \sa model
%coverage~\cite{pei2017deepxplore}. \Dxp is the first efficient
%white-box testing framework for large-scale \dl systems. It introduces
%neuron coverage of a set of test inputs as the ratio of the number of
%unique activated neurons and the total number of neurons in input \dl
%system. In this experiment, we compute the neuron coverage of a test
%cases from \tool and \Cklst on the fine-tuned \sw{first time mention fine-tuning: reader does not know how we fine-tune} \sa model of
%\bertsamodel and compare the coverage between \tool and \Cklst.

\paragraph{\textbf{RQ2 and RQ3}} As described in Section \ref{sec:approach},
\tool generates test cases in two steps: specification-based seed
generation and syntax-based sentence expansion. These automated steps
may generate seed/expanded sentences marked with incorrect sentiment
labels or categorized into wrong linguistic capabilities. For example,
the search rule and template defined in a linguistic capability may
not always generate seed sentences in that capability or with the
correct label.  To answer RQ2 and RQ3, we performed a manual study to
measure the correctness of the sentiment labels and linguistic
capabilities associated with the seed/expanded sentences, produced by
\tool.

In the manual study, we randomly sample 100 seed sentences, each of which has at least one expanded sentence, and divide these seeds to two sets (i.e., 50 in each set). For each sampled seed sentence, we randomly obtain one of its expanded sentences.
This forms the two sets of sentences (200 sentences in total) we use for this study, each with 50 seeds and 50 corresponding expanded sentences.
We recruited three participants of this study; all of them are graduate students with no knowledge about this work. 
2 of them were assigned one set of sentences and the third was assigned the other set.
Each participant was asked to provide two scores for each sentence.
\emph{(1) Relevancy score between sentence and its associated \lc}:
this score measures the correctness of \tool linguistic capability categorization.  The scores are discrete,
ranging from 1 (``strongly not relevant'') to 5 (``strongly relevant''). \emph{(2) sentiment score of the sentence}: this score measures the sentiment level of the sentence . It
is also discrete, ranging from 1 (``strongly
negative'') to 5 (``strongly positive''). 
We measure the following:

\begin{eqnarray}
  sent\_consistency &= \frac{1}{\#Sample\cdot~}\cdot \sum_{i} \delta(label_{S^2LCT}=label_{human}) \label{metric:srel} \\
  LC\_relevancy_{AVG} &= \frac{1}{\#Sample\cdot~}\cdot\sum_{i} Norm(LC\_relevancy_i) \label{metric:lcrel}
\end{eqnarray}

Equation~\ref{metric:srel} represents the percentage of the test cases that \tool and the participants produce the same sentiment labels. High value of this metric indicates \tool generates test cases with correct labels. Equation~\ref{metric:lcrel} represents the average of the
normalized relevancy score between a sentence and its associated
\lc. Higher average score means indicates the linguistic capability categorization by \tool is correct. We answer RQ2 and RQ3 using the metrics defined by
Equation~\ref{metric:srel} and Equation~\ref{metric:lcrel},
respectively.

\paragraph{\textbf{RQ4.}}

To answer RQ4, we conduct experiments to demonstrate that \tool can help developers understand the bugs in the NLP models.
Recall that \tool generates test cases by mutating seed sentences (\eg by expanding one token in the seed input). Still, it is unclear why mutating one token will cause the model to produce misclassified results.
We seek to help developers understand why such mutation will result in the misclassification. 
Existing work \cite{simin2020denas, lemna, lime} has demonstrated that the ML model prediction is dominated by a minimal set of input features (\ie tokens in input sentences). Motivated by such intuition, we identify a minimal set of input tokens that dominate the model prediction.

Formally, given a input sentence $x = [tk_1, tk_2, \cdots, tk_n]$, and the NLP model under test $f(\cdot)$, our goal is to find a masking template $T = [t_1, t_2, \cdots, t_n]$, where $t_i$ is 0 or 1, representing masking the $i^{th}$ token (\ie $tk_i$) in $x$ or not.
The template $T$ can mask some tokens in $x$ with attribute tokens, and the masked input has a high probability of retaining the original prediction $x$, denoted as
\begin{equation}
    P(f(T(x)) = f(x)) \ge P_{thresh}
    \label{eq:prob}
\end{equation}
To create such a template $T$, we first compute the contribution score of each input token using an existing explainable ML technique \cite{lemna}. We then begin with the full mask template (\ie, all tokens are masked); such full mask template definitely does not satisfy Equation \ref{eq:prob}.
We then iteratively shift one position from mask to non-mask based on the order of each token's contribution score, until the template $T$ satisfies Equation \ref{eq:prob}.
Because we iterate the size of the mask, the generated template $T$ will keep the minimum number of tokens in $x$. Moreover, since the input $x$ is an incorrect prediction, the generated template $T$ is likely to produce misclassification (i.e., the probability to be misclassified is larger than $P_{thresh}$).

% In particular, we apply the explainable ML techniques to visualize the contribution of each input token to the model predictions.
%

% \sw{@Simin: add the setup of the bug explanation case study here.}
%In addition to the detection of bugs \sw{have we defined ``bug" in this context?} in the model,
%explanation of the bugs is also important for debugging and
%repairing the model. Therefore, we answer {\bf RQ3} by analyzing the \sa model based on \tool test cases to find the root causes of the bugs.
%
%Specifically, we adapt \Denas{} for this experiment. \sw{Say more what DENAS does.} Rules generated from
%\Denas{} are interpretable functions mapping certain features of the
%input to the expected output of a deep \nn system, and the generated
%rules are considered as the behaviors of the deep \nn system. To
%analyze the bug of a \sa model, we generates rules over the test
%inputs using \Denas{} and identifies faulty rules for the failed test
%inputs.
%\sw{Previous sounds vague. Any specific adaptation of DENAS we did?}
%We manually identify the root of the faulty rules and find the
%root causes of bug in the end.

% \noindent\textbf{Implementation Details.}

% \sw{Missing: environment running these experiments.}

%% \MyPara{Seed Input Selection}
%% %
%% For each linguistic capability, we first search all sentences that
%% meet its requirement. Among found sentences, we randomly select 10
%% sentences due to memory constraint.

%% \MyPara{Word Sentiment}
%% %
%% we extract sentiments of words using the
%% \Swn~\cite{baccianella2010sentiwordnet}. The \Swn is a publicly
%% available lexical resource of words on Wordnet with three numerical
%% scores of objectivity, positivity and negativity. Sentiment word
%% labels from the scores are classified from the algorithm from Mihaela
%% \etal~\cite{mihaela2017sentiwordnetlabel}.

%% \MyPara{\Cfg Expansion}
%% %
%% We build a reference \Cfg of natural language from the English Penn
%% \Trb corpora~\cite{mitchell1993treebank,nltkTreebankCorporaWebPage}.
%% The corpus is sampled from 2,499 stories from a tree year \Wsj
%% collection The \Trb provides a parsed text corpus with annotation of
%% syntactic and semantic structure. In this experiment We implement the
%% \trb corpora available through \Nltk, which is a suite of libraries
%% and programs for \Nlp for English. In addition, we parse the seed
%% input using into its CFG using the Berkeley Neural
%% Parser~\cite{kitaev2018constituency, kitaev2019multilingual}, a
%% high-accuracy parser with models for 11 languages. The input is a raw
%% text in natural language and the output is the string representation of
%% parse tree. Next after comparing CFGs between reference and seed input,
%% we randomly select 10 expansions for generating templates due to
%% memory constraint.

%% \MyPara{Synonyms}
%% %
%% \Model searches synonyms of each token from synonym sets extracted
%% from \Wrdnt using \Spacy open-source library for NLP.

%% \MyPara{Models}
%% %
%% We evaluate the following \sa models via \Model:
%% \Bert~\cite{devlin2019bert}, \Roberta~\cite{liu2019roberta} and
%% \Dbert~\cite{sanh2019distilbert}. These models are fine-tuned on \Sstt
%% and their accuracies are \BertAcc, \RobertaAcc and \DbertAcc.

%% \MyPara{Retraining}
%% %
%% We retrain \sa models. we split \Model generated test cases into
%% train/validation/test sets with the ratio of 8:1:1. The number of
%% epochs and batch size for retraining are 1 and 16 respectively.

%\input{CM/coverage}

%\input{CM/explain}