### 2022-08-29 paper/ase22/tables/test-results-3-200-numbers.tex
add new tables and figs

%% Automatically generated by pyutil.latex

\DefMacro{test-results-3-200-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-3-200-lc7-num-seeds}{200}
\DefMacro{test-results-3-200-lc7-num-exps-avg}{6908.0}
\DefMacro{test-results-3-200-lc7-num-exps-med}{7022}
\DefMacro{test-results-3-200-lc7-num-exps-std}{978.245}
\DefMacro{test-results-3-200-model0-lc7-num-seed-fail-avg}{23.83}
\DefMacro{test-results-3-200-model0-lc7-num-seed-fail-med}{21.0}
\DefMacro{test-results-3-200-model0-lc7-num-seed-fail-std}{2.675}
\DefMacro{test-results-3-200-model0-lc7-num-exp-fail-avg}{25.44}
\DefMacro{test-results-3-200-model0-lc7-num-exp-fail-med}{25.12}
\DefMacro{test-results-3-200-model0-lc7-num-exp-fail-std}{2.355}
\DefMacro{test-results-3-200-model0-lc7-num-pass-to-fail-avg}{110.33}
\DefMacro{test-results-3-200-model0-lc7-num-pass-to-fail-med}{100}
\DefMacro{test-results-3-200-model0-lc7-num-pass-to-fail-std}{35.035}
\DefMacro{test-results-3-200-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-3-200-lc3-num-seeds}{67}
\DefMacro{test-results-3-200-lc3-num-exps-avg}{2845.67}
\DefMacro{test-results-3-200-lc3-num-exps-med}{2839}
\DefMacro{test-results-3-200-lc3-num-exps-std}{9.45}
\DefMacro{test-results-3-200-model0-lc3-num-seed-fail-avg}{89.55}
\DefMacro{test-results-3-200-model0-lc3-num-seed-fail-med}{89.55}
\DefMacro{test-results-3-200-model0-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model0-lc3-num-exp-fail-avg}{91.45}
\DefMacro{test-results-3-200-model0-lc3-num-exp-fail-med}{91.49}
\DefMacro{test-results-3-200-model0-lc3-num-exp-fail-std}{0.085}
\DefMacro{test-results-3-200-model0-lc3-num-pass-to-fail-avg}{81.33}
\DefMacro{test-results-3-200-model0-lc3-num-pass-to-fail-med}{81}
\DefMacro{test-results-3-200-model0-lc3-num-pass-to-fail-std}{0.765}
\DefMacro{test-results-3-200-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-3-200-lc4-num-seeds}{26}
\DefMacro{test-results-3-200-lc4-num-exps-avg}{1096.67}
\DefMacro{test-results-3-200-lc4-num-exps-med}{1105}
\DefMacro{test-results-3-200-lc4-num-exps-std}{10.395}
\DefMacro{test-results-3-200-model0-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-200-model0-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-200-model0-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model0-lc4-num-exp-fail-avg}{92.52}
\DefMacro{test-results-3-200-model0-lc4-num-exp-fail-med}{92.67}
\DefMacro{test-results-3-200-model0-lc4-num-exp-fail-std}{0.15}
\DefMacro{test-results-3-200-model0-lc4-num-pass-to-fail-avg}{65.67}
\DefMacro{test-results-3-200-model0-lc4-num-pass-to-fail-med}{65}
\DefMacro{test-results-3-200-model0-lc4-num-pass-to-fail-std}{0.575}
\DefMacro{test-results-3-200-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-3-200-lc6-num-seeds}{200}
\DefMacro{test-results-3-200-lc6-num-exps-avg}{4731.0}
\DefMacro{test-results-3-200-lc6-num-exps-med}{4927}
\DefMacro{test-results-3-200-lc6-num-exps-std}{205.36}
\DefMacro{test-results-3-200-model0-lc6-num-seed-fail-avg}{88.5}
\DefMacro{test-results-3-200-model0-lc6-num-seed-fail-med}{90.0}
\DefMacro{test-results-3-200-model0-lc6-num-seed-fail-std}{3.93}
\DefMacro{test-results-3-200-model0-lc6-num-exp-fail-avg}{92.94}
\DefMacro{test-results-3-200-model0-lc6-num-exp-fail-med}{96.43}
\DefMacro{test-results-3-200-model0-lc6-num-exp-fail-std}{3.34}
\DefMacro{test-results-3-200-model0-lc6-num-pass-to-fail-avg}{42.33}
\DefMacro{test-results-3-200-model0-lc6-num-pass-to-fail-med}{53}
\DefMacro{test-results-3-200-model0-lc6-num-pass-to-fail-std}{16.655}
\DefMacro{test-results-3-200-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-3-200-lc5-num-seeds}{200}
\DefMacro{test-results-3-200-lc5-num-exps-avg}{5389.67}
\DefMacro{test-results-3-200-lc5-num-exps-med}{5387}
\DefMacro{test-results-3-200-lc5-num-exps-std}{69.01}
\DefMacro{test-results-3-200-model0-lc5-num-seed-fail-avg}{99.5}
\DefMacro{test-results-3-200-model0-lc5-num-seed-fail-med}{99.5}
\DefMacro{test-results-3-200-model0-lc5-num-seed-fail-std}{0.25}
\DefMacro{test-results-3-200-model0-lc5-num-exp-fail-avg}{99.83}
\DefMacro{test-results-3-200-model0-lc5-num-exp-fail-med}{99.81}
\DefMacro{test-results-3-200-model0-lc5-num-exp-fail-std}{0.06}
\DefMacro{test-results-3-200-model0-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-200-model0-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-200-model0-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-200-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-3-200-lc9-num-seeds}{200}
\DefMacro{test-results-3-200-lc9-num-exps-avg}{5814.0}
\DefMacro{test-results-3-200-lc9-num-exps-med}{5937}
\DefMacro{test-results-3-200-lc9-num-exps-std}{281.335}
\DefMacro{test-results-3-200-model0-lc9-num-seed-fail-avg}{77.17}
\DefMacro{test-results-3-200-model0-lc9-num-seed-fail-med}{76.0}
\DefMacro{test-results-3-200-model0-lc9-num-seed-fail-std}{1.7}
\DefMacro{test-results-3-200-model0-lc9-num-exp-fail-avg}{75.48}
\DefMacro{test-results-3-200-model0-lc9-num-exp-fail-med}{76.47}
\DefMacro{test-results-3-200-model0-lc9-num-exp-fail-std}{0.905}
\DefMacro{test-results-3-200-model0-lc9-num-pass-to-fail-avg}{99.67}
\DefMacro{test-results-3-200-model0-lc9-num-pass-to-fail-med}{76}
\DefMacro{test-results-3-200-model0-lc9-num-pass-to-fail-std}{20.495}
\DefMacro{test-results-3-200-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-3-200-lc2-num-seeds}{200}
\DefMacro{test-results-3-200-lc2-num-exps-avg}{6245.33}
\DefMacro{test-results-3-200-lc2-num-exps-med}{6256}
\DefMacro{test-results-3-200-lc2-num-exps-std}{178.56}
\DefMacro{test-results-3-200-model0-lc2-num-seed-fail-avg}{20.67}
\DefMacro{test-results-3-200-model0-lc2-num-seed-fail-med}{19.5}
\DefMacro{test-results-3-200-model0-lc2-num-seed-fail-std}{1.465}
\DefMacro{test-results-3-200-model0-lc2-num-exp-fail-avg}{20.57}
\DefMacro{test-results-3-200-model0-lc2-num-exp-fail-med}{20.05}
\DefMacro{test-results-3-200-model0-lc2-num-exp-fail-std}{1.24}
\DefMacro{test-results-3-200-model0-lc2-num-pass-to-fail-avg}{199.33}
\DefMacro{test-results-3-200-model0-lc2-num-pass-to-fail-med}{198}
\DefMacro{test-results-3-200-model0-lc2-num-pass-to-fail-std}{13.51}
\DefMacro{test-results-3-200-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-3-200-lc0-num-seeds}{19}
\DefMacro{test-results-3-200-lc0-num-exps-avg}{235.33}
\DefMacro{test-results-3-200-lc0-num-exps-med}{243}
\DefMacro{test-results-3-200-lc0-num-exps-std}{9.36}
\DefMacro{test-results-3-200-model0-lc0-num-seed-fail-avg}{78.95}
\DefMacro{test-results-3-200-model0-lc0-num-seed-fail-med}{78.95}
\DefMacro{test-results-3-200-model0-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model0-lc0-num-exp-fail-avg}{89.5}
\DefMacro{test-results-3-200-model0-lc0-num-exp-fail-med}{89.25}
\DefMacro{test-results-3-200-model0-lc0-num-exp-fail-std}{0.385}
\DefMacro{test-results-3-200-model0-lc0-num-pass-to-fail-avg}{18.33}
\DefMacro{test-results-3-200-model0-lc0-num-pass-to-fail-med}{18}
\DefMacro{test-results-3-200-model0-lc0-num-pass-to-fail-std}{0.765}
\DefMacro{test-results-3-200-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-3-200-lc1-num-seeds}{200}
\DefMacro{test-results-3-200-lc1-num-exps-avg}{1951.0}
\DefMacro{test-results-3-200-lc1-num-exps-med}{1968}
\DefMacro{test-results-3-200-lc1-num-exps-std}{44.365}
\DefMacro{test-results-3-200-model0-lc1-num-seed-fail-avg}{1.5}
\DefMacro{test-results-3-200-model0-lc1-num-seed-fail-med}{1.5}
\DefMacro{test-results-3-200-model0-lc1-num-seed-fail-std}{0.25}
\DefMacro{test-results-3-200-model0-lc1-num-exp-fail-avg}{4.87}
\DefMacro{test-results-3-200-model0-lc1-num-exp-fail-med}{4.53}
\DefMacro{test-results-3-200-model0-lc1-num-exp-fail-std}{0.32}
\DefMacro{test-results-3-200-model0-lc1-num-pass-to-fail-avg}{38.67}
\DefMacro{test-results-3-200-model0-lc1-num-pass-to-fail-med}{42}
\DefMacro{test-results-3-200-model0-lc1-num-pass-to-fail-std}{11.095}
\DefMacro{test-results-3-200-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-3-200-lc8-num-seeds}{200}
\DefMacro{test-results-3-200-lc8-num-exps-avg}{5578.67}
\DefMacro{test-results-3-200-lc8-num-exps-med}{5578}
\DefMacro{test-results-3-200-lc8-num-exps-std}{65.5}
\DefMacro{test-results-3-200-model0-lc8-num-seed-fail-avg}{4.83}
\DefMacro{test-results-3-200-model0-lc8-num-seed-fail-med}{5.0}
\DefMacro{test-results-3-200-model0-lc8-num-seed-fail-std}{0.145}
\DefMacro{test-results-3-200-model0-lc8-num-exp-fail-avg}{5.48}
\DefMacro{test-results-3-200-model0-lc8-num-exp-fail-med}{5.66}
\DefMacro{test-results-3-200-model0-lc8-num-exp-fail-std}{0.56}
\DefMacro{test-results-3-200-model0-lc8-num-pass-to-fail-avg}{99.0}
\DefMacro{test-results-3-200-model0-lc8-num-pass-to-fail-med}{87}
\DefMacro{test-results-3-200-model0-lc8-num-pass-to-fail-std}{14.0}
\DefMacro{test-results-3-200-model1-lc7-num-seed-fail-avg}{15.0}
\DefMacro{test-results-3-200-model1-lc7-num-seed-fail-med}{16.5}
\DefMacro{test-results-3-200-model1-lc7-num-seed-fail-std}{2.46}
\DefMacro{test-results-3-200-model1-lc7-num-exp-fail-avg}{16.15}
\DefMacro{test-results-3-200-model1-lc7-num-exp-fail-med}{13.62}
\DefMacro{test-results-3-200-model1-lc7-num-exp-fail-std}{3.07}
\DefMacro{test-results-3-200-model1-lc7-num-pass-to-fail-avg}{171.33}
\DefMacro{test-results-3-200-model1-lc7-num-pass-to-fail-med}{224}
\DefMacro{test-results-3-200-model1-lc7-num-pass-to-fail-std}{53.14}
\DefMacro{test-results-3-200-model1-lc3-num-seed-fail-avg}{85.07}
\DefMacro{test-results-3-200-model1-lc3-num-seed-fail-med}{85.07}
\DefMacro{test-results-3-200-model1-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model1-lc3-num-exp-fail-avg}{88.59}
\DefMacro{test-results-3-200-model1-lc3-num-exp-fail-med}{88.63}
\DefMacro{test-results-3-200-model1-lc3-num-exp-fail-std}{0.08}
\DefMacro{test-results-3-200-model1-lc3-num-pass-to-fail-avg}{65.33}
\DefMacro{test-results-3-200-model1-lc3-num-pass-to-fail-med}{63}
\DefMacro{test-results-3-200-model1-lc3-num-pass-to-fail-std}{2.02}
\DefMacro{test-results-3-200-model1-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-200-model1-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-200-model1-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model1-lc4-num-exp-fail-avg}{93.13}
\DefMacro{test-results-3-200-model1-lc4-num-exp-fail-med}{93.12}
\DefMacro{test-results-3-200-model1-lc4-num-exp-fail-std}{0.06}
\DefMacro{test-results-3-200-model1-lc4-num-pass-to-fail-avg}{45.33}
\DefMacro{test-results-3-200-model1-lc4-num-pass-to-fail-med}{44}
\DefMacro{test-results-3-200-model1-lc4-num-pass-to-fail-std}{1.605}
\DefMacro{test-results-3-200-model1-lc6-num-seed-fail-avg}{74.17}
\DefMacro{test-results-3-200-model1-lc6-num-seed-fail-med}{73.5}
\DefMacro{test-results-3-200-model1-lc6-num-seed-fail-std}{6.005}
\DefMacro{test-results-3-200-model1-lc6-num-exp-fail-avg}{77.89}
\DefMacro{test-results-3-200-model1-lc6-num-exp-fail-med}{74.05}
\DefMacro{test-results-3-200-model1-lc6-num-exp-fail-std}{4.595}
\DefMacro{test-results-3-200-model1-lc6-num-pass-to-fail-avg}{112.0}
\DefMacro{test-results-3-200-model1-lc6-num-pass-to-fail-med}{97}
\DefMacro{test-results-3-200-model1-lc6-num-pass-to-fail-std}{31.43}
\DefMacro{test-results-3-200-model1-lc5-num-seed-fail-avg}{99.5}
\DefMacro{test-results-3-200-model1-lc5-num-seed-fail-med}{99.5}
\DefMacro{test-results-3-200-model1-lc5-num-seed-fail-std}{0.25}
\DefMacro{test-results-3-200-model1-lc5-num-exp-fail-avg}{99.58}
\DefMacro{test-results-3-200-model1-lc5-num-exp-fail-med}{99.67}
\DefMacro{test-results-3-200-model1-lc5-num-exp-fail-std}{0.145}
\DefMacro{test-results-3-200-model1-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-200-model1-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-200-model1-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-200-model1-lc9-num-seed-fail-avg}{86.33}
\DefMacro{test-results-3-200-model1-lc9-num-seed-fail-med}{86.0}
\DefMacro{test-results-3-200-model1-lc9-num-seed-fail-std}{1.505}
\DefMacro{test-results-3-200-model1-lc9-num-exp-fail-avg}{86.8}
\DefMacro{test-results-3-200-model1-lc9-num-exp-fail-med}{86.64}
\DefMacro{test-results-3-200-model1-lc9-num-exp-fail-std}{0.235}
\DefMacro{test-results-3-200-model1-lc9-num-pass-to-fail-avg}{46.33}
\DefMacro{test-results-3-200-model1-lc9-num-pass-to-fail-med}{52}
\DefMacro{test-results-3-200-model1-lc9-num-pass-to-fail-std}{6.25}
\DefMacro{test-results-3-200-model1-lc2-num-seed-fail-avg}{48.33}
\DefMacro{test-results-3-200-model1-lc2-num-seed-fail-med}{49.5}
\DefMacro{test-results-3-200-model1-lc2-num-seed-fail-std}{1.94}
\DefMacro{test-results-3-200-model1-lc2-num-exp-fail-avg}{47.94}
\DefMacro{test-results-3-200-model1-lc2-num-exp-fail-med}{47.1}
\DefMacro{test-results-3-200-model1-lc2-num-exp-fail-std}{1.72}
\DefMacro{test-results-3-200-model1-lc2-num-pass-to-fail-avg}{172.33}
\DefMacro{test-results-3-200-model1-lc2-num-pass-to-fail-med}{161}
\DefMacro{test-results-3-200-model1-lc2-num-pass-to-fail-std}{19.14}
\DefMacro{test-results-3-200-model1-lc0-num-seed-fail-avg}{89.47}
\DefMacro{test-results-3-200-model1-lc0-num-seed-fail-med}{89.47}
\DefMacro{test-results-3-200-model1-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model1-lc0-num-exp-fail-avg}{78.36}
\DefMacro{test-results-3-200-model1-lc0-num-exp-fail-med}{78.19}
\DefMacro{test-results-3-200-model1-lc0-num-exp-fail-std}{0.275}
\DefMacro{test-results-3-200-model1-lc0-num-pass-to-fail-avg}{6.0}
\DefMacro{test-results-3-200-model1-lc0-num-pass-to-fail-med}{5}
\DefMacro{test-results-3-200-model1-lc0-num-pass-to-fail-std}{0.865}
\DefMacro{test-results-3-200-model1-lc1-num-seed-fail-avg}{3.33}
\DefMacro{test-results-3-200-model1-lc1-num-seed-fail-med}{3.0}
\DefMacro{test-results-3-200-model1-lc1-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-200-model1-lc1-num-exp-fail-avg}{4.7}
\DefMacro{test-results-3-200-model1-lc1-num-exp-fail-med}{5.03}
\DefMacro{test-results-3-200-model1-lc1-num-exp-fail-std}{0.485}
\DefMacro{test-results-3-200-model1-lc1-num-pass-to-fail-avg}{42.67}
\DefMacro{test-results-3-200-model1-lc1-num-pass-to-fail-med}{40}
\DefMacro{test-results-3-200-model1-lc1-num-pass-to-fail-std}{11.06}
\DefMacro{test-results-3-200-model1-lc8-num-seed-fail-avg}{4.5}
\DefMacro{test-results-3-200-model1-lc8-num-seed-fail-med}{5.0}
\DefMacro{test-results-3-200-model1-lc8-num-seed-fail-std}{0.435}
\DefMacro{test-results-3-200-model1-lc8-num-exp-fail-avg}{5.33}
\DefMacro{test-results-3-200-model1-lc8-num-exp-fail-med}{4.33}
\DefMacro{test-results-3-200-model1-lc8-num-exp-fail-std}{0.96}
\DefMacro{test-results-3-200-model1-lc8-num-pass-to-fail-avg}{85.33}
\DefMacro{test-results-3-200-model1-lc8-num-pass-to-fail-med}{83}
\DefMacro{test-results-3-200-model1-lc8-num-pass-to-fail-std}{10.3}
\DefMacro{test-results-3-200-model2-lc7-num-seed-fail-avg}{29.5}
\DefMacro{test-results-3-200-model2-lc7-num-seed-fail-med}{23.0}
\DefMacro{test-results-3-200-model2-lc7-num-seed-fail-std}{7.66}
\DefMacro{test-results-3-200-model2-lc7-num-exp-fail-avg}{24.56}
\DefMacro{test-results-3-200-model2-lc7-num-exp-fail-med}{23.81}
\DefMacro{test-results-3-200-model2-lc7-num-exp-fail-std}{0.655}
\DefMacro{test-results-3-200-model2-lc7-num-pass-to-fail-avg}{166.33}
\DefMacro{test-results-3-200-model2-lc7-num-pass-to-fail-med}{197}
\DefMacro{test-results-3-200-model2-lc7-num-pass-to-fail-std}{57.555}
\DefMacro{test-results-3-200-model2-lc3-num-seed-fail-avg}{85.07}
\DefMacro{test-results-3-200-model2-lc3-num-seed-fail-med}{85.07}
\DefMacro{test-results-3-200-model2-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model2-lc3-num-exp-fail-avg}{86.41}
\DefMacro{test-results-3-200-model2-lc3-num-exp-fail-med}{86.37}
\DefMacro{test-results-3-200-model2-lc3-num-exp-fail-std}{0.16}
\DefMacro{test-results-3-200-model2-lc3-num-pass-to-fail-avg}{43.33}
\DefMacro{test-results-3-200-model2-lc3-num-pass-to-fail-med}{44}
\DefMacro{test-results-3-200-model2-lc3-num-pass-to-fail-std}{3.51}
\DefMacro{test-results-3-200-model2-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-200-model2-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-200-model2-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model2-lc4-num-exp-fail-avg}{95.72}
\DefMacro{test-results-3-200-model2-lc4-num-exp-fail-med}{95.68}
\DefMacro{test-results-3-200-model2-lc4-num-exp-fail-std}{0.085}
\DefMacro{test-results-3-200-model2-lc4-num-pass-to-fail-avg}{84.67}
\DefMacro{test-results-3-200-model2-lc4-num-pass-to-fail-med}{85}
\DefMacro{test-results-3-200-model2-lc4-num-pass-to-fail-std}{1.755}
\DefMacro{test-results-3-200-model2-lc6-num-seed-fail-avg}{88.5}
\DefMacro{test-results-3-200-model2-lc6-num-seed-fail-med}{88.5}
\DefMacro{test-results-3-200-model2-lc6-num-seed-fail-std}{3.75}
\DefMacro{test-results-3-200-model2-lc6-num-exp-fail-avg}{92.07}
\DefMacro{test-results-3-200-model2-lc6-num-exp-fail-med}{96.41}
\DefMacro{test-results-3-200-model2-lc6-num-exp-fail-std}{4.265}
\DefMacro{test-results-3-200-model2-lc6-num-pass-to-fail-avg}{45.0}
\DefMacro{test-results-3-200-model2-lc6-num-pass-to-fail-med}{36}
\DefMacro{test-results-3-200-model2-lc6-num-pass-to-fail-std}{15.74}
\DefMacro{test-results-3-200-model2-lc5-num-seed-fail-avg}{98.5}
\DefMacro{test-results-3-200-model2-lc5-num-seed-fail-med}{98.5}
\DefMacro{test-results-3-200-model2-lc5-num-seed-fail-std}{0.25}
\DefMacro{test-results-3-200-model2-lc5-num-exp-fail-avg}{98.5}
\DefMacro{test-results-3-200-model2-lc5-num-exp-fail-med}{98.92}
\DefMacro{test-results-3-200-model2-lc5-num-exp-fail-std}{0.66}
\DefMacro{test-results-3-200-model2-lc5-num-pass-to-fail-avg}{16.0}
\DefMacro{test-results-3-200-model2-lc5-num-pass-to-fail-med}{17}
\DefMacro{test-results-3-200-model2-lc5-num-pass-to-fail-std}{2.29}
\DefMacro{test-results-3-200-model2-lc9-num-seed-fail-avg}{95.67}
\DefMacro{test-results-3-200-model2-lc9-num-seed-fail-med}{96.0}
\DefMacro{test-results-3-200-model2-lc9-num-seed-fail-std}{1.26}
\DefMacro{test-results-3-200-model2-lc9-num-exp-fail-avg}{94.95}
\DefMacro{test-results-3-200-model2-lc9-num-exp-fail-med}{96.1}
\DefMacro{test-results-3-200-model2-lc9-num-exp-fail-std}{1.51}
\DefMacro{test-results-3-200-model2-lc9-num-pass-to-fail-avg}{31.33}
\DefMacro{test-results-3-200-model2-lc9-num-pass-to-fail-med}{16}
\DefMacro{test-results-3-200-model2-lc9-num-pass-to-fail-std}{18.715}
\DefMacro{test-results-3-200-model2-lc2-num-seed-fail-avg}{62.0}
\DefMacro{test-results-3-200-model2-lc2-num-seed-fail-med}{62.0}
\DefMacro{test-results-3-200-model2-lc2-num-seed-fail-std}{1.5}
\DefMacro{test-results-3-200-model2-lc2-num-exp-fail-avg}{62.25}
\DefMacro{test-results-3-200-model2-lc2-num-exp-fail-med}{62.44}
\DefMacro{test-results-3-200-model2-lc2-num-exp-fail-std}{1.95}
\DefMacro{test-results-3-200-model2-lc2-num-pass-to-fail-avg}{165.33}
\DefMacro{test-results-3-200-model2-lc2-num-pass-to-fail-med}{150}
\DefMacro{test-results-3-200-model2-lc2-num-pass-to-fail-std}{35.135}
\DefMacro{test-results-3-200-model2-lc0-num-seed-fail-avg}{100.0}
\DefMacro{test-results-3-200-model2-lc0-num-seed-fail-med}{100.0}
\DefMacro{test-results-3-200-model2-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-200-model2-lc0-num-exp-fail-avg}{95.36}
\DefMacro{test-results-3-200-model2-lc0-num-exp-fail-med}{95.98}
\DefMacro{test-results-3-200-model2-lc0-num-exp-fail-std}{0.665}
\DefMacro{test-results-3-200-model2-lc0-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-200-model2-lc0-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-200-model2-lc0-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-200-model2-lc1-num-seed-fail-avg}{1.5}
\DefMacro{test-results-3-200-model2-lc1-num-seed-fail-med}{1.5}
\DefMacro{test-results-3-200-model2-lc1-num-seed-fail-std}{0.25}
\DefMacro{test-results-3-200-model2-lc1-num-exp-fail-avg}{4.09}
\DefMacro{test-results-3-200-model2-lc1-num-exp-fail-med}{3.94}
\DefMacro{test-results-3-200-model2-lc1-num-exp-fail-std}{0.535}
\DefMacro{test-results-3-200-model2-lc1-num-pass-to-fail-avg}{41.33}
\DefMacro{test-results-3-200-model2-lc1-num-pass-to-fail-med}{37}
\DefMacro{test-results-3-200-model2-lc1-num-pass-to-fail-std}{9.44}
\DefMacro{test-results-3-200-model2-lc8-num-seed-fail-avg}{4.5}
\DefMacro{test-results-3-200-model2-lc8-num-seed-fail-med}{5.0}
\DefMacro{test-results-3-200-model2-lc8-num-seed-fail-std}{0.435}
\DefMacro{test-results-3-200-model2-lc8-num-exp-fail-avg}{5.39}
\DefMacro{test-results-3-200-model2-lc8-num-exp-fail-med}{6.33}
\DefMacro{test-results-3-200-model2-lc8-num-exp-fail-std}{1.285}
\DefMacro{test-results-3-200-model2-lc8-num-pass-to-fail-avg}{100.0}
\DefMacro{test-results-3-200-model2-lc8-num-pass-to-fail-med}{52}
\DefMacro{test-results-3-200-model2-lc8-num-pass-to-fail-std}{47.31}