### 2022-08-29 paper/ase22/tables/test-results-3-50-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-3-50-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-3-50-lc7-num-seeds}{50}
\DefMacro{test-results-3-50-lc7-num-exps-avg}{1564.67}
\DefMacro{test-results-3-50-lc7-num-exps-med}{1659}
\DefMacro{test-results-3-50-lc7-num-exps-std}{277.77}
\DefMacro{test-results-3-50-model0-lc7-num-seed-fail-avg}{24.0}
\DefMacro{test-results-3-50-model0-lc7-num-seed-fail-med}{24.0}
\DefMacro{test-results-3-50-model0-lc7-num-seed-fail-std}{5.0}
\DefMacro{test-results-3-50-model0-lc7-num-exp-fail-avg}{24.58}
\DefMacro{test-results-3-50-model0-lc7-num-exp-fail-med}{21.87}
\DefMacro{test-results-3-50-model0-lc7-num-exp-fail-std}{3.41}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-avg}{46.0}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-med}{51}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-std}{17.88}
\DefMacro{test-results-3-50-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-3-50-lc3-num-seeds}{50}
\DefMacro{test-results-3-50-lc3-num-exps-avg}{2122.0}
\DefMacro{test-results-3-50-lc3-num-exps-med}{2073}
\DefMacro{test-results-3-50-lc3-num-exps-std}{98.56}
\DefMacro{test-results-3-50-model0-lc3-num-seed-fail-avg}{89.33}
\DefMacro{test-results-3-50-model0-lc3-num-seed-fail-med}{88.0}
\DefMacro{test-results-3-50-model0-lc3-num-seed-fail-std}{1.155}
\DefMacro{test-results-3-50-model0-lc3-num-exp-fail-avg}{90.63}
\DefMacro{test-results-3-50-model0-lc3-num-exp-fail-med}{89.87}
\DefMacro{test-results-3-50-model0-lc3-num-exp-fail-std}{0.705}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-avg}{60.33}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-med}{58}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-std}{3.405}
\DefMacro{test-results-3-50-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-3-50-lc4-num-seeds}{26}
\DefMacro{test-results-3-50-lc4-num-exps-avg}{1091.67}
\DefMacro{test-results-3-50-lc4-num-exps-med}{1090}
\DefMacro{test-results-3-50-lc4-num-exps-std}{8.78}
\DefMacro{test-results-3-50-model0-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-50-model0-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-50-model0-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model0-lc4-num-exp-fail-avg}{93.04}
\DefMacro{test-results-3-50-model0-lc4-num-exp-fail-med}{93.02}
\DefMacro{test-results-3-50-model0-lc4-num-exp-fail-std}{0.035}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-avg}{62.67}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-med}{61}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-std}{3.33}
\DefMacro{test-results-3-50-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-3-50-lc6-num-seeds}{50}
\DefMacro{test-results-3-50-lc6-num-exps-avg}{1110.33}
\DefMacro{test-results-3-50-lc6-num-exps-med}{1137}
\DefMacro{test-results-3-50-lc6-num-exps-std}{56.69}
\DefMacro{test-results-3-50-model0-lc6-num-seed-fail-avg}{84.67}
\DefMacro{test-results-3-50-model0-lc6-num-seed-fail-med}{88.0}
\DefMacro{test-results-3-50-model0-lc6-num-seed-fail-std}{3.785}
\DefMacro{test-results-3-50-model0-lc6-num-exp-fail-avg}{88.51}
\DefMacro{test-results-3-50-model0-lc6-num-exp-fail-med}{89.86}
\DefMacro{test-results-3-50-model0-lc6-num-exp-fail-std}{4.55}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-avg}{8.0}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-med}{6}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-std}{2.18}
\DefMacro{test-results-3-50-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-3-50-lc5-num-seeds}{50}
\DefMacro{test-results-3-50-lc5-num-exps-avg}{1477.0}
\DefMacro{test-results-3-50-lc5-num-exps-med}{1345}
\DefMacro{test-results-3-50-lc5-num-exps-std}{128.42}
\DefMacro{test-results-3-50-model0-lc5-num-seed-fail-avg}{99.33}
\DefMacro{test-results-3-50-model0-lc5-num-seed-fail-med}{100.0}
\DefMacro{test-results-3-50-model0-lc5-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model0-lc5-num-exp-fail-avg}{99.73}
\DefMacro{test-results-3-50-model0-lc5-num-exp-fail-med}{99.94}
\DefMacro{test-results-3-50-model0-lc5-num-exp-fail-std}{0.205}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-3-50-lc9-num-seeds}{50}
\DefMacro{test-results-3-50-lc9-num-exps-avg}{1494.33}
\DefMacro{test-results-3-50-lc9-num-exps-med}{1417}
\DefMacro{test-results-3-50-lc9-num-exps-std}{121.215}
\DefMacro{test-results-3-50-model0-lc9-num-seed-fail-avg}{76.67}
\DefMacro{test-results-3-50-model0-lc9-num-seed-fail-med}{80.0}
\DefMacro{test-results-3-50-model0-lc9-num-seed-fail-std}{3.785}
\DefMacro{test-results-3-50-model0-lc9-num-exp-fail-avg}{75.51}
\DefMacro{test-results-3-50-model0-lc9-num-exp-fail-med}{79.39}
\DefMacro{test-results-3-50-model0-lc9-num-exp-fail-std}{3.86}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-avg}{38.33}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-med}{27}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-std}{15.78}
\DefMacro{test-results-3-50-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-3-50-lc2-num-seeds}{50}
\DefMacro{test-results-3-50-lc2-num-exps-avg}{1544.67}
\DefMacro{test-results-3-50-lc2-num-exps-med}{1523}
\DefMacro{test-results-3-50-lc2-num-exps-std}{34.065}
\DefMacro{test-results-3-50-model0-lc2-num-seed-fail-avg}{14.0}
\DefMacro{test-results-3-50-model0-lc2-num-seed-fail-med}{12.0}
\DefMacro{test-results-3-50-model0-lc2-num-seed-fail-std}{3.605}
\DefMacro{test-results-3-50-model0-lc2-num-exp-fail-avg}{15.77}
\DefMacro{test-results-3-50-model0-lc2-num-exp-fail-med}{13.7}
\DefMacro{test-results-3-50-model0-lc2-num-exp-fail-std}{5.195}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-avg}{60.67}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-med}{52}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-std}{27.26}
\DefMacro{test-results-3-50-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-3-50-lc0-num-seeds}{19}
\DefMacro{test-results-3-50-lc0-num-exps-avg}{232.33}
\DefMacro{test-results-3-50-lc0-num-exps-med}{224}
\DefMacro{test-results-3-50-lc0-num-exps-std}{7.655}
\DefMacro{test-results-3-50-model0-lc0-num-seed-fail-avg}{78.95}
\DefMacro{test-results-3-50-model0-lc0-num-seed-fail-med}{78.95}
\DefMacro{test-results-3-50-model0-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model0-lc0-num-exp-fail-avg}{87.8}
\DefMacro{test-results-3-50-model0-lc0-num-exp-fail-med}{88.0}
\DefMacro{test-results-3-50-model0-lc0-num-exp-fail-std}{0.36}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-avg}{17.33}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-med}{17}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-std}{0.29}
\DefMacro{test-results-3-50-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-3-50-lc1-num-seeds}{50}
\DefMacro{test-results-3-50-lc1-num-exps-avg}{454.0}
\DefMacro{test-results-3-50-lc1-num-exps-med}{467}
\DefMacro{test-results-3-50-lc1-num-exps-std}{42.625}
\DefMacro{test-results-3-50-model0-lc1-num-seed-fail-avg}{0.0}
\DefMacro{test-results-3-50-model0-lc1-num-seed-fail-med}{0.0}
\DefMacro{test-results-3-50-model0-lc1-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model0-lc1-num-exp-fail-avg}{1.38}
\DefMacro{test-results-3-50-model0-lc1-num-exp-fail-med}{1.28}
\DefMacro{test-results-3-50-model0-lc1-num-exp-fail-std}{0.25}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-avg}{6.0}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-med}{6}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-std}{0.5}
\DefMacro{test-results-3-50-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-3-50-lc8-num-seeds}{50}
\DefMacro{test-results-3-50-lc8-num-exps-avg}{1330.67}
\DefMacro{test-results-3-50-lc8-num-exps-med}{1319}
\DefMacro{test-results-3-50-lc8-num-exps-std}{33.63}
\DefMacro{test-results-3-50-model0-lc8-num-seed-fail-avg}{4.67}
\DefMacro{test-results-3-50-model0-lc8-num-seed-fail-med}{4.0}
\DefMacro{test-results-3-50-model0-lc8-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model0-lc8-num-exp-fail-avg}{5.87}
\DefMacro{test-results-3-50-model0-lc8-num-exp-fail-med}{6.67}
\DefMacro{test-results-3-50-model0-lc8-num-exp-fail-std}{1.25}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-avg}{27.33}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-med}{20}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-std}{8.605}
\DefMacro{test-results-3-50-model1-lc7-num-seed-fail-avg}{12.67}
\DefMacro{test-results-3-50-model1-lc7-num-seed-fail-med}{14.0}
\DefMacro{test-results-3-50-model1-lc7-num-seed-fail-std}{2.08}
\DefMacro{test-results-3-50-model1-lc7-num-exp-fail-avg}{20.97}
\DefMacro{test-results-3-50-model1-lc7-num-exp-fail-med}{15.97}
\DefMacro{test-results-3-50-model1-lc7-num-exp-fail-std}{5.215}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-avg}{59.67}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-med}{47}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-std}{31.48}
\DefMacro{test-results-3-50-model1-lc3-num-seed-fail-avg}{86.0}
\DefMacro{test-results-3-50-model1-lc3-num-seed-fail-med}{84.0}
\DefMacro{test-results-3-50-model1-lc3-num-seed-fail-std}{1.73}
\DefMacro{test-results-3-50-model1-lc3-num-exp-fail-avg}{88.14}
\DefMacro{test-results-3-50-model1-lc3-num-exp-fail-med}{87.15}
\DefMacro{test-results-3-50-model1-lc3-num-exp-fail-std}{1.31}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-avg}{53.67}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-med}{51}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-std}{3.685}
\DefMacro{test-results-3-50-model1-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-50-model1-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-50-model1-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model1-lc4-num-exp-fail-avg}{93.31}
\DefMacro{test-results-3-50-model1-lc4-num-exp-fail-med}{93.33}
\DefMacro{test-results-3-50-model1-lc4-num-exp-fail-std}{0.14}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-avg}{40.33}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-med}{41}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-std}{1.53}
\DefMacro{test-results-3-50-model1-lc6-num-seed-fail-avg}{74.67}
\DefMacro{test-results-3-50-model1-lc6-num-seed-fail-med}{74.0}
\DefMacro{test-results-3-50-model1-lc6-num-seed-fail-std}{3.51}
\DefMacro{test-results-3-50-model1-lc6-num-exp-fail-avg}{73.04}
\DefMacro{test-results-3-50-model1-lc6-num-exp-fail-med}{70.1}
\DefMacro{test-results-3-50-model1-lc6-num-exp-fail-std}{5.935}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-avg}{30.33}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-med}{26}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-std}{14.375}
\DefMacro{test-results-3-50-model1-lc5-num-seed-fail-avg}{99.33}
\DefMacro{test-results-3-50-model1-lc5-num-seed-fail-med}{100.0}
\DefMacro{test-results-3-50-model1-lc5-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model1-lc5-num-exp-fail-avg}{99.4}
\DefMacro{test-results-3-50-model1-lc5-num-exp-fail-med}{99.26}
\DefMacro{test-results-3-50-model1-lc5-num-exp-fail-std}{0.2}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-model1-lc9-num-seed-fail-avg}{78.0}
\DefMacro{test-results-3-50-model1-lc9-num-seed-fail-med}{78.0}
\DefMacro{test-results-3-50-model1-lc9-num-seed-fail-std}{5.0}
\DefMacro{test-results-3-50-model1-lc9-num-exp-fail-avg}{78.77}
\DefMacro{test-results-3-50-model1-lc9-num-exp-fail-med}{82.92}
\DefMacro{test-results-3-50-model1-lc9-num-exp-fail-std}{5.985}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-avg}{20.0}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-med}{13}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-std}{10.21}
\DefMacro{test-results-3-50-model1-lc2-num-seed-fail-avg}{52.0}
\DefMacro{test-results-3-50-model1-lc2-num-seed-fail-med}{50.0}
\DefMacro{test-results-3-50-model1-lc2-num-seed-fail-std}{1.73}
\DefMacro{test-results-3-50-model1-lc2-num-exp-fail-avg}{46.53}
\DefMacro{test-results-3-50-model1-lc2-num-exp-fail-med}{43.06}
\DefMacro{test-results-3-50-model1-lc2-num-exp-fail-std}{4.615}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-avg}{37.0}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-med}{31}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-std}{8.41}
\DefMacro{test-results-3-50-model1-lc0-num-seed-fail-avg}{89.47}
\DefMacro{test-results-3-50-model1-lc0-num-seed-fail-med}{89.47}
\DefMacro{test-results-3-50-model1-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model1-lc0-num-exp-fail-avg}{78.7}
\DefMacro{test-results-3-50-model1-lc0-num-exp-fail-med}{78.12}
\DefMacro{test-results-3-50-model1-lc0-num-exp-fail-std}{0.75}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-avg}{9.33}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-med}{9}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-std}{0.29}
\DefMacro{test-results-3-50-model1-lc1-num-seed-fail-avg}{2.67}
\DefMacro{test-results-3-50-model1-lc1-num-seed-fail-med}{2.0}
\DefMacro{test-results-3-50-model1-lc1-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model1-lc1-num-exp-fail-avg}{2.8}
\DefMacro{test-results-3-50-model1-lc1-num-exp-fail-med}{0.94}
\DefMacro{test-results-3-50-model1-lc1-num-exp-fail-std}{1.66}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-avg}{12.0}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-med}{3}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-std}{8.23}
\DefMacro{test-results-3-50-model1-lc8-num-seed-fail-avg}{5.33}
\DefMacro{test-results-3-50-model1-lc8-num-seed-fail-med}{6.0}
\DefMacro{test-results-3-50-model1-lc8-num-seed-fail-std}{1.53}
\DefMacro{test-results-3-50-model1-lc8-num-exp-fail-avg}{5.85}
\DefMacro{test-results-3-50-model1-lc8-num-exp-fail-med}{6.14}
\DefMacro{test-results-3-50-model1-lc8-num-exp-fail-std}{0.93}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-avg}{25.67}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-med}{27}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-std}{5.53}
\DefMacro{test-results-3-50-model2-lc7-num-seed-fail-avg}{32.0}
\DefMacro{test-results-3-50-model2-lc7-num-seed-fail-med}{24.0}
\DefMacro{test-results-3-50-model2-lc7-num-seed-fail-std}{10.585}
\DefMacro{test-results-3-50-model2-lc7-num-exp-fail-avg}{31.45}
\DefMacro{test-results-3-50-model2-lc7-num-exp-fail-med}{33.91}
\DefMacro{test-results-3-50-model2-lc7-num-exp-fail-std}{5.245}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-avg}{50.67}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-med}{49}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-std}{15.765}
\DefMacro{test-results-3-50-model2-lc3-num-seed-fail-avg}{83.33}
\DefMacro{test-results-3-50-model2-lc3-num-seed-fail-med}{84.0}
\DefMacro{test-results-3-50-model2-lc3-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model2-lc3-num-exp-fail-avg}{85.71}
\DefMacro{test-results-3-50-model2-lc3-num-exp-fail-med}{86.11}
\DefMacro{test-results-3-50-model2-lc3-num-exp-fail-std}{0.64}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-avg}{33.67}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-med}{34}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-std}{2.255}
\DefMacro{test-results-3-50-model2-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-50-model2-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-50-model2-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model2-lc4-num-exp-fail-avg}{95.66}
\DefMacro{test-results-3-50-model2-lc4-num-exp-fail-med}{95.41}
\DefMacro{test-results-3-50-model2-lc4-num-exp-fail-std}{0.25}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-avg}{84.0}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-med}{83}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-std}{2.29}
\DefMacro{test-results-3-50-model2-lc6-num-seed-fail-avg}{87.33}
\DefMacro{test-results-3-50-model2-lc6-num-seed-fail-med}{88.0}
\DefMacro{test-results-3-50-model2-lc6-num-seed-fail-std}{3.51}
\DefMacro{test-results-3-50-model2-lc6-num-exp-fail-avg}{91.1}
\DefMacro{test-results-3-50-model2-lc6-num-exp-fail-med}{92.9}
\DefMacro{test-results-3-50-model2-lc6-num-exp-fail-std}{4.19}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-avg}{4.33}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-med}{4}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-std}{1.26}
\DefMacro{test-results-3-50-model2-lc5-num-seed-fail-avg}{98.67}
\DefMacro{test-results-3-50-model2-lc5-num-seed-fail-med}{98.0}
\DefMacro{test-results-3-50-model2-lc5-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model2-lc5-num-exp-fail-avg}{98.44}
\DefMacro{test-results-3-50-model2-lc5-num-exp-fail-med}{99.11}
\DefMacro{test-results-3-50-model2-lc5-num-exp-fail-std}{0.99}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-avg}{1.33}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-std}{1.155}
\DefMacro{test-results-3-50-model2-lc9-num-seed-fail-avg}{80.67}
\DefMacro{test-results-3-50-model2-lc9-num-seed-fail-med}{84.0}
\DefMacro{test-results-3-50-model2-lc9-num-seed-fail-std}{9.61}
\DefMacro{test-results-3-50-model2-lc9-num-exp-fail-avg}{78.54}
\DefMacro{test-results-3-50-model2-lc9-num-exp-fail-med}{82.31}
\DefMacro{test-results-3-50-model2-lc9-num-exp-fail-std}{11.03}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-avg}{13.0}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-med}{4}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-std}{7.795}
\DefMacro{test-results-3-50-model2-lc2-num-seed-fail-avg}{54.67}
\DefMacro{test-results-3-50-model2-lc2-num-seed-fail-med}{58.0}
\DefMacro{test-results-3-50-model2-lc2-num-seed-fail-std}{2.885}
\DefMacro{test-results-3-50-model2-lc2-num-exp-fail-avg}{55.97}
\DefMacro{test-results-3-50-model2-lc2-num-exp-fail-med}{60.67}
\DefMacro{test-results-3-50-model2-lc2-num-exp-fail-std}{4.6}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-avg}{49.33}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-med}{60}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-std}{11.015}
\DefMacro{test-results-3-50-model2-lc0-num-seed-fail-avg}{100.0}
\DefMacro{test-results-3-50-model2-lc0-num-seed-fail-med}{100.0}
\DefMacro{test-results-3-50-model2-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model2-lc0-num-exp-fail-avg}{95.8}
\DefMacro{test-results-3-50-model2-lc0-num-exp-fail-med}{95.96}
\DefMacro{test-results-3-50-model2-lc0-num-exp-fail-std}{0.545}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-model2-lc1-num-seed-fail-avg}{0.0}
\DefMacro{test-results-3-50-model2-lc1-num-seed-fail-med}{0.0}
\DefMacro{test-results-3-50-model2-lc1-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-50-model2-lc1-num-exp-fail-avg}{0.92}
\DefMacro{test-results-3-50-model2-lc1-num-exp-fail-med}{0.94}
\DefMacro{test-results-3-50-model2-lc1-num-exp-fail-std}{0.24}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-avg}{4.0}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-med}{5}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-std}{0.865}
\DefMacro{test-results-3-50-model2-lc8-num-seed-fail-avg}{5.33}
\DefMacro{test-results-3-50-model2-lc8-num-seed-fail-med}{6.0}
\DefMacro{test-results-3-50-model2-lc8-num-seed-fail-std}{0.575}
\DefMacro{test-results-3-50-model2-lc8-num-exp-fail-avg}{6.76}
\DefMacro{test-results-3-50-model2-lc8-num-exp-fail-med}{6.07}
\DefMacro{test-results-3-50-model2-lc8-num-exp-fail-std}{1.515}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-avg}{25.33}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-med}{13}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-std}{12.005}