### 2022-08-29 paper/icse23/tables/test-results-3-100-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-3-100-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-3-100-lc7-num-seeds}{100}
\DefMacro{test-results-3-100-lc7-num-exps-avg}{3426.33}
\DefMacro{test-results-3-100-lc7-num-exps-med}{3435}
\DefMacro{test-results-3-100-lc7-num-exps-std}{361.02}
\DefMacro{test-results-3-100-model0-lc7-num-seed-fail-avg}{26.0}
\DefMacro{test-results-3-100-model0-lc7-num-seed-fail-med}{23.0}
\DefMacro{test-results-3-100-model0-lc7-num-seed-fail-std}{5.895}
\DefMacro{test-results-3-100-model0-lc7-num-exp-fail-avg}{25.68}
\DefMacro{test-results-3-100-model0-lc7-num-exp-fail-med}{22.15}
\DefMacro{test-results-3-100-model0-lc7-num-exp-fail-std}{3.105}
\DefMacro{test-results-3-100-model0-lc7-num-pass-to-fail-avg}{88.33}
\DefMacro{test-results-3-100-model0-lc7-num-pass-to-fail-med}{107}
\DefMacro{test-results-3-100-model0-lc7-num-pass-to-fail-std}{25.8}
\DefMacro{test-results-3-100-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-3-100-lc3-num-seeds}{67}
\DefMacro{test-results-3-100-lc3-num-exps-avg}{2860.67}
\DefMacro{test-results-3-100-lc3-num-exps-med}{2868}
\DefMacro{test-results-3-100-lc3-num-exps-std}{19.755}
\DefMacro{test-results-3-100-model0-lc3-num-seed-fail-avg}{89.55}
\DefMacro{test-results-3-100-model0-lc3-num-seed-fail-med}{89.55}
\DefMacro{test-results-3-100-model0-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model0-lc3-num-exp-fail-avg}{91.64}
\DefMacro{test-results-3-100-model0-lc3-num-exp-fail-med}{91.73}
\DefMacro{test-results-3-100-model0-lc3-num-exp-fail-std}{0.185}
\DefMacro{test-results-3-100-model0-lc3-num-pass-to-fail-avg}{77.67}
\DefMacro{test-results-3-100-model0-lc3-num-pass-to-fail-med}{79}
\DefMacro{test-results-3-100-model0-lc3-num-pass-to-fail-std}{1.605}
\DefMacro{test-results-3-100-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-3-100-lc4-num-seeds}{26}
\DefMacro{test-results-3-100-lc4-num-exps-avg}{1086.0}
\DefMacro{test-results-3-100-lc4-num-exps-med}{1090}
\DefMacro{test-results-3-100-lc4-num-exps-std}{3.905}
\DefMacro{test-results-3-100-model0-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-100-model0-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-100-model0-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model0-lc4-num-exp-fail-avg}{92.91}
\DefMacro{test-results-3-100-model0-lc4-num-exp-fail-med}{93.13}
\DefMacro{test-results-3-100-model0-lc4-num-exp-fail-std}{0.355}
\DefMacro{test-results-3-100-model0-lc4-num-pass-to-fail-avg}{63.33}
\DefMacro{test-results-3-100-model0-lc4-num-pass-to-fail-med}{65}
\DefMacro{test-results-3-100-model0-lc4-num-pass-to-fail-std}{1.445}
\DefMacro{test-results-3-100-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-3-100-lc6-num-seeds}{100}
\DefMacro{test-results-3-100-lc6-num-exps-avg}{2312.0}
\DefMacro{test-results-3-100-lc6-num-exps-med}{2296}
\DefMacro{test-results-3-100-lc6-num-exps-std}{53.945}
\DefMacro{test-results-3-100-model0-lc6-num-seed-fail-avg}{86.33}
\DefMacro{test-results-3-100-model0-lc6-num-seed-fail-med}{87.0}
\DefMacro{test-results-3-100-model0-lc6-num-seed-fail-std}{3.51}
\DefMacro{test-results-3-100-model0-lc6-num-exp-fail-avg}{91.58}
\DefMacro{test-results-3-100-model0-lc6-num-exp-fail-med}{94.64}
\DefMacro{test-results-3-100-model0-lc6-num-exp-fail-std}{3.155}
\DefMacro{test-results-3-100-model0-lc6-num-pass-to-fail-avg}{25.67}
\DefMacro{test-results-3-100-model0-lc6-num-pass-to-fail-med}{26}
\DefMacro{test-results-3-100-model0-lc6-num-pass-to-fail-std}{10.25}
\DefMacro{test-results-3-100-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-3-100-lc5-num-seeds}{100}
\DefMacro{test-results-3-100-lc5-num-exps-avg}{2845.0}
\DefMacro{test-results-3-100-lc5-num-exps-med}{2713}
\DefMacro{test-results-3-100-lc5-num-exps-std}{133.37}
\DefMacro{test-results-3-100-model0-lc5-num-seed-fail-avg}{99.33}
\DefMacro{test-results-3-100-model0-lc5-num-seed-fail-med}{99.0}
\DefMacro{test-results-3-100-model0-lc5-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-100-model0-lc5-num-exp-fail-avg}{99.63}
\DefMacro{test-results-3-100-model0-lc5-num-exp-fail-med}{99.67}
\DefMacro{test-results-3-100-model0-lc5-num-exp-fail-std}{0.165}
\DefMacro{test-results-3-100-model0-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-100-model0-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-100-model0-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-100-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-3-100-lc9-num-seeds}{100}
\DefMacro{test-results-3-100-lc9-num-exps-avg}{2830.33}
\DefMacro{test-results-3-100-lc9-num-exps-med}{2607}
\DefMacro{test-results-3-100-lc9-num-exps-std}{221.255}
\DefMacro{test-results-3-100-model0-lc9-num-seed-fail-avg}{74.33}
\DefMacro{test-results-3-100-model0-lc9-num-seed-fail-med}{76.0}
\DefMacro{test-results-3-100-model0-lc9-num-seed-fail-std}{2.845}
\DefMacro{test-results-3-100-model0-lc9-num-exp-fail-avg}{72.26}
\DefMacro{test-results-3-100-model0-lc9-num-exp-fail-med}{71.66}
\DefMacro{test-results-3-100-model0-lc9-num-exp-fail-std}{1.795}
\DefMacro{test-results-3-100-model0-lc9-num-pass-to-fail-avg}{66.0}
\DefMacro{test-results-3-100-model0-lc9-num-pass-to-fail-med}{56}
\DefMacro{test-results-3-100-model0-lc9-num-pass-to-fail-std}{10.44}
\DefMacro{test-results-3-100-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-3-100-lc2-num-seeds}{100}
\DefMacro{test-results-3-100-lc2-num-exps-avg}{3097.33}
\DefMacro{test-results-3-100-lc2-num-exps-med}{3151}
\DefMacro{test-results-3-100-lc2-num-exps-std}{60.86}
\DefMacro{test-results-3-100-model0-lc2-num-seed-fail-avg}{20.0}
\DefMacro{test-results-3-100-model0-lc2-num-seed-fail-med}{18.0}
\DefMacro{test-results-3-100-model0-lc2-num-seed-fail-std}{2.18}
\DefMacro{test-results-3-100-model0-lc2-num-exp-fail-avg}{19.77}
\DefMacro{test-results-3-100-model0-lc2-num-exp-fail-med}{20.11}
\DefMacro{test-results-3-100-model0-lc2-num-exp-fail-std}{3.26}
\DefMacro{test-results-3-100-model0-lc2-num-pass-to-fail-avg}{96.33}
\DefMacro{test-results-3-100-model0-lc2-num-pass-to-fail-med}{80}
\DefMacro{test-results-3-100-model0-lc2-num-pass-to-fail-std}{23.825}
\DefMacro{test-results-3-100-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-3-100-lc0-num-seeds}{19}
\DefMacro{test-results-3-100-lc0-num-exps-avg}{234.33}
\DefMacro{test-results-3-100-lc0-num-exps-med}{239}
\DefMacro{test-results-3-100-lc0-num-exps-std}{5.395}
\DefMacro{test-results-3-100-model0-lc0-num-seed-fail-avg}{78.95}
\DefMacro{test-results-3-100-model0-lc0-num-seed-fail-med}{78.95}
\DefMacro{test-results-3-100-model0-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model0-lc0-num-exp-fail-avg}{88.79}
\DefMacro{test-results-3-100-model0-lc0-num-exp-fail-med}{88.7}
\DefMacro{test-results-3-100-model0-lc0-num-exp-fail-std}{0.405}
\DefMacro{test-results-3-100-model0-lc0-num-pass-to-fail-avg}{18.67}
\DefMacro{test-results-3-100-model0-lc0-num-pass-to-fail-med}{19}
\DefMacro{test-results-3-100-model0-lc0-num-pass-to-fail-std}{0.765}
\DefMacro{test-results-3-100-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-3-100-lc1-num-seeds}{100}
\DefMacro{test-results-3-100-lc1-num-exps-avg}{973.33}
\DefMacro{test-results-3-100-lc1-num-exps-med}{969}
\DefMacro{test-results-3-100-lc1-num-exps-std}{8.95}
\DefMacro{test-results-3-100-model0-lc1-num-seed-fail-avg}{1.67}
\DefMacro{test-results-3-100-model0-lc1-num-seed-fail-med}{2.0}
\DefMacro{test-results-3-100-model0-lc1-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-100-model0-lc1-num-exp-fail-avg}{3.72}
\DefMacro{test-results-3-100-model0-lc1-num-exp-fail-med}{4.53}
\DefMacro{test-results-3-100-model0-lc1-num-exp-fail-std}{0.705}
\DefMacro{test-results-3-100-model0-lc1-num-pass-to-fail-avg}{11.67}
\DefMacro{test-results-3-100-model0-lc1-num-pass-to-fail-med}{12}
\DefMacro{test-results-3-100-model0-lc1-num-pass-to-fail-std}{4.75}
\DefMacro{test-results-3-100-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-3-100-lc8-num-seeds}{100}
\DefMacro{test-results-3-100-lc8-num-exps-avg}{2749.0}
\DefMacro{test-results-3-100-lc8-num-exps-med}{2755}
\DefMacro{test-results-3-100-lc8-num-exps-std}{44.575}
\DefMacro{test-results-3-100-model0-lc8-num-seed-fail-avg}{5.33}
\DefMacro{test-results-3-100-model0-lc8-num-seed-fail-med}{6.0}
\DefMacro{test-results-3-100-model0-lc8-num-seed-fail-std}{1.04}
\DefMacro{test-results-3-100-model0-lc8-num-exp-fail-avg}{4.93}
\DefMacro{test-results-3-100-model0-lc8-num-exp-fail-med}{4.93}
\DefMacro{test-results-3-100-model0-lc8-num-exp-fail-std}{0.44}
\DefMacro{test-results-3-100-model0-lc8-num-pass-to-fail-avg}{39.0}
\DefMacro{test-results-3-100-model0-lc8-num-pass-to-fail-med}{39}
\DefMacro{test-results-3-100-model0-lc8-num-pass-to-fail-std}{6.0}
\DefMacro{test-results-3-100-model1-lc7-num-seed-fail-avg}{13.33}
\DefMacro{test-results-3-100-model1-lc7-num-seed-fail-med}{11.0}
\DefMacro{test-results-3-100-model1-lc7-num-seed-fail-std}{3.405}
\DefMacro{test-results-3-100-model1-lc7-num-exp-fail-avg}{15.6}
\DefMacro{test-results-3-100-model1-lc7-num-exp-fail-med}{14.38}
\DefMacro{test-results-3-100-model1-lc7-num-exp-fail-std}{1.825}
\DefMacro{test-results-3-100-model1-lc7-num-pass-to-fail-avg}{121.33}
\DefMacro{test-results-3-100-model1-lc7-num-pass-to-fail-med}{148}
\DefMacro{test-results-3-100-model1-lc7-num-pass-to-fail-std}{43.075}
\DefMacro{test-results-3-100-model1-lc3-num-seed-fail-avg}{85.07}
\DefMacro{test-results-3-100-model1-lc3-num-seed-fail-med}{85.07}
\DefMacro{test-results-3-100-model1-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model1-lc3-num-exp-fail-avg}{88.99}
\DefMacro{test-results-3-100-model1-lc3-num-exp-fail-med}{88.72}
\DefMacro{test-results-3-100-model1-lc3-num-exp-fail-std}{0.335}
\DefMacro{test-results-3-100-model1-lc3-num-pass-to-fail-avg}{62.67}
\DefMacro{test-results-3-100-model1-lc3-num-pass-to-fail-med}{63}
\DefMacro{test-results-3-100-model1-lc3-num-pass-to-fail-std}{1.26}
\DefMacro{test-results-3-100-model1-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-100-model1-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-100-model1-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model1-lc4-num-exp-fail-avg}{93.55}
\DefMacro{test-results-3-100-model1-lc4-num-exp-fail-med}{93.41}
\DefMacro{test-results-3-100-model1-lc4-num-exp-fail-std}{0.305}
\DefMacro{test-results-3-100-model1-lc4-num-pass-to-fail-avg}{42.67}
\DefMacro{test-results-3-100-model1-lc4-num-pass-to-fail-med}{43}
\DefMacro{test-results-3-100-model1-lc4-num-pass-to-fail-std}{0.765}
\DefMacro{test-results-3-100-model1-lc6-num-seed-fail-avg}{73.0}
\DefMacro{test-results-3-100-model1-lc6-num-seed-fail-med}{69.0}
\DefMacro{test-results-3-100-model1-lc6-num-seed-fail-std}{4.36}
\DefMacro{test-results-3-100-model1-lc6-num-exp-fail-avg}{76.17}
\DefMacro{test-results-3-100-model1-lc6-num-exp-fail-med}{75.28}
\DefMacro{test-results-3-100-model1-lc6-num-exp-fail-std}{5.525}
\DefMacro{test-results-3-100-model1-lc6-num-pass-to-fail-avg}{58.0}
\DefMacro{test-results-3-100-model1-lc6-num-pass-to-fail-med}{42}
\DefMacro{test-results-3-100-model1-lc6-num-pass-to-fail-std}{14.73}
\DefMacro{test-results-3-100-model1-lc5-num-seed-fail-avg}{99.33}
\DefMacro{test-results-3-100-model1-lc5-num-seed-fail-med}{99.0}
\DefMacro{test-results-3-100-model1-lc5-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-100-model1-lc5-num-exp-fail-avg}{99.56}
\DefMacro{test-results-3-100-model1-lc5-num-exp-fail-med}{99.55}
\DefMacro{test-results-3-100-model1-lc5-num-exp-fail-std}{0.07}
\DefMacro{test-results-3-100-model1-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-100-model1-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-100-model1-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-100-model1-lc9-num-seed-fail-avg}{86.33}
\DefMacro{test-results-3-100-model1-lc9-num-seed-fail-med}{89.0}
\DefMacro{test-results-3-100-model1-lc9-num-seed-fail-std}{2.755}
\DefMacro{test-results-3-100-model1-lc9-num-exp-fail-avg}{87.16}
\DefMacro{test-results-3-100-model1-lc9-num-exp-fail-med}{86.91}
\DefMacro{test-results-3-100-model1-lc9-num-exp-fail-std}{1.285}
\DefMacro{test-results-3-100-model1-lc9-num-pass-to-fail-avg}{24.67}
\DefMacro{test-results-3-100-model1-lc9-num-pass-to-fail-med}{24}
\DefMacro{test-results-3-100-model1-lc9-num-pass-to-fail-std}{8.505}
\DefMacro{test-results-3-100-model1-lc2-num-seed-fail-avg}{50.33}
\DefMacro{test-results-3-100-model1-lc2-num-seed-fail-med}{52.0}
\DefMacro{test-results-3-100-model1-lc2-num-seed-fail-std}{2.845}
\DefMacro{test-results-3-100-model1-lc2-num-exp-fail-avg}{46.15}
\DefMacro{test-results-3-100-model1-lc2-num-exp-fail-med}{46.32}
\DefMacro{test-results-3-100-model1-lc2-num-exp-fail-std}{1.565}
\DefMacro{test-results-3-100-model1-lc2-num-pass-to-fail-avg}{78.33}
\DefMacro{test-results-3-100-model1-lc2-num-pass-to-fail-med}{76}
\DefMacro{test-results-3-100-model1-lc2-num-pass-to-fail-std}{17.28}
\DefMacro{test-results-3-100-model1-lc0-num-seed-fail-avg}{89.47}
\DefMacro{test-results-3-100-model1-lc0-num-seed-fail-med}{89.47}
\DefMacro{test-results-3-100-model1-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model1-lc0-num-exp-fail-avg}{78.95}
\DefMacro{test-results-3-100-model1-lc0-num-exp-fail-med}{78.83}
\DefMacro{test-results-3-100-model1-lc0-num-exp-fail-std}{0.665}
\DefMacro{test-results-3-100-model1-lc0-num-pass-to-fail-avg}{6.0}
\DefMacro{test-results-3-100-model1-lc0-num-pass-to-fail-med}{6}
\DefMacro{test-results-3-100-model1-lc0-num-pass-to-fail-std}{0.5}
\DefMacro{test-results-3-100-model1-lc1-num-seed-fail-avg}{2.67}
\DefMacro{test-results-3-100-model1-lc1-num-seed-fail-med}{3.0}
\DefMacro{test-results-3-100-model1-lc1-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-100-model1-lc1-num-exp-fail-avg}{3.66}
\DefMacro{test-results-3-100-model1-lc1-num-exp-fail-med}{3.76}
\DefMacro{test-results-3-100-model1-lc1-num-exp-fail-std}{0.31}
\DefMacro{test-results-3-100-model1-lc1-num-pass-to-fail-avg}{14.0}
\DefMacro{test-results-3-100-model1-lc1-num-pass-to-fail-med}{5}
\DefMacro{test-results-3-100-model1-lc1-num-pass-to-fail-std}{8.23}
\DefMacro{test-results-3-100-model1-lc8-num-seed-fail-avg}{5.0}
\DefMacro{test-results-3-100-model1-lc8-num-seed-fail-med}{5.0}
\DefMacro{test-results-3-100-model1-lc8-num-seed-fail-std}{0.5}
\DefMacro{test-results-3-100-model1-lc8-num-exp-fail-avg}{4.73}
\DefMacro{test-results-3-100-model1-lc8-num-exp-fail-med}{5.08}
\DefMacro{test-results-3-100-model1-lc8-num-exp-fail-std}{0.62}
\DefMacro{test-results-3-100-model1-lc8-num-pass-to-fail-avg}{42.67}
\DefMacro{test-results-3-100-model1-lc8-num-pass-to-fail-med}{49}
\DefMacro{test-results-3-100-model1-lc8-num-pass-to-fail-std}{8.22}
\DefMacro{test-results-3-100-model2-lc7-num-seed-fail-avg}{30.67}
\DefMacro{test-results-3-100-model2-lc7-num-seed-fail-med}{25.0}
\DefMacro{test-results-3-100-model2-lc7-num-seed-fail-std}{9.09}
\DefMacro{test-results-3-100-model2-lc7-num-exp-fail-avg}{26.16}
\DefMacro{test-results-3-100-model2-lc7-num-exp-fail-med}{28.7}
\DefMacro{test-results-3-100-model2-lc7-num-exp-fail-std}{2.495}
\DefMacro{test-results-3-100-model2-lc7-num-pass-to-fail-avg}{74.67}
\DefMacro{test-results-3-100-model2-lc7-num-pass-to-fail-med}{91}
\DefMacro{test-results-3-100-model2-lc7-num-pass-to-fail-std}{22.395}
\DefMacro{test-results-3-100-model2-lc3-num-seed-fail-avg}{85.07}
\DefMacro{test-results-3-100-model2-lc3-num-seed-fail-med}{85.07}
\DefMacro{test-results-3-100-model2-lc3-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model2-lc3-num-exp-fail-avg}{86.56}
\DefMacro{test-results-3-100-model2-lc3-num-exp-fail-med}{86.65}
\DefMacro{test-results-3-100-model2-lc3-num-exp-fail-std}{0.135}
\DefMacro{test-results-3-100-model2-lc3-num-pass-to-fail-avg}{40.33}
\DefMacro{test-results-3-100-model2-lc3-num-pass-to-fail-med}{40}
\DefMacro{test-results-3-100-model2-lc3-num-pass-to-fail-std}{3.255}
\DefMacro{test-results-3-100-model2-lc4-num-seed-fail-avg}{92.31}
\DefMacro{test-results-3-100-model2-lc4-num-seed-fail-med}{92.31}
\DefMacro{test-results-3-100-model2-lc4-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model2-lc4-num-exp-fail-avg}{95.43}
\DefMacro{test-results-3-100-model2-lc4-num-exp-fail-med}{95.26}
\DefMacro{test-results-3-100-model2-lc4-num-exp-fail-std}{0.24}
\DefMacro{test-results-3-100-model2-lc4-num-pass-to-fail-avg}{79.33}
\DefMacro{test-results-3-100-model2-lc4-num-pass-to-fail-med}{82}
\DefMacro{test-results-3-100-model2-lc4-num-pass-to-fail-std}{2.755}
\DefMacro{test-results-3-100-model2-lc6-num-seed-fail-avg}{87.33}
\DefMacro{test-results-3-100-model2-lc6-num-seed-fail-med}{86.0}
\DefMacro{test-results-3-100-model2-lc6-num-seed-fail-std}{3.055}
\DefMacro{test-results-3-100-model2-lc6-num-exp-fail-avg}{91.33}
\DefMacro{test-results-3-100-model2-lc6-num-exp-fail-med}{94.68}
\DefMacro{test-results-3-100-model2-lc6-num-exp-fail-std}{3.585}
\DefMacro{test-results-3-100-model2-lc6-num-pass-to-fail-avg}{11.0}
\DefMacro{test-results-3-100-model2-lc6-num-pass-to-fail-med}{9}
\DefMacro{test-results-3-100-model2-lc6-num-pass-to-fail-std}{3.605}
\DefMacro{test-results-3-100-model2-lc5-num-seed-fail-avg}{98.67}
\DefMacro{test-results-3-100-model2-lc5-num-seed-fail-med}{99.0}
\DefMacro{test-results-3-100-model2-lc5-num-seed-fail-std}{0.765}
\DefMacro{test-results-3-100-model2-lc5-num-exp-fail-avg}{98.42}
\DefMacro{test-results-3-100-model2-lc5-num-exp-fail-med}{98.99}
\DefMacro{test-results-3-100-model2-lc5-num-exp-fail-std}{0.77}
\DefMacro{test-results-3-100-model2-lc5-num-pass-to-fail-avg}{1.0}
\DefMacro{test-results-3-100-model2-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-100-model2-lc5-num-pass-to-fail-std}{0.865}
\DefMacro{test-results-3-100-model2-lc9-num-seed-fail-avg}{94.67}
\DefMacro{test-results-3-100-model2-lc9-num-seed-fail-med}{97.0}
\DefMacro{test-results-3-100-model2-lc9-num-seed-fail-std}{2.465}
\DefMacro{test-results-3-100-model2-lc9-num-exp-fail-avg}{94.43}
\DefMacro{test-results-3-100-model2-lc9-num-exp-fail-med}{96.41}
\DefMacro{test-results-3-100-model2-lc9-num-exp-fail-std}{2.69}
\DefMacro{test-results-3-100-model2-lc9-num-pass-to-fail-avg}{17.33}
\DefMacro{test-results-3-100-model2-lc9-num-pass-to-fail-med}{4}
\DefMacro{test-results-3-100-model2-lc9-num-pass-to-fail-std}{12.425}
\DefMacro{test-results-3-100-model2-lc2-num-seed-fail-avg}{58.67}
\DefMacro{test-results-3-100-model2-lc2-num-seed-fail-med}{58.0}
\DefMacro{test-results-3-100-model2-lc2-num-seed-fail-std}{3.015}
\DefMacro{test-results-3-100-model2-lc2-num-exp-fail-avg}{60.14}
\DefMacro{test-results-3-100-model2-lc2-num-exp-fail-med}{62.11}
\DefMacro{test-results-3-100-model2-lc2-num-exp-fail-std}{2.64}
\DefMacro{test-results-3-100-model2-lc2-num-pass-to-fail-avg}{79.33}
\DefMacro{test-results-3-100-model2-lc2-num-pass-to-fail-med}{82}
\DefMacro{test-results-3-100-model2-lc2-num-pass-to-fail-std}{9.075}
\DefMacro{test-results-3-100-model2-lc0-num-seed-fail-avg}{100.0}
\DefMacro{test-results-3-100-model2-lc0-num-seed-fail-med}{100.0}
\DefMacro{test-results-3-100-model2-lc0-num-seed-fail-std}{0.0}
\DefMacro{test-results-3-100-model2-lc0-num-exp-fail-avg}{95.75}
\DefMacro{test-results-3-100-model2-lc0-num-exp-fail-med}{95.82}
\DefMacro{test-results-3-100-model2-lc0-num-exp-fail-std}{0.34}
\DefMacro{test-results-3-100-model2-lc0-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-100-model2-lc0-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-100-model2-lc0-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-100-model2-lc1-num-seed-fail-avg}{1.33}
\DefMacro{test-results-3-100-model2-lc1-num-seed-fail-med}{1.0}
\DefMacro{test-results-3-100-model2-lc1-num-seed-fail-std}{0.29}
\DefMacro{test-results-3-100-model2-lc1-num-exp-fail-avg}{2.61}
\DefMacro{test-results-3-100-model2-lc1-num-exp-fail-med}{1.96}
\DefMacro{test-results-3-100-model2-lc1-num-exp-fail-std}{0.99}
\DefMacro{test-results-3-100-model2-lc1-num-pass-to-fail-avg}{8.0}
\DefMacro{test-results-3-100-model2-lc1-num-pass-to-fail-med}{8}
\DefMacro{test-results-3-100-model2-lc1-num-pass-to-fail-std}{1.5}
\DefMacro{test-results-3-100-model2-lc8-num-seed-fail-avg}{5.0}
\DefMacro{test-results-3-100-model2-lc8-num-seed-fail-med}{5.0}
\DefMacro{test-results-3-100-model2-lc8-num-seed-fail-std}{0.5}
\DefMacro{test-results-3-100-model2-lc8-num-exp-fail-avg}{5.97}
\DefMacro{test-results-3-100-model2-lc8-num-exp-fail-med}{5.01}
\DefMacro{test-results-3-100-model2-lc8-num-exp-fail-std}{0.935}
\DefMacro{test-results-3-100-model2-lc8-num-pass-to-fail-avg}{38.67}
\DefMacro{test-results-3-100-model2-lc8-num-pass-to-fail-med}{32}
\DefMacro{test-results-3-100-model2-lc8-num-pass-to-fail-std}{12.34}