### 2022-09-05 paper/ase22/tables/test-results-3-50-numbers.tex
final paper

%% Automatically generated by pyutil.latex

\DefMacro{test-results-3-50-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-3-50-lc7-num-seeds}{50}
\DefMacro{test-results-3-50-lc7-num-exps-avg}{1564.67}
\DefMacro{test-results-3-50-lc7-num-exps-med}{1659}
\DefMacro{test-results-3-50-lc7-num-exps-std}{277.77}
\DefMacro{test-results-3-50-model0-lc7-num-all-fail-avg}{375.0}
\DefMacro{test-results-3-50-model0-lc7-num-all-fail-med}{331}
\DefMacro{test-results-3-50-model0-lc7-num-all-fail-std}{38.54}
\DefMacro{test-results-3-50-model0-lc7-num-all-failrate-avg}{24.57}
\DefMacro{test-results-3-50-model0-lc7-num-all-failrate-med}{21.92}
\DefMacro{test-results-3-50-model0-lc7-num-all-failrate-std}{3.46}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-avg}{46.0}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-med}{51}
\DefMacro{test-results-3-50-model0-lc7-num-pass-to-fail-std}{17.88}
\DefMacro{test-results-3-50-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-3-50-lc3-num-seeds}{50}
\DefMacro{test-results-3-50-lc3-num-exps-avg}{2122.0}
\DefMacro{test-results-3-50-lc3-num-exps-med}{2073}
\DefMacro{test-results-3-50-lc3-num-exps-std}{98.56}
\DefMacro{test-results-3-50-model0-lc3-num-all-fail-avg}{1969.67}
\DefMacro{test-results-3-50-model0-lc3-num-all-fail-med}{1907}
\DefMacro{test-results-3-50-model0-lc3-num-all-fail-std}{105.065}
\DefMacro{test-results-3-50-model0-lc3-num-all-failrate-avg}{90.6}
\DefMacro{test-results-3-50-model0-lc3-num-all-failrate-med}{89.83}
\DefMacro{test-results-3-50-model0-lc3-num-all-failrate-std}{0.715}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-avg}{60.33}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-med}{58}
\DefMacro{test-results-3-50-model0-lc3-num-pass-to-fail-std}{3.405}
\DefMacro{test-results-3-50-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-3-50-lc4-num-seeds}{26}
\DefMacro{test-results-3-50-lc4-num-exps-avg}{1091.67}
\DefMacro{test-results-3-50-lc4-num-exps-med}{1090}
\DefMacro{test-results-3-50-lc4-num-exps-std}{8.78}
\DefMacro{test-results-3-50-model0-lc4-num-all-fail-avg}{1039.67}
\DefMacro{test-results-3-50-model0-lc4-num-all-fail-med}{1039}
\DefMacro{test-results-3-50-model0-lc4-num-all-fail-std}{8.005}
\DefMacro{test-results-3-50-model0-lc4-num-all-failrate-avg}{93.02}
\DefMacro{test-results-3-50-model0-lc4-num-all-failrate-med}{93.01}
\DefMacro{test-results-3-50-model0-lc4-num-all-failrate-std}{0.035}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-avg}{62.67}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-med}{61}
\DefMacro{test-results-3-50-model0-lc4-num-pass-to-fail-std}{3.33}
\DefMacro{test-results-3-50-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-3-50-lc6-num-seeds}{50}
\DefMacro{test-results-3-50-lc6-num-exps-avg}{1110.33}
\DefMacro{test-results-3-50-lc6-num-exps-med}{1137}
\DefMacro{test-results-3-50-lc6-num-exps-std}{56.69}
\DefMacro{test-results-3-50-model0-lc6-num-all-fail-avg}{1026.33}
\DefMacro{test-results-3-50-model0-lc6-num-all-fail-med}{934}
\DefMacro{test-results-3-50-model0-lc6-num-all-fail-std}{81.265}
\DefMacro{test-results-3-50-model0-lc6-num-all-failrate-avg}{88.35}
\DefMacro{test-results-3-50-model0-lc6-num-all-failrate-med}{89.86}
\DefMacro{test-results-3-50-model0-lc6-num-all-failrate-std}{4.5}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-avg}{8.0}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-med}{6}
\DefMacro{test-results-3-50-model0-lc6-num-pass-to-fail-std}{2.18}
\DefMacro{test-results-3-50-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-3-50-lc5-num-seeds}{50}
\DefMacro{test-results-3-50-lc5-num-exps-avg}{1477.0}
\DefMacro{test-results-3-50-lc5-num-exps-med}{1345}
\DefMacro{test-results-3-50-lc5-num-exps-std}{128.42}
\DefMacro{test-results-3-50-model0-lc5-num-all-fail-avg}{1523.0}
\DefMacro{test-results-3-50-model0-lc5-num-all-fail-med}{1384}
\DefMacro{test-results-3-50-model0-lc5-num-all-fail-std}{129.575}
\DefMacro{test-results-3-50-model0-lc5-num-all-failrate-avg}{99.72}
\DefMacro{test-results-3-50-model0-lc5-num-all-failrate-med}{99.95}
\DefMacro{test-results-3-50-model0-lc5-num-all-failrate-std}{0.22}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model0-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-3-50-lc9-num-seeds}{50}
\DefMacro{test-results-3-50-lc9-num-exps-avg}{1494.33}
\DefMacro{test-results-3-50-lc9-num-exps-med}{1417}
\DefMacro{test-results-3-50-lc9-num-exps-std}{121.215}
\DefMacro{test-results-3-50-model0-lc9-num-all-fail-avg}{1174.67}
\DefMacro{test-results-3-50-model0-lc9-num-all-fail-med}{1181}
\DefMacro{test-results-3-50-model0-lc9-num-all-fail-std}{135.78}
\DefMacro{test-results-3-50-model0-lc9-num-all-failrate-avg}{75.54}
\DefMacro{test-results-3-50-model0-lc9-num-all-failrate-med}{79.46}
\DefMacro{test-results-3-50-model0-lc9-num-all-failrate-std}{3.855}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-avg}{38.33}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-med}{27}
\DefMacro{test-results-3-50-model0-lc9-num-pass-to-fail-std}{15.78}
\DefMacro{test-results-3-50-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-3-50-lc2-num-seeds}{50}
\DefMacro{test-results-3-50-lc2-num-exps-avg}{1544.67}
\DefMacro{test-results-3-50-lc2-num-exps-med}{1523}
\DefMacro{test-results-3-50-lc2-num-exps-std}{34.065}
\DefMacro{test-results-3-50-model0-lc2-num-all-fail-avg}{251.0}
\DefMacro{test-results-3-50-model0-lc2-num-all-fail-med}{226}
\DefMacro{test-results-3-50-model0-lc2-num-all-fail-std}{80.48}
\DefMacro{test-results-3-50-model0-lc2-num-all-failrate-avg}{15.72}
\DefMacro{test-results-3-50-model0-lc2-num-all-failrate-med}{13.52}
\DefMacro{test-results-3-50-model0-lc2-num-all-failrate-std}{5.125}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-avg}{60.67}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-med}{52}
\DefMacro{test-results-3-50-model0-lc2-num-pass-to-fail-std}{27.26}
\DefMacro{test-results-3-50-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-3-50-lc0-num-seeds}{19}
\DefMacro{test-results-3-50-lc0-num-exps-avg}{232.33}
\DefMacro{test-results-3-50-lc0-num-exps-med}{224}
\DefMacro{test-results-3-50-lc0-num-exps-std}{7.655}
\DefMacro{test-results-3-50-model0-lc0-num-all-fail-avg}{219.0}
\DefMacro{test-results-3-50-model0-lc0-num-all-fail-med}{213}
\DefMacro{test-results-3-50-model0-lc0-num-all-fail-std}{7.0}
\DefMacro{test-results-3-50-model0-lc0-num-all-failrate-avg}{87.13}
\DefMacro{test-results-3-50-model0-lc0-num-all-failrate-med}{87.36}
\DefMacro{test-results-3-50-model0-lc0-num-all-failrate-std}{0.34}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-avg}{17.33}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-med}{17}
\DefMacro{test-results-3-50-model0-lc0-num-pass-to-fail-std}{0.29}
\DefMacro{test-results-3-50-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-3-50-lc1-num-seeds}{50}
\DefMacro{test-results-3-50-lc1-num-exps-avg}{454.0}
\DefMacro{test-results-3-50-lc1-num-exps-med}{467}
\DefMacro{test-results-3-50-lc1-num-exps-std}{42.625}
\DefMacro{test-results-3-50-model0-lc1-num-all-fail-avg}{6.0}
\DefMacro{test-results-3-50-model0-lc1-num-all-fail-med}{6}
\DefMacro{test-results-3-50-model0-lc1-num-all-fail-std}{0.5}
\DefMacro{test-results-3-50-model0-lc1-num-all-failrate-avg}{1.24}
\DefMacro{test-results-3-50-model0-lc1-num-all-failrate-med}{1.16}
\DefMacro{test-results-3-50-model0-lc1-num-all-failrate-std}{0.21}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-avg}{6.0}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-med}{6}
\DefMacro{test-results-3-50-model0-lc1-num-pass-to-fail-std}{0.5}
\DefMacro{test-results-3-50-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-3-50-lc8-num-seeds}{50}
\DefMacro{test-results-3-50-lc8-num-exps-avg}{1330.67}
\DefMacro{test-results-3-50-lc8-num-exps-med}{1319}
\DefMacro{test-results-3-50-lc8-num-exps-std}{33.63}
\DefMacro{test-results-3-50-model0-lc8-num-all-fail-avg}{79.33}
\DefMacro{test-results-3-50-model0-lc8-num-all-fail-med}{91}
\DefMacro{test-results-3-50-model0-lc8-num-all-fail-std}{15.12}
\DefMacro{test-results-3-50-model0-lc8-num-all-failrate-avg}{5.82}
\DefMacro{test-results-3-50-model0-lc8-num-all-failrate-med}{6.65}
\DefMacro{test-results-3-50-model0-lc8-num-all-failrate-std}{1.21}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-avg}{27.33}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-med}{20}
\DefMacro{test-results-3-50-model0-lc8-num-pass-to-fail-std}{8.605}
\DefMacro{test-results-3-50-model1-lc7-num-all-fail-avg}{297.33}
\DefMacro{test-results-3-50-model1-lc7-num-all-fail-med}{293}
\DefMacro{test-results-3-50-model1-lc7-num-all-fail-std}{13.875}
\DefMacro{test-results-3-50-model1-lc7-num-all-failrate-avg}{20.63}
\DefMacro{test-results-3-50-model1-lc7-num-all-failrate-med}{15.92}
\DefMacro{test-results-3-50-model1-lc7-num-all-failrate-std}{5.005}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-avg}{59.67}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-med}{47}
\DefMacro{test-results-3-50-model1-lc7-num-pass-to-fail-std}{31.48}
\DefMacro{test-results-3-50-model1-lc3-num-all-fail-avg}{1916.33}
\DefMacro{test-results-3-50-model1-lc3-num-all-fail-med}{1828}
\DefMacro{test-results-3-50-model1-lc3-num-all-fail-std}{114.34}
\DefMacro{test-results-3-50-model1-lc3-num-all-failrate-avg}{88.09}
\DefMacro{test-results-3-50-model1-lc3-num-all-failrate-med}{87.08}
\DefMacro{test-results-3-50-model1-lc3-num-all-failrate-std}{1.32}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-avg}{53.67}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-med}{51}
\DefMacro{test-results-3-50-model1-lc3-num-pass-to-fail-std}{3.685}
\DefMacro{test-results-3-50-model1-lc4-num-all-fail-avg}{1042.67}
\DefMacro{test-results-3-50-model1-lc4-num-all-fail-med}{1038}
\DefMacro{test-results-3-50-model1-lc4-num-all-fail-std}{7.765}
\DefMacro{test-results-3-50-model1-lc4-num-all-failrate-avg}{93.29}
\DefMacro{test-results-3-50-model1-lc4-num-all-failrate-med}{93.31}
\DefMacro{test-results-3-50-model1-lc4-num-all-failrate-std}{0.135}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-avg}{40.33}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-med}{41}
\DefMacro{test-results-3-50-model1-lc4-num-pass-to-fail-std}{1.53}
\DefMacro{test-results-3-50-model1-lc6-num-all-fail-avg}{839.33}
\DefMacro{test-results-3-50-model1-lc6-num-all-fail-med}{831}
\DefMacro{test-results-3-50-model1-lc6-num-all-fail-std}{23.53}
\DefMacro{test-results-3-50-model1-lc6-num-all-failrate-avg}{73.09}
\DefMacro{test-results-3-50-model1-lc6-num-all-failrate-med}{70.01}
\DefMacro{test-results-3-50-model1-lc6-num-all-failrate-std}{5.795}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-avg}{30.33}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-med}{26}
\DefMacro{test-results-3-50-model1-lc6-num-pass-to-fail-std}{14.375}
\DefMacro{test-results-3-50-model1-lc5-num-all-fail-avg}{1517.33}
\DefMacro{test-results-3-50-model1-lc5-num-all-fail-med}{1384}
\DefMacro{test-results-3-50-model1-lc5-num-all-fail-std}{125.56}
\DefMacro{test-results-3-50-model1-lc5-num-all-failrate-avg}{99.4}
\DefMacro{test-results-3-50-model1-lc5-num-all-failrate-med}{99.21}
\DefMacro{test-results-3-50-model1-lc5-num-all-failrate-std}{0.2}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model1-lc5-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-model1-lc9-num-all-fail-avg}{1225.33}
\DefMacro{test-results-3-50-model1-lc9-num-all-fail-med}{1117}
\DefMacro{test-results-3-50-model1-lc9-num-all-fail-std}{166.975}
\DefMacro{test-results-3-50-model1-lc9-num-all-failrate-avg}{78.74}
\DefMacro{test-results-3-50-model1-lc9-num-all-failrate-med}{82.74}
\DefMacro{test-results-3-50-model1-lc9-num-all-failrate-std}{5.94}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-avg}{20.0}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-med}{13}
\DefMacro{test-results-3-50-model1-lc9-num-pass-to-fail-std}{10.21}
\DefMacro{test-results-3-50-model1-lc2-num-all-fail-avg}{744.33}
\DefMacro{test-results-3-50-model1-lc2-num-all-fail-med}{723}
\DefMacro{test-results-3-50-model1-lc2-num-all-fail-std}{71.105}
\DefMacro{test-results-3-50-model1-lc2-num-all-failrate-avg}{46.7}
\DefMacro{test-results-3-50-model1-lc2-num-all-failrate-med}{43.27}
\DefMacro{test-results-3-50-model1-lc2-num-all-failrate-std}{4.525}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-avg}{37.0}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-med}{31}
\DefMacro{test-results-3-50-model1-lc2-num-pass-to-fail-std}{8.41}
\DefMacro{test-results-3-50-model1-lc0-num-all-fail-avg}{200.0}
\DefMacro{test-results-3-50-model1-lc0-num-all-fail-med}{192}
\DefMacro{test-results-3-50-model1-lc0-num-all-fail-std}{7.81}
\DefMacro{test-results-3-50-model1-lc0-num-all-failrate-avg}{79.52}
\DefMacro{test-results-3-50-model1-lc0-num-all-failrate-med}{79.01}
\DefMacro{test-results-3-50-model1-lc0-num-all-failrate-std}{0.67}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-avg}{9.33}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-med}{9}
\DefMacro{test-results-3-50-model1-lc0-num-pass-to-fail-std}{0.29}
\DefMacro{test-results-3-50-model1-lc1-num-all-fail-avg}{14.33}
\DefMacro{test-results-3-50-model1-lc1-num-all-fail-med}{6}
\DefMacro{test-results-3-50-model1-lc1-num-all-fail-std}{7.655}
\DefMacro{test-results-3-50-model1-lc1-num-all-failrate-avg}{2.81}
\DefMacro{test-results-3-50-model1-lc1-num-all-failrate-med}{1.21}
\DefMacro{test-results-3-50-model1-lc1-num-all-failrate-std}{1.465}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-avg}{12.0}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-med}{3}
\DefMacro{test-results-3-50-model1-lc1-num-pass-to-fail-std}{8.23}
\DefMacro{test-results-3-50-model1-lc8-num-all-fail-avg}{81.0}
\DefMacro{test-results-3-50-model1-lc8-num-all-fail-med}{81}
\DefMacro{test-results-3-50-model1-lc8-num-all-fail-std}{14.5}
\DefMacro{test-results-3-50-model1-lc8-num-all-failrate-avg}{5.84}
\DefMacro{test-results-3-50-model1-lc8-num-all-failrate-med}{6.14}
\DefMacro{test-results-3-50-model1-lc8-num-all-failrate-std}{0.95}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-avg}{25.67}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-med}{27}
\DefMacro{test-results-3-50-model1-lc8-num-pass-to-fail-std}{5.53}
\DefMacro{test-results-3-50-model2-lc7-num-all-fail-avg}{490.67}
\DefMacro{test-results-3-50-model2-lc7-num-all-fail-med}{420}
\DefMacro{test-results-3-50-model2-lc7-num-all-fail-std}{98.38}
\DefMacro{test-results-3-50-model2-lc7-num-all-failrate-avg}{31.59}
\DefMacro{test-results-3-50-model2-lc7-num-all-failrate-med}{33.68}
\DefMacro{test-results-3-50-model2-lc7-num-all-failrate-std}{5.43}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-avg}{50.67}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-med}{49}
\DefMacro{test-results-3-50-model2-lc7-num-pass-to-fail-std}{15.765}
\DefMacro{test-results-3-50-model2-lc3-num-all-fail-avg}{1862.0}
\DefMacro{test-results-3-50-model2-lc3-num-all-fail-med}{1827}
\DefMacro{test-results-3-50-model2-lc3-num-all-fail-std}{96.94}
\DefMacro{test-results-3-50-model2-lc3-num-all-failrate-avg}{85.66}
\DefMacro{test-results-3-50-model2-lc3-num-all-failrate-med}{86.06}
\DefMacro{test-results-3-50-model2-lc3-num-all-failrate-std}{0.64}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-avg}{33.67}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-med}{34}
\DefMacro{test-results-3-50-model2-lc3-num-pass-to-fail-std}{2.255}
\DefMacro{test-results-3-50-model2-lc4-num-all-fail-avg}{1068.33}
\DefMacro{test-results-3-50-model2-lc4-num-all-fail-med}{1073}
\DefMacro{test-results-3-50-model2-lc4-num-all-fail-std}{8.735}
\DefMacro{test-results-3-50-model2-lc4-num-all-failrate-avg}{95.59}
\DefMacro{test-results-3-50-model2-lc4-num-all-failrate-med}{95.33}
\DefMacro{test-results-3-50-model2-lc4-num-all-failrate-std}{0.245}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-avg}{84.0}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-med}{83}
\DefMacro{test-results-3-50-model2-lc4-num-pass-to-fail-std}{2.29}
\DefMacro{test-results-3-50-model2-lc6-num-all-fail-avg}{1056.0}
\DefMacro{test-results-3-50-model2-lc6-num-all-fail-med}{972}
\DefMacro{test-results-3-50-model2-lc6-num-all-fail-std}{76.675}
\DefMacro{test-results-3-50-model2-lc6-num-all-failrate-avg}{90.95}
\DefMacro{test-results-3-50-model2-lc6-num-all-failrate-med}{92.95}
\DefMacro{test-results-3-50-model2-lc6-num-all-failrate-std}{4.125}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-avg}{4.33}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-med}{4}
\DefMacro{test-results-3-50-model2-lc6-num-pass-to-fail-std}{1.26}
\DefMacro{test-results-3-50-model2-lc5-num-all-fail-avg}{1500.0}
\DefMacro{test-results-3-50-model2-lc5-num-all-fail-med}{1382}
\DefMacro{test-results-3-50-model2-lc5-num-all-fail-std}{110.52}
\DefMacro{test-results-3-50-model2-lc5-num-all-failrate-avg}{98.45}
\DefMacro{test-results-3-50-model2-lc5-num-all-failrate-med}{99.07}
\DefMacro{test-results-3-50-model2-lc5-num-all-failrate-std}{0.97}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-avg}{1.33}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model2-lc5-num-pass-to-fail-std}{1.155}
\DefMacro{test-results-3-50-model2-lc9-num-all-fail-avg}{1235.67}
\DefMacro{test-results-3-50-model2-lc9-num-all-fail-med}{1112}
\DefMacro{test-results-3-50-model2-lc9-num-all-fail-std}{251.03}
\DefMacro{test-results-3-50-model2-lc9-num-all-failrate-avg}{78.61}
\DefMacro{test-results-3-50-model2-lc9-num-all-failrate-med}{82.37}
\DefMacro{test-results-3-50-model2-lc9-num-all-failrate-std}{10.985}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-avg}{13.0}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-med}{4}
\DefMacro{test-results-3-50-model2-lc9-num-pass-to-fail-std}{7.795}
\DefMacro{test-results-3-50-model2-lc2-num-all-fail-avg}{895.0}
\DefMacro{test-results-3-50-model2-lc2-num-all-fail-med}{953}
\DefMacro{test-results-3-50-model2-lc2-num-all-fail-std}{86.715}
\DefMacro{test-results-3-50-model2-lc2-num-all-failrate-avg}{55.93}
\DefMacro{test-results-3-50-model2-lc2-num-all-failrate-med}{60.58}
\DefMacro{test-results-3-50-model2-lc2-num-all-failrate-std}{4.545}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-avg}{49.33}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-med}{60}
\DefMacro{test-results-3-50-model2-lc2-num-pass-to-fail-std}{11.015}
\DefMacro{test-results-3-50-model2-lc0-num-all-fail-avg}{241.67}
\DefMacro{test-results-3-50-model2-lc0-num-all-fail-med}{233}
\DefMacro{test-results-3-50-model2-lc0-num-all-fail-std}{8.385}
\DefMacro{test-results-3-50-model2-lc0-num-all-failrate-avg}{96.12}
\DefMacro{test-results-3-50-model2-lc0-num-all-failrate-med}{96.28}
\DefMacro{test-results-3-50-model2-lc0-num-all-failrate-std}{0.495}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-avg}{0.0}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-med}{0}
\DefMacro{test-results-3-50-model2-lc0-num-pass-to-fail-std}{0.0}
\DefMacro{test-results-3-50-model2-lc1-num-all-fail-avg}{4.0}
\DefMacro{test-results-3-50-model2-lc1-num-all-fail-med}{5}
\DefMacro{test-results-3-50-model2-lc1-num-all-fail-std}{0.865}
\DefMacro{test-results-3-50-model2-lc1-num-all-failrate-avg}{0.82}
\DefMacro{test-results-3-50-model2-lc1-num-all-failrate-med}{0.86}
\DefMacro{test-results-3-50-model2-lc1-num-all-failrate-std}{0.205}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-avg}{4.0}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-med}{5}
\DefMacro{test-results-3-50-model2-lc1-num-pass-to-fail-std}{0.865}
\DefMacro{test-results-3-50-model2-lc8-num-all-fail-avg}{91.33}
\DefMacro{test-results-3-50-model2-lc8-num-all-fail-med}{82}
\DefMacro{test-results-3-50-model2-lc8-num-all-fail-std}{17.96}
\DefMacro{test-results-3-50-model2-lc8-num-all-failrate-avg}{6.7}
\DefMacro{test-results-3-50-model2-lc8-num-all-failrate-med}{5.99}
\DefMacro{test-results-3-50-model2-lc8-num-all-failrate-std}{1.465}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-avg}{25.33}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-med}{13}
\DefMacro{test-results-3-50-model2-lc8-num-pass-to-fail-std}{12.005}