### 2022-10-26 paper/icse23/tables/test-results-hs-all-numbers.tex
add hs table

%% Automatically generated by pyutil.latex

\DefMacro{test-results-hs-lc2}{Dehumanisation (explicit)}
\DefMacro{test-results-hs-lc2-num-seeds}{140}
\DefMacro{test-results-hs-lc2-num-exps}{3124}
\DefMacro{test-results-hs-model0-lc2-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc2-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc2-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc2-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc2-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc1}{Description using very negative attributes (explicit)}
\DefMacro{test-results-hs-lc1-num-seeds}{140}
\DefMacro{test-results-hs-lc1-num-exps}{2959}
\DefMacro{test-results-hs-model0-lc1-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc1-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc1-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc1-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc1-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc4}{Direct threat}
\DefMacro{test-results-hs-lc4-num-seeds}{133}
\DefMacro{test-results-hs-lc4-num-exps}{2689}
\DefMacro{test-results-hs-model0-lc4-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc4-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc4-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc4-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc0}{Expression of strong negative emotions (explicit)}
\DefMacro{test-results-hs-lc0-num-seeds}{140}
\DefMacro{test-results-hs-lc0-num-exps}{799}
\DefMacro{test-results-hs-model0-lc0-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc0-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc0-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc0-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc0-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc8}{Hate expressed using profanity}
\DefMacro{test-results-hs-lc8-num-seeds}{1842}
\DefMacro{test-results-hs-lc8-num-exps}{49743}
\DefMacro{test-results-hs-model0-lc8-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc8-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc8-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc8-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc8-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc6}{Hate expressed using slur}
\DefMacro{test-results-hs-lc6-num-seeds}{805}
\DefMacro{test-results-hs-lc6-num-exps}{17318}
\DefMacro{test-results-hs-model0-lc6-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc6-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc6-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc6-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc6-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc3}{Implicit derogation}
\DefMacro{test-results-hs-lc3-num-seeds}{140}
\DefMacro{test-results-hs-lc3-num-exps}{5664}
\DefMacro{test-results-hs-model0-lc3-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc3-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc3-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc3-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc3-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc9}{Non-Hateful use of profanity}
\DefMacro{test-results-hs-lc9-num-seeds}{701}
\DefMacro{test-results-hs-lc9-num-exps}{15761}
\DefMacro{test-results-hs-model0-lc9-num-all-fail}{16,462}
\DefMacro{test-results-hs-model0-lc9-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc9-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc9-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc9-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc7}{Non-hateful use of slur}
\DefMacro{test-results-hs-lc7-num-seeds}{395}
\DefMacro{test-results-hs-lc7-num-exps}{7881}
\DefMacro{test-results-hs-model0-lc7-num-all-fail}{8,276}
\DefMacro{test-results-hs-model0-lc7-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc7-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc7-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc7-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc5}{Threat as normative statement}
\DefMacro{test-results-hs-lc5-num-seeds}{140}
\DefMacro{test-results-hs-lc5-num-exps}{4163}
\DefMacro{test-results-hs-model0-lc5-num-all-fail}{0}
\DefMacro{test-results-hs-model0-lc5-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc5-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc5-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc5-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc2-num-all-fail}{3,210}
\DefMacro{test-results-hs-model1-lc2-num-seed-failrate}{98.57}
\DefMacro{test-results-hs-model1-lc2-num-exp-failrate}{98.34}
\DefMacro{test-results-hs-model1-lc2-num-all-failrate}{98.35}
\DefMacro{test-results-hs-model1-lc2-num-pass-to-fail}{11}
\DefMacro{test-results-hs-model1-lc1-num-all-fail}{3,011}
\DefMacro{test-results-hs-model1-lc1-num-seed-failrate}{97.86}
\DefMacro{test-results-hs-model1-lc1-num-exp-failrate}{97.13}
\DefMacro{test-results-hs-model1-lc1-num-all-failrate}{97.16}
\DefMacro{test-results-hs-model1-lc1-num-pass-to-fail}{30}
\DefMacro{test-results-hs-model1-lc4-num-all-fail}{2,750}
\DefMacro{test-results-hs-model1-lc4-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc4-num-exp-failrate}{97.32}
\DefMacro{test-results-hs-model1-lc4-num-all-failrate}{97.45}
\DefMacro{test-results-hs-model1-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc0-num-all-fail}{845}
\DefMacro{test-results-hs-model1-lc0-num-seed-failrate}{95.71}
\DefMacro{test-results-hs-model1-lc0-num-exp-failrate}{88.99}
\DefMacro{test-results-hs-model1-lc0-num-all-failrate}{89.99}
\DefMacro{test-results-hs-model1-lc0-num-pass-to-fail}{19}
\DefMacro{test-results-hs-model1-lc8-num-all-fail}{49,739}
\DefMacro{test-results-hs-model1-lc8-num-seed-failrate}{96.15}
\DefMacro{test-results-hs-model1-lc8-num-exp-failrate}{96.43}
\DefMacro{test-results-hs-model1-lc8-num-all-failrate}{96.42}
\DefMacro{test-results-hs-model1-lc8-num-pass-to-fail}{188}
\DefMacro{test-results-hs-model1-lc6-num-all-fail}{17,212}
\DefMacro{test-results-hs-model1-lc6-num-seed-failrate}{96.40}
\DefMacro{test-results-hs-model1-lc6-num-exp-failrate}{94.91}
\DefMacro{test-results-hs-model1-lc6-num-all-failrate}{94.97}
\DefMacro{test-results-hs-model1-lc6-num-pass-to-fail}{81}
\DefMacro{test-results-hs-model1-lc3-num-all-fail}{5,526}
\DefMacro{test-results-hs-model1-lc3-num-seed-failrate}{96.43}
\DefMacro{test-results-hs-model1-lc3-num-exp-failrate}{95.18}
\DefMacro{test-results-hs-model1-lc3-num-all-failrate}{95.21}
\DefMacro{test-results-hs-model1-lc3-num-pass-to-fail}{56}
\DefMacro{test-results-hs-model1-lc9-num-all-fail}{541}
\DefMacro{test-results-hs-model1-lc9-num-seed-failrate}{3.28}
\DefMacro{test-results-hs-model1-lc9-num-exp-failrate}{3.29}
\DefMacro{test-results-hs-model1-lc9-num-all-failrate}{3.29}
\DefMacro{test-results-hs-model1-lc9-num-pass-to-fail}{51}
\DefMacro{test-results-hs-model1-lc7-num-all-fail}{318}
\DefMacro{test-results-hs-model1-lc7-num-seed-failrate}{2.03}
\DefMacro{test-results-hs-model1-lc7-num-exp-failrate}{3.93}
\DefMacro{test-results-hs-model1-lc7-num-all-failrate}{3.84}
\DefMacro{test-results-hs-model1-lc7-num-pass-to-fail}{20}
\DefMacro{test-results-hs-model1-lc5-num-all-fail}{4,198}
\DefMacro{test-results-hs-model1-lc5-num-seed-failrate}{97.14}
\DefMacro{test-results-hs-model1-lc5-num-exp-failrate}{97.57}
\DefMacro{test-results-hs-model1-lc5-num-all-failrate}{97.56}
\DefMacro{test-results-hs-model1-lc5-num-pass-to-fail}{12}
\DefMacro{test-results-hs-model2-lc2-num-all-fail}{3,220}
\DefMacro{test-results-hs-model2-lc2-num-seed-failrate}{98.57}
\DefMacro{test-results-hs-model2-lc2-num-exp-failrate}{98.66}
\DefMacro{test-results-hs-model2-lc2-num-all-failrate}{98.65}
\DefMacro{test-results-hs-model2-lc2-num-pass-to-fail}{18}
\DefMacro{test-results-hs-model2-lc1-num-all-fail}{3,067}
\DefMacro{test-results-hs-model2-lc1-num-seed-failrate}{98.57}
\DefMacro{test-results-hs-model2-lc1-num-exp-failrate}{98.99}
\DefMacro{test-results-hs-model2-lc1-num-all-failrate}{98.97}
\DefMacro{test-results-hs-model2-lc1-num-pass-to-fail}{23}
\DefMacro{test-results-hs-model2-lc4-num-all-fail}{2,770}
\DefMacro{test-results-hs-model2-lc4-num-seed-failrate}{97.74}
\DefMacro{test-results-hs-model2-lc4-num-exp-failrate}{98.18}
\DefMacro{test-results-hs-model2-lc4-num-all-failrate}{98.16}
\DefMacro{test-results-hs-model2-lc4-num-pass-to-fail}{9}
\DefMacro{test-results-hs-model2-lc0-num-all-fail}{920}
\DefMacro{test-results-hs-model2-lc0-num-seed-failrate}{97.14}
\DefMacro{test-results-hs-model2-lc0-num-exp-failrate}{98.12}
\DefMacro{test-results-hs-model2-lc0-num-all-failrate}{97.98}
\DefMacro{test-results-hs-model2-lc0-num-pass-to-fail}{22}
\DefMacro{test-results-hs-model2-lc8-num-all-fail}{50,158}
\DefMacro{test-results-hs-model2-lc8-num-seed-failrate}{97.34}
\DefMacro{test-results-hs-model2-lc8-num-exp-failrate}{97.23}
\DefMacro{test-results-hs-model2-lc8-num-all-failrate}{97.23}
\DefMacro{test-results-hs-model2-lc8-num-pass-to-fail}{164}
\DefMacro{test-results-hs-model2-lc6-num-all-fail}{17,597}
\DefMacro{test-results-hs-model2-lc6-num-seed-failrate}{97.14}
\DefMacro{test-results-hs-model2-lc6-num-exp-failrate}{97.10}
\DefMacro{test-results-hs-model2-lc6-num-all-failrate}{97.10}
\DefMacro{test-results-hs-model2-lc6-num-pass-to-fail}{70}
\DefMacro{test-results-hs-model2-lc3-num-all-fail}{5,730}
\DefMacro{test-results-hs-model2-lc3-num-seed-failrate}{97.86}
\DefMacro{test-results-hs-model2-lc3-num-exp-failrate}{98.75}
\DefMacro{test-results-hs-model2-lc3-num-all-failrate}{98.73}
\DefMacro{test-results-hs-model2-lc3-num-pass-to-fail}{30}
\DefMacro{test-results-hs-model2-lc9-num-all-fail}{549}
\DefMacro{test-results-hs-model2-lc9-num-seed-failrate}{3.14}
\DefMacro{test-results-hs-model2-lc9-num-exp-failrate}{3.34}
\DefMacro{test-results-hs-model2-lc9-num-all-failrate}{3.33}
\DefMacro{test-results-hs-model2-lc9-num-pass-to-fail}{58}
\DefMacro{test-results-hs-model2-lc7-num-all-fail}{222}
\DefMacro{test-results-hs-model2-lc7-num-seed-failrate}{2.28}
\DefMacro{test-results-hs-model2-lc7-num-exp-failrate}{2.70}
\DefMacro{test-results-hs-model2-lc7-num-all-failrate}{2.68}
\DefMacro{test-results-hs-model2-lc7-num-pass-to-fail}{26}
\DefMacro{test-results-hs-model2-lc5-num-all-fail}{4,261}
\DefMacro{test-results-hs-model2-lc5-num-seed-failrate}{99.29}
\DefMacro{test-results-hs-model2-lc5-num-exp-failrate}{99.02}
\DefMacro{test-results-hs-model2-lc5-num-all-failrate}{99.02}
\DefMacro{test-results-hs-model2-lc5-num-pass-to-fail}{1}