### 2022-10-26 paper/icse23/tables/test-results-hs-bl-table.tex

%% Automatically generated by pyutil.latex

\begin{table*}[htbp]
\begin{small}
\begin{center}
\caption{\TestResultsHsBlTableCaption}
\resizebox{0.75\textwidth}{!}{
\begin{tabular}{p{8cm}||cclll}
\toprule
\tLc & \parbox{1cm}{\tNumSeeds} & \parbox{1.5cm}{\centering\tNumFail} & \parbox{1.5cm}{\centering\tFailRate}\\
\midrule
\multirow{3}{*}{\parbox{8cm}{LC1: \UseMacro{test-results-hs-bl-lc0}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc0-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc0-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc0-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc0-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc0-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc0-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc0-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC2: \UseMacro{test-results-hs-bl-lc1}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc1-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc1-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc1-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc1-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc1-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc1-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc1-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC3: \UseMacro{test-results-hs-bl-lc2}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc2-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc2-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc2-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc2-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc2-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc2-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc2-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC4: \UseMacro{test-results-hs-bl-lc3}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc3-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc3-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc3-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc3-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc3-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc3-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc3-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC5: \UseMacro{test-results-hs-bl-lc4}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc4-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc4-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc4-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc4-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc4-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc4-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc4-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC6: \UseMacro{test-results-hs-bl-lc5}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc5-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc5-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc5-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc5-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc5-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc5-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc5-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC7: \UseMacro{test-results-hs-bl-lc6}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc6-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc6-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc6-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc6-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc6-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc6-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc6-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC8: \UseMacro{test-results-hs-bl-lc7}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc7-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc7-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc7-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc7-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc7-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc7-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc7-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC9: \UseMacro{test-results-hs-bl-lc8}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc8-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc8-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc8-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc8-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc8-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc8-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc8-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC10: \UseMacro{test-results-hs-bl-lc9}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc9-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc9-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc9-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc9-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc9-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc9-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc9-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC11: \UseMacro{test-results-hs-bl-lc10}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc10-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc10-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc10-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc10-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc10-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc10-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc10-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC12: \UseMacro{test-results-hs-bl-lc11}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc11-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc11-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc11-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc11-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc11-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc11-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc11-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC13: \UseMacro{test-results-hs-bl-lc12}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc12-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc12-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc12-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc12-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc12-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc12-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc12-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC14: \UseMacro{test-results-hs-bl-lc13}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc13-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc13-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc13-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc13-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc13-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc13-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc13-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC15: \UseMacro{test-results-hs-bl-lc14}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc14-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc14-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc14-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc14-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc14-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc14-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc14-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC16: \UseMacro{test-results-hs-bl-lc15}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc15-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc15-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc15-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc15-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc15-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc15-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc15-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC17: \UseMacro{test-results-hs-bl-lc16}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc16-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc16-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc16-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc16-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc16-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc16-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc16-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC18: \UseMacro{test-results-hs-bl-lc17}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc17-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc17-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc17-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc17-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc17-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc17-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc17-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC19: \UseMacro{test-results-hs-bl-lc18}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc18-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc18-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc18-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc18-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc18-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc18-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc18-num-failrate}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC20: \UseMacro{test-results-hs-bl-lc19}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-hs-bl-lc19-num-tcs}}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc19-num-fail}
 & BERT$\colon$\UseMacro{test-results-hs-bl-model0-lc19-num-failrate}\\
 & & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc19-num-fail}
 & daELECTRA$\colon$\UseMacro{test-results-hs-bl-model1-lc19-num-failrate}\\
 & & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc19-num-fail}
 & daBERT$\colon$\UseMacro{test-results-hs-bl-model2-lc19-num-failrate}\\
\hline
\bottomrule
\end{tabular}}
\end{center}
\end{small}
\vspace{\TestResultsTableVSpace}
\end{table*}