### 2022-10-26 paper/icse23/tables/test-results-all-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-all-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-all-lc3-num-seeds}{62}
\DefMacro{test-results-all-lc3-num-exps}{503}
\DefMacro{test-results-all-model0-lc3-num-fail}{518}
\DefMacro{test-results-all-model0-lc3-num-failrate}{91.68}
\DefMacro{test-results-all-model0-lc3-num-pass-to-fail}{20}
\DefMacro{test-results-all-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-all-lc4-num-seeds}{24}
\DefMacro{test-results-all-lc4-num-exps}{194}
\DefMacro{test-results-all-model0-lc4-num-fail}{205}
\DefMacro{test-results-all-model0-lc4-num-failrate}{94.04}
\DefMacro{test-results-all-model0-lc4-num-pass-to-fail}{11}
\DefMacro{test-results-all-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-all-lc6-num-seeds}{24,328}
\DefMacro{test-results-all-lc6-num-exps}{190,723}
\DefMacro{test-results-all-model0-lc6-num-fail}{195,525}
\DefMacro{test-results-all-model0-lc6-num-failrate}{90.92}
\DefMacro{test-results-all-model0-lc6-num-pass-to-fail}{2,014}
\DefMacro{test-results-all-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-all-lc5-num-seeds}{15,031}
\DefMacro{test-results-all-lc5-num-exps}{102,273}
\DefMacro{test-results-all-model0-lc5-num-fail}{116,894}
\DefMacro{test-results-all-model0-lc5-num-failrate}{99.65}
\DefMacro{test-results-all-model0-lc5-num-pass-to-fail}{71}
\DefMacro{test-results-all-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-all-lc0-num-seeds}{15}
\DefMacro{test-results-all-lc0-num-exps}{51}
\DefMacro{test-results-all-model0-lc0-num-fail}{57}
\DefMacro{test-results-all-model0-lc0-num-failrate}{86.36}
\DefMacro{test-results-all-model0-lc0-num-pass-to-fail}{9}
\DefMacro{test-results-all-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-all-lc1-num-seeds}{83}
\DefMacro{test-results-all-lc1-num-exps}{258}
\DefMacro{test-results-all-model0-lc1-num-fail}{25}
\DefMacro{test-results-all-model0-lc1-num-failrate}{7.33}
\DefMacro{test-results-all-model0-lc1-num-pass-to-fail}{5}
\DefMacro{test-results-all-model1-lc3-num-fail}{493}
\DefMacro{test-results-all-model1-lc3-num-failrate}{87.26}
\DefMacro{test-results-all-model1-lc3-num-pass-to-fail}{9}
\DefMacro{test-results-all-model1-lc4-num-fail}{202}
\DefMacro{test-results-all-model1-lc4-num-failrate}{92.66}
\DefMacro{test-results-all-model1-lc4-num-pass-to-fail}{6}
\DefMacro{test-results-all-model1-lc6-num-fail}{157,635}
\DefMacro{test-results-all-model1-lc6-num-failrate}{73.30}
\DefMacro{test-results-all-model1-lc6-num-pass-to-fail}{7,289}
\DefMacro{test-results-all-model1-lc5-num-fail}{116,514}
\DefMacro{test-results-all-model1-lc5-num-failrate}{99.33}
\DefMacro{test-results-all-model1-lc5-num-pass-to-fail}{93}
\DefMacro{test-results-all-model1-lc0-num-fail}{51}
\DefMacro{test-results-all-model1-lc0-num-failrate}{77.27}
\DefMacro{test-results-all-model1-lc0-num-pass-to-fail}{2}
\DefMacro{test-results-all-model1-lc1-num-fail}{34}
\DefMacro{test-results-all-model1-lc1-num-failrate}{9.97}
\DefMacro{test-results-all-model1-lc1-num-pass-to-fail}{14}
\DefMacro{test-results-all-model2-lc3-num-fail}{489}
\DefMacro{test-results-all-model2-lc3-num-failrate}{86.55}
\DefMacro{test-results-all-model2-lc3-num-pass-to-fail}{6}
\DefMacro{test-results-all-model2-lc4-num-fail}{211}
\DefMacro{test-results-all-model2-lc4-num-failrate}{96.79}
\DefMacro{test-results-all-model2-lc4-num-pass-to-fail}{10}
\DefMacro{test-results-all-model2-lc6-num-fail}{180,865}
\DefMacro{test-results-all-model2-lc6-num-failrate}{84.10}
\DefMacro{test-results-all-model2-lc6-num-pass-to-fail}{5,130}
\DefMacro{test-results-all-model2-lc5-num-fail}{115,407}
\DefMacro{test-results-all-model2-lc5-num-failrate}{98.38}
\DefMacro{test-results-all-model2-lc5-num-pass-to-fail}{334}
\DefMacro{test-results-all-model2-lc0-num-fail}{64}
\DefMacro{test-results-all-model2-lc0-num-failrate}{96.97}
\DefMacro{test-results-all-model2-lc0-num-pass-to-fail}{0}
\DefMacro{test-results-all-model2-lc1-num-fail}{17}
\DefMacro{test-results-all-model2-lc1-num-failrate}{4.99}
\DefMacro{test-results-all-model2-lc1-num-pass-to-fail}{10}