### 2022-11-18 paper/icse23/tables/test-results-bl-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-bl-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-bl-lc7-num-seeds}{8,528}
\DefMacro{test-results-bl-model0-lc7-num-seed-fail}{3,741}
\DefMacro{test-results-bl-model0-lc7-num-seed-failrate}{43.87}
\DefMacro{test-results-bl-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-bl-lc3-num-seeds}{6,786}
\DefMacro{test-results-bl-model0-lc3-num-seed-fail}{799}
\DefMacro{test-results-bl-model0-lc3-num-seed-failrate}{11.77}
\DefMacro{test-results-bl-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-bl-lc4-num-seeds}{2,496}
\DefMacro{test-results-bl-model0-lc4-num-seed-fail}{2,427}
\DefMacro{test-results-bl-model0-lc4-num-seed-failrate}{97.24}
\DefMacro{test-results-bl-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-bl-lc6-num-seeds}{1,000}
\DefMacro{test-results-bl-model0-lc6-num-seed-fail}{860}
\DefMacro{test-results-bl-model0-lc6-num-seed-failrate}{86.00}
\DefMacro{test-results-bl-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-bl-lc5-num-seeds}{2,124}
\DefMacro{test-results-bl-model0-lc5-num-seed-fail}{1,871}
\DefMacro{test-results-bl-model0-lc5-num-seed-failrate}{88.09}
\DefMacro{test-results-bl-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-bl-lc2-num-seeds}{8,000}
\DefMacro{test-results-bl-model0-lc2-num-seed-fail}{1,680}
\DefMacro{test-results-bl-model0-lc2-num-seed-failrate}{21.00}
\DefMacro{test-results-bl-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-bl-lc0-num-seeds}{1,716}
\DefMacro{test-results-bl-model0-lc0-num-seed-fail}{1,330}
\DefMacro{test-results-bl-model0-lc0-num-seed-failrate}{77.51}
\DefMacro{test-results-bl-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-bl-lc1-num-seeds}{8,658}
\DefMacro{test-results-bl-model0-lc1-num-seed-fail}{26}
\DefMacro{test-results-bl-model0-lc1-num-seed-failrate}{0.30}
\DefMacro{test-results-bl-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-bl-lc9-num-seeds}{7,644}
\DefMacro{test-results-bl-model0-lc9-num-seed-fail}{4,056}
\DefMacro{test-results-bl-model0-lc9-num-seed-failrate}{53.06}
\DefMacro{test-results-bl-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-bl-lc8-num-seeds}{9,204}
\DefMacro{test-results-bl-model0-lc8-num-seed-fail}{1,793}
\DefMacro{test-results-bl-model0-lc8-num-seed-failrate}{19.48}
\DefMacro{test-results-bl-model1-lc7-num-seed-fail}{2,692}
\DefMacro{test-results-bl-model1-lc7-num-seed-failrate}{31.57}
\DefMacro{test-results-bl-model1-lc3-num-seed-fail}{218}
\DefMacro{test-results-bl-model1-lc3-num-seed-failrate}{3.21}
\DefMacro{test-results-bl-model1-lc4-num-seed-fail}{2,304}
\DefMacro{test-results-bl-model1-lc4-num-seed-failrate}{92.31}
\DefMacro{test-results-bl-model1-lc6-num-seed-fail}{416}
\DefMacro{test-results-bl-model1-lc6-num-seed-failrate}{41.60}
\DefMacro{test-results-bl-model1-lc5-num-seed-fail}{445}
\DefMacro{test-results-bl-model1-lc5-num-seed-failrate}{20.95}
\DefMacro{test-results-bl-model1-lc2-num-seed-fail}{829}
\DefMacro{test-results-bl-model1-lc2-num-seed-failrate}{10.36}
\DefMacro{test-results-bl-model1-lc0-num-seed-fail}{1,391}
\DefMacro{test-results-bl-model1-lc0-num-seed-failrate}{81.06}
\DefMacro{test-results-bl-model1-lc1-num-seed-fail}{139}
\DefMacro{test-results-bl-model1-lc1-num-seed-failrate}{1.61}
\DefMacro{test-results-bl-model1-lc9-num-seed-fail}{4,576}
\DefMacro{test-results-bl-model1-lc9-num-seed-failrate}{59.86}
\DefMacro{test-results-bl-model1-lc8-num-seed-fail}{1,262}
\DefMacro{test-results-bl-model1-lc8-num-seed-failrate}{13.71}
\DefMacro{test-results-bl-model2-lc7-num-seed-fail}{3,535}
\DefMacro{test-results-bl-model2-lc7-num-seed-failrate}{41.45}
\DefMacro{test-results-bl-model2-lc3-num-seed-fail}{734}
\DefMacro{test-results-bl-model2-lc3-num-seed-failrate}{10.82}
\DefMacro{test-results-bl-model2-lc4-num-seed-fail}{2,450}
\DefMacro{test-results-bl-model2-lc4-num-seed-failrate}{98.16}
\DefMacro{test-results-bl-model2-lc6-num-seed-fail}{865}
\DefMacro{test-results-bl-model2-lc6-num-seed-failrate}{86.50}
\DefMacro{test-results-bl-model2-lc5-num-seed-fail}{2,124}
\DefMacro{test-results-bl-model2-lc5-num-seed-failrate}{100.00}
\DefMacro{test-results-bl-model2-lc2-num-seed-fail}{2,532}
\DefMacro{test-results-bl-model2-lc2-num-seed-failrate}{31.65}
\DefMacro{test-results-bl-model2-lc0-num-seed-fail}{1,661}
\DefMacro{test-results-bl-model2-lc0-num-seed-failrate}{96.79}
\DefMacro{test-results-bl-model2-lc1-num-seed-fail}{125}
\DefMacro{test-results-bl-model2-lc1-num-seed-failrate}{1.44}
\DefMacro{test-results-bl-model2-lc9-num-seed-fail}{6,440}
\DefMacro{test-results-bl-model2-lc9-num-seed-failrate}{84.25}
\DefMacro{test-results-bl-model2-lc8-num-seed-fail}{1,557}
\DefMacro{test-results-bl-model2-lc8-num-seed-failrate}{16.92}