### 2023-01-09 paper/icse23/tables/test-results-hs-bl-all-numbers.tex
debug

%% Automatically generated by pyutil.latex

\DefMacro{test-results-hs-bl-lc19}{Counter speech::Denouncements of hate that make direct reference to it}
\DefMacro{test-results-hs-bl-lc19-num-tcs}{141}
\DefMacro{test-results-hs-bl-model0-lc19-num-fail}{20}
\DefMacro{test-results-hs-bl-model0-lc19-num-failrate}{14.18}
\DefMacro{test-results-hs-bl-lc18}{Counter speech::Denouncements of hate that quote it}
\DefMacro{test-results-hs-bl-lc18-num-tcs}{173}
\DefMacro{test-results-hs-bl-model0-lc18-num-fail}{18}
\DefMacro{test-results-hs-bl-model0-lc18-num-failrate}{10.40}
\DefMacro{test-results-hs-bl-lc2}{Derogation::Dehumanisation (explicit)}
\DefMacro{test-results-hs-bl-lc2-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc2-num-fail}{138}
\DefMacro{test-results-hs-bl-model0-lc2-num-failrate}{98.57}
\DefMacro{test-results-hs-bl-lc1}{Derogation::Description using very negative attributes (explicit)}
\DefMacro{test-results-hs-bl-lc1-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc1-num-fail}{138}
\DefMacro{test-results-hs-bl-model0-lc1-num-failrate}{98.57}
\DefMacro{test-results-hs-bl-lc0}{Derogation::Expression of strong negative emotions (explicit)}
\DefMacro{test-results-hs-bl-lc0-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc0-num-fail}{136}
\DefMacro{test-results-hs-bl-model0-lc0-num-failrate}{97.14}
\DefMacro{test-results-hs-bl-lc3}{Derogation::Implicit derogation}
\DefMacro{test-results-hs-bl-lc3-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc3-num-fail}{137}
\DefMacro{test-results-hs-bl-model0-lc3-num-failrate}{97.86}
\DefMacro{test-results-hs-bl-lc12}{Negation::Hate expressed using negated positive statement}
\DefMacro{test-results-hs-bl-lc12-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc12-num-fail}{138}
\DefMacro{test-results-hs-bl-model0-lc12-num-failrate}{98.57}
\DefMacro{test-results-hs-bl-lc13}{Negation::Non-hate expressed using negated hateful statement}
\DefMacro{test-results-hs-bl-lc13-num-tcs}{133}
\DefMacro{test-results-hs-bl-model0-lc13-num-fail}{9}
\DefMacro{test-results-hs-bl-model0-lc13-num-failrate}{6.77}
\DefMacro{test-results-hs-bl-lc16}{Non-hate grp. ident.::Neutral statements using protected group identifiers}
\DefMacro{test-results-hs-bl-lc16-num-tcs}{126}
\DefMacro{test-results-hs-bl-model0-lc16-num-fail}{8}
\DefMacro{test-results-hs-bl-model0-lc16-num-failrate}{6.35}
\DefMacro{test-results-hs-bl-lc17}{Non-hate grp. ident.::Positive statements using protected group identifiers}
\DefMacro{test-results-hs-bl-lc17-num-tcs}{189}
\DefMacro{test-results-hs-bl-model0-lc17-num-fail}{10}
\DefMacro{test-results-hs-bl-model0-lc17-num-failrate}{5.29}
\DefMacro{test-results-hs-bl-lc15}{Phrasing::Hate phrased as a opinion}
\DefMacro{test-results-hs-bl-lc15-num-tcs}{133}
\DefMacro{test-results-hs-bl-model0-lc15-num-fail}{131}
\DefMacro{test-results-hs-bl-model0-lc15-num-failrate}{98.50}
\DefMacro{test-results-hs-bl-lc14}{Phrasing::Hate phrased as a question}
\DefMacro{test-results-hs-bl-lc14-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc14-num-fail}{139}
\DefMacro{test-results-hs-bl-model0-lc14-num-failrate}{99.29}
\DefMacro{test-results-hs-bl-lc8}{Profanity usage::Hate expressed using profanity}
\DefMacro{test-results-hs-bl-lc8-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc8-num-fail}{138}
\DefMacro{test-results-hs-bl-model0-lc8-num-failrate}{98.57}
\DefMacro{test-results-hs-bl-lc9}{Profanity usage::Non-Hateful use of profanity}
\DefMacro{test-results-hs-bl-lc9-num-tcs}{100}
\DefMacro{test-results-hs-bl-model0-lc9-num-fail}{30}
\DefMacro{test-results-hs-bl-model0-lc9-num-failrate}{30.00}
\DefMacro{test-results-hs-bl-lc10}{Pronoun reference::Hate expressed through reference in subsequent clauses}
\DefMacro{test-results-hs-bl-lc10-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc10-num-fail}{138}
\DefMacro{test-results-hs-bl-model0-lc10-num-failrate}{98.57}
\DefMacro{test-results-hs-bl-lc11}{Pronoun reference::Hate expressed through reference in subsequent sentences}
\DefMacro{test-results-hs-bl-lc11-num-tcs}{133}
\DefMacro{test-results-hs-bl-model0-lc11-num-fail}{133}
\DefMacro{test-results-hs-bl-model0-lc11-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-lc6}{Slur usage::Hate expressed using slur}
\DefMacro{test-results-hs-bl-lc6-num-tcs}{144}
\DefMacro{test-results-hs-bl-model0-lc6-num-fail}{144}
\DefMacro{test-results-hs-bl-model0-lc6-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-lc7}{Slur usage::Non-hateful use of slur}
\DefMacro{test-results-hs-bl-lc7-num-tcs}{111}
\DefMacro{test-results-hs-bl-model0-lc7-num-fail}{5}
\DefMacro{test-results-hs-bl-model0-lc7-num-failrate}{4.50}
\DefMacro{test-results-hs-bl-lc4}{Threatening language::Direct threat}
\DefMacro{test-results-hs-bl-lc4-num-tcs}{133}
\DefMacro{test-results-hs-bl-model0-lc4-num-fail}{133}
\DefMacro{test-results-hs-bl-model0-lc4-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-lc5}{Threatening language::Threat as normative statement}
\DefMacro{test-results-hs-bl-lc5-num-tcs}{140}
\DefMacro{test-results-hs-bl-model0-lc5-num-fail}{139}
\DefMacro{test-results-hs-bl-model0-lc5-num-failrate}{99.29}
\DefMacro{test-results-hs-bl-model1-lc19-num-fail}{141}
\DefMacro{test-results-hs-bl-model1-lc19-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc18-num-fail}{173}
\DefMacro{test-results-hs-bl-model1-lc18-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc2-num-fail}{1}
\DefMacro{test-results-hs-bl-model1-lc2-num-failrate}{0.71}
\DefMacro{test-results-hs-bl-model1-lc1-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc1-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc0-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc0-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc3-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc3-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc12-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc12-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc13-num-fail}{133}
\DefMacro{test-results-hs-bl-model1-lc13-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc16-num-fail}{126}
\DefMacro{test-results-hs-bl-model1-lc16-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc17-num-fail}{189}
\DefMacro{test-results-hs-bl-model1-lc17-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc15-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc15-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc14-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc14-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc8-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc8-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc9-num-fail}{100}
\DefMacro{test-results-hs-bl-model1-lc9-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc10-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc10-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc11-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc11-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc6-num-fail}{11}
\DefMacro{test-results-hs-bl-model1-lc6-num-failrate}{7.64}
\DefMacro{test-results-hs-bl-model1-lc7-num-fail}{111}
\DefMacro{test-results-hs-bl-model1-lc7-num-failrate}{100.00}
\DefMacro{test-results-hs-bl-model1-lc4-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc4-num-failrate}{0.00}
\DefMacro{test-results-hs-bl-model1-lc5-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc5-num-failrate}{0.00}