### 2023-01-10 paper/fse23/main.tex

\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
%% Use macros
\input{macros}

\begin{document}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\title{\tool{}: Specification- and Syntax-based Automated Testing Linguistic Capabilities of NLP Models}

%% \title{\tool{}: Specification- and Syntax-based Automated Testing Linguistic Capabilities of NLP Models}
\maketitle

%% Abstract
\input{abstract}

%% Contents
\input{intro}
\input{background}
\input{approach}
%\input{rq}
\input{experiment}
\input{result}
\input{threats}
\input{related}
\input{concl}
\clearpage

\bibliographystyle{IEEEtran}
\bibliography{bib}

%% Appendix
%\input{appendix}

%% Text of appendix \ldots

\end{document}