### 2023-01-10 paper/fse23/tables/test-results-hs-all-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-hs-lc2}{Dehumanisation (explicit)}
\DefMacro{test-results-hs-lc2-num-seeds}{140}
\DefMacro{test-results-hs-lc2-num-exps}{3,124}
\DefMacro{test-results-hs-model0-lc2-num-all-fail}{3,220}
\DefMacro{test-results-hs-model0-lc2-num-seed-failrate}{98.57}
\DefMacro{test-results-hs-model0-lc2-num-exp-failrate}{98.66}
\DefMacro{test-results-hs-model0-lc2-num-all-failrate}{98.65}
\DefMacro{test-results-hs-model0-lc2-num-pass-to-fail}{18}
\DefMacro{test-results-hs-lc18}{Denouncements of hate that quote it}
\DefMacro{test-results-hs-lc18-num-seeds}{45,616}
\DefMacro{test-results-hs-lc18-num-exps}{300,704}
\DefMacro{test-results-hs-model0-lc18-num-all-fail}{48,804}
\DefMacro{test-results-hs-model0-lc18-num-seed-failrate}{13.78}
\DefMacro{test-results-hs-model0-lc18-num-exp-failrate}{14.14}
\DefMacro{test-results-hs-model0-lc18-num-all-failrate}{14.09}
\DefMacro{test-results-hs-model0-lc18-num-pass-to-fail}{4,831}
\DefMacro{test-results-hs-lc1}{Description using very negative attributes (explicit)}
\DefMacro{test-results-hs-lc1-num-seeds}{140}
\DefMacro{test-results-hs-lc1-num-exps}{2,959}
\DefMacro{test-results-hs-model0-lc1-num-all-fail}{3,067}
\DefMacro{test-results-hs-model0-lc1-num-seed-failrate}{98.57}
\DefMacro{test-results-hs-model0-lc1-num-exp-failrate}{98.99}
\DefMacro{test-results-hs-model0-lc1-num-all-failrate}{98.97}
\DefMacro{test-results-hs-model0-lc1-num-pass-to-fail}{23}
\DefMacro{test-results-hs-lc4}{Direct threat}
\DefMacro{test-results-hs-lc4-num-seeds}{133}
\DefMacro{test-results-hs-lc4-num-exps}{2,689}
\DefMacro{test-results-hs-model0-lc4-num-all-fail}{2,770}
\DefMacro{test-results-hs-model0-lc4-num-seed-failrate}{97.74}
\DefMacro{test-results-hs-model0-lc4-num-exp-failrate}{98.18}
\DefMacro{test-results-hs-model0-lc4-num-all-failrate}{98.16}
\DefMacro{test-results-hs-model0-lc4-num-pass-to-fail}{9}
\DefMacro{test-results-hs-lc0}{Expression of strong negative emotions (explicit)}
\DefMacro{test-results-hs-lc0-num-seeds}{140}
\DefMacro{test-results-hs-lc0-num-exps}{799}
\DefMacro{test-results-hs-model0-lc0-num-all-fail}{920}
\DefMacro{test-results-hs-model0-lc0-num-seed-failrate}{97.14}
\DefMacro{test-results-hs-model0-lc0-num-exp-failrate}{98.12}
\DefMacro{test-results-hs-model0-lc0-num-all-failrate}{97.98}
\DefMacro{test-results-hs-model0-lc0-num-pass-to-fail}{22}
\DefMacro{test-results-hs-lc10}{Hate expressed through reference in subsequent clauses}
\DefMacro{test-results-hs-lc10-num-seeds}{11,968}
\DefMacro{test-results-hs-lc10-num-exps}{44,890}
\DefMacro{test-results-hs-model0-lc10-num-all-fail}{49,929}
\DefMacro{test-results-hs-model0-lc10-num-seed-failrate}{60.60}
\DefMacro{test-results-hs-model0-lc10-num-exp-failrate}{95.07}
\DefMacro{test-results-hs-model0-lc10-num-all-failrate}{87.81}
\DefMacro{test-results-hs-model0-lc10-num-pass-to-fail}{1,169}
\DefMacro{test-results-hs-lc11}{Hate expressed through reference in subsequent sentences}
\DefMacro{test-results-hs-lc11-num-seeds}{11,968}
\DefMacro{test-results-hs-lc11-num-exps}{42,840}
\DefMacro{test-results-hs-model0-lc11-num-all-fail}{47,294}
\DefMacro{test-results-hs-model0-lc11-num-seed-failrate}{58.92}
\DefMacro{test-results-hs-model0-lc11-num-exp-failrate}{93.94}
\DefMacro{test-results-hs-model0-lc11-num-all-failrate}{86.29}
\DefMacro{test-results-hs-model0-lc11-num-pass-to-fail}{1,726}
\DefMacro{test-results-hs-lc12}{Hate expressed using negated positive statement}
\DefMacro{test-results-hs-lc12-num-seeds}{23,379}
\DefMacro{test-results-hs-lc12-num-exps}{126,742}
\DefMacro{test-results-hs-model0-lc12-num-all-fail}{145,761}
\DefMacro{test-results-hs-model0-lc12-num-seed-failrate}{97.03}
\DefMacro{test-results-hs-model0-lc12-num-exp-failrate}{97.11}
\DefMacro{test-results-hs-model0-lc12-num-all-failrate}{97.10}
\DefMacro{test-results-hs-model0-lc12-num-pass-to-fail}{2,290}
\DefMacro{test-results-hs-lc8}{Hate expressed using profanity}
\DefMacro{test-results-hs-lc8-num-seeds}{1,842}
\DefMacro{test-results-hs-lc8-num-exps}{49,743}
\DefMacro{test-results-hs-model0-lc8-num-all-fail}{50,158}
\DefMacro{test-results-hs-model0-lc8-num-seed-failrate}{97.34}
\DefMacro{test-results-hs-model0-lc8-num-exp-failrate}{97.23}
\DefMacro{test-results-hs-model0-lc8-num-all-failrate}{97.23}
\DefMacro{test-results-hs-model0-lc8-num-pass-to-fail}{164}
\DefMacro{test-results-hs-lc6}{Hate expressed using slur}
\DefMacro{test-results-hs-lc6-num-seeds}{805}
\DefMacro{test-results-hs-lc6-num-exps}{17,318}
\DefMacro{test-results-hs-model0-lc6-num-all-fail}{17,597}
\DefMacro{test-results-hs-model0-lc6-num-seed-failrate}{97.14}
\DefMacro{test-results-hs-model0-lc6-num-exp-failrate}{97.10}
\DefMacro{test-results-hs-model0-lc6-num-all-failrate}{97.10}
\DefMacro{test-results-hs-model0-lc6-num-pass-to-fail}{70}
\DefMacro{test-results-hs-lc15}{Hate phrased as a opinion}
\DefMacro{test-results-hs-lc15-num-seeds}{22,808}
\DefMacro{test-results-hs-lc15-num-exps}{160,927}
\DefMacro{test-results-hs-model0-lc15-num-all-fail}{178,058}
\DefMacro{test-results-hs-model0-lc15-num-seed-failrate}{97.25}
\DefMacro{test-results-hs-model0-lc15-num-exp-failrate}{96.86}
\DefMacro{test-results-hs-model0-lc15-num-all-failrate}{96.91}
\DefMacro{test-results-hs-model0-lc15-num-pass-to-fail}{3,078}
\DefMacro{test-results-hs-lc14}{Hate phrased as a question}
\DefMacro{test-results-hs-lc14-num-seeds}{22,808}
\DefMacro{test-results-hs-lc14-num-exps}{185,267}
\DefMacro{test-results-hs-model0-lc14-num-all-fail}{202,023}
\DefMacro{test-results-hs-model0-lc14-num-seed-failrate}{97.37}
\DefMacro{test-results-hs-model0-lc14-num-exp-failrate}{97.06}
\DefMacro{test-results-hs-model0-lc14-num-all-failrate}{97.09}
\DefMacro{test-results-hs-model0-lc14-num-pass-to-fail}{2,956}
\DefMacro{test-results-hs-lc3}{Implicit derogation}
\DefMacro{test-results-hs-lc3-num-seeds}{140}
\DefMacro{test-results-hs-lc3-num-exps}{5,664}
\DefMacro{test-results-hs-model0-lc3-num-all-fail}{5,730}
\DefMacro{test-results-hs-model0-lc3-num-seed-failrate}{97.86}
\DefMacro{test-results-hs-model0-lc3-num-exp-failrate}{98.75}
\DefMacro{test-results-hs-model0-lc3-num-all-failrate}{98.73}
\DefMacro{test-results-hs-model0-lc3-num-pass-to-fail}{30}
\DefMacro{test-results-hs-lc16}{Neutral statements using protected group identifiers}
\DefMacro{test-results-hs-lc16-num-seeds}{5}
\DefMacro{test-results-hs-lc16-num-exps}{9}
\DefMacro{test-results-hs-model0-lc16-num-all-fail}{1}
\DefMacro{test-results-hs-model0-lc16-num-seed-failrate}{20.00}
\DefMacro{test-results-hs-model0-lc16-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model0-lc16-num-all-failrate}{7.14}
\DefMacro{test-results-hs-model0-lc16-num-pass-to-fail}{0}
\DefMacro{test-results-hs-lc9}{Non-Hateful use of profanity}
\DefMacro{test-results-hs-lc9-num-seeds}{701}
\DefMacro{test-results-hs-lc9-num-exps}{15,761}
\DefMacro{test-results-hs-model0-lc9-num-all-fail}{11,939}
\DefMacro{test-results-hs-model0-lc9-num-seed-failrate}{71.04}
\DefMacro{test-results-hs-model0-lc9-num-exp-failrate}{72.59}
\DefMacro{test-results-hs-model0-lc9-num-all-failrate}{72.52}
\DefMacro{test-results-hs-model0-lc9-num-pass-to-fail}{75}
\DefMacro{test-results-hs-lc13}{Non-hate expressed using negated hateful statement}
\DefMacro{test-results-hs-lc13-num-seeds}{34,212}
\DefMacro{test-results-hs-lc13-num-exps}{243,739}
\DefMacro{test-results-hs-model0-lc13-num-all-fail}{77,191}
\DefMacro{test-results-hs-model0-lc13-num-seed-failrate}{26.90}
\DefMacro{test-results-hs-model0-lc13-num-exp-failrate}{27.89}
\DefMacro{test-results-hs-model0-lc13-num-all-failrate}{27.77}
\DefMacro{test-results-hs-model0-lc13-num-pass-to-fail}{5,222}
\DefMacro{test-results-hs-lc7}{Non-hateful use of slur}
\DefMacro{test-results-hs-lc7-num-seeds}{395}
\DefMacro{test-results-hs-lc7-num-exps}{7,881}
\DefMacro{test-results-hs-model0-lc7-num-all-fail}{2,739}
\DefMacro{test-results-hs-model0-lc7-num-seed-failrate}{30.89}
\DefMacro{test-results-hs-model0-lc7-num-exp-failrate}{33.21}
\DefMacro{test-results-hs-model0-lc7-num-all-failrate}{33.10}
\DefMacro{test-results-hs-model0-lc7-num-pass-to-fail}{44}
\DefMacro{test-results-hs-lc17}{Positive statements using protected group identifiers}
\DefMacro{test-results-hs-lc17-num-seeds}{41}
\DefMacro{test-results-hs-lc17-num-exps}{182}
\DefMacro{test-results-hs-model0-lc17-num-all-fail}{14}
\DefMacro{test-results-hs-model0-lc17-num-seed-failrate}{4.88}
\DefMacro{test-results-hs-model0-lc17-num-exp-failrate}{6.59}
\DefMacro{test-results-hs-model0-lc17-num-all-failrate}{6.28}
\DefMacro{test-results-hs-model0-lc17-num-pass-to-fail}{1}
\DefMacro{test-results-hs-lc5}{Threat as normative statement}
\DefMacro{test-results-hs-lc5-num-seeds}{140}
\DefMacro{test-results-hs-lc5-num-exps}{4,163}
\DefMacro{test-results-hs-model0-lc5-num-all-fail}{4,261}
\DefMacro{test-results-hs-model0-lc5-num-seed-failrate}{99.29}
\DefMacro{test-results-hs-model0-lc5-num-exp-failrate}{99.02}
\DefMacro{test-results-hs-model0-lc5-num-all-failrate}{99.02}
\DefMacro{test-results-hs-model0-lc5-num-pass-to-fail}{1}
\DefMacro{test-results-hs-model1-lc2-num-all-fail}{87}
\DefMacro{test-results-hs-model1-lc2-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc2-num-exp-failrate}{2.78}
\DefMacro{test-results-hs-model1-lc2-num-all-failrate}{2.67}
\DefMacro{test-results-hs-model1-lc2-num-pass-to-fail}{31}
\DefMacro{test-results-hs-model1-lc18-num-all-fail}{346,320}
\DefMacro{test-results-hs-model1-lc18-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc18-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc18-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc18-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc1-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc1-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc1-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc1-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc1-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc4-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc4-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc4-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc4-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc4-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc0-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc0-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc0-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc0-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc0-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc10-num-all-fail}{5,984}
\DefMacro{test-results-hs-model1-lc10-num-seed-failrate}{50.00}
\DefMacro{test-results-hs-model1-lc10-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc10-num-all-failrate}{10.52}
\DefMacro{test-results-hs-model1-lc10-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc11-num-all-fail}{5,984}
\DefMacro{test-results-hs-model1-lc11-num-seed-failrate}{50.00}
\DefMacro{test-results-hs-model1-lc11-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc11-num-all-failrate}{10.92}
\DefMacro{test-results-hs-model1-lc11-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc12-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc12-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc12-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc12-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc12-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc8-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc8-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc8-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc8-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc8-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc6-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc6-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc6-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc6-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc6-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc15-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc15-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc15-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc15-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc15-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc14-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc14-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc14-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc14-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc14-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc3-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc3-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc3-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc3-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc3-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc16-num-all-fail}{14}
\DefMacro{test-results-hs-model1-lc16-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc16-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc16-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc16-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc9-num-all-fail}{16,462}
\DefMacro{test-results-hs-model1-lc9-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc9-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc9-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc9-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc13-num-all-fail}{277,951}
\DefMacro{test-results-hs-model1-lc13-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc13-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc13-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc13-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc7-num-all-fail}{8,276}
\DefMacro{test-results-hs-model1-lc7-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc7-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc7-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc7-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc17-num-all-fail}{223}
\DefMacro{test-results-hs-model1-lc17-num-seed-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc17-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc17-num-all-failrate}{100.00}
\DefMacro{test-results-hs-model1-lc17-num-pass-to-fail}{0}
\DefMacro{test-results-hs-model1-lc5-num-all-fail}{0}
\DefMacro{test-results-hs-model1-lc5-num-seed-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc5-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc5-num-all-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc5-num-pass-to-fail}{0}