### 2023-01-11 paper/ase22/tables/test-results-bl-hs-all-numbers.tex
related work

%% Automatically generated by pyutil.latex

\DefMacro{test-results-bl-lc19}{Counter speech::Denouncements of hate that make direct reference to it}
\DefMacro{test-results-bl-lc19-num-tcs}{141}
\DefMacro{test-results-bl-model0-lc19-num-fail}{[141]}
\DefMacro{test-results-bl-model0-lc19-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc18}{Counter speech::Denouncements of hate that quote it}
\DefMacro{test-results-bl-lc18-num-tcs}{173}
\DefMacro{test-results-bl-model0-lc18-num-fail}{[173]}
\DefMacro{test-results-bl-model0-lc18-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc2}{Derogation::Dehumanisation (explicit)}
\DefMacro{test-results-bl-lc2-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc2-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc2-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc1}{Derogation::Description using very negative attributes (explicit)}
\DefMacro{test-results-bl-lc1-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc1-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc1-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc0}{Derogation::Expression of strong negative emotions (explicit)}
\DefMacro{test-results-bl-lc0-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc0-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc0-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc3}{Derogation::Implicit derogation}
\DefMacro{test-results-bl-lc3-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc3-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc3-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc12}{Negation::Hate expressed using negated positive statement}
\DefMacro{test-results-bl-lc12-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc12-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc12-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc13}{Negation::Non-hate expressed using negated hateful statement}
\DefMacro{test-results-bl-lc13-num-tcs}{133}
\DefMacro{test-results-bl-model0-lc13-num-fail}{[133]}
\DefMacro{test-results-bl-model0-lc13-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc16}{Non-hate grp. ident.::Neutral statements using protected group identifiers}
\DefMacro{test-results-bl-lc16-num-tcs}{126}
\DefMacro{test-results-bl-model0-lc16-num-fail}{[126]}
\DefMacro{test-results-bl-model0-lc16-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc17}{Non-hate grp. ident.::Positive statements using protected group identifiers}
\DefMacro{test-results-bl-lc17-num-tcs}{189}
\DefMacro{test-results-bl-model0-lc17-num-fail}{[189]}
\DefMacro{test-results-bl-model0-lc17-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc14}{Phrasing::Hate phrased as a question}
\DefMacro{test-results-bl-lc14-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc14-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc14-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc8}{Profanity usage::Hate expressed using profanity}
\DefMacro{test-results-bl-lc8-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc8-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc8-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc9}{Profanity usage::Non-Hateful use of profanity}
\DefMacro{test-results-bl-lc9-num-tcs}{100}
\DefMacro{test-results-bl-model0-lc9-num-fail}{[100]}
\DefMacro{test-results-bl-model0-lc9-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc10}{Pronoun reference::Hate expressed through reference in subsequent clauses}
\DefMacro{test-results-bl-lc10-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc10-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc10-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc11}{Pronoun reference::Hate expressed through reference in subsequent sentences}
\DefMacro{test-results-bl-lc11-num-tcs}{133}
\DefMacro{test-results-bl-model0-lc11-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc11-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc6}{Slur usage::Hate expressed using slur}
\DefMacro{test-results-bl-lc6-num-tcs}{144}
\DefMacro{test-results-bl-model0-lc6-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc6-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc7}{Slur usage::Non-hateful use of slur}
\DefMacro{test-results-bl-lc7-num-tcs}{111}
\DefMacro{test-results-bl-model0-lc7-num-fail}{[111]}
\DefMacro{test-results-bl-model0-lc7-num-failrate}{[100.0]}
\DefMacro{test-results-bl-lc4}{Threatening language::Direct threat}
\DefMacro{test-results-bl-lc4-num-tcs}{133}
\DefMacro{test-results-bl-model0-lc4-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc4-num-failrate}{[0.0]}
\DefMacro{test-results-bl-lc5}{Threatening language::Threat as normative statement}
\DefMacro{test-results-bl-lc5-num-tcs}{140}
\DefMacro{test-results-bl-model0-lc5-num-fail}{[0]}
\DefMacro{test-results-bl-model0-lc5-num-failrate}{[0.0]}
\DefMacro{test-results-bl-model1-lc19-num-fail}{[5]}
\DefMacro{test-results-bl-model1-lc19-num-failrate}{[3.5460992907801416]}
\DefMacro{test-results-bl-model1-lc18-num-fail}{[4]}
\DefMacro{test-results-bl-model1-lc18-num-failrate}{[2.3121387283236996]}
\DefMacro{test-results-bl-model1-lc2-num-fail}{[138]}
\DefMacro{test-results-bl-model1-lc2-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model1-lc1-num-fail}{[136]}
\DefMacro{test-results-bl-model1-lc1-num-failrate}{[97.14285714285714]}
\DefMacro{test-results-bl-model1-lc0-num-fail}{[134]}
\DefMacro{test-results-bl-model1-lc0-num-failrate}{[95.71428571428571]}
\DefMacro{test-results-bl-model1-lc3-num-fail}{[135]}
\DefMacro{test-results-bl-model1-lc3-num-failrate}{[96.42857142857143]}
\DefMacro{test-results-bl-model1-lc12-num-fail}{[136]}
\DefMacro{test-results-bl-model1-lc12-num-failrate}{[97.14285714285714]}
\DefMacro{test-results-bl-model1-lc13-num-fail}{[5]}
\DefMacro{test-results-bl-model1-lc13-num-failrate}{[3.7593984962406015]}
\DefMacro{test-results-bl-model1-lc16-num-fail}{[3]}
\DefMacro{test-results-bl-model1-lc16-num-failrate}{[2.380952380952381]}
\DefMacro{test-results-bl-model1-lc17-num-fail}{[8]}
\DefMacro{test-results-bl-model1-lc17-num-failrate}{[4.232804232804233]}
\DefMacro{test-results-bl-model1-lc14-num-fail}{[137]}
\DefMacro{test-results-bl-model1-lc14-num-failrate}{[97.85714285714286]}
\DefMacro{test-results-bl-model1-lc8-num-fail}{[135]}
\DefMacro{test-results-bl-model1-lc8-num-failrate}{[96.42857142857143]}
\DefMacro{test-results-bl-model1-lc9-num-fail}{[1]}
\DefMacro{test-results-bl-model1-lc9-num-failrate}{[1.0]}
\DefMacro{test-results-bl-model1-lc10-num-fail}{[132]}
\DefMacro{test-results-bl-model1-lc10-num-failrate}{[94.28571428571429]}
\DefMacro{test-results-bl-model1-lc11-num-fail}{[127]}
\DefMacro{test-results-bl-model1-lc11-num-failrate}{[95.48872180451127]}
\DefMacro{test-results-bl-model1-lc6-num-fail}{[142]}
\DefMacro{test-results-bl-model1-lc6-num-failrate}{[98.61111111111111]}
\DefMacro{test-results-bl-model1-lc7-num-fail}{[2]}
\DefMacro{test-results-bl-model1-lc7-num-failrate}{[1.8018018018018018]}
\DefMacro{test-results-bl-model1-lc4-num-fail}{[132]}
\DefMacro{test-results-bl-model1-lc4-num-failrate}{[99.24812030075188]}
\DefMacro{test-results-bl-model1-lc5-num-fail}{[136]}
\DefMacro{test-results-bl-model1-lc5-num-failrate}{[97.14285714285714]}
\DefMacro{test-results-bl-model2-lc19-num-fail}{[1]}
\DefMacro{test-results-bl-model2-lc19-num-failrate}{[0.7092198581560284]}
\DefMacro{test-results-bl-model2-lc18-num-fail}{[4]}
\DefMacro{test-results-bl-model2-lc18-num-failrate}{[2.3121387283236996]}
\DefMacro{test-results-bl-model2-lc2-num-fail}{[138]}
\DefMacro{test-results-bl-model2-lc2-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model2-lc1-num-fail}{[138]}
\DefMacro{test-results-bl-model2-lc1-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model2-lc0-num-fail}{[136]}
\DefMacro{test-results-bl-model2-lc0-num-failrate}{[97.14285714285714]}
\DefMacro{test-results-bl-model2-lc3-num-fail}{[137]}
\DefMacro{test-results-bl-model2-lc3-num-failrate}{[97.85714285714286]}
\DefMacro{test-results-bl-model2-lc12-num-fail}{[138]}
\DefMacro{test-results-bl-model2-lc12-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model2-lc13-num-fail}{[2]}
\DefMacro{test-results-bl-model2-lc13-num-failrate}{[1.5037593984962405]}
\DefMacro{test-results-bl-model2-lc16-num-fail}{[0]}
\DefMacro{test-results-bl-model2-lc16-num-failrate}{[0.0]}
\DefMacro{test-results-bl-model2-lc17-num-fail}{[3]}
\DefMacro{test-results-bl-model2-lc17-num-failrate}{[1.5873015873015872]}
\DefMacro{test-results-bl-model2-lc14-num-fail}{[139]}
\DefMacro{test-results-bl-model2-lc14-num-failrate}{[99.28571428571429]}
\DefMacro{test-results-bl-model2-lc8-num-fail}{[138]}
\DefMacro{test-results-bl-model2-lc8-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model2-lc9-num-fail}{[4]}
\DefMacro{test-results-bl-model2-lc9-num-failrate}{[4.0]}
\DefMacro{test-results-bl-model2-lc10-num-fail}{[138]}
\DefMacro{test-results-bl-model2-lc10-num-failrate}{[98.57142857142857]}
\DefMacro{test-results-bl-model2-lc11-num-fail}{[133]}
\DefMacro{test-results-bl-model2-lc11-num-failrate}{[100.0]}
\DefMacro{test-results-bl-model2-lc6-num-fail}{[144]}
\DefMacro{test-results-bl-model2-lc6-num-failrate}{[100.0]}
\DefMacro{test-results-bl-model2-lc7-num-fail}{[1]}
\DefMacro{test-results-bl-model2-lc7-num-failrate}{[0.9009009009009009]}
\DefMacro{test-results-bl-model2-lc4-num-fail}{[133]}
\DefMacro{test-results-bl-model2-lc4-num-failrate}{[100.0]}
\DefMacro{test-results-bl-model2-lc5-num-fail}{[139]}
\DefMacro{test-results-bl-model2-lc5-num-failrate}{[99.28571428571429]}