### 2023-01-11 paper/ase22/tables/test-results-hs-all-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-lc2}{Dehumanisation (explicit)}
\DefMacro{test-results-lc2-num-seeds}{140}
\DefMacro{test-results-lc2-num-exps}{3124}
\DefMacro{test-results-model0-lc2-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc2-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc2-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc1}{Description using very negative attributes (explicit)}
\DefMacro{test-results-lc1-num-seeds}{140}
\DefMacro{test-results-lc1-num-exps}{2959}
\DefMacro{test-results-model0-lc1-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc1-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc1-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc4}{Direct threat}
\DefMacro{test-results-lc4-num-seeds}{133}
\DefMacro{test-results-lc4-num-exps}{2689}
\DefMacro{test-results-model0-lc4-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc4-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc4-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc0}{Expression of strong negative emotions (explicit)}
\DefMacro{test-results-lc0-num-seeds}{140}
\DefMacro{test-results-lc0-num-exps}{799}
\DefMacro{test-results-model0-lc0-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc0-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc0-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc8}{Hate expressed using profanity}
\DefMacro{test-results-lc8-num-seeds}{1842}
\DefMacro{test-results-lc8-num-exps}{49743}
\DefMacro{test-results-model0-lc8-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc8-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc8-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc6}{Hate expressed using slur}
\DefMacro{test-results-lc6-num-seeds}{805}
\DefMacro{test-results-lc6-num-exps}{17318}
\DefMacro{test-results-model0-lc6-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc6-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc6-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc3}{Implicit derogation}
\DefMacro{test-results-lc3-num-seeds}{140}
\DefMacro{test-results-lc3-num-exps}{5664}
\DefMacro{test-results-model0-lc3-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc3-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc3-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc9}{Non-Hateful use of profanity}
\DefMacro{test-results-lc9-num-seeds}{701}
\DefMacro{test-results-lc9-num-exps}{15761}
\DefMacro{test-results-model0-lc9-num-all-fail}{[16462]}
\DefMacro{test-results-model0-lc9-num-all-failrate}{[100.0]}
\DefMacro{test-results-model0-lc9-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc7}{Non-hateful use of slur}
\DefMacro{test-results-lc7-num-seeds}{395}
\DefMacro{test-results-lc7-num-exps}{7881}
\DefMacro{test-results-model0-lc7-num-all-fail}{[8276]}
\DefMacro{test-results-model0-lc7-num-all-failrate}{[100.0]}
\DefMacro{test-results-model0-lc7-num-pass-to-fail}{[0]}
\DefMacro{test-results-lc5}{Threat as normative statement}
\DefMacro{test-results-lc5-num-seeds}{140}
\DefMacro{test-results-lc5-num-exps}{4163}
\DefMacro{test-results-model0-lc5-num-all-fail}{[0]}
\DefMacro{test-results-model0-lc5-num-all-failrate}{[0.0]}
\DefMacro{test-results-model0-lc5-num-pass-to-fail}{[0]}
\DefMacro{test-results-model1-lc2-num-all-fail}{[3210]}
\DefMacro{test-results-model1-lc2-num-all-failrate}{[98.34558823529412]}
\DefMacro{test-results-model1-lc2-num-pass-to-fail}{[11]}
\DefMacro{test-results-model1-lc1-num-all-fail}{[3011]}
\DefMacro{test-results-model1-lc1-num-all-failrate}{[97.16037431429493]}
\DefMacro{test-results-model1-lc1-num-pass-to-fail}{[30]}
\DefMacro{test-results-model1-lc4-num-all-fail}{[2750]}
\DefMacro{test-results-model1-lc4-num-all-failrate}{[97.44861800141743]}
\DefMacro{test-results-model1-lc4-num-pass-to-fail}{[0]}
\DefMacro{test-results-model1-lc0-num-all-fail}{[845]}
\DefMacro{test-results-model1-lc0-num-all-failrate}{[89.98935037273695]}
\DefMacro{test-results-model1-lc0-num-pass-to-fail}{[19]}
\DefMacro{test-results-model1-lc8-num-all-fail}{[49739]}
\DefMacro{test-results-model1-lc8-num-all-failrate}{[96.42144034118445]}
\DefMacro{test-results-model1-lc8-num-pass-to-fail}{[188]}
\DefMacro{test-results-model1-lc6-num-all-fail}{[17212]}
\DefMacro{test-results-model1-lc6-num-all-failrate}{[94.97323842630911]}
\DefMacro{test-results-model1-lc6-num-pass-to-fail}{[81]}
\DefMacro{test-results-model1-lc3-num-all-fail}{[5526]}
\DefMacro{test-results-model1-lc3-num-all-failrate}{[95.21019986216402]}
\DefMacro{test-results-model1-lc3-num-pass-to-fail}{[56]}
\DefMacro{test-results-model1-lc9-num-all-fail}{[541]}
\DefMacro{test-results-model1-lc9-num-all-failrate}{[3.28635645729559]}
\DefMacro{test-results-model1-lc9-num-pass-to-fail}{[51]}
\DefMacro{test-results-model1-lc7-num-all-fail}{[318]}
\DefMacro{test-results-model1-lc7-num-all-failrate}{[3.8424359594006767]}
\DefMacro{test-results-model1-lc7-num-pass-to-fail}{[20]}
\DefMacro{test-results-model1-lc5-num-all-fail}{[4198]}
\DefMacro{test-results-model1-lc5-num-all-failrate}{[97.5598419707181]}
\DefMacro{test-results-model1-lc5-num-pass-to-fail}{[12]}
\DefMacro{test-results-model2-lc2-num-all-fail}{[3220]}
\DefMacro{test-results-model2-lc2-num-all-failrate}{[98.65196078431373]}
\DefMacro{test-results-model2-lc2-num-pass-to-fail}{[18]}
\DefMacro{test-results-model2-lc1-num-all-fail}{[3067]}
\DefMacro{test-results-model2-lc1-num-all-failrate}{[98.96740884156179]}
\DefMacro{test-results-model2-lc1-num-pass-to-fail}{[23]}
\DefMacro{test-results-model2-lc4-num-all-fail}{[2770]}
\DefMacro{test-results-model2-lc4-num-all-failrate}{[98.15733522324592]}
\DefMacro{test-results-model2-lc4-num-pass-to-fail}{[9]}
\DefMacro{test-results-model2-lc0-num-all-fail}{[920]}
\DefMacro{test-results-model2-lc0-num-all-failrate}{[97.9765708200213]}
\DefMacro{test-results-model2-lc0-num-pass-to-fail}{[22]}
\DefMacro{test-results-model2-lc8-num-all-fail}{[50158]}
\DefMacro{test-results-model2-lc8-num-all-failrate}{[97.23369196471843]}
\DefMacro{test-results-model2-lc8-num-pass-to-fail}{[164]}
\DefMacro{test-results-model2-lc6-num-all-fail}{[17597]}
\DefMacro{test-results-model2-lc6-num-all-failrate}{[97.09761077084367]}
\DefMacro{test-results-model2-lc6-num-pass-to-fail}{[70]}
\DefMacro{test-results-model2-lc3-num-all-fail}{[5730]}
\DefMacro{test-results-model2-lc3-num-all-failrate}{[98.7250172294969]}
\DefMacro{test-results-model2-lc3-num-pass-to-fail}{[30]}
\DefMacro{test-results-model2-lc9-num-all-fail}{[549]}
\DefMacro{test-results-model2-lc9-num-all-failrate}{[3.334953225610497]}
\DefMacro{test-results-model2-lc9-num-pass-to-fail}{[58]}
\DefMacro{test-results-model2-lc7-num-all-fail}{[222]}
\DefMacro{test-results-model2-lc7-num-all-failrate}{[2.682455292411793]}
\DefMacro{test-results-model2-lc7-num-pass-to-fail}{[26]}
\DefMacro{test-results-model2-lc5-num-all-fail}{[4261]}
\DefMacro{test-results-model2-lc5-num-all-failrate}{[99.02393678828724]}
\DefMacro{test-results-model2-lc5-num-pass-to-fail}{[1]}