### 2023-01-17 paper/fse23/tables/test-results-all-table.tex

%% Automatically generated by pyutil.latex

\begin{table*}[t]
\begin{small}
\begin{center}
\caption{\TestResultsAllTableCaption}
\resizebox{0.9\textwidth}{!}{
\begin{tabular}{p{8cm}||ccclll}
\toprule
\tLc & \parbox{1.5cm}{\tNumBlTcs} & \parbox{1cm}{\tNumSeeds} & \parbox{1cm}{\tNumExps} & \parbox{1.5cm}{\centering\tNumFail} & \parbox{1.5cm}{\centering\tFailRate} & \parbox{1.5cm}{\centering\tNumPasstoFail}\\
\midrule
\multirow{3}{*}{\parbox{8cm}{LC1: \UseMacro{test-results-all-lc0}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc0-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc0-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc0-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc0-num-fail}/\UseMacro{test-results-bl-model0-lc0-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc0-num-failrate}/\UseMacro{test-results-bl-model0-lc0-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc0-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc0-num-fail}/\UseMacro{test-results-bl-model1-lc0-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc0-num-failrate}/\UseMacro{test-results-bl-model1-lc0-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc0-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc0-num-fail}/\UseMacro{test-results-bl-model2-lc0-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc0-num-failrate}/\UseMacro{test-results-bl-model2-lc0-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc0-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC2: \UseMacro{test-results-all-lc1}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc1-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc1-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc1-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc1-num-fail}/\UseMacro{test-results-bl-model0-lc1-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc1-num-failrate}/\UseMacro{test-results-bl-model0-lc1-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc1-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc1-num-fail}/\UseMacro{test-results-bl-model1-lc1-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc1-num-failrate}/\UseMacro{test-results-bl-model1-lc1-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc1-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc1-num-fail}/\UseMacro{test-results-bl-model2-lc1-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc1-num-failrate}/\UseMacro{test-results-bl-model2-lc1-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc1-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC3: \UseMacro{test-results-all-lc2}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc2-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc2-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc2-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc2-num-fail}/\UseMacro{test-results-bl-model0-lc2-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc2-num-failrate}/\UseMacro{test-results-bl-model0-lc2-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc2-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc2-num-fail}/\UseMacro{test-results-bl-model1-lc2-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc2-num-failrate}/\UseMacro{test-results-bl-model1-lc2-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc2-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc2-num-fail}/\UseMacro{test-results-bl-model2-lc2-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc2-num-failrate}/\UseMacro{test-results-bl-model2-lc2-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc2-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC4: \UseMacro{test-results-all-lc3}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc3-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc3-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc3-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc3-num-fail}/\UseMacro{test-results-bl-model0-lc3-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc3-num-failrate}/\UseMacro{test-results-bl-model0-lc3-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc3-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc3-num-fail}/\UseMacro{test-results-bl-model1-lc3-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc3-num-failrate}/\UseMacro{test-results-bl-model1-lc3-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc3-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc3-num-fail}/\UseMacro{test-results-bl-model2-lc3-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc3-num-failrate}/\UseMacro{test-results-bl-model2-lc3-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc3-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC5: \UseMacro{test-results-all-lc4}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc4-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc4-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc4-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc4-num-fail}/\UseMacro{test-results-bl-model0-lc4-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc4-num-failrate}/\UseMacro{test-results-bl-model0-lc4-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc4-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc4-num-fail}/\UseMacro{test-results-bl-model1-lc4-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc4-num-failrate}/\UseMacro{test-results-bl-model1-lc4-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc4-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc4-num-fail}/\UseMacro{test-results-bl-model2-lc4-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc4-num-failrate}/\UseMacro{test-results-bl-model2-lc4-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc4-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC6: \UseMacro{test-results-all-lc5}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc5-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc5-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc5-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc5-num-fail}/\UseMacro{test-results-bl-model0-lc5-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc5-num-failrate}/\UseMacro{test-results-bl-model0-lc5-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc5-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc5-num-fail}/\UseMacro{test-results-bl-model1-lc5-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc5-num-failrate}/\UseMacro{test-results-bl-model1-lc5-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc5-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc5-num-fail}/\UseMacro{test-results-bl-model2-lc5-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc5-num-failrate}/\UseMacro{test-results-bl-model2-lc5-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc5-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC7: \UseMacro{test-results-all-lc6}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc6-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc6-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc6-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc6-num-fail}/\UseMacro{test-results-bl-model0-lc6-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc6-num-failrate}/\UseMacro{test-results-bl-model0-lc6-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc6-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc6-num-fail}/\UseMacro{test-results-bl-model1-lc6-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc6-num-failrate}/\UseMacro{test-results-bl-model1-lc6-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc6-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc6-num-fail}/\UseMacro{test-results-bl-model2-lc6-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc6-num-failrate}/\UseMacro{test-results-bl-model2-lc6-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc6-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC8: \UseMacro{test-results-all-lc7}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc7-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc7-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc7-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc7-num-fail}/\UseMacro{test-results-bl-model0-lc7-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc7-num-failrate}/\UseMacro{test-results-bl-model0-lc7-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc7-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc7-num-fail}/\UseMacro{test-results-bl-model1-lc7-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc7-num-failrate}/\UseMacro{test-results-bl-model1-lc7-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc7-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc7-num-fail}/\UseMacro{test-results-bl-model2-lc7-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc7-num-failrate}/\UseMacro{test-results-bl-model2-lc7-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc7-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC9: \UseMacro{test-results-all-lc8}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc8-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc8-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc8-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc8-num-fail}/\UseMacro{test-results-bl-model0-lc8-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc8-num-failrate}/\UseMacro{test-results-bl-model0-lc8-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc8-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc8-num-fail}/\UseMacro{test-results-bl-model1-lc8-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc8-num-failrate}/\UseMacro{test-results-bl-model1-lc8-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc8-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc8-num-fail}/\UseMacro{test-results-bl-model2-lc8-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc8-num-failrate}/\UseMacro{test-results-bl-model2-lc8-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc8-num-pass-to-fail}\\
\hline
\multirow{3}{*}{\parbox{8cm}{LC10: \UseMacro{test-results-all-lc9}}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-bl-lc9-num-tcs}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc9-num-seeds}}
 & \multirow{3}{*}{\centering\UseMacro{test-results-all-lc9-num-exps}}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc9-num-fail}/\UseMacro{test-results-bl-model0-lc9-num-fail}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc9-num-failrate}/\UseMacro{test-results-bl-model0-lc9-num-failrate}
 & BERT$\colon$\UseMacro{test-results-all-model0-lc9-num-pass-to-fail}\\
 & & & & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc9-num-fail}/\UseMacro{test-results-bl-model1-lc9-num-fail}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc9-num-failrate}/\UseMacro{test-results-bl-model1-lc9-num-failrate}
 & RoBERTa$\colon$\UseMacro{test-results-all-model1-lc9-num-pass-to-fail}\\
 & & & & dstBERT$\colon$\UseMacro{test-results-all-model2-lc9-num-fail}/\UseMacro{test-results-bl-model2-lc9-num-fail}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc9-num-failrate}/\UseMacro{test-results-bl-model2-lc9-num-failrate}
 & dstBERT$\colon$\UseMacro{test-results-all-model2-lc9-num-pass-to-fail}\\
\hline
\bottomrule
\end{tabular}}
\end{center}
\end{small}
\vspace{\TestResultsTableVSpace}
\end{table*}