### 2023-01-18 paper/fse23/tables/test-results-hs-all-numbers.tex
tabnle

%% Automatically generated by pyutil.latex

\DefMacro{test-results-hs-lc2}{Dehumanisation (explicit)}
\DefMacro{test-results-hs-lc2-num-seeds}{140}
\DefMacro{test-results-hs-lc2-num-exps}{3,124}
\DefMacro{test-results-hs-bl-lc2-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc2-num-all-fail}{2,646}
\DefMacro{test-results-hs-model0-lc2-num-seed-failrate}{85.00}
\DefMacro{test-results-hs-model0-lc2-num-exp-failrate}{80.89}
\DefMacro{test-results-hs-model0-lc2-num-all-failrate}{81.07}
\DefMacro{test-results-hs-model0-lc2-num-pass-to-fail}{15}
\DefMacro{test-results-hs-bl-model0-lc2-num-fail}{119}
\DefMacro{test-results-hs-bl-model0-lc2-num-failrate}{85.00}
\DefMacro{test-results-hs-lc18}{Denouncements of hate that quote it}
\DefMacro{test-results-hs-lc18-num-seeds}{45,616}
\DefMacro{test-results-hs-lc18-num-exps}{300,704}
\DefMacro{test-results-hs-bl-lc18-num-tcs}{173}
\DefMacro{test-results-hs-model0-lc18-num-all-fail}{87,197}
\DefMacro{test-results-hs-model0-lc18-num-seed-failrate}{23.04}
\DefMacro{test-results-hs-model0-lc18-num-exp-failrate}{25.50}
\DefMacro{test-results-hs-model0-lc18-num-all-failrate}{25.18}
\DefMacro{test-results-hs-model0-lc18-num-pass-to-fail}{3,379}
\DefMacro{test-results-hs-bl-model0-lc18-num-fail}{36}
\DefMacro{test-results-hs-bl-model0-lc18-num-failrate}{20.81}
\DefMacro{test-results-hs-lc1}{Description using very negative attributes (explicit)}
\DefMacro{test-results-hs-lc1-num-seeds}{140}
\DefMacro{test-results-hs-lc1-num-exps}{2,959}
\DefMacro{test-results-hs-bl-lc1-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc1-num-all-fail}{2,731}
\DefMacro{test-results-hs-model0-lc1-num-seed-failrate}{87.86}
\DefMacro{test-results-hs-model0-lc1-num-exp-failrate}{88.14}
\DefMacro{test-results-hs-model0-lc1-num-all-failrate}{88.13}
\DefMacro{test-results-hs-model0-lc1-num-pass-to-fail}{31}
\DefMacro{test-results-hs-bl-model0-lc1-num-fail}{123}
\DefMacro{test-results-hs-bl-model0-lc1-num-failrate}{87.86}
\DefMacro{test-results-hs-lc4}{Direct threat}
\DefMacro{test-results-hs-lc4-num-seeds}{133}
\DefMacro{test-results-hs-lc4-num-exps}{2,689}
\DefMacro{test-results-hs-bl-lc4-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc4-num-all-fail}{2,056}
\DefMacro{test-results-hs-model0-lc4-num-seed-failrate}{69.17}
\DefMacro{test-results-hs-model0-lc4-num-exp-failrate}{73.04}
\DefMacro{test-results-hs-model0-lc4-num-all-failrate}{72.86}
\DefMacro{test-results-hs-model0-lc4-num-pass-to-fail}{49}
\DefMacro{test-results-hs-bl-model0-lc4-num-fail}{90}
\DefMacro{test-results-hs-bl-model0-lc4-num-failrate}{67.67}
\DefMacro{test-results-hs-lc0}{Expression of strong negative emotions (explicit)}
\DefMacro{test-results-hs-lc0-num-seeds}{140}
\DefMacro{test-results-hs-lc0-num-exps}{799}
\DefMacro{test-results-hs-bl-lc0-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc0-num-all-fail}{823}
\DefMacro{test-results-hs-model0-lc0-num-seed-failrate}{87.86}
\DefMacro{test-results-hs-model0-lc0-num-exp-failrate}{87.61}
\DefMacro{test-results-hs-model0-lc0-num-all-failrate}{87.65}
\DefMacro{test-results-hs-model0-lc0-num-pass-to-fail}{12}
\DefMacro{test-results-hs-bl-model0-lc0-num-fail}{123}
\DefMacro{test-results-hs-bl-model0-lc0-num-failrate}{87.86}
\DefMacro{test-results-hs-lc10}{Hate expressed through reference in subsequent clauses}
\DefMacro{test-results-hs-lc10-num-seeds}{11,968}
\DefMacro{test-results-hs-lc10-num-exps}{44,890}
\DefMacro{test-results-hs-bl-lc10-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc10-num-all-fail}{33,230}
\DefMacro{test-results-hs-model0-lc10-num-seed-failrate}{48.53}
\DefMacro{test-results-hs-model0-lc10-num-exp-failrate}{61.09}
\DefMacro{test-results-hs-model0-lc10-num-all-failrate}{58.44}
\DefMacro{test-results-hs-model0-lc10-num-pass-to-fail}{1,434}
\DefMacro{test-results-hs-bl-model0-lc10-num-fail}{114}
\DefMacro{test-results-hs-bl-model0-lc10-num-failrate}{81.43}
\DefMacro{test-results-hs-lc11}{Hate expressed through reference in subsequent sentences}
\DefMacro{test-results-hs-lc11-num-seeds}{11,968}
\DefMacro{test-results-hs-lc11-num-exps}{42,840}
\DefMacro{test-results-hs-bl-lc11-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc11-num-all-fail}{32,167}
\DefMacro{test-results-hs-model0-lc11-num-seed-failrate}{50.70}
\DefMacro{test-results-hs-model0-lc11-num-exp-failrate}{60.92}
\DefMacro{test-results-hs-model0-lc11-num-all-failrate}{58.69}
\DefMacro{test-results-hs-model0-lc11-num-pass-to-fail}{1,410}
\DefMacro{test-results-hs-bl-model0-lc11-num-fail}{104}
\DefMacro{test-results-hs-bl-model0-lc11-num-failrate}{78.20}
\DefMacro{test-results-hs-lc12}{Hate expressed using negated positive statement}
\DefMacro{test-results-hs-lc12-num-seeds}{23,379}
\DefMacro{test-results-hs-lc12-num-exps}{126,742}
\DefMacro{test-results-hs-bl-lc12-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc12-num-all-fail}{130,886}
\DefMacro{test-results-hs-model0-lc12-num-seed-failrate}{86.56}
\DefMacro{test-results-hs-model0-lc12-num-exp-failrate}{87.30}
\DefMacro{test-results-hs-model0-lc12-num-all-failrate}{87.19}
\DefMacro{test-results-hs-model0-lc12-num-pass-to-fail}{1,888}
\DefMacro{test-results-hs-bl-model0-lc12-num-fail}{123}
\DefMacro{test-results-hs-bl-model0-lc12-num-failrate}{87.86}
\DefMacro{test-results-hs-lc8}{Hate expressed using profanity}
\DefMacro{test-results-hs-lc8-num-seeds}{1,842}
\DefMacro{test-results-hs-lc8-num-exps}{49,743}
\DefMacro{test-results-hs-bl-lc8-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc8-num-all-fail}{38,673}
\DefMacro{test-results-hs-model0-lc8-num-seed-failrate}{77.31}
\DefMacro{test-results-hs-model0-lc8-num-exp-failrate}{74.88}
\DefMacro{test-results-hs-model0-lc8-num-all-failrate}{74.97}
\DefMacro{test-results-hs-model0-lc8-num-pass-to-fail}{177}
\DefMacro{test-results-hs-bl-model0-lc8-num-fail}{119}
\DefMacro{test-results-hs-bl-model0-lc8-num-failrate}{85.00}
\DefMacro{test-results-hs-lc6}{Hate expressed using slur}
\DefMacro{test-results-hs-lc6-num-seeds}{805}
\DefMacro{test-results-hs-lc6-num-exps}{17,318}
\DefMacro{test-results-hs-bl-lc6-num-tcs}{144}
\DefMacro{test-results-hs-model0-lc6-num-all-fail}{13,383}
\DefMacro{test-results-hs-model0-lc6-num-seed-failrate}{75.16}
\DefMacro{test-results-hs-model0-lc6-num-exp-failrate}{73.78}
\DefMacro{test-results-hs-model0-lc6-num-all-failrate}{73.85}
\DefMacro{test-results-hs-model0-lc6-num-pass-to-fail}{92}
\DefMacro{test-results-hs-bl-model0-lc6-num-fail}{129}
\DefMacro{test-results-hs-bl-model0-lc6-num-failrate}{89.58}
\DefMacro{test-results-hs-lc15}{Hate phrased as a opinion}
\DefMacro{test-results-hs-lc15-num-seeds}{22,808}
\DefMacro{test-results-hs-lc15-num-exps}{160,927}
\DefMacro{test-results-hs-bl-lc15-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc15-num-all-fail}{142,434}
\DefMacro{test-results-hs-model0-lc15-num-seed-failrate}{79.31}
\DefMacro{test-results-hs-model0-lc15-num-exp-failrate}{77.27}
\DefMacro{test-results-hs-model0-lc15-num-all-failrate}{77.52}
\DefMacro{test-results-hs-model0-lc15-num-pass-to-fail}{3,544}
\DefMacro{test-results-hs-bl-model0-lc15-num-fail}{108}
\DefMacro{test-results-hs-bl-model0-lc15-num-failrate}{81.20}
\DefMacro{test-results-hs-lc14}{Hate phrased as a question}
\DefMacro{test-results-hs-lc14-num-seeds}{22,808}
\DefMacro{test-results-hs-lc14-num-exps}{185,267}
\DefMacro{test-results-hs-bl-lc14-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc14-num-all-fail}{163,340}
\DefMacro{test-results-hs-model0-lc14-num-seed-failrate}{80.37}
\DefMacro{test-results-hs-model0-lc14-num-exp-failrate}{78.27}
\DefMacro{test-results-hs-model0-lc14-num-all-failrate}{78.50}
\DefMacro{test-results-hs-model0-lc14-num-pass-to-fail}{3,435}
\DefMacro{test-results-hs-bl-model0-lc14-num-fail}{117}
\DefMacro{test-results-hs-bl-model0-lc14-num-failrate}{83.57}
\DefMacro{test-results-hs-lc3}{Implicit derogation}
\DefMacro{test-results-hs-lc3-num-seeds}{140}
\DefMacro{test-results-hs-lc3-num-exps}{5,664}
\DefMacro{test-results-hs-bl-lc3-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc3-num-all-fail}{5,029}
\DefMacro{test-results-hs-model0-lc3-num-seed-failrate}{84.29}
\DefMacro{test-results-hs-model0-lc3-num-exp-failrate}{86.71}
\DefMacro{test-results-hs-model0-lc3-num-all-failrate}{86.65}
\DefMacro{test-results-hs-model0-lc3-num-pass-to-fail}{77}
\DefMacro{test-results-hs-bl-model0-lc3-num-fail}{118}
\DefMacro{test-results-hs-bl-model0-lc3-num-failrate}{84.29}
\DefMacro{test-results-hs-lc16}{Neutral statements using protected group identifiers}
\DefMacro{test-results-hs-lc16-num-seeds}{5}
\DefMacro{test-results-hs-lc16-num-exps}{9}
\DefMacro{test-results-hs-bl-lc16-num-tcs}{126}
\DefMacro{test-results-hs-model0-lc16-num-all-fail}{13}
\DefMacro{test-results-hs-model0-lc16-num-seed-failrate}{80.00}
\DefMacro{test-results-hs-model0-lc16-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc16-num-all-failrate}{92.86}
\DefMacro{test-results-hs-model0-lc16-num-pass-to-fail}{0}
\DefMacro{test-results-hs-bl-model0-lc16-num-fail}{12}
\DefMacro{test-results-hs-bl-model0-lc16-num-failrate}{9.52}
\DefMacro{test-results-hs-lc9}{Non-Hateful use of profanity}
\DefMacro{test-results-hs-lc9-num-seeds}{701}
\DefMacro{test-results-hs-lc9-num-exps}{15,761}
\DefMacro{test-results-hs-bl-lc9-num-tcs}{100}
\DefMacro{test-results-hs-model0-lc9-num-all-fail}{3,583}
\DefMacro{test-results-hs-model0-lc9-num-seed-failrate}{17.69}
\DefMacro{test-results-hs-model0-lc9-num-exp-failrate}{21.95}
\DefMacro{test-results-hs-model0-lc9-num-all-failrate}{21.77}
\DefMacro{test-results-hs-model0-lc9-num-pass-to-fail}{44}
\DefMacro{test-results-hs-bl-model0-lc9-num-fail}{6}
\DefMacro{test-results-hs-bl-model0-lc9-num-failrate}{6.00}
\DefMacro{test-results-hs-lc13}{Non-hate expressed using negated hateful statement}
\DefMacro{test-results-hs-lc13-num-seeds}{34,212}
\DefMacro{test-results-hs-lc13-num-exps}{243,739}
\DefMacro{test-results-hs-bl-lc13-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc13-num-all-fail}{86,576}
\DefMacro{test-results-hs-model0-lc13-num-seed-failrate}{29.45}
\DefMacro{test-results-hs-model0-lc13-num-exp-failrate}{31.39}
\DefMacro{test-results-hs-model0-lc13-num-all-failrate}{31.15}
\DefMacro{test-results-hs-model0-lc13-num-pass-to-fail}{3,134}
\DefMacro{test-results-hs-bl-model0-lc13-num-fail}{25}
\DefMacro{test-results-hs-bl-model0-lc13-num-failrate}{18.80}
\DefMacro{test-results-hs-lc7}{Non-hateful use of slur}
\DefMacro{test-results-hs-lc7-num-seeds}{395}
\DefMacro{test-results-hs-lc7-num-exps}{7,881}
\DefMacro{test-results-hs-bl-lc7-num-tcs}{111}
\DefMacro{test-results-hs-model0-lc7-num-all-fail}{5,600}
\DefMacro{test-results-hs-model0-lc7-num-seed-failrate}{73.16}
\DefMacro{test-results-hs-model0-lc7-num-exp-failrate}{67.39}
\DefMacro{test-results-hs-model0-lc7-num-all-failrate}{67.67}
\DefMacro{test-results-hs-model0-lc7-num-pass-to-fail}{11}
\DefMacro{test-results-hs-bl-model0-lc7-num-fail}{22}
\DefMacro{test-results-hs-bl-model0-lc7-num-failrate}{19.82}
\DefMacro{test-results-hs-lc17}{Positive statements using protected group identifiers}
\DefMacro{test-results-hs-lc17-num-seeds}{41}
\DefMacro{test-results-hs-lc17-num-exps}{182}
\DefMacro{test-results-hs-bl-lc17-num-tcs}{189}
\DefMacro{test-results-hs-model0-lc17-num-all-fail}{137}
\DefMacro{test-results-hs-model0-lc17-num-seed-failrate}{65.85}
\DefMacro{test-results-hs-model0-lc17-num-exp-failrate}{60.44}
\DefMacro{test-results-hs-model0-lc17-num-all-failrate}{61.43}
\DefMacro{test-results-hs-model0-lc17-num-pass-to-fail}{1}
\DefMacro{test-results-hs-bl-model0-lc17-num-fail}{32}
\DefMacro{test-results-hs-bl-model0-lc17-num-failrate}{16.93}
\DefMacro{test-results-hs-lc5}{Threat as normative statement}
\DefMacro{test-results-hs-lc5-num-seeds}{140}
\DefMacro{test-results-hs-lc5-num-exps}{4,163}
\DefMacro{test-results-hs-bl-lc5-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc5-num-all-fail}{3,207}
\DefMacro{test-results-hs-model0-lc5-num-seed-failrate}{77.14}
\DefMacro{test-results-hs-model0-lc5-num-exp-failrate}{74.44}
\DefMacro{test-results-hs-model0-lc5-num-all-failrate}{74.53}
\DefMacro{test-results-hs-model0-lc5-num-pass-to-fail}{31}
\DefMacro{test-results-hs-bl-model0-lc5-num-fail}{108}
\DefMacro{test-results-hs-bl-model0-lc5-num-failrate}{77.14}
\DefMacro{test-results-hs-model1-lc2-num-all-fail}{2,854}
\DefMacro{test-results-hs-model1-lc2-num-seed-failrate}{88.57}
\DefMacro{test-results-hs-model1-lc2-num-exp-failrate}{87.39}
\DefMacro{test-results-hs-model1-lc2-num-all-failrate}{87.44}
\DefMacro{test-results-hs-model1-lc2-num-pass-to-fail}{96}
\DefMacro{test-results-hs-bl-model1-lc2-num-fail}{124}
\DefMacro{test-results-hs-bl-model1-lc2-num-failrate}{88.57}
\DefMacro{test-results-hs-model1-lc18-num-all-fail}{200,285}
\DefMacro{test-results-hs-model1-lc18-num-seed-failrate}{56.83}
\DefMacro{test-results-hs-model1-lc18-num-exp-failrate}{57.98}
\DefMacro{test-results-hs-model1-lc18-num-all-failrate}{57.83}
\DefMacro{test-results-hs-model1-lc18-num-pass-to-fail}{5,259}
\DefMacro{test-results-hs-bl-model1-lc18-num-fail}{8}
\DefMacro{test-results-hs-bl-model1-lc18-num-failrate}{4.62}
\DefMacro{test-results-hs-model1-lc1-num-all-fail}{2,465}
\DefMacro{test-results-hs-model1-lc1-num-seed-failrate}{77.14}
\DefMacro{test-results-hs-model1-lc1-num-exp-failrate}{79.66}
\DefMacro{test-results-hs-model1-lc1-num-all-failrate}{79.54}
\DefMacro{test-results-hs-model1-lc1-num-pass-to-fail}{140}
\DefMacro{test-results-hs-bl-model1-lc1-num-fail}{108}
\DefMacro{test-results-hs-bl-model1-lc1-num-failrate}{77.14}
\DefMacro{test-results-hs-model1-lc4-num-all-fail}{2,455}
\DefMacro{test-results-hs-model1-lc4-num-seed-failrate}{93.23}
\DefMacro{test-results-hs-model1-lc4-num-exp-failrate}{86.69}
\DefMacro{test-results-hs-model1-lc4-num-all-failrate}{87.00}
\DefMacro{test-results-hs-model1-lc4-num-pass-to-fail}{25}
\DefMacro{test-results-hs-bl-model1-lc4-num-fail}{121}
\DefMacro{test-results-hs-bl-model1-lc4-num-failrate}{90.98}
\DefMacro{test-results-hs-model1-lc0-num-all-fail}{829}
\DefMacro{test-results-hs-model1-lc0-num-seed-failrate}{88.57}
\DefMacro{test-results-hs-model1-lc0-num-exp-failrate}{88.24}
\DefMacro{test-results-hs-model1-lc0-num-all-failrate}{88.29}
\DefMacro{test-results-hs-model1-lc0-num-pass-to-fail}{38}
\DefMacro{test-results-hs-bl-model1-lc0-num-fail}{124}
\DefMacro{test-results-hs-bl-model1-lc0-num-failrate}{88.57}
\DefMacro{test-results-hs-model1-lc10-num-all-fail}{44,034}
\DefMacro{test-results-hs-model1-lc10-num-seed-failrate}{67.54}
\DefMacro{test-results-hs-model1-lc10-num-exp-failrate}{80.09}
\DefMacro{test-results-hs-model1-lc10-num-all-failrate}{77.45}
\DefMacro{test-results-hs-model1-lc10-num-pass-to-fail}{1,441}
\DefMacro{test-results-hs-bl-model1-lc10-num-fail}{122}
\DefMacro{test-results-hs-bl-model1-lc10-num-failrate}{87.14}
\DefMacro{test-results-hs-model1-lc11-num-all-fail}{43,256}
\DefMacro{test-results-hs-model1-lc11-num-seed-failrate}{68.57}
\DefMacro{test-results-hs-model1-lc11-num-exp-failrate}{81.81}
\DefMacro{test-results-hs-model1-lc11-num-all-failrate}{78.92}
\DefMacro{test-results-hs-model1-lc11-num-pass-to-fail}{1,348}
\DefMacro{test-results-hs-bl-model1-lc11-num-fail}{102}
\DefMacro{test-results-hs-bl-model1-lc11-num-failrate}{76.69}
\DefMacro{test-results-hs-model1-lc12-num-all-fail}{126,635}
\DefMacro{test-results-hs-model1-lc12-num-seed-failrate}{82.76}
\DefMacro{test-results-hs-model1-lc12-num-exp-failrate}{84.65}
\DefMacro{test-results-hs-model1-lc12-num-all-failrate}{84.36}
\DefMacro{test-results-hs-model1-lc12-num-pass-to-fail}{4,361}
\DefMacro{test-results-hs-bl-model1-lc12-num-fail}{124}
\DefMacro{test-results-hs-bl-model1-lc12-num-failrate}{88.57}
\DefMacro{test-results-hs-model1-lc8-num-all-fail}{46,325}
\DefMacro{test-results-hs-model1-lc8-num-seed-failrate}{90.88}
\DefMacro{test-results-hs-model1-lc8-num-exp-failrate}{89.76}
\DefMacro{test-results-hs-model1-lc8-num-all-failrate}{89.80}
\DefMacro{test-results-hs-model1-lc8-num-pass-to-fail}{145}
\DefMacro{test-results-hs-bl-model1-lc8-num-fail}{114}
\DefMacro{test-results-hs-bl-model1-lc8-num-failrate}{81.43}
\DefMacro{test-results-hs-model1-lc6-num-all-fail}{15,710}
\DefMacro{test-results-hs-model1-lc6-num-seed-failrate}{89.07}
\DefMacro{test-results-hs-model1-lc6-num-exp-failrate}{86.57}
\DefMacro{test-results-hs-model1-lc6-num-all-failrate}{86.69}
\DefMacro{test-results-hs-model1-lc6-num-pass-to-fail}{68}
\DefMacro{test-results-hs-bl-model1-lc6-num-fail}{135}
\DefMacro{test-results-hs-bl-model1-lc6-num-failrate}{93.75}
\DefMacro{test-results-hs-model1-lc15-num-all-fail}{158,127}
\DefMacro{test-results-hs-model1-lc15-num-seed-failrate}{85.90}
\DefMacro{test-results-hs-model1-lc15-num-exp-failrate}{86.09}
\DefMacro{test-results-hs-model1-lc15-num-all-failrate}{86.06}
\DefMacro{test-results-hs-model1-lc15-num-pass-to-fail}{4,584}
\DefMacro{test-results-hs-bl-model1-lc15-num-fail}{121}
\DefMacro{test-results-hs-bl-model1-lc15-num-failrate}{90.98}
\DefMacro{test-results-hs-model1-lc14-num-all-fail}{182,688}
\DefMacro{test-results-hs-model1-lc14-num-seed-failrate}{87.13}
\DefMacro{test-results-hs-model1-lc14-num-exp-failrate}{87.88}
\DefMacro{test-results-hs-model1-lc14-num-all-failrate}{87.80}
\DefMacro{test-results-hs-model1-lc14-num-pass-to-fail}{5,531}
\DefMacro{test-results-hs-bl-model1-lc14-num-fail}{131}
\DefMacro{test-results-hs-bl-model1-lc14-num-failrate}{93.57}
\DefMacro{test-results-hs-model1-lc3-num-all-fail}{5,481}
\DefMacro{test-results-hs-model1-lc3-num-seed-failrate}{95.00}
\DefMacro{test-results-hs-model1-lc3-num-exp-failrate}{94.42}
\DefMacro{test-results-hs-model1-lc3-num-all-failrate}{94.43}
\DefMacro{test-results-hs-model1-lc3-num-pass-to-fail}{19}
\DefMacro{test-results-hs-bl-model1-lc3-num-fail}{133}
\DefMacro{test-results-hs-bl-model1-lc3-num-failrate}{95.00}
\DefMacro{test-results-hs-model1-lc16-num-all-fail}{2}
\DefMacro{test-results-hs-model1-lc16-num-seed-failrate}{20.00}
\DefMacro{test-results-hs-model1-lc16-num-exp-failrate}{11.11}
\DefMacro{test-results-hs-model1-lc16-num-all-failrate}{14.29}
\DefMacro{test-results-hs-model1-lc16-num-pass-to-fail}{1}
\DefMacro{test-results-hs-bl-model1-lc16-num-fail}{1}
\DefMacro{test-results-hs-bl-model1-lc16-num-failrate}{0.79}
\DefMacro{test-results-hs-model1-lc9-num-all-fail}{7,267}
\DefMacro{test-results-hs-model1-lc9-num-seed-failrate}{58.35}
\DefMacro{test-results-hs-model1-lc9-num-exp-failrate}{43.51}
\DefMacro{test-results-hs-model1-lc9-num-all-failrate}{44.14}
\DefMacro{test-results-hs-model1-lc9-num-pass-to-fail}{60}
\DefMacro{test-results-hs-bl-model1-lc9-num-fail}{3}
\DefMacro{test-results-hs-bl-model1-lc9-num-failrate}{3.00}
\DefMacro{test-results-hs-model1-lc13-num-all-fail}{203,241}
\DefMacro{test-results-hs-model1-lc13-num-seed-failrate}{74.12}
\DefMacro{test-results-hs-model1-lc13-num-exp-failrate}{72.98}
\DefMacro{test-results-hs-model1-lc13-num-all-failrate}{73.12}
\DefMacro{test-results-hs-model1-lc13-num-pass-to-fail}{3,274}
\DefMacro{test-results-hs-bl-model1-lc13-num-fail}{33}
\DefMacro{test-results-hs-bl-model1-lc13-num-failrate}{24.81}
\DefMacro{test-results-hs-model1-lc7-num-all-fail}{6,815}
\DefMacro{test-results-hs-model1-lc7-num-seed-failrate}{89.62}
\DefMacro{test-results-hs-model1-lc7-num-exp-failrate}{81.98}
\DefMacro{test-results-hs-model1-lc7-num-all-failrate}{82.35}
\DefMacro{test-results-hs-model1-lc7-num-pass-to-fail}{26}
\DefMacro{test-results-hs-bl-model1-lc7-num-fail}{76}
\DefMacro{test-results-hs-bl-model1-lc7-num-failrate}{68.47}
\DefMacro{test-results-hs-model1-lc17-num-all-fail}{63}
\DefMacro{test-results-hs-model1-lc17-num-seed-failrate}{39.02}
\DefMacro{test-results-hs-model1-lc17-num-exp-failrate}{25.82}
\DefMacro{test-results-hs-model1-lc17-num-all-failrate}{28.25}
\DefMacro{test-results-hs-model1-lc17-num-pass-to-fail}{5}
\DefMacro{test-results-hs-bl-model1-lc17-num-fail}{26}
\DefMacro{test-results-hs-bl-model1-lc17-num-failrate}{13.76}
\DefMacro{test-results-hs-model1-lc5-num-all-fail}{3,591}
\DefMacro{test-results-hs-model1-lc5-num-seed-failrate}{84.29}
\DefMacro{test-results-hs-model1-lc5-num-exp-failrate}{83.43}
\DefMacro{test-results-hs-model1-lc5-num-all-failrate}{83.45}
\DefMacro{test-results-hs-model1-lc5-num-pass-to-fail}{50}
\DefMacro{test-results-hs-bl-model1-lc5-num-fail}{118}
\DefMacro{test-results-hs-bl-model1-lc5-num-failrate}{84.29}