### 2023-01-30 paper/fse23/tables/manual-study-numbers.tex
add table/plot

%% Automatically generated by pyutil.latex

\DefMacro{manual-study-sa-seed-num-sents}{78}
\DefMacro{manual-study-sa-seed-label-consistency}{0.85}
\DefMacro{manual-study-sa-seed-lc-relevancy}{0.89}
\DefMacro{manual-study-sa-exp-num-sents}{78}
\DefMacro{manual-study-sa-exp-label-consistency}{0.85}
\DefMacro{manual-study-sa-exp-lc-relevancy}{0.89}
\DefMacro{manual-study-hs-seed-num-sents}{100}
\DefMacro{manual-study-hs-seed-label-consistency}{0.79}
\DefMacro{manual-study-hs-seed-lc-relevancy}{0.88}
\DefMacro{manual-study-hs-exp-num-sents}{100}
\DefMacro{manual-study-hs-exp-label-consistency}{0.79}
\DefMacro{manual-study-hs-exp-lc-relevancy}{0.88}