### 2023-01-30 paper/fse23/tables/manual-study-table.tex

%% Automatically generated by pyutil.latex

\begin{table}[htbp]
\begin{small}
\begin{center}
\caption{\ManualStudyTableCaption}
\resizebox{0.47\textwidth}{!}{
\begin{tabular}{l|l|ccc}
\toprule
\tTask & \tSentType & \tNumTestCases & \tAvgLabelCons & \tAvgLCRel \\
\midrule
\multirow{2}{*}{SA} 
 & SEED & \UseMacro{manual-study-sa-seed-num-sents}
 & \UseMacro{manual-study-sa-seed-label-consistency}
 & \UseMacro{manual-study-sa-seed-lc-relevancy}\\
 & EXP & \UseMacro{manual-study-sa-exp-num-sents}
 & \UseMacro{manual-study-sa-exp-label-consistency}
 & \UseMacro{manual-study-sa-exp-lc-relevancy}\\
\hline
\multirow{2}{*}{HSD} 
 & SEED & \UseMacro{manual-study-hs-seed-num-sents}
 & \UseMacro{manual-study-hs-seed-label-consistency}
 & \UseMacro{manual-study-hs-seed-lc-relevancy}\\
 & EXP & \UseMacro{manual-study-hs-exp-num-sents}
 & \UseMacro{manual-study-hs-exp-label-consistency}
 & \UseMacro{manual-study-hs-exp-lc-relevancy}\\
\bottomrule
\end{tabular}}
\end{center}
\end{small}
\vspace{\ManualStudyTableVSpace}
\end{table}