### 2023-01-30 paper/fse23/tables/mtnlp-comp-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{mtnlp-comp-sa-ours-selfbleu-num-mutate}{50}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-avg}{0.15}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-med}{0.15}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-std}{0.02}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-num-mutate}{32}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-avg}{0.93}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-med}{0.93}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-std}{0.0}
\DefMacro{mtnlp-comp-sa-ours-pdr-num-mutate}{50}
\DefMacro{mtnlp-comp-sa-ours-pdr-avg}{164.7}
\DefMacro{mtnlp-comp-sa-ours-pdr-med}{160.5}
\DefMacro{mtnlp-comp-sa-ours-pdr-std}{10.07}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-num-mutate}{32}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-avg}{129.0}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-med}{129.0}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-std}{0.0}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-num-mutate}{50}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-avg}{0.15}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-med}{0.15}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-std}{0.01}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-num-mutate}{40}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-avg}{0.93}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-med}{0.93}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-std}{0.0}
\DefMacro{mtnlp-comp-hs-ours-pdr-num-mutate}{50}
\DefMacro{mtnlp-comp-hs-ours-pdr-avg}{206.6}
\DefMacro{mtnlp-comp-hs-ours-pdr-med}{208.0}
\DefMacro{mtnlp-comp-hs-ours-pdr-std}{7.12}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-num-mutate}{40}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-avg}{131.0}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-med}{131.0}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-std}{0.0}