### 2023-03-16 paper/fse23/tables/adv-comp-table.tex

%% Automatically generated by pyutil.latex

\begin{table}[htbp]
\begin{small}
\begin{center}
\caption{\AdvCompTableCaption}
\resizebox{0.47\textwidth}{!}{
\begin{tabular}{l|ccc}
\toprule
\tApproach & \tNummut & \tSelfbleu & \tPdrcov \\
\midrule
 \tool & \UseMacro{adv-comp-sa-ours-num-adv}
 & \UseMacro{adv-comp-sa-ours-selfbleu-avg}\pm\UseMacro{adv-comp-sa-ours-selfbleu-std}
 & \UseMacro{adv-comp-sa-ours-pdrcov-avg}\pm\UseMacro{adv-comp-sa-ours-pdrcov-std}\\
Alzantot & \UseMacro{adv-comp-sa-alzantot-num-adv}
 & \UseMacro{adv-comp-sa-alzantot-selfbleu-avg}\pm\UseMacro{adv-comp-sa-alzantot-selfbleu-std}
 & \UseMacro{adv-comp-sa-alzantot-pdrcov-avg}\pm\UseMacro{adv-comp-sa-alzantot-pdrcov-std}\\
BERT-Attack & \UseMacro{adv-comp-sa-bert-attack-num-adv}
 & \UseMacro{adv-comp-sa-bert-attack-selfbleu-avg}\pm\UseMacro{adv-comp-sa-bert-attack-selfbleu-std}
 & \UseMacro{adv-comp-sa-bert-attack-pdrcov-avg}\pm\UseMacro{adv-comp-sa-bert-attack-pdrcov-std}\\
PSO & \UseMacro{adv-comp-sa-pso-num-adv}
 & \UseMacro{adv-comp-sa-pso-selfbleu-avg}\pm\UseMacro{adv-comp-sa-pso-selfbleu-std}
 & \UseMacro{adv-comp-sa-pso-pdrcov-avg}\pm\UseMacro{adv-comp-sa-pso-pdrcov-std}\\
\bottomrule
\end{tabular}}
\end{center}
\end{small}
\vspace{\AdvCompTableVSpace}
\end{table*}