### 2023-03-16 paper/icse24/CM/coverage.tex

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs/nlp_coverage.pdf}
    \vspace{-6mm}
    \caption{Coverage results of \tool and \Cklst.}
    \label{fig:coverage}
\end{figure}

\MyPara{Neuron Coverage} 
Figure \ref{fig:coverage} shows the coverage results of \tool and \Cklst test cases. The red line represents \tool coverage and the black line represents \Cklst coverage. Each column in Figure \ref{fig:coverage} represents the results for one \sa model. The first row is the \textit{BoundCov} results and the second row is the \textit{SActCov} results.
We made three observations from the results.
First, for \emph{all} experimental settings (i.e., NLP model and coverage metric), \tool achieves higher coverage than \Cklst. Recall that a higher coverage implies the test cases are more diverse and do not have a similar statistical distribution to the model training data. As a result, a test suite with greater coverage complements the model training data distribution (\ie holdout testing data) better.
For example, for the first NLP model under test, \tool can achieve a higher coverage than \Cklst with only half the number of test cases.
This result confirms that \tool can generate more diverse test cases to complement the holdout dataset for testing NLP models.
Second, as the number of test cases increases, the test suite can achieve better coverage. Such observation is intuitive. However, generating a more extensive test suite is not easy, particularly  for \Cklst, which is a manually template-based approach.
Third, for each NLP model, there is no fixed relationship between \textit{BoundCov} and \textit{SActCov}. In other words, while a test suite may produce higher \textit{BoundCov} for some models, the same test suite may get higher \textit{SActCov} for other NLP models.
Recall that \textit{BoundCov} measures both the upper and lower corner neurons and \textit{SActCov} measures only the upper corner neurons. 
Such observation implies that the upper and lower corner neurons are distributed unevenly, and measuring only one of them is not enough.