### 2023-03-16 paper/icse24/abstract.tex

\begin{abstract}

\Nlp (NLP) has gained widespread adoption in the development 
of diverse real-world applications. However, the black-box nature of the backend neural networks in NLP applications poses a challenge when evaluating their performance, let alone ensuring it. 
Recent research has proposed testing techniques to enhance the trustworthiness of NLP-based applications. 
However, existing works either generate invalid test cases (\ie those with incorrect grammar) or rely on manually designed test cases, which may limit their diversity.
To address this limitation, we propose \tool, an automated testing technique for validating NLP applications based on their \lcs. 
\tool first transforms test cases from existing real-world datasets into seed inputs  that belong to individual \lcs. \tool then generates a sufficient number of test cases by expanding syntax of the seed inputs based on their context-free grammar (CFG). 
In the syntax expansion process, \tool validates that the expanded test cases maintain the same semantic and capability as the seed inputs, thereby automatically  constructing a test oracle and \lc for the expanded test cases.
We evaluate \tool on two widely adopted NLP tasks, \sa and \hsd, to determine its  performance in terms of diversity, effectiveness, and consistency.
The results showed that \tool generates test cases that are at least 88\% more diverse than those produced by \sota techniques, as measured by \selfbleu and \pdr metrics. % two neuron coverage metrics.
Additionally, \tool was found to produce a larger number of NLP model failures.
% \sw{I'm not sure what 7 and 13 were referring to.}
% over 7 \lcs for \sa and 14 \lcs for \hsd.
% Our evaluation also revealed that \tool facilitates the identification of critical failures and the sources of these failures within NLP models. 
% \sw{Double check numbers and update evaluation results when finalized.}

\end{abstract}