### 2023-03-16 paper/icse24/macros.tex

%%% macros.tex
%%% Import packages, define utility commands and define macros
%%% Version 1.2.0

%%%----------
%%% Imports

%%% Text Formatting
% \usepackage{cite}
% \usepackage{amssymb}
% \usepackage{amsfonts} 
% \usepackage{amsmath}
\usepackage[compact]{titlesec}
\usepackage[noend]{algorithmic}
\usepackage{textcomp}
\usepackage{microtype}
\usepackage{xspace}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage[normalem]{ulem} % normalem do not replace \emph to underline
\usepackage{breakcites}
\usepackage{enumitem}
\usepackage{adjustbox}

%%% Reference, Citation and Link
\usepackage{hyperref}
%\usepackage{url}
\usepackage{breakurl}
%\usepackage{cite}

%%% Tables
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{ragged2e}

%%% Plots
\usepackage{caption}
%\usepackage{subcaption}
%\usepackage{subfloat}
%\usepackage{wrapfig}
\usepackage{subfig}

% Code
\usepackage{algorithm}
% \usepackage{algpseudocode} 
\usepackage{listings}

% Tikz Figs
\usepackage{tikz}
%\usepackage{pgf-umlsd} % for flow diagram
\usetikzlibrary{calc}
\usetikzlibrary{shapes.geometric}
\usetikzlibrary{decorations.pathreplacing}
\usetikzlibrary{positioning}
%\usetikzlibrary{shapes}

%%% Miscellaneous
%\usepackage{flushend} % balance the end of double column document
\usepackage{datetime} % for getting current date time

\input{defs-logic} % for logic symbols
\input{defs-code} % for code listings

%%%----------
%%% Utility Commands
\newcommand{\Cong}[1]{\textbf{\color{blue}Cong:{ #1} }}

\newcommand{\XSpace}[1]{}
\newcommand{\XComment}[1]{}
\newcommand{\Fix}[1]{\textcolor{red}{#1}}
%\newcommand{\CR}[2]{\textcolor{blue}{#1}\textcolor{brown}{\sout{#2}}}
\newcommand{\CR}[2]{#1}
\newcommand{\EditAdd}[1]{\textcolor{green}{[#1]}}
\newcommand{\EditRm}[1]{\textcolor{red}{[\sout{#1}]}}
\newcommand{\EditMod}[2]{\textcolor{red}{[\sout{#1}]}\textcolor{green}{[#2]}}
\newcommand{\DefMacro}[2]{\expandafter\newcommand\csname rmk-#1\endcsname{#2}}
\newcommand{\UseMacro}[1]{\csname rmk-#1\endcsname}
\newcommand{\AlgCompTableVSpace}{-5pt}
\newcommand{\CfgStmtTableCaptionVSpace}{-5pt}
\newcommand{\ResultTableCaptionVSpace}{-5pt}
\newcommand{\ReqTableVSpace}{-5pt}
\newcommand{\HsdReqTableVSpace}{-5pt}
\newcommand{\SelfBleuTableVSpace}{-5pt}
\newcommand{\RetrainDebugTableVSpace}{-5pt}
\newcommand{\TestResultsTableVSpace}{-5pt}
\newcommand{\ManualStudyTableVSpace}{-5pt}
\newcommand{\MtnlpCompTableVSpace}{-5pt}
\newcommand{\AdvCompTableVSpace}{-15pt}

\newcommand{\romnum}[1]{\uppercase\expandafter{\romannumeral #1\relax}}

\newcommand{\jl}[1]{\textcolor{teal}{JL: #1}}
\newcommand{\sw}[1]{\textcolor{purple}{Shiyi: #1}}
\newcommand{\wy}[1]{\textcolor{blue}{WY: #1}}
\newcommand{\CM}[1]{\textcolor{blue}{Simin: #1}}
\newcommand{\TODO}[1]{\textcolor{red}{TODO: #1}}

\newcommand{\MyPara}[1]{\vspace{0pt}\noindent\textbf{#1}.}
\newcommand{\MyParaOnly}[1]{\noindent\textbf{#1}}
\newcommand{\reducedstrut}{\vrule width 0pt height .9\ht\strutbox depth .9\dp\strutbox\relax}
\newcommand{\InputWithSpace}[1]{\bgroup\def\arraystretch{1.2}\input{#1}\egroup}
\newcommand{\Code}[1]{{\ifmmode{\mathtt{#1}}\else$\mathtt{#1}$\fi}}
\newcommand{\CodeIn}[1]{{\ifmmode{\mathtt{#1}}\else$\mathtt{#1}$\fi}}
\newcommand{\ColorBack}[1]{%
  \begingroup
  \setlength{\fboxsep}{0pt}%
  \colorbox{purple!20}{\reducedstrut#1\/}%
  \endgroup
}

% for table headers
\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}
\newcolumntype{R}[1]{>{\RaggedLeft\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\RaggedRight\arraybackslash}p{#1}}
\def\clap#1{\hbox to 0pt{\hss#1\hss}}

%%% TEXT
\newcommand{\Title}{Programming Language Representation with Semantic-level Structure}

\newcommand{\TreeSitterURL}{https://github.com/tree-sitter/tree-sitter}
\newcommand{\Dxp}{DeepXplore}
\newcommand{\Denas}{DENAS}

\newcommand{\tool}{S$^2$LCT\xspace}

% Sorted alphabetically, ignoring case
\newcommand{\Model}{Auto-CHECKLIST\xspace}
\newcommand{\Lc}{Linguistic capability\xspace}
\newcommand{\lc}{linguistic capability\xspace}
\newcommand{\lcs}{linguistic capabilities\xspace}
\newcommand{\Cfg}{Context-free grammar\xspace}
\newcommand{\cfg}{context-free grammar\xspace}
\newcommand{\Pcfg}{Probabilistic context-free grammar\xspace}
\newcommand{\pcfg}{probabilistic context-free grammar\xspace}
\newcommand{\lm}{langugae model\xspace}
\newcommand{\lms}{langugae models\xspace}
\newcommand{\Lm}{Langugae Model\xspace}
\newcommand{\dl}{deep learning\xspace}
\newcommand{\Dl}{Deep Learning\xspace}
\newcommand{\Hsd}{Hate speech detection\xspace}
\newcommand{\hsd}{hate speech detection\xspace}
\newcommand{\ml}{machine learning\xspace}
\newcommand{\ML}{Machine Learning\xspace}
\newcommand{\nl}{natural language\xspace}
\newcommand{\Nl}{Natural language\xspace}
\newcommand{\nlp}{natural language processing\xspace}
\newcommand{\Nlp}{Natural language processing\xspace}
\newcommand{\Sa}{Sentiment analysis\xspace}
\newcommand{\sa}{sentiment analysis\xspace}
\newcommand{\se}{software engineering\xspace}
\newcommand{\Se}{Software Engineering\xspace}
\newcommand{\sota}{state-of-the-art\xspace}
\newcommand{\sent}{sentence\xspace}
\newcommand{\sents}{sentences\xspace}
\newcommand{\Sent}{Sentence\xspace}
%\newcommand{\Sw}{Software\xspace}
%\newcommand{\sw}{software\xspace}
\newcommand{\pos}{part-of-speech\xspace}
\newcommand{\ho}{holdout\xspace}
\newcommand{\ie}{i.e.\xspace}
\newcommand{\eg}{e.g.\xspace}
\newcommand{\etal}{et al.\xspace}

\newcommand{\Req}{Requirement\xspace}
\newcommand{\req}{requirement\xspace}
\newcommand{\Prodr}{Production rule\xspace}
\newcommand{\prodr}{production rule\xspace}
\newcommand{\prodrs}{production rules\xspace}
\newcommand{\lhs}{left-hand side\xspace}
\newcommand{\rhs}{right-hand side\xspace}
\newcommand{\mask}{\{MASK\}}
\newcommand{\SareqExOne}{``Short sentences with neutral adjectives and nouns''\xspace}
\newcommand{\SareqExTwo}{``Negated positive with neutral content in the middle''\xspace}
\newcommand{\reqs}{requirements\xspace}
\newcommand{\bl}{baseline\xspace}
\newcommand{\Bl}{Baseline\xspace}
\newcommand{\bls}{baselines\xspace}
\newcommand{\Bls}{Baselines\xspace}
\newcommand{\Pstv}{Positive\xspace}
\newcommand{\pstv}{positive\xspace}
\newcommand{\Neu}{Neutral\xspace}
\newcommand{\neu}{neutral\xspace}
\newcommand{\Ngtv}{Negative\xspace}
\newcommand{\ngtv}{negative\xspace}
\newcommand{\Noun}{Noun\xspace}
\newcommand{\noun}{noun\xspace}
\newcommand{\nns}{nouns\xspace}
\newcommand{\Vb}{Verb\xspace}
\newcommand{\vb}{verb\xspace}
\newcommand{\vbs}{verbs\xspace}
\newcommand{\Adj}{Adjective\xspace}
\newcommand{\adj}{adjective\xspace}
\newcommand{\adjs}{adjectives\xspace}
\newcommand{\bleu}{BLEU\xspace}
\newcommand{\ph}{non-terminal\xspace}
\newcommand{\phs}{non-terminals\xspace}
\newcommand{\Ph}{Non-terminal\xspace}
\newcommand{\Phs}{Non-terminals\xspace}
\newcommand{\selfbleu}{Self-BLEU\xspace}
\newcommand{\Pdr}{Production Rule Coverage\xspace}
\newcommand{\pdr}{production rule coverage\xspace}

\newcommand{\Nn}{neural network\xspace}
\newcommand{\nn}{neural network\xspace}
\newcommand{\Bert}{BERT-base\xspace}
\newcommand{\BertAcc}{92.43\%\xspace}
\newcommand{\Roberta}{RoBERTa-base\xspace}
\newcommand{\RobertaAcc}{94.04\%\xspace}
\newcommand{\Dbert}{DistilBERT-base\xspace}
\newcommand{\DbertAcc}{91.3\%\xspace}
\newcommand{\hsdBert}{dehate-BERT\xspace}
\newcommand{\hsdRoBerta}{tweeter-RoBERTa\xspace}
\newcommand{\Cklst}{CHECKLIST\xspace}
\newcommand{\hck}{Hatecheck\xspace}
\newcommand{\mtnlp}{MT-NLP\xspace}
\newcommand{\Trb}{Treebank\xspace}
\newcommand{\trb}{treebank\xspace}
\newcommand{\Wsj}{Wall Street Journal\xspace}
\newcommand{\Nltk}{NLTK\xspace}
\newcommand{\Spacy}{Spacy\xspace}
\newcommand{\Wrdnt}{WordNet\xspace}
\newcommand{\Swn}{SentiWordNet\xspace}
\newcommand{\Sst}{SST\xspace}
\newcommand{\Hatexp}{HateXplain\xspace}
\newcommand{\spacy}{SpaCy\xspace}
\newcommand{\bertsamodel}{textattack/bert-base-uncased-SST-2\xspace}
\newcommand{\robertasamodel}{textattack/bert-base-uncased-SST-2\xspace}
\newcommand{\disbertsamodel}{distilbert-base-uncased\xspace}
\newcommand{\berthsdmodel}{Hate-speech-CNERG/dehatebert-mono-english\xspace}
\newcommand{\robertahsdmodel}{cardiffnlp/twitter-roberta-base-hate\xspace}
\newcommand{\Ptf}{Pass-to-Fail\xspace}

%% Table
\newcommand{\tStatement}{\textbf{Statement Types}}
\newcommand{\tModel}{\textbf{Models}}
\newcommand{\tMetric}{\textbf{Metrics}}
\newcommand{\tLc}{\textbf{Linguistic capability}}
\newcommand{\tRules}{\textbf{Search rule and transformation template}}
\newcommand{\tOurNumData}{\textbf{\sharp~Data (\tool)}}
\newcommand{\tChecklistNumData}{\textbf{\sharp~Data (\Cklst)}}
\newcommand{\tOurSelfBleuScore}{\textbf{\selfbleu score (\tool)}}
\newcommand{\tChecklistSelfBleuScore}{\textbf{\selfbleu score (\Cklst)}}
\newcommand{\tApproach}{\textbf{Approach}}
\newcommand{\tRetrainlc}{\textbf{RetrainLC}}
\newcommand{\tEvallc}{\textbf{EvalLC}}
\newcommand{\tNumfailpass}{\textbf{\#Fail2Pass}}
\newcommand{\tNumfailbeforeretrain}{\textbf{\#Fail(BeforeRetrain)}}
\newcommand{\tNumpassbeforeretrain}{\textbf{\#Pass(BeforeRetrain)}}
\newcommand{\tNumfailafterretrain}{\textbf{\#Fail(AfterRetrain)}}
\newcommand{\tNumpassafterretrain}{\textbf{\#Pass(AfterRetrain)}}
\newcommand{\tTask}{\textbf{Task}}
\newcommand{\tSentType}{\textbf{Type}}
\newcommand{\tNumTestCases}{\textbf{\#TC}}
\newcommand{\tAvgLabelCons}{\textbf{LabelCons}}
\newcommand{\tAvgLCRel}{\textbf{LCRel}}
\newcommand{\tAvgExpVal}{\textbf{ExpValidity}}
\newcommand{\tNumBlSents}{\textbf{Cklst \#TCs}}
\newcommand{\tNumBlFail}{\textbf{Cklst \#Fails[\%]}}
\newcommand{\tNumBlTcs}{\textbf{Cklst \#TCs}}
\newcommand{\tNumSeeds}{\textbf{\tool\#Seeds}}
\newcommand{\tNumExps}{\textbf{\tool\#Exps}}
\newcommand{\tNumFail}{\textbf{\tool/Cklst \#Fail}}
\newcommand{\tFailRate}{\textbf{\tool/Cklst Fail rate[\%]}}
\newcommand{\tNumPasstoFail}{\textbf{\tool\#PassToFail}}
\newcommand{\tNumHsBlTcs}{\textbf{Htck \#TCs}}
\newcommand{\tNumHsFail}{\textbf{\tool/Htck \#Fail}}
\newcommand{\tHsFailRate}{\textbf{\tool/Htck Fail rate[\%]}}
\newcommand{\tNummut}{\textbf{\#Gen}}
\newcommand{\tSelfbleu}{\textbf{\selfbleu}}
\newcommand{\tPdrcov}{\textbf{PDRCov}}

%% Fig Caption
\newcommand{\OverviewFigCaption}{Overview of \tool.\label{fig:overview}\vspace{-5pt}}
\newcommand{\SearchRequirementExampleFigCaption}{Search requirement
on the linguistic capability of \SareqExOne.\label{fig:SearchReqEx}\vspace{-5pt}}
\newcommand{\TransformRequirementExampleSubFigCaption}{Transform requirement.\label{fig:TransformReqSubEx}\vspace{-5pt}}
\newcommand{\TransformTemplateExampleSubFigCaption}{Transform template.\label{fig:TransformTempSubEx}\vspace{-5pt}}
\newcommand{\TransformRequirementExampleFigCaption}{Transform
requirement (\ref{fig:TransformReqSubEx}) and its template
(\ref{fig:TransformTempSubEx}) on the linguistic capability of \SareqExTwo.\label{fig:TransformReqEx}\vspace{-5pt}}
\newcommand{\ProdDiffAlgCaption}{Pseudocode of Production differentiation\label{code:ProdDiffAlg}\vspace{-5pt}}
\newcommand{\RunningExCaption}{Example of sentence expansion.\label{fig:ExpEx}\vspace{-5pt}}
\newcommand{\CklstTemplateFigCaption}{
% Example of \Cklst templates on the \lc of \SareqExThree.
Example of CHECKLIST template\label{code:TempEx}\vspace{-5pt}}
\newcommand{\PdrSelfbleuFigCaption}{Results of \selfbleu(left) and \Pdr(right) of \tool and capability-based testing baselines for \sa (SA) and \hsd (HS). Use of only \tool seed \sents and all \tool \sents are denoted as SEED and SEED+EXP respectively.\label{fig:PdrSelfbleu}\vspace{-10pt}}
% \newcommand{\PdrBarplotFigCaption}{Results of \pdr between \tool (seed and expanded cases) and \Cklst test cases for \sa.\label{fig:PdrBarplot}\vspace{-5pt}}
\newcommand{\PdrSelfbleuBarplotFigCaption}{Results of \selfbleu(left) and \Pdr(right) between original \sents of capability-based testing baselines and \tool generated \sents from the original \sents.\label{fig:PdrSelfbleuBarplot}\vspace{-5pt}}
% \newcommand{\PdrBarplotHsFigCaption}{Results of \pdr between \tool (seed and expanded cases) and \Cklst test cases for \hsd.\label{fig:HsPdrBarplot}\vspace{-5pt}}
\newcommand{\PdrBarplotHsFigCaption}{Results of \pdr between \tool (\hck and its expanded cases) and \hck test cases for \hsd.\label{fig:HsPdrBarplot}\vspace{-5pt}}
\newcommand{\SelfbleuBarplotFigCaption}{Results of \selfbleu between \tool (\Cklst and its expanded cases) and \Cklst test cases for \sa.\label{fig:SbBarplot}\vspace{-5pt}}
\newcommand{\SelfbleuBarplotHsFigCaption}{Results of \selfbleu between \tool (\hck and its expanded cases) and \hck test cases for \hsd.\label{fig:HsSbBarplot}\vspace{-5pt}}
\newcommand{\PdrFigCaption}{Results of \pdr between \tool (seed and expanded cases) and \Cklst.\label{fig:Pdr}\vspace{-5pt}}

\newcommand{\NumFailModelsSubFigCaption}{Number of fail cases from \tool test cases over numbers of seeds.\label{fig:NumFail}\vspace{-5pt}}
\newcommand{\FailRateModelsSubFigCaption}{Number of Pass-to-Fail cases from \tool test cases over numbers of seeds.\label{fig:FailRate}\vspace{-5pt}}
\newcommand{\FailModelsFigCaption}{Number of failure (left) and \Ptf(right) results of NLP models on \tool and \Cklst test cases over numbers of seeds.\label{fig:FailModels}\vspace{-5pt}}

\newcommand{\SareqExThree}{``Sentiment change over time, present should prevail"}

%% Table Caption
\newcommand{\ReqTableCaption}{Search-based and enumerate predicates for two linguistic
capabilities of \sa task.\label{tab:specification}\vspace{\ReqTableVSpace}}
\newcommand{\HsdReqTableCaption}{Search rules and transformation
templates for linguistic
capabilities of \hsd task. The \textit{slur}, \textit{profanity} in LC7-LC10 are the collections of terms that express slur and profanity. The \textit{identity} in and LC11-LC12 and LC17-18 is a list of names that used to describe social groups. In this work, we use the terms obtained from \hck~\cite{rottger2020hatecheck}.  \label{tab:hsd-specification}\vspace{\HsdReqTableVSpace}}
\newcommand{\ResultTableCaption}{Results on \Nlp tasks
of \Ccd, \Ctr, \Cref respectively.\label{table:Result}\vspace{-5pt}}
\newcommand{\SelfBleuTableCaption}{\selfbleu scores over testcases
from \tool and \Cklst}
\newcommand{\RetrainDebugTableCaption}{Result of retraining on
each \lc and the evaluation of the retrained model.\label{table:RetrainDebug}}
\newcommand{\ManualStudyTableCaption}{
Consistency Results
% Manual
% study results of label consistency, \lc relevancy and expansion validity of \tool 
% test cases.
% compared to human annotations for \sa and \hsd. \Sa and \hsd are denoted as SA and HSD respectively.
\label{table:ManualStudy}}
\newcommand{\TestResultsTableCaption}{Results of \Bert, \Roberta and
\Dbert \sa models on \tool test cases using 50 seeds. \Bert, \Roberta and
\Dbert models are denoted as BERT, RoBERTa and dstBERT, respectively.\label{table:TestResult}}
\newcommand{\TestResultsAllTableCaption}{Results of \Bert, \Roberta and
\Dbert \sa models on \tool test cases using all seeds. \Bert, \Roberta and
\Dbert models are denoted as BERT, RoBERTa and dstBERT, respectively.\label{table:TestResultAll}}
\newcommand{\TestResultsHsTableCaption}{Results of \hsdBert and \hsdRoBerta \hsd models on \tool test cases using all seeds. \hsdBert and \hsdRoBerta are denoted as BERT, RoBERTa respectively.\label{table:TestResultHsd}}
\newcommand{\MtnlpCompTableCaption}{Results of \sent generation from 100 seeds using \mtnlp and \tool expansion phase.
%using  100 sample size over 5 trials. \Sa and \hsd are denoted as SA and HSD respectively. The number of generated sentences, \selfbleu and \pdr are denoted as \tNummut, \selfbleu and PDRCov respectively. 
\label{table:MtnlpComp}}

\newcommand{\AdvCompTableCaption}{Results of \sent generation from 50 seeds using \tool expansion and adversarial text generation \bls. 
% and  using 50 \sa seeds, statistics of \selfbleu and \pdr using 50 sample size over 5 trials. The number of generated sentences, \selfbleu and \pdr are denoted as \tNummut, \selfbleu and PDRCov respectively. 
\label{table:AdvComp}}

%%% NUMBRERS
\input{tables/lc-requirement-numbers}
\input{tables/lc-hsd-requirement-numbers}
%% \input{tables/selfbleu-numbers}
%% \input{tables/retrain-debug-numbers}
\input{tables/manual-study-numbers}
\input{tables/test-results-numbers}
\input{tables/test-results-3-50-numbers}
\input{tables/test-results-3-100-numbers}
\input{tables/test-results-3-200-numbers}
\input{tables/test-results-all-numbers}
\input{tables/test-results-hs-all-numbers}
\input{tables/mtnlp-comp-numbers}
\input{tables/adv-comp-numbers}