### 2023-03-16 paper/icse24/related.tex

\section{Related Work}

In addition to the capability-based testing works discussed in Section \ref{sec:background}, we review other related works in this section.

% In addition to the existing capability-based testing works that are discussed in Section \ref{sec:background}, our work is related to works that (1) present new metrics and methodologies for NLP testing, (2) automated text augmentation, and (3) explanation and debugging of NLP model results.

\MyPara{NLP Algorithms \& Applications} Deep neural networks (DNNs) have significantly improved various natural language processing (NLP) applications \cite{mikolov2013efficient, sutskever2014sequence, vaswani2017attention, devlin2018bert, radford2018improving, gehring2017convolutional, pennington2014glove, joulin2016bag, schuster1997bidirectional, chen2022nicgslowdown}, including reading comprehension, hate speech detection, and machine translation.
For instance, Word2vec \cite{mikolov2013efficient} distributes the semantic of words into numeric vectors, which are then utilized to train neural networks for classification tasks. 
Meanwhile, Seq2Seq \cite{sutskever2014sequence} presents an encoder-decoder neural network architecture that has been widely adopted for modeling the sequence generation task, particularly in machine translation applications.
In recent years, Google \cite{vaswani2017attention} has introduced the attention mechanism, which, when combined with Seq2Seq, can greatly enhance the accuracy of the generated texts.

\MyPara{Machine Learning Testing \& NLP Testing} Machine learning has shown great potential in various real-world applications. Nonetheless, despite the high accuracy rates of ML models, there have been instances where ML models can generate inferior results, leading to fatal accidents \cite{lambert2016understanding, levin2018tesla}. Therefore, researchers have developed a series of techniques \cite{pei2017deepxplore, tian2018deeptest, ma2018deepgauge, hendrycks2021natural, du2019deepstellar, gambi2019automatically, kim2019guiding, ma2018deepmutation, pham2019cradle, xie2019deephunter, chen2022nicgslowdown, zhang2020machine, chen2022deepperform, token_nlp2, character_nlp1, character_nlp2, character_nlp3, nmt_se4} to test ML-based applications. 
% For example, DeepXplore \cite{pei2017deepxplore} developed a differential testing framework to generate test inputs that can expose inconsistencies between multiple (ML) models, and they introduced neuron coverage as a metric to measure the diversity of the generated test suite.
% Along this direction, DeepRoad \cite{zhang2018deeproad} proposed a GAN-based method for automatically transforming the style of scenes for testing autonomous driving systems. 
%Meanwhile, DeepGauge \cite{ma2018deepgauge} proposed several neuron coverage metrics to test ML models at multiple levels of granularity.
In recent years, researchers have investigated the occurrence of bugs produced by neural networks in NLP applications \cite{nmt_se1, nmt_se2, nmt_se3, token_nlp3, token_nlp1, token_nlp2, character_nlp1, character_nlp2, character_nlp3, nmt_se4, li2018textbugger, morris2020textattack, chen2022nmtsloth, sentence_nlp1, maIJCAI2020mtnlp}, inspired by the work on adversarial examples in computer vision.
% TestBugger \cite{li2018textbugger} proposes a gradient-guided approach to generate test inputs for identifying bugs in NLP models used for classification tasks, while SIT \cite{pinjia2020structinvtestingnmt} proposes a method to detect errors in neural machine translation (NMT) systems by constructing the structure invariant as an oracle. 
% In contrast, NMTSloth \cite{chen2022nmtsloth} focuses on generating test inputs that can expose performance bugs in NMT systems.
Our approach differs from existing work in that we concentrate on  testing the linguistic capabilities of NLP applications in an automatic manner, a topic that has yet to be explored.

% \MyPara{Linguistic Capability Evaluation} 
% \sw{This paragraph (1) is incomplete (missing hatecheck), and (2) may not be necessary (the intro/motivation should have already covered these). So I suggest removing this paragraph but make sure these works (including hatecheck) have been well discussed in intro/motivation sections.} 
% Wang \textit{\etal}~\cite{wang2018glue, wang2019superglue} proposed
% multiple diagnostic datasets to evaluate NLP models 
% % . These datasets
% % evaluate NLP model's ability to understand input sentence
% via natural language inference problems. More recently, \Cklst proposed an evaluation
% method of input-output behavior defined as \lcs. \Cklst generates
% behavior-guided inputs for validating the
% behaviors.~\cite{marcoACL2020checklist}. Unlike prior work that relies on manual
% test case generation, we used structural information in text to
% generate test cases automatically.