### 2023-03-16 paper/icse24/tables/adv-comp-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{adv-comp-sa-model-under-test}{textattack/bert-base-uncased-SST-2}
\DefMacro{adv-comp-sa-alzantot-num-adv}{20}
\DefMacro{adv-comp-sa-alzantot-num-samples}{20}
\DefMacro{adv-comp-sa-alzantot-selfbleu-med}{\textbf{0.373}}
\DefMacro{adv-comp-sa-alzantot-selfbleu-avg}{\textbf{0.373}}
\DefMacro{adv-comp-sa-alzantot-selfbleu-std}{\textbf{0.0}}
\DefMacro{adv-comp-sa-alzantot-pdrcov-med}{170}
\DefMacro{adv-comp-sa-alzantot-pdrcov-avg}{170.0}
\DefMacro{adv-comp-sa-alzantot-pdrcov-std}{0.0}
\DefMacro{adv-comp-sa-bert-attack-num-adv}{25}
\DefMacro{adv-comp-sa-bert-attack-num-samples}{25}
\DefMacro{adv-comp-sa-bert-attack-selfbleu-med}{0.438}
\DefMacro{adv-comp-sa-bert-attack-selfbleu-avg}{0.438}
\DefMacro{adv-comp-sa-bert-attack-selfbleu-std}{0.0}
\DefMacro{adv-comp-sa-bert-attack-pdrcov-med}{178}
\DefMacro{adv-comp-sa-bert-attack-pdrcov-avg}{178.0}
\DefMacro{adv-comp-sa-bert-attack-pdrcov-std}{0.0}
\DefMacro{adv-comp-sa-pso-num-adv}{25}
\DefMacro{adv-comp-sa-pso-num-samples}{25}
\DefMacro{adv-comp-sa-pso-selfbleu-med}{0.411}
\DefMacro{adv-comp-sa-pso-selfbleu-avg}{0.411}
\DefMacro{adv-comp-sa-pso-selfbleu-std}{0.0}
\DefMacro{adv-comp-sa-pso-pdrcov-med}{178}
\DefMacro{adv-comp-sa-pso-pdrcov-avg}{178.0}
\DefMacro{adv-comp-sa-pso-pdrcov-std}{0.0}
\DefMacro{adv-comp-sa-ours-num-adv}{\textbf{323}}
\DefMacro{adv-comp-sa-ours-num-samples}{50}
\DefMacro{adv-comp-sa-ours-selfbleu-med}{0.437}
\DefMacro{adv-comp-sa-ours-selfbleu-avg}{0.435}
\DefMacro{adv-comp-sa-ours-selfbleu-std}{0.005}
\DefMacro{adv-comp-sa-ours-pdrcov-med}{\textbf{264}}
\DefMacro{adv-comp-sa-ours-pdrcov-avg}{\textbf{262.0}}
\DefMacro{adv-comp-sa-ours-pdrcov-std}{\textbf{2.739}}