### 2023-03-16 paper/icse24/tables/lc-hsd-requirement-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{hsd_lc_1_desc}{Expression of strong negative emotions (explicit)}
\DefMacro{hsd_lc_1_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_1_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_2_desc}{Description using very negative attributes (explicit)}
\DefMacro{hsd_lc_2_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_2_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_3_desc}{Dehumanisation (explicit)}
\DefMacro{hsd_lc_3_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_3_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_4_desc}{Implicit derogation}
\DefMacro{hsd_lc_4_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_4_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_5_desc}{Direct threat}
\DefMacro{hsd_lc_5_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_5_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_6_desc}{Threat as normative statement}
\DefMacro{hsd_lc_6_search}{\textbf{Search}~\textit{hatecheck}}
\DefMacro{hsd_lc_6_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_7_desc}{Hate expressed using slur}
\DefMacro{hsd_lc_7_search}{\textbf{Search}~seed=\{include: \textit{slur}; label: toxic\}}
\DefMacro{hsd_lc_7_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_8_desc}{Non-hateful use of slur}
\DefMacro{hsd_lc_8_search}{\textbf{Search}~seed=\{include: \textit{slur}; label: non-toxic\}}
\DefMacro{hsd_lc_8_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_9_desc}{Hate expressed using profanity}
\DefMacro{hsd_lc_9_search}{\textbf{Search}~seed=\{include: \textit{profanity}; label: toxic\}}
\DefMacro{hsd_lc_9_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_10_desc}{Non-Hateful use of profanity}
\DefMacro{hsd_lc_10_search}{\textbf{Search}~seed=\{include: \textit{profanity}; label: non-toxic\}}
\DefMacro{hsd_lc_10_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_11_desc}{Hate expressed through reference in subsequent clauses}
\DefMacro{hsd_lc_11_search}{\textbf{Search}~toxic\_sent=\{include: \textit{identity}; label: toxic\}, non\_toxic\_sent=\{include: \textit{identity}; label: non-toxic\}}
\DefMacro{hsd_lc_11_transform}{\textbf{Transform}~seed=\{['people say', 'I heard that']+[toxic\_sent]+['and']+['i hate them too', 'i also hate them']\} $|$ \{['people say', 'I heard that']+[toxic\_sent]+['but']+['i do not hate them', 'i like them']\} $|$ \{['people say', 'I heard that']+[non\_toxic\_sent]+['but']+['i hate them', 'i dislike them']\} $|$ \{['people say', 'I heard that']+[non\_toxic\_sent]+['and']+['i also do not hate them', 'i do not hate them too']\}}
\DefMacro{hsd_lc_12_desc}{Hate expressed through reference in subsequent sentences}
\DefMacro{hsd_lc_12_search}{\textbf{Search}~toxic\_sent=\{include: \textit{identity}; label: toxic\}, non\_toxic\_sent=\{include: \textit{identity}; label: non-toxic\}}
\DefMacro{hsd_lc_12_transform}{\textbf{Transform}~seed=\{['people say', 'I heard that']+[toxic\_sent]+['.']+['i hate them too', 'i also hate them']\} $|$ \{['people say', 'I heard that']+[toxic\_sent]+['.']+['however i do not hate them', 'regardless i like them']\} $|$ \{['people say', 'I heard that']+[non\_toxic\_sent]+['.']+['however i hate them', 'i dislike them though']\} $|$ \{['people say', 'I heard that']+[non\_toxic\_sent]+['.']+['i also do not hate them', 'i do not hate them too']\}}
\DefMacro{hsd_lc_13_desc}{Hate expressed using negated positive statement}
\DefMacro{hsd_lc_13_search}{\textbf{Search}~non\_toxic\_sent=\{label: non-toxic\}}
\DefMacro{hsd_lc_13_transform}{\textbf{Transform}~seed=\{['I do not think that', 'I do not agree with that', 'I do not believe that']+[non\_toxic\_sent]\}}
\DefMacro{hsd_lc_14_desc}{Non-hate expressed using negated hateful statement}
\DefMacro{hsd_lc_14_search}{\textbf{Search}~toxic\_sent=\{label: toxic\}}
\DefMacro{hsd_lc_14_transform}{\textbf{Transform}~seed=\{['I do not think that', 'I do not agree with that', 'I do not believe that']+[toxic\_sent]\}}
\DefMacro{hsd_lc_15_desc}{Hate phrased as a question}
\DefMacro{hsd_lc_15_search}{\textbf{Search}~toxic\_sent=\{label: toxic\}}
\DefMacro{hsd_lc_15_transform}{\textbf{Transform}~seed=\{['why do you agree with that', 'why do you like to say that']+[toxic\_sent]+['?']\}}
\DefMacro{hsd_lc_16_desc}{Hate phrased as a opinion}
\DefMacro{hsd_lc_16_search}{\textbf{Search}~toxic\_sent=\{label: toxic\}}
\DefMacro{hsd_lc_16_transform}{\textbf{Transform}~seed=\{['i think that', 'i believe that']+[toxic\_sent]\}}
\DefMacro{hsd_lc_17_desc}{Neutral statements using protected group identifiers}
\DefMacro{hsd_lc_17_search}{\textbf{Search}~seed=\{include: \textit{identity}; exclude: pos adjs \& neg adjs \& pos nouns \& neg nouns \& pos verbs \& neg verbs; label: non-toxic\}}
\DefMacro{hsd_lc_17_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_18_desc}{Positive statements using protected group identifiers}
\DefMacro{hsd_lc_18_search}{\textbf{Search}~seed=\{include: \textit{identity}; exclude: neg adjs \& neg nouns \& neg verbs; label: non-toxic\}}
\DefMacro{hsd_lc_18_transform}{\textbf{Transform}~N/A}
\DefMacro{hsd_lc_19_desc}{Denouncements of hate that quote it}
\DefMacro{hsd_lc_19_search}{\textbf{Search}~toxic\_sent=\{label: toxic\}}
\DefMacro{hsd_lc_19_transform}{\textbf{Transform}~seed=\{['statements like']+[toxic\_sent]+['are hurtful', 'are harmful', 'are deleterious']\} $|$ \{['you should stop saying like']+[toxic\_sent]\}}