### 2023-03-16 paper/icse24/tables/lc-requirement-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{lc_1_desc}{Short sentences with neutral adjectives and nouns}
\DefMacro{lc_1_search}{\textbf{Search-based:}~\textbf{hasattr}(length,<10)\land
\textbf{hasattr}(label,neutral)\land(\textbf{contain}(tag\_of\_pos,neutral\_adj)\lor
\textbf{contain}(tag\_of\_pos,neutral\_noun))\land~\neg(\textbf{contain}(tag\_of\_pos,positive\_adj)\lor
\textbf{contain}(tag\_of\_pos,positive\_noun)\lor~\textbf{contain}(tag\_of\_pos,negative\_adj)\lor
\textbf{contain}(tag\_of\_pos,negative\_noun))}
\DefMacro{lc_1_transform}{\textbf{Enumerative:}~N/A}
\DefMacro{lc_2_desc}{Short sentences with sentiment-laden adjectives}
\DefMacro{lc_2_search}{\textbf{Search-based:}~\textbf{hasattr}(length,$<$10)$\land$((\textbf{contain}
(pos,positive$\_$adj)$\lor$\textbf{contain}(pos,positive$\_$noun)$\land$\textbf{hasattr}(label,positive))$\lor$(\textbf{contain}(pos,negative$\_$adj)$\lor$\textbf{contain}(pos,negative$\_$noun)$\land$\textbf{hasattr}(label,negative)))}
\DefMacro{lc_2_transform}{\textbf{Enumerative:}~N/A}
\DefMacro{lc_3_desc}{Sentiment change over time, present should prevail}
\DefMacro{lc_3_search}{\textbf{Search-based:}~pos\_sent\rightarrow\textbf{hasattr}(length,<20)\land\textbf{hasattr}(label,positive), neg\_sent\rightarrow\textbf{hasattr}(length,<20)\land\textbf{hasattr}(label,negative)}
\DefMacro{lc_3_transform}{
\textbf{Enumerative:}~E1=\textbf{combine}(prefix,sent,infix,postfix) where prefix=(Previously, I used to like it saying that|Last time, I agreed with saying that|I liked it much as to say that), sent=pos\_sent, infix=(but|although|on the other hand), postfix=(now I don't like it.|now I hate it.)

E2=\textbf{combine}(prefix,sent,infix,postfix) where prefix=(I used to disagree with saying that|Last time, I didn't like it saying that|I hated it much as to say that), sent=neg\_sent, infix=(but|although|on the other hand), postfix=(now I like it.)}
\DefMacro{lc_4_desc}{Negated negative should be positive or neutral}
\DefMacro{lc_4_search}{\textbf{Search-based:}~\textbf{hasattr}(word,`(This|That|These|Those)~(is|are)')\land\textbf{hasattr}(label,negative)}
\DefMacro{lc_4_transform}{\textbf{Transform:}~\textbf{replace}(is->is not,isn't'|are->are not,aren't)}
\DefMacro{lc_5_desc}{Negated neutral should still be neutral}
\DefMacro{lc_5_search}{\textbf{Search-based:}~\textbf{hasattr}(word,`(This|That|These|Those)~(is|are)')\land\textbf{hasattr}(label,neutral)}
\DefMacro{lc_5_transform}{\textbf{Transform:}~\textbf{replace}(is->is not,isn't'|are->are not,aren't)}
\DefMacro{lc_6_desc}{Negation of negative at the end, should be positive or neutral}
\DefMacro{lc_6_search}{\textbf{Search-based:}~$neg\_sent\rightarrow\textbf{hasattr}(label,negative)$}
\DefMacro{lc_6_transform}{\textbf{Enumerative:}~\textbf{combine}(prefix,sent,postfix) where prefix=(I agreed that$|$I thought that), sent=neg$\_$sent, postfix=(but it wasn't|but I didn't)}
\DefMacro{lc_7_desc}{Negated positive with neutral content in the middle}
\DefMacro{lc_7_search}{\textbf{Search-based:}~pos\_sent\rightarrow\textbf{hasattr}(length,<20)\land\textbf{hasattr}(label,positive),
neutral\_sent\rightarrow\textbf{hasattr}(length,<20)\land\textbf{hasattr}(label,neutral)}
\DefMacro{lc_7_transform}{
\textbf{Enumerative:}~\textbf{combine}(prefix,sent,infix,postfix) where prefix=(I$\,$wouldn't$\,$say,|I$\,$do$\,$not$\,$think,|I$\,$don't$\,$agree$\,$with,), sent=neutral\_sent, infix=`,', postfix=pos\_sent}
\DefMacro{lc_8_desc}{Author sentiment is more important than of others}
\DefMacro{lc_8_search}{\textbf{Search-based:}~pos\_sent\rightarrow\textbf{hasattr}(label,positive), neg\_sent\rightarrow\textbf{hasattr}(label,negative)}
\DefMacro{lc_8_transform}{\textbf{Enumerative:}~T1=\textbf{combine}(prefix,sent,infix,postfix) where prefix=(Some people think that|Many people agree with that|They think that|You agree with that), sent=pos\_sent, infix=(but I think that), postfix=neg\_sent,

T2=\textbf{combine}(prefix,sent,infix,postfix) where prefix=(Some people think that|Many people agree with that|They think that|You agree with that), sent=neg\_sent, infix=(but I think that), postfix=pos\_sent}
\DefMacro{lc_9_desc}{Parsing sentiment in (question, yes) form}
\DefMacro{lc_9_search}{\textbf{Search-based:}~pos\_sent\rightarrow\textbf{hasattr}(label,positive), neg\_sent\rightarrow~\textbf{hasattr}(label,negative)}
\DefMacro{lc_9_transform}{\textbf{Enumerative:}~\textbf{combine}(prefix,sent,postfix) where prefix=(Do I think that|Do I agree that), sent=(pos\_sent|neg\_sent), postfix=`? yes'}
\DefMacro{lc_10_desc}{Parsing sentiment in (question, no) form}
\DefMacro{lc_10_search}{\textbf{Search-based:}~pos\_sent\rightarrow\textbf{hasattr}(label,positive), neg\_sent\rightarrow~\textbf{hasattr}(label,negative)}
\DefMacro{lc_10_transform}{\textbf{Enumerative:}~\textbf{combine}(prefix,sent,postfix) where prefix=(Do I think that|Do I agree that), sent=(pos\_sent|neg\_sent), postfix=`? no'}