### 2023-03-16 paper/icse24/tables/manual-study-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{manual-study-sa-seed-num-sents}{100}
\DefMacro{manual-study-sa-seed-label-consistency}{0.83}
\DefMacro{manual-study-sa-seed-lc-relevancy}{0.92}
\DefMacro{manual-study-sa-seed-validity}{-}
\DefMacro{manual-study-sa-exp-num-sents}{100}
\DefMacro{manual-study-sa-exp-label-consistency}{0.83}
\DefMacro{manual-study-sa-exp-lc-relevancy}{0.92}
\DefMacro{manual-study-sa-exp-validity}{1.00}
\DefMacro{manual-study-hs-seed-num-sents}{100}
\DefMacro{manual-study-hs-seed-label-consistency}{0.80}
\DefMacro{manual-study-hs-seed-lc-relevancy}{0.84}
\DefMacro{manual-study-hs-seed-validity}{-}
\DefMacro{manual-study-hs-exp-num-sents}{100}
\DefMacro{manual-study-hs-exp-label-consistency}{0.79}
\DefMacro{manual-study-hs-exp-lc-relevancy}{0.84}
\DefMacro{manual-study-hs-exp-validity}{0.97}