### 2023-03-16 paper/icse24/tables/model.tex

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{The NLP model used in our evaluation}
  \vspace{-3mm}
  \resizebox{0.45\textwidth}{!}{
    \begin{tabular}{l|l|l|r}
    \toprule
    \textbf{Tasks} & \textbf{Model Name} & \textbf{API URL} & \multicolumn{1}{l}{\textbf{\#Downloads}} \\
    \midrule
    Semantic Analysis & BERT-base &
    \href{https://huggingface.co/textattack/bert-base-uncased-SST-2}{bert-base-uncased-SST-2} & 109,143 \\
    Semantic Analysis & RoBERTa-base &
    \href{https://huggingface.co/textattack/roberta-base-SST-2}{roberta-base-SST-2} & 6,099 \\
    Semantic Analysis & DistilBERT-base &
    \href{https://huggingface.co/textattack/distilbert-base-uncased-SST-2}{distilbert-base-uncased-SST-2} & 31 \\
    Hate Speech Detection & dehate-BERT & \href{https://huggingface.co/Hate-speech-CNERG/dehatebert-mono-english}{dehatebert-mono-english}  & 1,366 \\
    Hate Speech Detection & twitter-RoBERTa &
    \href{https://huggingface.co/cardiffnlp/twitter-roberta-base-hate}{twitter-roberta-base-hate} & 40,097 \\
    \bottomrule
    \end{tabular}%
    }
  \label{tab:model}%
\end{table}%