### 2023-03-16 paper/icse24/tables/mtnlp-comp-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{mtnlp-comp-sa-ours-selfbleu-num-mutate}{\textbf{606}}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-avg}{\textbf{0.75}}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-med}{\textbf{0.76}}
\DefMacro{mtnlp-comp-sa-ours-selfbleu-std}{\textbf{0.01}}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-num-mutate}{23}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-avg}{0.91}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-med}{0.91}
\DefMacro{mtnlp-comp-sa-mtnlp-selfbleu-std}{0.0}
\DefMacro{mtnlp-comp-sa-ours-pdr-num-mutate}{\textbf{606}}
\DefMacro{mtnlp-comp-sa-ours-pdr-avg}{\textbf{338.8}}
\DefMacro{mtnlp-comp-sa-ours-pdr-med}{\textbf{338}}
\DefMacro{mtnlp-comp-sa-ours-pdr-std}{\textbf{12.03}}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-num-mutate}{23}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-avg}{96.0}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-med}{96}
\DefMacro{mtnlp-comp-sa-mtnlp-pdr-std}{0.0}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-num-mutate}{\textbf{800}}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-avg}{\textbf{0.69}}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-med}{\textbf{0.68}}
\DefMacro{mtnlp-comp-hs-ours-selfbleu-std}{\textbf{0.02}}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-num-mutate}{211}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-avg}{0.79}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-med}{0.78}
\DefMacro{mtnlp-comp-hs-mtnlp-selfbleu-std}{0.02}
\DefMacro{mtnlp-comp-hs-ours-pdr-num-mutate}{\textbf{800}}
\DefMacro{mtnlp-comp-hs-ours-pdr-avg}{\textbf{400.4}}
\DefMacro{mtnlp-comp-hs-ours-pdr-med}{\textbf{392}}
\DefMacro{mtnlp-comp-hs-ours-pdr-std}{\textbf{17.21}}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-num-mutate}{211}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-avg}{344.0}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-med}{348}
\DefMacro{mtnlp-comp-hs-mtnlp-pdr-std}{15.86}