### 2023-03-16 paper/icse24/tables/mtnlp-comp-table.tex

%% Automatically generated by pyutil.latex

\begin{table}[htbp]
\begin{small}
\begin{center}
\caption{Comparison results against MT-NLP}
\label{table:MtnlpComp}
\vspace{-3mm}
\resizebox{0.35\textwidth}{!}{
\begin{tabular}{l|l|ccc}
\toprule
\tTask & \tApproach & \tNummut & \tSelfbleu & \tPdrcov \\
\midrule
\multirow{2}{*}{SA} 
 & $S^{2}LCT$ & \UseMacro{mtnlp-comp-sa-ours-pdr-num-mutate}
 & \UseMacro{mtnlp-comp-sa-ours-selfbleu-avg}$\pm$
\UseMacro{mtnlp-comp-sa-ours-selfbleu-std}
 & \UseMacro{mtnlp-comp-sa-ours-pdr-avg}$\pm$\UseMacro{mtnlp-comp-sa-ours-pdr-std}\\
 & MT-NLP & \UseMacro{mtnlp-comp-sa-mtnlp-pdr-num-mutate}
 & \UseMacro{mtnlp-comp-sa-mtnlp-selfbleu-avg}$\pm$
\UseMacro{mtnlp-comp-sa-mtnlp-selfbleu-std}
 & \UseMacro{mtnlp-comp-sa-mtnlp-pdr-avg}$\pm$\UseMacro{mtnlp-comp-sa-mtnlp-pdr-std}\\
\hline
\multirow{2}{*}{HSD} 
 & $S^{2}LCT$ & \UseMacro{mtnlp-comp-hs-ours-pdr-num-mutate}
 & \UseMacro{mtnlp-comp-hs-ours-selfbleu-avg}$\pm$
\UseMacro{mtnlp-comp-hs-ours-selfbleu-std}
 & \UseMacro{mtnlp-comp-hs-ours-pdr-avg}$\pm$\UseMacro{mtnlp-comp-hs-ours-pdr-std}\\
 & MT-NLP & \UseMacro{mtnlp-comp-hs-mtnlp-pdr-num-mutate}
 & \UseMacro{mtnlp-comp-hs-mtnlp-selfbleu-avg}$\pm$
\UseMacro{mtnlp-comp-hs-mtnlp-selfbleu-std}
 & \UseMacro{mtnlp-comp-hs-mtnlp-pdr-avg}$\pm$\UseMacro{mtnlp-comp-hs-mtnlp-pdr-std}\\
\bottomrule
\end{tabular}
}
\end{center}
\end{small}
\vspace{\MtnlpCompTableVSpace}
\end{table}