### 2023-03-16 paper/icse24/tables/test-results-all-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-all-lc7}{Author sentiment is more important than of others}
\DefMacro{test-results-all-lc7-num-seeds}{68,284}
\DefMacro{test-results-all-lc7-num-exps}{465,291}
\DefMacro{test-results-bl-lc7-num-tcs}{8,528}
\DefMacro{test-results-all-model0-lc7-num-fail}{152,009}
\DefMacro{test-results-all-model0-lc7-num-failrate}{28.49}
\DefMacro{test-results-all-model0-lc7-num-pass-to-fail}{8,878}
\DefMacro{test-results-bl-model0-lc7-num-fail}{3,741}
\DefMacro{test-results-bl-model0-lc7-num-failrate}{43.87}
\DefMacro{test-results-all-lc3}{Negated negative should be positive or neutral}
\DefMacro{test-results-all-lc3-num-seeds}{67}
\DefMacro{test-results-all-lc3-num-exps}{503}
\DefMacro{test-results-bl-lc3-num-tcs}{6,786}
\DefMacro{test-results-all-model0-lc3-num-fail}{523}
\DefMacro{test-results-all-model0-lc3-num-failrate}{91.75}
\DefMacro{test-results-all-model0-lc3-num-pass-to-fail}{20}
\DefMacro{test-results-bl-model0-lc3-num-fail}{799}
\DefMacro{test-results-bl-model0-lc3-num-failrate}{11.77}
\DefMacro{test-results-all-lc4}{Negated neutral should still be neutral}
\DefMacro{test-results-all-lc4-num-seeds}{26}
\DefMacro{test-results-all-lc4-num-exps}{194}
\DefMacro{test-results-bl-lc4-num-tcs}{2,496}
\DefMacro{test-results-all-model0-lc4-num-fail}{207}
\DefMacro{test-results-all-model0-lc4-num-failrate}{94.09}
\DefMacro{test-results-all-model0-lc4-num-pass-to-fail}{11}
\DefMacro{test-results-bl-model0-lc4-num-fail}{2,427}
\DefMacro{test-results-bl-model0-lc4-num-failrate}{97.24}
\DefMacro{test-results-all-lc6}{Negated positive with neutral content in the middle}
\DefMacro{test-results-all-lc6-num-seeds}{24,328}
\DefMacro{test-results-all-lc6-num-exps}{184,328}
\DefMacro{test-results-bl-lc6-num-tcs}{1,000}
\DefMacro{test-results-all-model0-lc6-num-fail}{189,935}
\DefMacro{test-results-all-model0-lc6-num-failrate}{91.03}
\DefMacro{test-results-all-model0-lc6-num-pass-to-fail}{1,972}
\DefMacro{test-results-bl-model0-lc6-num-fail}{860}
\DefMacro{test-results-bl-model0-lc6-num-failrate}{86.00}
\DefMacro{test-results-all-lc5}{Negation of negative at the end, should be positive or neutral}
\DefMacro{test-results-all-lc5-num-seeds}{18,576}
\DefMacro{test-results-all-lc5-num-exps}{97,897}
\DefMacro{test-results-bl-lc5-num-tcs}{2,124}
\DefMacro{test-results-all-model0-lc5-num-fail}{116,049}
\DefMacro{test-results-all-model0-lc5-num-failrate}{99.64}
\DefMacro{test-results-all-model0-lc5-num-pass-to-fail}{67}
\DefMacro{test-results-bl-model0-lc5-num-fail}{1,871}
\DefMacro{test-results-bl-model0-lc5-num-failrate}{88.09}
\DefMacro{test-results-all-lc9}{Parsing sentiment in (question, no) form}
\DefMacro{test-results-all-lc9-num-seeds}{15,483}
\DefMacro{test-results-all-lc9-num-exps}{102,214}
\DefMacro{test-results-bl-lc9-num-tcs}{7,644}
\DefMacro{test-results-all-model0-lc9-num-fail}{89,155}
\DefMacro{test-results-all-model0-lc9-num-failrate}{75.75}
\DefMacro{test-results-all-model0-lc9-num-pass-to-fail}{1,722}
\DefMacro{test-results-bl-model0-lc9-num-fail}{4,056}
\DefMacro{test-results-bl-model0-lc9-num-failrate}{53.06}
\DefMacro{test-results-all-lc2}{Sentiment change over time, present should prevail}
\DefMacro{test-results-all-lc2-num-seeds}{75,159}
\DefMacro{test-results-all-lc2-num-exps}{343,214}
\DefMacro{test-results-bl-lc2-num-tcs}{8,000}
\DefMacro{test-results-all-model0-lc2-num-fail}{99,312}
\DefMacro{test-results-all-model0-lc2-num-failrate}{23.74}
\DefMacro{test-results-all-model0-lc2-num-pass-to-fail}{10,357}
\DefMacro{test-results-bl-model0-lc2-num-fail}{1,680}
\DefMacro{test-results-bl-model0-lc2-num-failrate}{21.00}
\DefMacro{test-results-all-lc0}{Short sentences with neutral adjectives and nouns}
\DefMacro{test-results-all-lc0-num-seeds}{19}
\DefMacro{test-results-all-lc0-num-exps}{51}
\DefMacro{test-results-bl-lc0-num-tcs}{1,716}
\DefMacro{test-results-all-model0-lc0-num-fail}{60}
\DefMacro{test-results-all-model0-lc0-num-failrate}{85.71}
\DefMacro{test-results-all-model0-lc0-num-pass-to-fail}{9}
\DefMacro{test-results-bl-model0-lc0-num-fail}{1,330}
\DefMacro{test-results-bl-model0-lc0-num-failrate}{77.51}
\DefMacro{test-results-all-lc1}{Short sentences with sentiment-laden adjectives}
\DefMacro{test-results-all-lc1-num-seeds}{160}
\DefMacro{test-results-all-lc1-num-exps}{262}
\DefMacro{test-results-bl-lc1-num-tcs}{8,658}
\DefMacro{test-results-all-model0-lc1-num-fail}{25}
\DefMacro{test-results-all-model0-lc1-num-failrate}{5.92}
\DefMacro{test-results-all-model0-lc1-num-pass-to-fail}{5}
\DefMacro{test-results-bl-model0-lc1-num-fail}{26}
\DefMacro{test-results-bl-model0-lc1-num-failrate}{0.30}
\DefMacro{test-results-all-lc8}{Parsing sentiment in (question, yes) form}
\DefMacro{test-results-all-lc8-num-seeds}{15,465}
\DefMacro{test-results-all-lc8-num-exps}{102,203}
\DefMacro{test-results-bl-lc8-num-tcs}{7,644}
\DefMacro{test-results-all-model0-lc8-num-fail}{7,097}
\DefMacro{test-results-all-model0-lc8-num-failrate}{6.03}
\DefMacro{test-results-all-model0-lc8-num-pass-to-fail}{1,590}
\DefMacro{test-results-bl-model0-lc8-num-fail}{253}
\DefMacro{test-results-bl-model0-lc8-num-failrate}{3.31}
\DefMacro{test-results-all-model1-lc7-num-fail}{105,152}
\DefMacro{test-results-all-model1-lc7-num-failrate}{19.71}
\DefMacro{test-results-all-model1-lc7-num-pass-to-fail}{8,487}
\DefMacro{test-results-bl-model1-lc7-num-fail}{2,693}
\DefMacro{test-results-bl-model1-lc7-num-failrate}{31.58}
\DefMacro{test-results-all-model1-lc3-num-fail}{498}
\DefMacro{test-results-all-model1-lc3-num-failrate}{87.37}
\DefMacro{test-results-all-model1-lc3-num-pass-to-fail}{9}
\DefMacro{test-results-bl-model1-lc3-num-fail}{218}
\DefMacro{test-results-bl-model1-lc3-num-failrate}{3.21}
\DefMacro{test-results-all-model1-lc4-num-fail}{204}
\DefMacro{test-results-all-model1-lc4-num-failrate}{92.73}
\DefMacro{test-results-all-model1-lc4-num-pass-to-fail}{6}
\DefMacro{test-results-bl-model1-lc4-num-fail}{2,304}
\DefMacro{test-results-bl-model1-lc4-num-failrate}{92.31}
\DefMacro{test-results-all-model1-lc6-num-fail}{153,686}
\DefMacro{test-results-all-model1-lc6-num-failrate}{73.66}
\DefMacro{test-results-all-model1-lc6-num-pass-to-fail}{7,007}
\DefMacro{test-results-bl-model1-lc6-num-fail}{416}
\DefMacro{test-results-bl-model1-lc6-num-failrate}{41.60}
\DefMacro{test-results-all-model1-lc5-num-fail}{115,676}
\DefMacro{test-results-all-model1-lc5-num-failrate}{99.32}
\DefMacro{test-results-all-model1-lc5-num-pass-to-fail}{90}
\DefMacro{test-results-bl-model1-lc5-num-fail}{445}
\DefMacro{test-results-bl-model1-lc5-num-failrate}{20.95}
\DefMacro{test-results-all-model1-lc9-num-fail}{100,351}
\DefMacro{test-results-all-model1-lc9-num-failrate}{85.26}
\DefMacro{test-results-all-model1-lc9-num-pass-to-fail}{1,452}
\DefMacro{test-results-bl-model1-lc9-num-fail}{4,576}
\DefMacro{test-results-bl-model1-lc9-num-failrate}{59.86}
\DefMacro{test-results-all-model1-lc2-num-fail}{208,313}
\DefMacro{test-results-all-model1-lc2-num-failrate}{49.79}
\DefMacro{test-results-all-model1-lc2-num-pass-to-fail}{11,472}
\DefMacro{test-results-bl-model1-lc2-num-fail}{829}
\DefMacro{test-results-bl-model1-lc2-num-failrate}{10.36}
\DefMacro{test-results-all-model1-lc0-num-fail}{55}
\DefMacro{test-results-all-model1-lc0-num-failrate}{78.57}
\DefMacro{test-results-all-model1-lc0-num-pass-to-fail}{2}
\DefMacro{test-results-bl-model1-lc0-num-fail}{1,391}
\DefMacro{test-results-bl-model1-lc0-num-failrate}{81.06}
\DefMacro{test-results-all-model1-lc1-num-fail}{39}
\DefMacro{test-results-all-model1-lc1-num-failrate}{9.24}
\DefMacro{test-results-all-model1-lc1-num-pass-to-fail}{14}
\DefMacro{test-results-bl-model1-lc1-num-fail}{139}
\DefMacro{test-results-bl-model1-lc1-num-failrate}{1.61}
\DefMacro{test-results-all-model1-lc8-num-fail}{6,226}
\DefMacro{test-results-all-model1-lc8-num-failrate}{5.29}
\DefMacro{test-results-all-model1-lc8-num-pass-to-fail}{1,489}
\DefMacro{test-results-bl-model1-lc8-num-fail}{32}
\DefMacro{test-results-bl-model1-lc8-num-failrate}{0.42}
\DefMacro{test-results-all-model2-lc7-num-fail}{162,426}
\DefMacro{test-results-all-model2-lc7-num-failrate}{30.44}
\DefMacro{test-results-all-model2-lc7-num-pass-to-fail}{12,729}
\DefMacro{test-results-bl-model2-lc7-num-fail}{3,535}
\DefMacro{test-results-bl-model2-lc7-num-failrate}{41.45}
\DefMacro{test-results-all-model2-lc3-num-fail}{494}
\DefMacro{test-results-all-model2-lc3-num-failrate}{86.67}
\DefMacro{test-results-all-model2-lc3-num-pass-to-fail}{6}
\DefMacro{test-results-bl-model2-lc3-num-fail}{734}
\DefMacro{test-results-bl-model2-lc3-num-failrate}{10.82}
\DefMacro{test-results-all-model2-lc4-num-fail}{213}
\DefMacro{test-results-all-model2-lc4-num-failrate}{96.82}
\DefMacro{test-results-all-model2-lc4-num-pass-to-fail}{10}
\DefMacro{test-results-bl-model2-lc4-num-fail}{2,450}
\DefMacro{test-results-bl-model2-lc4-num-failrate}{98.16}
\DefMacro{test-results-all-model2-lc6-num-fail}{175,323}
\DefMacro{test-results-all-model2-lc6-num-failrate}{84.02}
\DefMacro{test-results-all-model2-lc6-num-pass-to-fail}{5,003}
\DefMacro{test-results-bl-model2-lc6-num-fail}{865}
\DefMacro{test-results-bl-model2-lc6-num-failrate}{86.50}
\DefMacro{test-results-all-model2-lc5-num-fail}{114,556}
\DefMacro{test-results-all-model2-lc5-num-failrate}{98.35}
\DefMacro{test-results-all-model2-lc5-num-pass-to-fail}{325}
\DefMacro{test-results-bl-model2-lc5-num-fail}{2,124}
\DefMacro{test-results-bl-model2-lc5-num-failrate}{100.00}
\DefMacro{test-results-all-model2-lc9-num-fail}{111,874}
\DefMacro{test-results-all-model2-lc9-num-failrate}{95.05}
\DefMacro{test-results-all-model2-lc9-num-pass-to-fail}{575}
\DefMacro{test-results-bl-model2-lc9-num-fail}{6,440}
\DefMacro{test-results-bl-model2-lc9-num-failrate}{84.25}
\DefMacro{test-results-all-model2-lc2-num-fail}{262,994}
\DefMacro{test-results-all-model2-lc2-num-failrate}{62.86}
\DefMacro{test-results-all-model2-lc2-num-pass-to-fail}{9,808}
\DefMacro{test-results-bl-model2-lc2-num-fail}{2,532}
\DefMacro{test-results-bl-model2-lc2-num-failrate}{31.65}
\DefMacro{test-results-all-model2-lc0-num-fail}{68}
\DefMacro{test-results-all-model2-lc0-num-failrate}{97.14}
\DefMacro{test-results-all-model2-lc0-num-pass-to-fail}{0}
\DefMacro{test-results-bl-model2-lc0-num-fail}{1,661}
\DefMacro{test-results-bl-model2-lc0-num-failrate}{96.79}
\DefMacro{test-results-all-model2-lc1-num-fail}{18}
\DefMacro{test-results-all-model2-lc1-num-failrate}{4.27}
\DefMacro{test-results-all-model2-lc1-num-pass-to-fail}{10}
\DefMacro{test-results-bl-model2-lc1-num-fail}{125}
\DefMacro{test-results-bl-model2-lc1-num-failrate}{1.44}
\DefMacro{test-results-all-model2-lc8-num-fail}{5,470}
\DefMacro{test-results-all-model2-lc8-num-failrate}{4.65}
\DefMacro{test-results-all-model2-lc8-num-pass-to-fail}{1,151}
\DefMacro{test-results-bl-model2-lc8-num-fail}{52}
\DefMacro{test-results-bl-model2-lc8-num-failrate}{0.68}