### 2023-03-16 paper/icse24/tables/test-results-hs-all-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{test-results-hs-lc13}{Denouncements of hate that make direct reference to it}
\DefMacro{test-results-hs-lc13-num-seeds}{17,796}
\DefMacro{test-results-hs-lc13-num-exps}{127,067}
\DefMacro{test-results-hs-bl-lc13-num-tcs}{141}
\DefMacro{test-results-hs-model0-lc13-num-all-fail}{17,060}
\DefMacro{test-results-hs-model0-lc13-num-seed-failrate}{10.60}
\DefMacro{test-results-hs-model0-lc13-num-exp-failrate}{11.94}
\DefMacro{test-results-hs-model0-lc13-num-all-failrate}{11.78}
\DefMacro{test-results-hs-model0-lc13-num-pass-to-fail}{1,070}
\DefMacro{test-results-hs-bl-model0-lc13-num-fail}{4}
\DefMacro{test-results-hs-bl-model0-lc13-num-failrate}{2.84}
\DefMacro{test-results-hs-lc12}{Denouncements of hate that quote it}
\DefMacro{test-results-hs-lc12-num-seeds}{23,728}
\DefMacro{test-results-hs-lc12-num-exps}{167,404}
\DefMacro{test-results-hs-bl-lc12-num-tcs}{173}
\DefMacro{test-results-hs-model0-lc12-num-all-fail}{20,511}
\DefMacro{test-results-hs-model0-lc12-num-seed-failrate}{9.47}
\DefMacro{test-results-hs-model0-lc12-num-exp-failrate}{10.91}
\DefMacro{test-results-hs-model0-lc12-num-all-failrate}{10.73}
\DefMacro{test-results-hs-model0-lc12-num-pass-to-fail}{1,229}
\DefMacro{test-results-hs-bl-model0-lc12-num-fail}{17}
\DefMacro{test-results-hs-bl-model0-lc12-num-failrate}{9.83}
\DefMacro{test-results-hs-lc4}{Hate expressed through reference in subsequent clauses}
\DefMacro{test-results-hs-lc4-num-seeds}{11,968}
\DefMacro{test-results-hs-lc4-num-exps}{43,641}
\DefMacro{test-results-hs-bl-lc4-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc4-num-all-fail}{37,022}
\DefMacro{test-results-hs-model0-lc4-num-seed-failrate}{45.10}
\DefMacro{test-results-hs-model0-lc4-num-exp-failrate}{72.47}
\DefMacro{test-results-hs-model0-lc4-num-all-failrate}{66.58}
\DefMacro{test-results-hs-model0-lc4-num-pass-to-fail}{793}
\DefMacro{test-results-hs-bl-model0-lc4-num-fail}{108}
\DefMacro{test-results-hs-bl-model0-lc4-num-failrate}{77.14}
\DefMacro{test-results-hs-lc5}{Hate expressed through reference in subsequent sentences}
\DefMacro{test-results-hs-lc5-num-seeds}{11,968}
\DefMacro{test-results-hs-lc5-num-exps}{42,416}
\DefMacro{test-results-hs-bl-lc5-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc5-num-all-fail}{35,958}
\DefMacro{test-results-hs-model0-lc5-num-seed-failrate}{45.19}
\DefMacro{test-results-hs-model0-lc5-num-exp-failrate}{72.02}
\DefMacro{test-results-hs-model0-lc5-num-all-failrate}{66.12}
\DefMacro{test-results-hs-model0-lc5-num-pass-to-fail}{783}
\DefMacro{test-results-hs-bl-model0-lc5-num-fail}{101}
\DefMacro{test-results-hs-bl-model0-lc5-num-failrate}{75.94}
\DefMacro{test-results-hs-lc6}{Hate expressed using negated positive statement}
\DefMacro{test-results-hs-lc6-num-seeds}{39,783}
\DefMacro{test-results-hs-lc6-num-exps}{220,483}
\DefMacro{test-results-hs-bl-lc6-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc6-num-all-fail}{222,574}
\DefMacro{test-results-hs-model0-lc6-num-seed-failrate}{85.77}
\DefMacro{test-results-hs-model0-lc6-num-exp-failrate}{85.47}
\DefMacro{test-results-hs-model0-lc6-num-all-failrate}{85.52}
\DefMacro{test-results-hs-model0-lc6-num-pass-to-fail}{2,457}
\DefMacro{test-results-hs-bl-model0-lc6-num-fail}{109}
\DefMacro{test-results-hs-bl-model0-lc6-num-failrate}{77.86}
\DefMacro{test-results-hs-lc2}{Hate expressed using profanity}
\DefMacro{test-results-hs-lc2-num-seeds}{1,064}
\DefMacro{test-results-hs-lc2-num-exps}{6,394}
\DefMacro{test-results-hs-bl-lc2-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc2-num-all-fail}{5,869}
\DefMacro{test-results-hs-model0-lc2-num-seed-failrate}{80.55}
\DefMacro{test-results-hs-model0-lc2-num-exp-failrate}{78.39}
\DefMacro{test-results-hs-model0-lc2-num-all-failrate}{78.69}
\DefMacro{test-results-hs-model0-lc2-num-pass-to-fail}{51}
\DefMacro{test-results-hs-bl-model0-lc2-num-fail}{98}
\DefMacro{test-results-hs-bl-model0-lc2-num-failrate}{70.00}
\DefMacro{test-results-hs-lc0}{Hate expressed using slur}
\DefMacro{test-results-hs-lc0-num-seeds}{203}
\DefMacro{test-results-hs-lc0-num-exps}{1,171}
\DefMacro{test-results-hs-bl-lc0-num-tcs}{144}
\DefMacro{test-results-hs-model0-lc0-num-all-fail}{435}
\DefMacro{test-results-hs-model0-lc0-num-seed-failrate}{26.60}
\DefMacro{test-results-hs-model0-lc0-num-exp-failrate}{32.54}
\DefMacro{test-results-hs-model0-lc0-num-all-failrate}{31.66}
\DefMacro{test-results-hs-model0-lc0-num-pass-to-fail}{16}
\DefMacro{test-results-hs-bl-model0-lc0-num-fail}{108}
\DefMacro{test-results-hs-bl-model0-lc0-num-failrate}{75.00}
\DefMacro{test-results-hs-lc9}{Hate phrased as a opinion}
\DefMacro{test-results-hs-lc9-num-seeds}{11,864}
\DefMacro{test-results-hs-lc9-num-exps}{87,996}
\DefMacro{test-results-hs-bl-lc9-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc9-num-all-fail}{84,221}
\DefMacro{test-results-hs-model0-lc9-num-seed-failrate}{85.45}
\DefMacro{test-results-hs-model0-lc9-num-exp-failrate}{84.19}
\DefMacro{test-results-hs-model0-lc9-num-all-failrate}{84.34}
\DefMacro{test-results-hs-model0-lc9-num-pass-to-fail}{999}
\DefMacro{test-results-hs-bl-model0-lc9-num-fail}{100}
\DefMacro{test-results-hs-bl-model0-lc9-num-failrate}{75.19}
\DefMacro{test-results-hs-lc8}{Hate phrased as a question}
\DefMacro{test-results-hs-lc8-num-seeds}{11,864}
\DefMacro{test-results-hs-lc8-num-exps}{101,569}
\DefMacro{test-results-hs-bl-lc8-num-tcs}{140}
\DefMacro{test-results-hs-model0-lc8-num-all-fail}{98,879}
\DefMacro{test-results-hs-model0-lc8-num-seed-failrate}{88.01}
\DefMacro{test-results-hs-model0-lc8-num-exp-failrate}{87.07}
\DefMacro{test-results-hs-model0-lc8-num-all-failrate}{87.17}
\DefMacro{test-results-hs-model0-lc8-num-pass-to-fail}{961}
\DefMacro{test-results-hs-bl-model0-lc8-num-fail}{107}
\DefMacro{test-results-hs-bl-model0-lc8-num-failrate}{76.43}
\DefMacro{test-results-hs-lc10}{Neutral statements using protected group identifiers}
\DefMacro{test-results-hs-lc10-num-seeds}{6}
\DefMacro{test-results-hs-lc10-num-exps}{12}
\DefMacro{test-results-hs-bl-lc10-num-tcs}{126}
\DefMacro{test-results-hs-model0-lc10-num-all-fail}{16}
\DefMacro{test-results-hs-model0-lc10-num-seed-failrate}{66.67}
\DefMacro{test-results-hs-model0-lc10-num-exp-failrate}{100.00}
\DefMacro{test-results-hs-model0-lc10-num-all-failrate}{88.89}
\DefMacro{test-results-hs-model0-lc10-num-pass-to-fail}{0}
\DefMacro{test-results-hs-bl-model0-lc10-num-fail}{9}
\DefMacro{test-results-hs-bl-model0-lc10-num-failrate}{7.14}
\DefMacro{test-results-hs-lc3}{Non-Hateful use of profanity}
\DefMacro{test-results-hs-lc3-num-seeds}{1,478}
\DefMacro{test-results-hs-lc3-num-exps}{7,709}
\DefMacro{test-results-hs-bl-lc3-num-tcs}{100}
\DefMacro{test-results-hs-model0-lc3-num-all-fail}{1,683}
\DefMacro{test-results-hs-model0-lc3-num-seed-failrate}{18.00}
\DefMacro{test-results-hs-model0-lc3-num-exp-failrate}{18.38}
\DefMacro{test-results-hs-model0-lc3-num-all-failrate}{18.32}
\DefMacro{test-results-hs-model0-lc3-num-pass-to-fail}{49}
\DefMacro{test-results-hs-bl-model0-lc3-num-fail}{1}
\DefMacro{test-results-hs-bl-model0-lc3-num-failrate}{1.00}
\DefMacro{test-results-hs-lc7}{Non-hate expressed using negated hateful statement}
\DefMacro{test-results-hs-lc7-num-seeds}{17,796}
\DefMacro{test-results-hs-lc7-num-exps}{133,756}
\DefMacro{test-results-hs-bl-lc7-num-tcs}{133}
\DefMacro{test-results-hs-model0-lc7-num-all-fail}{23,027}
\DefMacro{test-results-hs-model0-lc7-num-seed-failrate}{14.06}
\DefMacro{test-results-hs-model0-lc7-num-exp-failrate}{15.35}
\DefMacro{test-results-hs-model0-lc7-num-all-failrate}{15.19}
\DefMacro{test-results-hs-model0-lc7-num-pass-to-fail}{1,265}
\DefMacro{test-results-hs-bl-model0-lc7-num-fail}{13}
\DefMacro{test-results-hs-bl-model0-lc7-num-failrate}{9.77}
\DefMacro{test-results-hs-lc1}{Non-hateful use of slur}
\DefMacro{test-results-hs-lc1-num-seeds}{997}
\DefMacro{test-results-hs-lc1-num-exps}{4,422}
\DefMacro{test-results-hs-bl-lc1-num-tcs}{111}
\DefMacro{test-results-hs-model0-lc1-num-all-fail}{3,835}
\DefMacro{test-results-hs-model0-lc1-num-seed-failrate}{71.82}
\DefMacro{test-results-hs-model0-lc1-num-exp-failrate}{70.53}
\DefMacro{test-results-hs-model0-lc1-num-all-failrate}{70.77}
\DefMacro{test-results-hs-model0-lc1-num-pass-to-fail}{29}
\DefMacro{test-results-hs-bl-model0-lc1-num-fail}{18}
\DefMacro{test-results-hs-bl-model0-lc1-num-failrate}{16.22}
\DefMacro{test-results-hs-lc11}{Positive statements using protected group identifiers}
\DefMacro{test-results-hs-lc11-num-seeds}{57}
\DefMacro{test-results-hs-lc11-num-exps}{246}
\DefMacro{test-results-hs-bl-lc11-num-tcs}{189}
\DefMacro{test-results-hs-model0-lc11-num-all-fail}{151}
\DefMacro{test-results-hs-model0-lc11-num-seed-failrate}{52.63}
\DefMacro{test-results-hs-model0-lc11-num-exp-failrate}{49.19}
\DefMacro{test-results-hs-model0-lc11-num-all-failrate}{49.83}
\DefMacro{test-results-hs-model0-lc11-num-pass-to-fail}{7}
\DefMacro{test-results-hs-bl-model0-lc11-num-fail}{23}
\DefMacro{test-results-hs-bl-model0-lc11-num-failrate}{12.17}
\DefMacro{test-results-hs-model1-lc13-num-all-fail}{100,848}
\DefMacro{test-results-hs-model1-lc13-num-seed-failrate}{69.93}
\DefMacro{test-results-hs-model1-lc13-num-exp-failrate}{69.57}
\DefMacro{test-results-hs-model1-lc13-num-all-failrate}{69.62}
\DefMacro{test-results-hs-model1-lc13-num-pass-to-fail}{1,594}
\DefMacro{test-results-hs-bl-model1-lc13-num-fail}{7}
\DefMacro{test-results-hs-bl-model1-lc13-num-failrate}{4.96}
\DefMacro{test-results-hs-model1-lc12-num-all-fail}{117,788}
\DefMacro{test-results-hs-model1-lc12-num-seed-failrate}{61.67}
\DefMacro{test-results-hs-model1-lc12-num-exp-failrate}{61.62}
\DefMacro{test-results-hs-model1-lc12-num-all-failrate}{61.63}
\DefMacro{test-results-hs-model1-lc12-num-pass-to-fail}{2,440}
\DefMacro{test-results-hs-bl-model1-lc12-num-fail}{5}
\DefMacro{test-results-hs-bl-model1-lc12-num-failrate}{2.89}
\DefMacro{test-results-hs-model1-lc4-num-all-fail}{30,276}
\DefMacro{test-results-hs-model1-lc4-num-seed-failrate}{51.40}
\DefMacro{test-results-hs-model1-lc4-num-exp-failrate}{55.28}
\DefMacro{test-results-hs-model1-lc4-num-all-failrate}{54.44}
\DefMacro{test-results-hs-model1-lc4-num-pass-to-fail}{855}
\DefMacro{test-results-hs-bl-model1-lc4-num-fail}{93}
\DefMacro{test-results-hs-bl-model1-lc4-num-failrate}{66.43}
\DefMacro{test-results-hs-model1-lc5-num-all-fail}{31,195}
\DefMacro{test-results-hs-model1-lc5-num-seed-failrate}{52.92}
\DefMacro{test-results-hs-model1-lc5-num-exp-failrate}{58.61}
\DefMacro{test-results-hs-model1-lc5-num-all-failrate}{57.36}
\DefMacro{test-results-hs-model1-lc5-num-pass-to-fail}{721}
\DefMacro{test-results-hs-bl-model1-lc5-num-fail}{69}
\DefMacro{test-results-hs-bl-model1-lc5-num-failrate}{51.88}
\DefMacro{test-results-hs-model1-lc6-num-all-fail}{152,929}
\DefMacro{test-results-hs-model1-lc6-num-seed-failrate}{55.55}
\DefMacro{test-results-hs-model1-lc6-num-exp-failrate}{59.34}
\DefMacro{test-results-hs-model1-lc6-num-all-failrate}{58.76}
\DefMacro{test-results-hs-model1-lc6-num-pass-to-fail}{4,365}
\DefMacro{test-results-hs-bl-model1-lc6-num-fail}{116}
\DefMacro{test-results-hs-bl-model1-lc6-num-failrate}{82.86}
\DefMacro{test-results-hs-model1-lc2-num-all-fail}{1,115}
\DefMacro{test-results-hs-model1-lc2-num-seed-failrate}{14.00}
\DefMacro{test-results-hs-model1-lc2-num-exp-failrate}{15.11}
\DefMacro{test-results-hs-model1-lc2-num-all-failrate}{14.95}
\DefMacro{test-results-hs-model1-lc2-num-pass-to-fail}{69}
\DefMacro{test-results-hs-bl-model1-lc2-num-fail}{93}
\DefMacro{test-results-hs-bl-model1-lc2-num-failrate}{66.43}
\DefMacro{test-results-hs-model1-lc0-num-all-fail}{26}
\DefMacro{test-results-hs-model1-lc0-num-seed-failrate}{1.48}
\DefMacro{test-results-hs-model1-lc0-num-exp-failrate}{1.96}
\DefMacro{test-results-hs-model1-lc0-num-all-failrate}{1.89}
\DefMacro{test-results-hs-model1-lc0-num-pass-to-fail}{12}
\DefMacro{test-results-hs-bl-model1-lc0-num-fail}{56}
\DefMacro{test-results-hs-bl-model1-lc0-num-failrate}{38.89}
\DefMacro{test-results-hs-model1-lc9-num-all-fail}{27,637}
\DefMacro{test-results-hs-model1-lc9-num-seed-failrate}{27.08}
\DefMacro{test-results-hs-model1-lc9-num-exp-failrate}{27.76}
\DefMacro{test-results-hs-model1-lc9-num-all-failrate}{27.68}
\DefMacro{test-results-hs-model1-lc9-num-pass-to-fail}{1,348}
\DefMacro{test-results-hs-bl-model1-lc9-num-fail}{109}
\DefMacro{test-results-hs-bl-model1-lc9-num-failrate}{81.95}
\DefMacro{test-results-hs-model1-lc8-num-all-fail}{33,589}
\DefMacro{test-results-hs-model1-lc8-num-seed-failrate}{28.51}
\DefMacro{test-results-hs-model1-lc8-num-exp-failrate}{29.74}
\DefMacro{test-results-hs-model1-lc8-num-all-failrate}{29.61}
\DefMacro{test-results-hs-model1-lc8-num-pass-to-fail}{1,305}
\DefMacro{test-results-hs-bl-model1-lc8-num-fail}{123}
\DefMacro{test-results-hs-bl-model1-lc8-num-failrate}{87.86}
\DefMacro{test-results-hs-model1-lc10-num-all-fail}{1}
\DefMacro{test-results-hs-model1-lc10-num-seed-failrate}{16.67}
\DefMacro{test-results-hs-model1-lc10-num-exp-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc10-num-all-failrate}{5.56}
\DefMacro{test-results-hs-model1-lc10-num-pass-to-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc10-num-fail}{0}
\DefMacro{test-results-hs-bl-model1-lc10-num-failrate}{0.00}
\DefMacro{test-results-hs-model1-lc3-num-all-fail}{5,160}
\DefMacro{test-results-hs-model1-lc3-num-seed-failrate}{63.19}
\DefMacro{test-results-hs-model1-lc3-num-exp-failrate}{54.82}
\DefMacro{test-results-hs-model1-lc3-num-all-failrate}{56.17}
\DefMacro{test-results-hs-model1-lc3-num-pass-to-fail}{120}
\DefMacro{test-results-hs-bl-model1-lc3-num-fail}{1}
\DefMacro{test-results-hs-bl-model1-lc3-num-failrate}{1.00}
\DefMacro{test-results-hs-model1-lc7-num-all-fail}{113,265}
\DefMacro{test-results-hs-model1-lc7-num-seed-failrate}{75.11}
\DefMacro{test-results-hs-model1-lc7-num-exp-failrate}{74.69}
\DefMacro{test-results-hs-model1-lc7-num-all-failrate}{74.74}
\DefMacro{test-results-hs-model1-lc7-num-pass-to-fail}{1,626}
\DefMacro{test-results-hs-bl-model1-lc7-num-fail}{26}
\DefMacro{test-results-hs-bl-model1-lc7-num-failrate}{19.55}
\DefMacro{test-results-hs-model1-lc1-num-all-fail}{4,484}
\DefMacro{test-results-hs-model1-lc1-num-seed-failrate}{87.26}
\DefMacro{test-results-hs-model1-lc1-num-exp-failrate}{81.73}
\DefMacro{test-results-hs-model1-lc1-num-all-failrate}{82.75}
\DefMacro{test-results-hs-model1-lc1-num-pass-to-fail}{70}
\DefMacro{test-results-hs-bl-model1-lc1-num-fail}{68}
\DefMacro{test-results-hs-bl-model1-lc1-num-failrate}{61.26}
\DefMacro{test-results-hs-model1-lc11-num-all-fail}{73}
\DefMacro{test-results-hs-model1-lc11-num-seed-failrate}{38.60}
\DefMacro{test-results-hs-model1-lc11-num-exp-failrate}{20.73}
\DefMacro{test-results-hs-model1-lc11-num-all-failrate}{24.09}
\DefMacro{test-results-hs-model1-lc11-num-pass-to-fail}{1}
\DefMacro{test-results-hs-bl-model1-lc11-num-fail}{16}
\DefMacro{test-results-hs-bl-model1-lc11-num-failrate}{8.47}