### 2023-03-16 paper/icse24/threats.tex

\section{Threats to Validity}
\label{sec:threats}

There are two potential threats to validity.
First, the dataset
used in our study might not be representative of all
English grammatical structures and word sentiments. We
mitigate this threat by using widely used dataset in the NLP
domain~\cite{husnain2021swnvalidity}.
Second, using ranges of the sentiment scores in \Sst and majority voting 
method in \Hatexp to determine the sentiment labels of search dataset may 
introduce incorrect labels. We performed a manual study to confirm that 
this is rare. 
% Use of publicly available datasets poses an external validity threat
% with respect to the generalizability of our results. We implemented
% domain-specific knowledge on publicly available word sentiments
% dataset~\cite{baccianella2010sentiwordnet}. We also constructed
% reference CFG from publicly available corpus dataset
% \cite{mitchell1993treebank, nltkTreebankCorporaWebPage}.