### 2023-06-08 paper/icse24/tables/lc-requirement-numbers.tex

%% Automatically generated by pyutil.latex

\DefMacro{lc_1_desc}{Short sentences with neutral adjectives and nouns}
\DefMacro{lc_1_predicates}{$Seed\leftarrow$\textbf{hasattr}(length,<10)$\land$
\textbf{hasattr}(label,neutral)$\land$(\textbf{contain}(tag\_of\_pos,neutral\_adj)$\lor$
\textbf{contain}(tag\_of\_pos,neutral\_noun))$\land$~\neg(\textbf{contain}(tag\_of\_pos,positive\_adj)$\lor$
\textbf{contain}(tag\_of\_pos,positive\_noun)$\lor$~\textbf{contain}(tag\_of\_pos,negative\_adj)$\lor$
\textbf{contain}(tag\_of\_pos,negative\_noun))}
\DefMacro{lc_2_desc}{Short sentences with sentiment-laden adjectives}
\DefMacro{lc_2_predicates}{$Seed\leftarrow$\textbf{hasattr}(length,$<$10)$\land$((\textbf{contain}
(tag\_of\_pos,positive$\_$adj)$\lor$\textbf{contain}
(tag\_of\_pos,positive$\_$noun)
$\land$\textbf{hasattr}(label,positive))$\lor$(\textbf{contain}
(tag\_of\_pos,negative$\_$adj)$\lor$\textbf{contain}(tag\_of\_pos,negative$\_$noun)$\land$\textbf{hasattr}(label,negative)))}
\DefMacro{lc_3_desc}{Sentiment change over time, present should prevail}
\DefMacro{lc_3_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix1,$S_{pos}$,infix,postfix1)$+$\textbf{combine}(prefix2,$S_{neg}$,infix,postfix2)}
\DefMacro{lc_3_predicates2}{prefix1$:=$``Previously, I used to like it saying that''|``Last time, I agreed with saying that''|``I liked it much as to say that''}
\DefMacro{lc_3_predicates3}{$S_{pos}\leftarrow$\textbf{hasattr}(length,<20)$\land$\textbf{hasattr}(label,positive)}
\DefMacro{lc_3_predicates4}{infix$:=$``but''|``although''|``on the other hand''}
\DefMacro{lc_3_predicates5}{postfix1$:=$``now I don't like it.''|``now I hate it.''}
\DefMacro{lc_3_predicates6}{prefix2$:=$``I used to disagree with saying that''|``Last time, I didn't like it saying that''|``I hated it much as to say that''}
\DefMacro{lc_3_predicates7}{$S_{neg}\leftarrow$\textbf{hasattr}(length,<20)$\land$\textbf{hasattr}(label,negative)}
\DefMacro{lc_3_predicates8}{postfix2$:=$``now I like it.''}
\DefMacro{lc_4_desc}{Negated negative should be positive or neutral}
\DefMacro{lc_4_predicates1}{$Seed\leftarrow$\textbf{replace}($S$, ``is'', [``is not'',``isn't''])$+$\textbf{replace}($S$, ``are'', [``are not'',``aren't''])}
\DefMacro{lc_4_predicates2}{$S\leftarrow$\textbf{hasattr}(clause,`(This|That|These|Those)~(is|are)')$\land$\textbf{hasattr}(label,negative)}
\DefMacro{lc_5_desc}{Negated neutral should still be neutral}
\DefMacro{lc_5_predicates1}{$Seed\leftarrow$\textbf{replace}($S$, ``is'', [``is not'',``isn't''])$+$\textbf{replace}($S$, ``are'', [``are not'',``aren't''])}
\DefMacro{lc_5_predicates2}{$S\leftarrow$\textbf{hasattr}(clause,`(This|That|These|Those)~(is|are)')$\land$\textbf{hasattr}(label,neutral)}
\DefMacro{lc_6_desc}{Negation of negative at the end, should be positive or neutral}
\DefMacro{lc_6_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix,$S_{neg}$,postfix)}
\DefMacro{lc_6_predicates2}{prefix$:=$``I agreed that''|``I thought that''}
\DefMacro{lc_6_predicates3}{$S_{neg}\leftarrow$\textbf{hasattr}(label,negative)}
\DefMacro{lc_6_predicates4}{postfix$:=$``but it wasn't''|``but I didn't''}
\DefMacro{lc_7_desc}{Negated positive with neutral content in the middle}
\DefMacro{lc_7_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix,$S_{neutral}$,infix,$S_{pos}$)}
\DefMacro{lc_7_predicates2}{prefix$:=$``I wouldn't say,''|``I do not think,''|``I don't agree with''}
\DefMacro{lc_7_predicates3}{$S_{neutral}\leftarrow$\textbf{hasattr}(length,<20)$\land$\textbf{hasattr}(label,neutral)}
\DefMacro{lc_7_predicates4}{infix$:=$`,'}
\DefMacro{lc_7_predicates5}{$S_{pos}\leftarrow$\textbf{hasattr}(length,<20)$\land$\textbf{hasattr}(label,positive)}
\DefMacro{lc_8_desc}{Author sentiment is more important than of others}
\DefMacro{lc_8_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix1,infix,$S_{neg}$)$+$\textbf{combine}(prefix2,infix,$S_{pos}$)}
\DefMacro{lc_8_predicates2}{prefix1$:=$``Some people think that''|``Many people agree with that''|``They think that''|``You agree with that''}
\DefMacro{lc_8_predicates3}{infix$:=$``but I think that''}
\DefMacro{lc_8_predicates4}{$S_{neg}\leftarrow$\textbf{hasattr}(label,negative)}
\DefMacro{lc_8_predicates5}{prefix2$:=$``Some people think that''|``Many people agree with that''|``They think that''|``You agree with that''}
\DefMacro{lc_8_predicates6}{$S_{pos}\leftarrow$\textbf{hasattr}(label,positive)}
\DefMacro{lc_9_desc}{Parsing sentiment in (question, yes) form}
\DefMacro{lc_9_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix,$S_{pos}$,postfix)$+$\textbf{combine}(prefix,$S_{neg}$,postfix)}
\DefMacro{lc_9_predicates2}{prefix$:=$``Do I think that''|``Do I agree that''}
\DefMacro{lc_9_predicates3}{$S_{pos}\leftarrow$\textbf{hasattr}(label,positive)}
\DefMacro{lc_9_predicates4}{postfix$:=$`? yes'}
\DefMacro{lc_9_predicates5}{$S_{neg}\leftarrow$~\textbf{hasattr}(label,negative)}
\DefMacro{lc_10_desc}{Parsing sentiment in (question, no) form}
\DefMacro{lc_10_predicates1}{$Seed\leftarrow$\textbf{combine}(prefix,$S_{pos}$,postfix)$+$\textbf{combine}(prefix,$S_{neg}$,postfix)}
\DefMacro{lc_10_predicates2}{prefix$:=$``Do I think that''|``Do I agree that''}
\DefMacro{lc_10_predicates3}{$S_{pos}\leftarrow$\textbf{hasattr}(label,positive)}
\DefMacro{lc_10_predicates4}{postfix$:=$`? no'}
\DefMacro{lc_10_predicates5}{$S_{neg}\leftarrow$~\textbf{hasattr}(label,negative)}